### 2021-10-05 paper/GHRVis.tex
feat: add a paper skeleton

% ---------------------------------------------------------------------------
% Author guideline and sample document for EG publication using LaTeX2e input
% D.Fellner, v1.15, Dec 14, 2018

\documentclass{egpubl}
\usepackage{eurovis2021}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
\SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
% \WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

% --- for EuroVis 
% for full papers use \SpecialIssuePaper
% \STAREurovis   % for EuroVis additional material 
% \EuroVisPoster % for EuroVis additional material 
% \EuroVisShort  % for EuroVis additional material

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------
\usepackage[T1]{fontenc}
\usepackage{dfadobe}

% \usepackage{cite}  % comment out for biblatex with backend=biber
% ---------------------------
\biberVersion
\BibtexOrBiblatex
\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true]{biblatex}
%\addbibresource{egbibsample.bib}
% ---------------------------  
\electronicVersion
\PrintedOrElectronic
% for including postscript figures
% mind: package option 'draft' will replace PS figure by a filename within a frame
\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
\else \usepackage[dvips]{graphicx} \fi

\usepackage{egweblnk}
% end of prologue

\addbibresource{LetterVis.bib}

\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}

\usepackage{dblfloatfix}

\usepackage{enumitem}

\usepackage[table, dvipsnames]{xcolor}

% \author[Q. Wang \& R. S. Laramee]
% {\parbox{\textwidth}{\centering Q. Wang
%   and R. S. Laramee
%     }
%     \\
% \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
% }

% \author[Q. Wang \& R. S. Laramee \& A. S. Lacey \& W. O. Pickrell]
% {\parbox{\textwidth}
%   {\centering Q. Wang$^{1}$, R. S. Laramee$^{1}$, A. Lacey$^{2,3}$ \& W. O. Pickrell$^{2,4}$}
%   \\
%   {\parbox{\textwidth}
%     {\centering $^1$School of Computer Science, University of Nottingham, UK\\
%         $^2$Neurology and Molecular Neuroscience Group, Institute of Life Science, Swansea University, UK\\
%         $^3$Population Data Science, Swansea University Medical School, Swansea University, UK\\
%         $^4$Neurology Department, Morriston Hospital, Swansea Bay University Healthboard, UK
%     }
%   }
% }

\title{GHRVis:}

\begin{document}

\pagestyle{plain}

% auto labeling of sections
\let\origsection=\section
\renewcommand\section[1]{\origsection{#1}\label{sec:{#1}}}

\let\origsubsection=\subsection
\renewcommand\subsection[1]{\origsubsection{#1}\label{subsec:{#1}}}

\let\origsubsubsection=\subsubsection
\renewcommand\subsubsection[1]{\origsubsubsection{#1}\label{subsubsec:{#1}}}

\newcommand{\figureref}[1]{Figure \ref{#1}}
\newcommand{\tableref}[1]{Table \ref{#1}}
\newcommand{\phantomref}[1]{$\hyperref[#1]{\ref{#1}}$}

% Bob's dream paragraph style
\newcommand{\bobgraph}[1]{ \noindent\textbf{#1} \phantomsection \label{bobgraph:{#1}}}

\definecolor{Mycolor1}{HTML}{4C5866} %#4C5866
\definecolor{Mycolor2}{HTML}{A6B6D2} %#A6B6D2
\definecolor{Mycolor3}{HTML}{C1D4E8} %#C1D4E8
\definecolor{Mycolor4}{HTML}{F7F7FE} %#F7F7FE
\definecolor{Mycolor5}{HTML}{d8f3dc} %#d8f3dc
\definecolor{Mycolor6}{HTML}{EC6A5C} %#EC6A5C
\newcommand{\mycolorbox}[2]{\begingroup\setlength{\fboxsep}{2pt}\colorbox{#1}{#2}\endgroup}

\newcommand{\note}[1]{\textcolor{blue}{#1}}

\newcommand{\expertcomment}[1]{\textit{``#1''}}

\newcommand{\task}[1]{\textbf{T#1}}
\newcommand{\req}[1]{\textbf{R#1}}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction and Motivation}

\section{Related Work}

\section{Data Description}

\section{GHRVis}

\section{Evaluation}

\section{Limitations and Future Work}

\section{Conclusions}

\section{Acknowledgments}
This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC). EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.

% reset \section command for section title "References"
\let\section=\origsection
\printbibliography

\end{document}

### 2021-12-05 paper/section/related.tex
feat: add a table for dorling papers

\section{Related Work}
\begin{table*}[!tb]
	\centering
	\resizebox{\textwidth}{!}{
		\begin{tabulary}{\textwidth}{|*{5}{c|}}
			\hhline{~|*{3}{-}|~}
			\multicolumn{1}{c|}{\textbf{Literature}} &
			\textbf{Title} &
			\textbf{Geographic Region} &
			\textbf{Number of Nodes} &
			\multicolumn{1}{c}{\textbf{Year}} \\
			\hline

			\citea{dorling2011Area} & \citetitlea{dorling2011Area} &  &   \cellcolor{Mycolor6} & 2021 \\
			\hline

			% \multicolumn{1}{c|}{\textbf{Total unique papers: 51}} & 24&15&12& \multicolumn{1}{c}{}  \\
			% \hhline{~|*{3}{-}|~}

		\end{tabulary}
	}
	\caption{}
	\label{table:challenge table}
\end{table*}

### 2021-12-09 paper/section/related.tex
feat: update dorling table

@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabulary}{\textwidth}{|*{5}{c|}}
+		\begin{tabulary}{\textwidth}{|*{3}{l|}r|c|}
 			\hhline{~|*{3}{-}|~}
@@
 
-			\citea{dorling2011Area} & \citetitlea{dorling2011Area} &  &   \cellcolor{Mycolor6} & 2021 \\
+			% \citea{auber2007Geographical} & \citetitlea{auber2007Geographical} & US & 19 & 2007 \\
+			% \hline
+			\citea{warf2008Geography} & \citetitlea{warf2008Geography} & US & 3,142 & \citeyear{warf2008Geography} \\
+			\hline
+			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+			\hline
+			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+			\hline
+			\citea{tong2018cartograms} & \citetitlea{tong2018cartograms} & England & 209 & \citeyear{tong2018cartograms} \\
+			\hline
+			\citea{gao2020Visualising} & \citetitlea{gao2020Visualising} & China & 34 & \citeyear{gao2020Visualising} \\
+			\hline
+			\citea{nusrat2020Recognition} & \citetitlea{nusrat2020Recognition} & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
 			\hline
@@
 	}
-	\caption{}
+	\caption{Discuss Demers cartogram}
 	\label{table:challenge table}

### 2022-01-02 paper/section/related.tex
draft: related work

@@
 \section{Related Work}
+
+Cartograms combine statistical and geographical information in thematic maps.
+We follow the categorization of cartograms by \citeauthor{vankreveld2004Rectangular} \cite{vankreveld2004Rectangular}: contiguous, non-contiguous, rectangular, and Dorling.
+\citeauthor{nusrat2016State} \cite{nusrat2016State} summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+Each cartogram design may make accuracy trade-offs between dimensions, we provide an overview in Table \ref{table:accuracy}.
+
+{
+\renewcommand{\arraystretch}{1.5}
+\begin{table*}[!tb]
+	\centering
+	\resizebox{\textwidth}{!}{
+		\begin{tabulary}{\textwidth}{|*{5}{l|}}
+			\hhline{~|*{3}{-}|~}
+			\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{\cellcolor{Mycolor2} \textbf{Accuracy}} \\
+			\hhline{~|*{4}{-}}
+			\multicolumn{1}{c|}{\textbf{Catogram Type}} &
+			\textbf{Statistical} &
+			\textbf{Geographical} &
+			\textbf{Topological} &
+			\textbf{Contiguity} \\
+			\hline
+			Contiguous & \cellcolor{Mycolor3}Varies & \cellcolor{Mycolor3}Varies & Accurate & Yes \\
+			\hline
+			Non-contiguous & \cellcolor{Mycolor5}\textcolor{white}{Accurate} & \cellcolor{Mycolor5}\textcolor{white}{Shape is accurate} &\cellcolor{Mycolor6}Inaccurate & No   \\
+			\hline
+			Rectangular & \cellcolor{Mycolor3}Varies & Shape is inaccurate & \cellcolor{Mycolor3}Varies & Yes   \\
+			\hline
+			Dorling & Accurate & Inaccurate & Inaccurate & No  \\
+			\hline
+		\end{tabulary}
+	}
+	\caption{\textcolor{Mycolor3}{\textbf{Trade-off between dimensions}}.
+\textcolor{Mycolor6}{\textbf{Dimension sacrificed}} in order to improve \textcolor{Mycolor5}{\textbf{target dimension}}'s accuracy.}
+	\label{table:accuracy}
+\end{table*}
+}
+
+Dorling cartograms are non-contiguous and do not preserve geography and topology.
+A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size.
+In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology.
+Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.
+Table \ref{table:region vs node} includes a list of literature that adopts Dorling cartograms with corresponding geographical regions and node counts.
+
+{
+\renewcommand{\arraystretch}{1.5}
 \begin{table*}[!tb]
@@
 	}
-	\caption{Discuss Demers cartogram}
-	\label{table:challenge table}
+	\caption{}
+	\label{table:region vs node}
 \end{table*}
+}
+
+Rectangular cartograms are contiguous and do not preserve geographical accuracy. Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.
+
+Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy. Some variants are able to preserve topological accuracy as well.

### 2022-01-14 src/router/index.js
feat: add CCG2

@@ -13,10 +13,12 @@ const routes = [
   {
     path: "/ccg",
     name: "CCG",
-    // route level code-splitting
-    // this generates a separate chunk (about.[hash].js) for this route
-    // which is lazy-loaded when the route is visited.
     component: () => import(/* webpackChunkName: "US" */ "../views/CCG.vue"),
+  },
+  {
+    path: "/ccg2",
+    name: "CCG2",
+    component: () => import(/* webpackChunkName: "US" */ "../views/CCG2.vue"),
   },
 ];
 

### 2022-01-16 paper/section/related.tex
draft: related work

--- 2021-12-05 paper/section/related.tex
@@
 \section{Related Work}
+
+Cartograms combine statistical and geographical information in thematic maps.
+We follow the categorization of cartograms by \citea{vankreveld2004Rectangular}: contiguous, non-contiguous, rectangular, and Dorling.
+\citea{nusrat2016State} summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+Each cartogram design may make accuracy trade-offs between dimensions, we provide an overview in Table \ref{table:accuracy}.
+
+{
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[!b]
+	\centering
+	\resizebox{\columnwidth}{!}{
+		\begin{tabulary}{\columnwidth}{|*{5}{l|}}
+			\hhline{~|*{3}{-}|~}
+			\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{\cellcolor{Mycolor2} \textbf{Accuracy}} \\
+			\hhline{~|*{4}{-}}
+			\multicolumn{1}{c|}{\textbf{Catogram Type}} &
+			\textbf{Statistical} &
+			\textbf{Geographical} &
+			\textbf{Topological} &
+			\textbf{Contiguity} \\
+			\hline
+			Contiguous & \cellcolor{Mycolor3}Variable & \cellcolor{Mycolor3}Variable & Accurate & Yes \\
+			\hline
+			Non-contiguous & \cellcolor{Mycolor5}\textcolor{white}{Accurate} & \cellcolor{Mycolor5}\textcolor{white}{Shape is accurate} &\cellcolor{Mycolor6}Inaccurate & No   \\
+			\hline
+			Rectangular & \cellcolor{Mycolor3}Variable & Shape is inaccurate & \cellcolor{Mycolor3}Variable & Yes   \\
+			\hline
+			Dorling & Accurate & Inaccurate & Inaccurate & No  \\
+			\hline
+		\end{tabulary}
+	}
+	\caption{\textcolor{Mycolor3}{\textbf{Trade-off between dimensions}}.
+\textcolor{Mycolor6}{\textbf{Dimension sacrificed}} in order to improve \textcolor{Mycolor5}{\textbf{target dimension}}'s accuracy.}
+	\label{table:accuracy}
+\end{table}
+}
+
+Dorling cartograms are non-contiguous and do not preserve geography and topology.
+A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size \cite{dorling2011Area}.
+In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
+Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.
+
+Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular}.
+Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.
+
+Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+Some variants are able to preserve topological accuracy as well.
+
+\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
+\citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms.
+\citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
+\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
+\citea{tong2018cartograms} use a Dorling cartogram to visualize health-related date by regions in England.
+\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, Agriculture, and Retail data in the US.
+See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+
+{
+\renewcommand{\arraystretch}{1.5}
 \begin{table*}[!tb]
@@
 	\resizebox{\textwidth}{!}{
-		\begin{tabulary}{\textwidth}{|*{5}{c|}}
-			\hhline{~|*{3}{-}|~}
+		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
+			\hhline{~|*{4}{-}|~}
 			\multicolumn{1}{c|}{\textbf{Literature}} &
 			\textbf{Title} &
-			\textbf{Geographic Region} &
+			\textbf{Cartogram Type(s)} &
+			\textbf{Geographic Region(s)} &
 			\textbf{Number of Nodes} &
@@
 
-			\citea{dorling2011Area} & \citetitlea{dorling2011Area} &  &   \cellcolor{Mycolor6} & 2021 \\
+			% \citea{auber2007Geographical} & \citetitlea{auber2007Geographical} & US & 19 & 2007 \\
+			% \hline
+			\citea{warf2008Geography} & \citetitlea{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
+			\hline
+			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & Dorling, Mosaic, Contiguous & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+			\hline
+			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Dorling, Non-contiguous, Contiguous & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+			\hline
+			\citea{tong2018cartograms} & \citetitlea{tong2018cartograms} & Demers & England & 209 & \citeyear{tong2018cartograms} \\
+			\hline
+			\citea{gao2020Visualising} & \citetitlea{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
+			\hline
+			\citea{nusrat2020Recognition} & \citetitlea{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
 			\hline
@@
 	\caption{}
-	\label{table:challenge table}
+	\label{table:region vs node}
 \end{table*}
+}

### 2022-01-16 paper/section/related.tex
draft: related work: media

@@
 \section{Related Work}
+
+This section introduces the definition of cartogram types, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
+
+\subsection{Definitions}
 
@@
 
+\subsection{Peer-reviewed Applications}
+
 \citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
@@
 \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Dorling cartogram to visualize health-related date by regions in England.
-\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, Agriculture, and Retail data in the US.
+\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
 See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
@@
 }
+
+\subsection{Other Applications}
+
+Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news.
+\citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
+The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
+The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
+\citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

### 2022-01-23 paper/GHRVis.tex
feat: move all nodes and move back nodeX
feat: add a table for dorling papers
draft: related work
draft: related work: media
draft: data

@@
-% ---------------------------------------------------------------------------
-% Author guideline and sample document for EG publication using LaTeX2e input
-% D.Fellner, v1.15, Dec 14, 2018
-
 \documentclass{egpubl}
@@
 % ------------------------------------------------------------------------
+
 \usepackage[T1]{fontenc}
@@
 \BibtexOrBiblatex
-\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true]{biblatex}
+\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true,url=false]{biblatex}
 %\addbibresource{egbibsample.bib}
+
+\renewbibmacro*{doi+eprint+url}{%
+    \printfield{doi}%
+    \newunit\newblock%
+    \iftoggle{bbx:eprint}{%
+        \usebibmacro{eprint}%
+    }{}%
+    \newunit\newblock%
+    \iffieldundef{doi}{%
+        \usebibmacro{url+urldate}}%
+        {}%
+    }
+
+\addbibresource{GeoVis.bib}
+
+\newcommand{\citetitlea}[1]{\citefield{#1}{title}}
+
 % ---------------------------  
@@
 
-\addbibresource{LetterVis.bib}
+\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}
 
-\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}
+% code listing
+\usepackage{listings}
+\lstdefinestyle{customc}{
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  xleftmargin=\parindent,
+  language=C,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  keywordstyle=\bfseries\color{green!40!black},
+  commentstyle=\itshape\color{purple!40!black},
+  identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+
+\lstdefinestyle{customasm}{
+  belowcaptionskip=1\baselineskip,
+  xleftmargin=\parindent,
+  language=[x86masm]Assembler,
+  basicstyle=\footnotesize\ttfamily,
+  commentstyle=\itshape\color{purple!40!black},
+}
+
+\lstset{escapechar=@,style=customc}
 
@@
 
-\usepackage{enumitem}
-
 \usepackage[table, dvipsnames]{xcolor}
 
-% \author[Q. Wang \& R. S. Laramee]
-% {\parbox{\textwidth}{\centering Q. Wang
-%   and R. S. Laramee
-%     }
-%     \\
-% \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
-% }
-
-% \author[Q. Wang \& R. S. Laramee \& A. S. Lacey \& W. O. Pickrell]
-% {\parbox{\textwidth}
-%   {\centering Q. Wang$^{1}$, R. S. Laramee$^{1}$, A. Lacey$^{2,3}$ \& W. O. Pickrell$^{2,4}$}
-%   \\
-%   {\parbox{\textwidth}
-%     {\centering $^1$School of Computer Science, University of Nottingham, UK\\
-%         $^2$Neurology and Molecular Neuroscience Group, Institute of Life Science, Swansea University, UK\\
-%         $^3$Population Data Science, Swansea University Medical School, Swansea University, UK\\
-%         $^4$Neurology Department, Morriston Hospital, Swansea Bay University Healthboard, UK
-%     }
-%   }
-% }
+\author[Q. Wang \& R. S. Laramee]
+{\parbox{\textwidth}{\centering Q. Wang
+  and R. S. Laramee
+    }
+    \\
+\parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
+}
 
@@
 \definecolor{Mycolor2}{HTML}{A6B6D2} %#A6B6D2
-\definecolor{Mycolor3}{HTML}{C1D4E8} %#C1D4E8
+\definecolor{Mycolor3}{HTML}{67daff} %#67daff
 \definecolor{Mycolor4}{HTML}{F7F7FE} %#F7F7FE
-\definecolor{Mycolor5}{HTML}{d8f3dc} %#d8f3dc
-\definecolor{Mycolor6}{HTML}{EC6A5C} %#EC6A5C
+\definecolor{Mycolor5}{HTML}{666ad1} %#666ad1
+\definecolor{Mycolor6}{HTML}{ffbb93} %#ffbb93
 \newcommand{\mycolorbox}[2]{\begingroup\setlength{\fboxsep}{2pt}\colorbox{#1}{#2}\endgroup}
+
+% table cite
+\newcommand{\citea}[1]{\citeauthor{#1}\cite{#1}}
+
+% customized table headers
+\newcommand{\theaderR}[1]{\multicolumn{1}{c|}{\cellcolor{Mycolor3} \textbf{\rotatebox{90}{#1 \phantom{-} }}}}
+\newcommand{\theaderP}[1]{\cellcolor{Mycolor3} \textbf{#1}}
 
@@
 
-\section{Introduction and Motivation}
+\input{section/introduction}
 
-\section{Related Work}
+\input{section/related}
 
-\section{Data Description}
+\input{section/data}
 
 \section{GHRVis}
+
+\subsection{VPSC}
 

### 2022-01-23 paper/section/data.tex
draft: data

\section{Data Description}

Obtaining the data used in this paper is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}. The data acquisition involves multiple complicated steps. The first step is to obtain both geospaital data and EHR data that are compatible for projection. The second step is to pre-process the data for removing empty and erroneous values. The final step is to transform the data into a format that is suitable for visualization.

\subsection{Geospatial Data}

Geospatial data, commonly known as shapefiles, were obtained from the following sources. After acquiring these data, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile. Finally, mapshaper \cite{blochMapshaper} is used to converted the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size in order to improve the performance of our system.

\subsubsection{Clinical Commissioning Groups}

Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs is changing each year due to the NHS annual re-organization, the most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 as at the time of writing, there is no EHR data published based on the latest CCG re-organization took place in April 2021.

\subsubsection{River Thames}

We first obtained the relation ID (2263653) for River Thames from OpenStreetMap \cite{openstreetmapRelation}, then downloaded the entire shapefile via a query (See Listing~\ref{overpass}) through the Overpass Turbo \cite{overpassturboOverpass}.

\begin{lstlisting}[caption={The Overpass Turbo query that downloads the shapefile of River Thames from OpenStreetMap.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    out skel;
\end{lstlisting}

\subsection{EHR Data}

We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}. The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.

### 2022-02-02 paper/section/data.tex
draft: EHR data description

@@
 
-\subsubsection{River Thames}
+\subsubsection{Rivers}
 
-We first obtained the relation ID (2263653) for River Thames from OpenStreetMap \cite{openstreetmapRelation}, then downloaded the entire shapefile via a query (See Listing~\ref{overpass}) through the Overpass Turbo \cite{overpassturboOverpass}.
+We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Ouse in England.
+These rivers were chosen as they pass through regions with clusters of CCGs, and provide informative geographical and topological cues.
 
-\begin{lstlisting}[caption={The Overpass Turbo query that downloads the shapefile of River Thames from OpenStreetMap.}, label={overpass},captionpos=b]
+We first obtained a relation ID for the shapefile from OpenStreetMap, then downloaded the entire shapefile via a query (See Listing~\ref{overpass}) through Overpass Turbo \cite{overpassturboOverpass}.
+
+\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via Overpass Turbo API.}, label={overpass},captionpos=b]
     relation(2263653);>>;
@@
 The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.
+Some datasets include:
+\begin{itemize}
+    \item Under 75 mortality
+    \begin{multicols}{2}
+        \begin{itemize}
+            \item Cardiovascular disease
+            \item Respiratory disease
+            \item Liver disease
+            \item Cancer
+        \end{itemize}
+    \end{multicols}
+    \item Emergency hospital admission
+        \begin{itemize}
+            \item Stoke
+            \item Alcohol-specific admission and readmission
+            \item Coronary heart disease
+            \item Readmissions within 30 days of discharge
+            \item Children with lower respiratory tract infections
+        \end{itemize}
+\end{itemize}
+
+For all datasets, a spreadsheet including the following data is provided:
+
+\begin{itemize}
+    \item Reporting period: Calendar year of registration
+    \item Period of coverage: Start and end date or reporting period
+    \item Breakdown: Organization type
+    \item ONS code: UK Office for National Statistics CCG code
+    \item Level: CCG Code
+    \item Level description: CCG Name
+    \item Gender
+    \item Indicator value: Directly standardized mortality rate
+    \item CI lower: lower 95\% confidence interval
+    \item CI upper: upper 95\% confidence interval
+    \item Denominator: The count of registered patients
+    \item Numerator: Number of deaths
+\end{itemize}

### 2022-02-10 paper/section/data.tex
draft: add algorithms

@@
 After acquiring these data, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile.
-Finally, mapshaper \cite{blochMapshaper} is used to converted the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size in order to improve the performance of our system.
+Finally, mapshaper \cite{blochMapshaper} is used to converted the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size in order to improve the performance of our software.
 
@@
 
-We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Ouse in England.
+We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
 These rivers were chosen as they pass through regions with clusters of CCGs, and provide informative geographical and topological cues.
 
-We first obtained a relation ID for the shapefile from OpenStreetMap, then downloaded the entire shapefile via a query (See Listing~\ref{overpass}) through Overpass Turbo \cite{overpassturboOverpass}.
+We first obtained a relation ID by searching a river, e.g.
+River Thames, on OpenStreetMap.
+The obtained relation ID was used to construct a query (See Listing~\ref{overpass}) which enabled use to download the entire river shapefile through Overpass Turbo \cite{overpassturboOverpass}.
 

### 2022-02-10 paper/section/main.tex
draft: add algorithms

\section{\software}

\subsection{Node Overlap Removal Algorithm}

% \begin{noindent}
\begin{algorithm}[hbt!]
    \caption{An overview of all steps in \software.}\label{alg:overview}
    \textbf{Variables:} \\
    $NodeList \gets$ a list of nodes representing CCGs \\
    $Node_{size}\gets$ the current size of all nodes \\
    $Node_{inc} \gets$ the increment node size of each iteration \\
    $Node_{max} \gets$ the maximum size of a node \\

    \begin{algorithmic}[1]
        \While{$ Node_{size} < Node_{max} $}
            \State RemoveOverlap ($ NodeList $)
            \State $Node_{size} \gets Node_{size} + Node_{inc}$
        \EndWhile
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to remove node overlaps.}\label{alg:remove overlap}
    \textbf{Variables:} \\
    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
    \begin{algorithmic}[1]
        \Function{RemoveOverlap}{$ NodeList $}
        \State $ recurse \gets False$
        \State $ nodeList_{ora} \gets $ RunOverlapRemoval ($ NodeList $)
            \For{$ node^i_{ora} \in nodeList_{ora} $}
                \If {MoveNode ($ Node^i, node^i_{ora} $)}
                    \State $ cross \gets $ CheckRiverCrossing ($ Node^i, node^i_{ora} $)
                    \If{$ cross = True $}
                        \State $recurse \gets True$
                        \State $ Node^i.stale \gets Node^i.stale + 1 $
                        \If{$ Node^i.stale < Node_{stale} $}
                            \State MoveNode ($ Node^i $)
                        \Else
                            \State DeriveCorridor ($ Node^i, node^i_{ora} $)
                        \EndIf
                    \EndIf
                \EndIf
            \EndFor
            \If{$ recurse = True $}
                \State RemoveOverlap ($ NodeList $)
            \EndIf
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to update a node's position.}\label{alg:move position}
    \begin{algorithmic}[1]
        \Function{MoveNode}{$ Node, Node_{ora} $}
        \If{$ Node.x = Node_{ora}.x $ and $ Node.y = Node_{ora}.y $}
            \State \Return{$ False $}
        \EndIf

        \State init. $ x,y $

        \If{$ Node_{ora}$ is $ undefined $}
            \State $previous \gets Node.history.pop() $
            \State $ x \gets previous.x,~y \gets previous.y $
        \Else
            \State $ x \gets Node_{ora}.x,~y \gets Node_{ora}.y $
            \State $ Node.history.push(x, y) $
        \EndIf

        \State $ Node.x \gets x $
        \State $ Node.y \gets y $
        \State \Return{$ True $}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to check if a node crosses a river.}\label{alg:check river crossing}
    \textbf{Variables:} \\
    $RiverEdges \gets$ a list of river edges \\

    \begin{algorithmic}[1]
        \Function{CheckRiverCrossing}{$ Node, Node_{ora} $}
            \For{$ edge \in RiverEdges $}
                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
                \If{$ collide = True $}
                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
                    \State $ edge_l \gets $ GenerateLine ($ edge $)
                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
                    \If{$ intersect = True $}
                        \State \Return{$ True $}
                    \EndIf
                \EndIf
            \EndFor
            \State \Return{$ False $}
        \EndFunction
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[hbt!]
    \caption{A function to derive a corridor to include and move neighboring nodes. On a SVG canvas, the point of origin (0,0) is at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
    \textbf{Variables:} \\
    $C_l \gets$ the length of a corridor \\
    $C_{w} \gets$ the width of a corridor \\
    $Node_{size}\gets$ the current size of all nodes \\
    \begin{algorithmic}[1]
        \Function{DeriveCorridor}{$ Node, Node_{ora} $}
        \State MoveNode ($ Node $)

        \State $ \Delta x \gets Node.x - Node_{ora}.x,~ \Delta y \gets Node.y - Node_{ora}.y $

        \State $ \vec{diff} \gets
            \begin{pmatrix}
                \Delta x \\
                \Delta y
            \end{pmatrix},~ \vec{Node} \gets
            \begin{pmatrix}
                    Node.x \\
                    Node.y
            \end{pmatrix} $

        \State $ \vec{p} \gets
            \vec{Node} + \sqrt{
                C_{l}^2 \cdot (\Delta x + \Delta y))
            } \cdot \vec{diff} $

        \State $ shift \gets
                \sqrt{ 
                    \frac{C_{w}^2}{2}
                },~ \vec{shift_a} \gets \begin{pmatrix}
                    shift \\
                    - shift
                \end{pmatrix},~ \vec{shift_b} \gets \begin{pmatrix}
                    - shift \\
                    +shift
                \end{pmatrix} $

        \State $ corridor \gets
            \begin{bmatrix}
                \vec{Node} + \vec{shift_a} &
                \vec{p}+ \vec{shift_a} \\

                \vec{Node} + \vec{shift_b} &
                \vec{p}+ \vec{shift_b} \\
            \end{bmatrix} $

        \For{$ node^i $ inside $ corridor $}
            \State MoveNode ($ node^i,  $)
        \EndFor

        \EndFunction
    \end{algorithmic}
\end{algorithm}
%\end{noindent}

### 2022-02-16 paper/section/main.tex
draft: update main section

@@
 
-\subsection{Node Overlap Removal Algorithm}
+\algoref{alg:UpdateNodePosition} describes the overview of \software.
+We first load all processed data from \secref{sec:{Data Description}} and apply VPSC \cite{dwyer2006fast} recursively to remove overlaps.
+We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+During each recursion, we check node trajectories (See \algoref{alg:check river crossing}) and move back nodes that crossed a river.
+A recursion ends when 1) no overlaps are present; 2) no nodes crossed a river.
+We then increase the node size and repeat the recursion until the entire map canvas is filled.
 
 % \begin{noindent}
-\begin{algorithm}[hbt!]
-    \caption{An overview of all steps in \software.}\label{alg:overview}
+
+\begin{algorithm}[tb!]
+    \caption{The procedure to update node positions by removing overlaps and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
     \textbf{Variables:} \\
@@
     $Node_{max} \gets$ the maximum size of a node \\
+    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
+    \begin{algorithmic}[1]
+        \Procedure{UpdateNodePosition}{}
+        \While{$ Node_{size} < Node_{max} $}
 
-    \begin{algorithmic}[1]
-        \While{$ Node_{size} < Node_{max} $}
-            \State RemoveOverlap ($ NodeList $)
+        \State $ nodeList_{ora} \gets $ RunVPSC ($ NodeList $)
+
+            \For{$ node^i_{ora} \in nodeList_{ora} $}
+
+                \If {MoveNode ($ Node^i, node^i_{ora} $)}
+                    \State $ cross \gets $ \Call{CheckRiverCrossing}{$ Node^i, node^i_{ora} $}
+
+                    \If{$ cross = True $}
+                        \State $ Node^i.stale \gets Node^i.stale + 1 $
+
+                        \If{$ Node^i.stale < Node_{stale} $}
+                            \State \Call{MoveNode}{$ Node^i $}
+                        \Else
+                            \State \Call{DeriveCorridor}{$ Node^i, node^i_{ora} $}
+                        \EndIf
+
+                    \EndIf
+
+                \EndIf
+
+            \EndFor
+
+        \EndWhile
+
         \State $Node_{size} \gets Node_{size} + Node_{inc}$
-        \EndWhile
+
+        \EndProcedure
     \end{algorithmic}
@@
 
-\begin{algorithm}[hbt!]
-    \caption{A function to remove node overlaps.}\label{alg:remove overlap}
-    \textbf{Variables:} \\
-    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
+\begin{algorithm}[tb!]
+    \caption{The procedure to update a node's position.}\label{alg:move position}
     \begin{algorithmic}[1]
-        \Function{RemoveOverlap}{$ NodeList $}
-        \State $ recurse \gets False$
-        \State $ nodeList_{ora} \gets $ RunOverlapRemoval ($ NodeList $)
-            \For{$ node^i_{ora} \in nodeList_{ora} $}
-                \If {MoveNode ($ Node^i, node^i_{ora} $)}
-                    \State $ cross \gets $ CheckRiverCrossing ($ Node^i, node^i_{ora} $)
-                    \If{$ cross = True $}
-                        \State $recurse \gets True$
-                        \State $ Node^i.stale \gets Node^i.stale + 1 $
-                        \If{$ Node^i.stale < Node_{stale} $}
-                            \State MoveNode ($ Node^i $)
-                        \Else
-                            \State DeriveCorridor ($ Node^i, node^i_{ora} $)
-                        \EndIf
-                    \EndIf
-                \EndIf
-            \EndFor
-            \If{$ recurse = True $}
-                \State RemoveOverlap ($ NodeList $)
-            \EndIf
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}[hbt!]
-    \caption{A function to update a node's position.}\label{alg:move position}
-    \begin{algorithmic}[1]
-        \Function{MoveNode}{$ Node, Node_{ora} $}
-        \If{$ Node.x = Node_{ora}.x $ and $ Node.y = Node_{ora}.y $}
+        \Procedure{MoveNode}{$ Node, Node_{new} $}
+        \If{$ Node.x = Node_{new}.x $ and $ Node.y = Node_{new}.y $}
             \State \Return{$ False $}
@@
 
-        \If{$ Node_{ora}$ is $ undefined $}
+        \If{$ Node_{new}$ is $ undefined $}
             \State $previous \gets Node.history.pop() $
@@
         \Else
-            \State $ x \gets Node_{ora}.x,~y \gets Node_{ora}.y $
+            \State $ x \gets Node_{new}.x,~y \gets Node_{new}.y $
             \State $ Node.history.push(x, y) $
@@
         \State \Return{$ True $}
-        \EndFunction
+        \EndProcedure
     \end{algorithmic}
@@
 
-\begin{algorithm}[hbt!]
-    \caption{A function to check if a node crosses a river.}\label{alg:check river crossing}
+%\end{noindent}
+
+\subsection{River Cross Checking}
+
+We use rivers as topological boundaries to prevent nodes from crossing them.
+When a node's position changes, we check if the node's trajectory intersects any segment of a river, see \algoref{alg:check river crossing}.
+A boundary collision detection is performed to reduce the number of line intersection detections required.
+
+% \begin{noindent}
+\begin{algorithm}[tb!]
+    \caption{The procedure to check if a node crosses a river.}\label{alg:check river crossing}
     \textbf{Variables:} \\
-    $RiverEdges \gets$ a list of river edges \\
+    $RiverEdges \gets$ a list of river edges formed by two points \\
 
     \begin{algorithmic}[1]
-        \Function{CheckRiverCrossing}{$ Node, Node_{ora} $}
+        \Procedure{CheckRiverCrossing}{$ Node, Node_{ora} $}
             \For{$ edge \in RiverEdges $}
@@
             \State \Return{$ False $}
-        \EndFunction
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+%\end{noindent}
+
+\subsection{Corridor Derivation}
+
+As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly traveling between two positions due to the lack of available space, creating a stalemate situation for recursions, as shown in \figref{fig:stalemate}.
+In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ Node_{stale} $, to break the stalemate: if a node is traveling between two positions for more than $ Node_{stale} $ iterations, it is considered to be in a stalemate.
+A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/stalemate.png}
+    \caption{A stalemate situation is when a node repeatedly travels between two positions (A and B) for more than $Node_{stale}$ iterations. }
+    \label{fig:stalemate}
+\end{figure}
+}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/corridor.png}
+    \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}.
+All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \vec{BA} $.
+}
+    \label{fig:corridor}
+\end{figure}
+}
+
+% \begin{noindent}
+
+\begin{algorithm}[tb!]
+    \caption{The procedure to derive a corridor to move included nodes.
+We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
+    \textbf{Variables:} \\
+    $\mathcal{C}_l \gets$ the length of a corridor \\
+    $\mathcal{C}_w \gets$ the width of a corridor \\
+    $\mathcal{D} \gets$ the distance to move nodes within the corridor \\
+    $Node_{size}\gets$ the current size of all nodes \\
+    \begin{algorithmic}[1]
+        \Procedure{DeriveCorridor}{$ Node, Node_{old} $}
+        \State MoveNode ($ Node $)
+
+        \State $ \Delta x \gets Node.x - Node_{old}.x,~ \Delta y \gets Node.y - Node_{old}.y $
+
+        \State $ slope \gets \frac{\Delta x}{\Delta y}$
+
+        \State $ pLine = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ Node $, $ \mathcal{C}_w $}
+        \State $ sideLine_1 = $ \Call{DerivePoint}{$ slope $, $ pLine.a $, $ \mathcal{C}_l $, $ True $}
+        \State $ sideLine_2 = $ \Call{DerivePoint}{$ slope $, $ pLine.b $, $ \mathcal{C}_l $, $ True $}
+        \State $ corridor \gets
+            \begin{bmatrix}
+                pLine.a &
+                pLine.b \\
+
+                sideLine_1.a &
+                sideLine_2.b \\
+            \end{bmatrix} $
+
+        \For{$ node^i $ inside $ corridor $}
+            \State $ destination = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ node^i $, $ \mathcal{D} $}
+
+            \State MoveNode ($ node^i, destination.b $)
+        \EndFor
+
+        \EndProcedure
+
+        \item[]
+
+        \Procedure{DerivePoint}{$ Slope, Point, Length, IsSideLine $}
+
+        \State init. $ a $, $ b $
+        \Switch{$Slope$}
+            \Case{$0$}
+                \If{$ IsSideLine = True $}
+                    \State $ a.x \gets Point.x $
+                \Else
+                    \State $ a.x \gets Point.x + Length $
+                \EndIf
+                \State $ a.y \gets Point.y $
+                \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
+            \EndCase
+            \Case{$ \infty $ or $ -\infty $}
+                \State $ a.x \gets Point.x $
+
+                \If{$ IsSideLine = True $}
+                    \State $ a.y \gets Point.y $
+                \Else
+                    \State $ a.y \gets Point.y + Length $
+                \EndIf
+
+                \State $ b.x \gets Point.x $,~$ b.y \gets Point.y - Length $
+            \EndCase
+            \Default
+                \State $ dx = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ dy = Slope \cdot dx $
+                \If{$ IsSideLine = True $}
+                    \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
+                \Else
+                    \State $ a.x \gets Point.x + dx $,~$ a.y \gets Point.y + dy $
+                \EndIf
+
+                \State $ b.x \gets Point.x - dx $,~$ b.y \gets Point.y - dy $
+            \EndDefault
+
+        \EndSwitch
+
+        \State \Return{$ a, b $}
+
+        \EndProcedure
+
     \end{algorithmic}
@@
 
-\begin{algorithm}[hbt!]
-    \caption{A function to derive a corridor to include and move neighboring nodes.
-On a SVG canvas, the point of origin (0,0) is at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
-    \textbf{Variables:} \\
-    $C_l \gets$ the length of a corridor \\
-    $C_{w} \gets$ the width of a corridor \\
-    $Node_{size}\gets$ the current size of all nodes \\
-    \begin{algorithmic}[1]
-        \Function{DeriveCorridor}{$ Node, Node_{ora} $}
-        \State MoveNode ($ Node $)
-
-        \State $ \Delta x \gets Node.x - Node_{ora}.x,~ \Delta y \gets Node.y - Node_{ora}.y $
-
-        \State $ \vec{diff} \gets
-            \begin{pmatrix}
-                \Delta x \\
-                \Delta y
-            \end{pmatrix},~ \vec{Node} \gets
-            \begin{pmatrix}
-                    Node.x \\
-                    Node.y
-            \end{pmatrix} $
-
-        \State $ \vec{p} \gets
-            \vec{Node} + \sqrt{
-                C_{l}^2 \cdot (\Delta x + \Delta y))
-            } \cdot \vec{diff} $
-
-        \State $ shift \gets
-                \sqrt{ 
-                    \frac{C_{w}^2}{2}
-                },~ \vec{shift_a} \gets \begin{pmatrix}
-                    shift \\
-                    - shift
-                \end{pmatrix},~ \vec{shift_b} \gets \begin{pmatrix}
-                    - shift \\
-                    +shift
-                \end{pmatrix} $
-
-        \State $ corridor \gets
-            \begin{bmatrix}
-                \vec{Node} + \vec{shift_a} &
-                \vec{p}+ \vec{shift_a} \\
-
-                \vec{Node} + \vec{shift_b} &
-                \vec{p}+ \vec{shift_b} \\
-            \end{bmatrix} $
-
-        \For{$ node^i $ inside $ corridor $}
-            \State MoveNode ($ node^i,  $)
-        \EndFor
-
-        \EndFunction
-    \end{algorithmic}
-\end{algorithm}
 %\end{noindent}

### 2022-02-16 paper/section/related.tex
draft: data
feat: 3 UK rivers with correct sorting
draft: EHR data description
draft: update main section

@@
 
-This section introduces the definition of cartogram types, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
+This section introduces the definition of various cartograms, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
 
@@
 
-Dorling cartograms are non-contiguous and do not preserve geography and topology.
-A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size \cite{dorling2011Area}.
-In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
-Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.
-
-Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular}.
-Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.
-
-Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
-Some variants are able to preserve topological accuracy as well.
-
-\subsection{Peer-reviewed Applications}
-
-\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
-\citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms.
-\citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
-\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy.
-\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
-See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
-
 {
 	\renewcommand{\arraystretch}{1.5}
-\begin{table*}[!tb]
+	\begin{table*}[!htb]
 		\centering
@@
 		}
-	\caption{}
+		\caption{region vs node}
 		\label{table:region vs node}
@@
 
+Dorling cartograms are non-contiguous and do not preserve geography and topology.
+A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size \cite{dorling2011Area}.
+In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
+Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.
+
+Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular}.
+Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.
+
+Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+Some variants are able to preserve topological accuracy as well.
+
+\subsection{Peer-reviewed Applications}
+
+\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
+\citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms.
+\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
+\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
+See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+
 \subsection{Other Applications}
@@
 Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news.
-\citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
 The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
 The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
+\cite{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
 \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.
+
+% \subsection{Contiguous Applications}
+
+% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
+
+Compared to previous work on Demers Cartogram, we introduce a topological feature, river, aiming to improve the geographical and topological accuracy.

### 2022-02-18 paper/section/data.tex
draft: update

@@
 
-Obtaining the data used in this paper is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}.
-The data acquisition involves multiple complicated steps.
-The first step is to obtain both geospaital data and EHR data that are compatible for projection.
-The second step is to pre-process the data for removing empty and erroneous values.
+Obtaining the data is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}.
+The data acquisition involves multiple steps.
+The first step is to obtain both geospatial boundaries and EHR data that match one another.
+The second step is to pre-process the EHR data for removing empty and erroneous values.
 The final step is to transform the data into a format that is suitable for visualization.
@@
 
-Geospatial data, commonly known as shapefiles, were obtained from the following sources.
-After acquiring these data, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile.
-Finally, mapshaper \cite{blochMapshaper} is used to converted the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size in order to improve the performance of our software.
+Geospatial boundaries, or shapefiles, were obtained from the following sources.
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile.
+Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size and improve the performance of our implementation.
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
+    \caption{A map of 106 CCGs in England as of April 2021, obtained from the Open Geography portalx \cite{opengeographyportalxOpen}.}
+    \label{fig:ccg}
+\end{figure}
+}
 
@@
 Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
-The number of CCGs is changing each year due to the NHS annual re-organization, the most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 as at the time of writing, there is no EHR data published based on the latest CCG re-organization took place in April 2021.
+The number of CCGs changes each year due to the NHS re-organization.
+The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
+We decided to use the CCG shapefile from 2020 at the time of writing, there is no public EHR data published based on the latest CCG re-organization that took place in April 2021.
 
@@
 We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
-These rivers were chosen as they pass through regions with clusters of CCGs, and provide informative geographical and topological cues.
+These rivers were chosen as they pass through regions with dense population, and provide informative geographical and topological cues.
 
-We first obtained a relation ID by searching a river, e.g.
+We first obtain a relation ID by searching for a river, e.g.
 River Thames, on OpenStreetMap.
-The obtained relation ID was used to construct a query (See Listing~\ref{overpass}) which enabled use to download the entire river shapefile through Overpass Turbo \cite{overpassturboOverpass}.
+The relation ID is used to construct a query (See Listing~\ref{overpass}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
 
-\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via Overpass Turbo API.}, label={overpass},captionpos=b]
+\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
     relation(2263653);>>;

### 2022-02-18 paper/section/main.tex
draft: update

@@
-\section{\software}
+\section{Cartograms with Features}
+
+{
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{An overview of \software. Figure TBA.}
+    \label{fig:overview}
+\end{figure}
+}
 
 \algoref{alg:UpdateNodePosition} describes the overview of \software.
-We first load all processed data from \secref{sec:{Data Description}} and apply VPSC \cite{dwyer2006fast} recursively to remove overlaps.
+We first load and render the CCG geospatial boundaries, for each CCG we compute the centroid where we render a node with $ \mathcal{S} $ as its size.
+We then apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} recursively to remove overlaps.
 We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-During each recursion, we check node trajectories (See \algoref{alg:check river crossing}) and move back nodes that crossed a river.
-A recursion ends when 1) no overlaps are present; 2) no nodes crossed a river.
-We then increase the node size and repeat the recursion until the entire map canvas is filled.
+During each iteration, we test node trajectories (See \algoref{alg:check river crossing}) and move nodes that cross a river back to their previous position.
+A procedure ends when 1) no overlaps are present; 2) no nodes cross a river.
+We then increase $ \mathcal{S} $ by one pixel and repeat the algorithm until the entire map canvas is filled.
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to update node positions by removing overlaps and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
-    \textbf{Variables:} \\
-    $NodeList \gets$ a list of nodes representing CCGs \\
-    $Node_{size}\gets$ the current size of all nodes \\
-    $Node_{inc} \gets$ the increment node size of each iteration \\
-    $Node_{max} \gets$ the maximum size of a node \\
-    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
+    \caption{The procedure to update node positions by removing overlap and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
+    \textbf{Global variables:} \\
+    $ \mathbcal{L} \gets $ a list of nodes representing CCGs \\
+    $ \mathcal{S} \gets $ the current size of all nodes \\
+    $ \mathcal{S}_{max} \gets $ the maximum size of a node \\
+    $ \mathcal{I} \gets $ the maximum number of iterations before a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \mathbcal{L}_{vpsc} \gets $ a list of nodes representing CCGs after running VPSC \\
+    $ cross \gets $ the boolean for whether a node crossed a river \\
+
     \begin{algorithmic}[1]
         \Procedure{UpdateNodePosition}{}
-        \While{$ Node_{size} < Node_{max} $}
+        \While{$ \mathcal{S} < \mathcal{S}_{max} $}
 
-        \State $ nodeList_{ora} \gets $ RunVPSC ($ NodeList $)
+        \State $ \mathbcal{L}_{vpsc} \gets $ RunVPSC ($ \mathbcal{L} $)
 
-            \For{$ node^i_{ora} \in nodeList_{ora} $}
+            \ForEach {$ \mathbcal{L}^i_{vpsc} \in \mathbcal{L}_{vpsc} $}
 
-                \If {MoveNode ($ Node^i, node^i_{ora} $)}
-                    \State $ cross \gets $ \Call{CheckRiverCrossing}{$ Node^i, node^i_{ora} $}
+                \If {TranslateNode ($ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $)}
+                    \State $ cross \gets $ \Call{TestRiverCross}{$ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $}
 
                     \If{$ cross = True $}
-                        \State $ Node^i.stale \gets Node^i.stale + 1 $
+                        \State $ \mathcal{I} \gets \mathcal{I} + 1 $
 
-                        \If{$ Node^i.stale < Node_{stale} $}
-                            \State \Call{MoveNode}{$ Node^i $}
+                        \If{$ \mathcal{I} < \mathcal{I} $}
+                            \State \Call{TranslateNode}{$ \mathbcal{L}^i $}
                         \Else
-                            \State \Call{DeriveCorridor}{$ Node^i, node^i_{ora} $}
+                            \State \Call{DeriveCorridor}{$ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $}
                         \EndIf
@@
 
-        \State $Node_{size} \gets Node_{size} + Node_{inc}$
+        \State $ \mathcal{S} \gets \mathcal{S} ++$
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to update a node's position.}\label{alg:move position}
+    \caption{The procedure to update a single node's position.}\label{alg:move position}
+
+    \textbf{Input:} \\
+    $ \mathcal{N} \gets $ required, the node \\
+    $ \mathcal{N}_{vpsc} \gets $ optional, the node after running VPSC \\
+
+    \textbf{Output:} \\
+    A boolean for whether the node's position is updated. \\
+
+    \textbf{Local variables:} \\
+    $ x,y \gets $ variables used to store coordinates \\
+
     \begin{algorithmic}[1]
-        \Procedure{MoveNode}{$ Node, Node_{new} $}
-        \If{$ Node.x = Node_{new}.x $ and $ Node.y = Node_{new}.y $}
+        \Procedure{TranslateNode}{$ \mathcal{N}, \mathcal{N}_{vpsc} $}
+        \If{$ \mathcal{N}.x = \mathcal{N}_{vpsc}.x $ and $ \mathcal{N}.y = \mathcal{N}_{vpsc}.y $}
+            \item[] \Comment{Node position remains unchanged}
             \State \Return{$ False $}
@@
 
-        \State init. $ x,y $
-
-        \If{$ Node_{new}$ is $ undefined $}
-            \State $previous \gets Node.history.pop() $
-            \State $ x \gets previous.x,~y \gets previous.y $
+        \If{$ \mathcal{N}_{vpsc}$ is undefined } 
+        \item[] \Comment{When optional parameter $ \mathcal{N}_{vpsc} $ is undefined}
+        \item[] \Comment{Move the node back to its previous position}
+            \State $ x \gets \mathcal{N}.previous.x,~y \gets \mathcal{N}.previous.y $
         \Else
-            \State $ x \gets Node_{new}.x,~y \gets Node_{new}.y $
-            \State $ Node.history.push(x, y) $
+            \State $ x \gets \mathcal{N}_{vpsc}.x,~y \gets \mathcal{N}_{vpsc}.y $
+            \State $ \mathcal{N}.history.push(x, y) $
         \EndIf
 
-        \State $ Node.x \gets x $
-        \State $ Node.y \gets y $
+        \State $ \mathcal{N}.x \gets x $
+        \State $ \mathcal{N}.y \gets y $
         \State \Return{$ True $}
@@
 
-\subsection{River Cross Checking}
+\subsection{River Cross Testing}
 
-We use rivers as topological boundaries to prevent nodes from crossing them.
-When a node's position changes, we check if the node's trajectory intersects any segment of a river, see \algoref{alg:check river crossing}.
-A boundary collision detection is performed to reduce the number of line intersection detections required.
+We use rivers as topological boundaries and prevent nodes from crossing them.
+When a node's position changes, we test if the node's trajectory intersects any segment of a river.
+See \algoref{alg:check river crossing}.
+A bounding box intersection test can be performed to reduce the number of line intersection detections required.
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to check if a node crosses a river.}\label{alg:check river crossing}
-    \textbf{Variables:} \\
-    $RiverEdges \gets$ a list of river edges formed by two points \\
+    \caption{The procedure to test if a node crosses a river.}\label{alg:check river crossing}
+    \textbf{Input:} \\
+    $ \mathcal{N} \gets $ required, the node \\
+    $ \mathcal{N}_{vpsc} \gets $ required, the node after running VPSC \\
+
+    \textbf{Output:} \\
+    A boolean for whether the node crosses a river. \\
+
+    \textbf{Global variables:} \\
+    $ \mathbcal{R} \gets $ a list of river edges \\
+
+    \textbf{Local variables:} \\
+    $ b_\mathcal{N}, b_{\mathcal{R}^i} \gets $ the bounding boxes for $\mathcal{N}$ and $\mathcal{R}^i$ \\
+    $ b_{intersect} \gets $ a boolean for whether $b_\mathcal{N}$ intersects $b_\mathcal{R}^i$ \\
+    $ e_\mathcal{N}, b_{\mathcal{R}^i} \gets $ the edges for $\mathcal{N}$ and $\mathcal{R}^i$ \\
+    $ e_{intersect} \gets $ a boolean for whether $e_\mathcal{N}$ intersects $e_\mathcal{R}^i$ \\
 
     \begin{algorithmic}[1]
-        \Procedure{CheckRiverCrossing}{$ Node, Node_{ora} $}
-            \For{$ edge \in RiverEdges $}
-                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
-                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
-                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
-                \If{$ collide = True $}
-                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
-                    \State $ edge_l \gets $ GenerateLine ($ edge $)
-                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
-                    \If{$ intersect = True $}
+        \Procedure{TestRiverCross}{$ \mathcal{N}, \mathcal{N}_{vpsc} $}
+
+        \ForEach{$ \mathcal{R}^i \in \mathbcal{R} $}
+            \State $ b_{\mathcal{N}} \gets $ GenerateBoundingBox ($ \mathcal{N}, \mathcal{N}_{vpsc} $)
+            \State $ b_{\mathcal{R}^i} \gets $ GenerateBoundingBox ($ \mathcal{R}^i $)
+            \State $ b_{intersect} \gets $ TestBoundingIntersection ($ b_{\mathcal{N}}, b_{\mathcal{R}^i} $)
+
+            \If{$ b_{intersect} = True $}
+                \State $ e_{\mathcal{N}} \gets $ GenerateEdge ($ \mathcal{N}, \mathcal{N}_{vpsc} $)
+                \State $ e_{\mathcal{R}^i} \gets $ GenerateEdge ($ \mathcal{R}^i $)
+                \State $ e_{intersect} \gets $ TestEdgeIntersection ($ e_\mathcal{N}, e_{\mathcal{R}^i} $)
+
+                \If{$ e_{intersect} = True $}
                     \State \Return{$ True $}
                 \EndIf
+
                 \EndIf
         \EndFor
+
         \State \Return{$ False $}
@@
 
-As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly traveling between two positions due to the lack of available space, creating a stalemate situation for recursions, as shown in \figref{fig:stalemate}.
-In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ Node_{stale} $, to break the stalemate: if a node is traveling between two positions for more than $ Node_{stale} $ iterations, it is considered to be in a stalemate.
+As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly translated between two positions due to the lack of available space, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ \mathcal{I} $, to test for stalemate.
+If a node is translated between two positions for more than $ \mathcal{I} $ iterations, it is considered to be in a stalemate.
 A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node repeatedly travels between two positions (A and B) for more than $Node_{stale}$ iterations. }
+    \caption{A stalemate situation is when a node repeatedly translates between two positions (A and B) for more than $\mathcal{I}$ iterations. }
     \label{fig:stalemate}
@@
     \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}.
-All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \vec{BA} $.
+All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \protect\overrightarrow{BA\strut} $.
 }
@@
 We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
-    \textbf{Variables:} \\
-    $\mathcal{C}_l \gets$ the length of a corridor \\
-    $\mathcal{C}_w \gets$ the width of a corridor \\
-    $\mathcal{D} \gets$ the distance to move nodes within the corridor \\
-    $Node_{size}\gets$ the current size of all nodes \\
+
+    \textbf{Input:} \\
+    $ \mathcal{N} \gets $ required, the node current state \\
+    $ \mathcal{N}_{p} \gets $ required, the node's previous state \\
+
+    \textbf{Output:} \\
+    A boolean for whether the node's position is updated. \\
+
+    \textbf{Global variables:} \\
+    $ \mathcal{C}_l \gets $ the length of a corridor \\
+    $ \mathcal{C}_w \gets $ the width of a corridor \\
+    $ \mathcal{D} \gets $ the distance to move nodes within the corridor \\
+    $ \mathcal{S} \gets $ the current size of all nodes \\
+
+    \textbf{Local variables:} \\
+    $ slope \gets $ the slope of $ \overrightarrow{\mathcal{N}\mathcal{N}_{p}\strut} $ \\
+    $ e_{p} \gets $ an edge perpendicular at $ \mathcal{N} $ on $ \overrightarrow{\mathcal{N}\mathcal{N}_{p}\strut} $ \\
+    $ e_{s^1} \gets $ a perpendicular edge at the start of $ e_{p} $ \\
+    $ e_{s^2} \gets $ a perpendicular edge at the end of $ e_{p} $ \\
+    $ corridor \gets $ a rectangular formed by $ e_p $, $ e_{s^1} $ and $ e_{s^2} $ \\
+    $ n_{in} \gets $ a node located inside $ corridor $ \\
+    $ e_{d} \gets $ an edge connecting $ \mathcal{N} $ and its destination \\
+
     \begin{algorithmic}[1]
-        \Procedure{DeriveCorridor}{$ Node, Node_{old} $}
-        \State MoveNode ($ Node $)
+        \Procedure{DeriveCorridor}{$ \mathcal{N}, \mathcal{N}_{p} $}
+            \State TranslateNode ($ \mathcal{N} $)
 
-        \State $ \Delta x \gets Node.x - Node_{old}.x,~ \Delta y \gets Node.y - Node_{old}.y $
+            \State $ \Delta x \gets \mathcal{N}.x - \mathcal{N}_{p}.x,~ \Delta y \gets \mathcal{N}.y - \mathcal{N}_{p}.y $
 
@@
 
-        \State $ pLine = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ Node $, $ \mathcal{C}_w $}
-        \State $ sideLine_1 = $ \Call{DerivePoint}{$ slope $, $ pLine.a $, $ \mathcal{C}_l $, $ True $}
-        \State $ sideLine_2 = $ \Call{DerivePoint}{$ slope $, $ pLine.b $, $ \mathcal{C}_l $, $ True $}
+            \State $ e_{p} = $ \Call{DeriveEdge}{$ \frac{-1}{ slope } $, $ \mathcal{N} $, $ \mathcal{C}_w $}
+            \State $ e_{s^1} = $ \Call{DeriveEdge}{$ slope $, $ e_{p}.a $, $ \mathcal{C}_l $, $ True $}
+            \State $ e_{s^2} = $ \Call{DeriveEdge}{$ slope $, $ e_{p}.b $, $ \mathcal{C}_l $, $ True $}
             \State $ corridor \gets
                 \begin{bmatrix}
-                pLine.a &
-                pLine.b \\
+                    e_{p}.a &
+                    e_{p}.b \\
 
-                sideLine_1.a &
-                sideLine_2.b \\
+                    e_{s^1}.a &
+                    e_{s^2}.b \\
                 \end{bmatrix} $
 
-        \For{$ node^i $ inside $ corridor $}
-            \State $ destination = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ node^i $, $ \mathcal{D} $}
+            \ForEach{$ {n}_{in}^i $ inside $ corridor $}
+                \State $ e_{d} = $ \Call{DeriveEdge}{$ slope $, $ {n}_{in}^i $, $ \mathcal{D} $}
 
-            \State MoveNode ($ node^i, destination.b $)
+                \State TranslateNode ($ {n}_{in}^i, e_{d}.b $)
             \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
 
-        \EndProcedure
+%\end{noindent}
 
-        \item[]
+% \begin{noindent}
 
-        \Procedure{DerivePoint}{$ Slope, Point, Length, IsSideLine $}
+\begin{algorithm}[tb!]
+    \caption{The procedure to derive an edge based on a slope and a point.}\label{alg:derive corridor point}
+    \textbf{Input:} \\
+    $ Slope \gets $ the slope used to derive an edge \\
+    $ Point \gets $ the point used to derive an edge \\
+    $ Length \gets $ the length of the derived edge \\
+    $ IsSideEdge \gets $ the boolean used to determine the edge's type \\
 
-        \State init. $ a $, $ b $
+    \textbf{Output:} \\
+    An edge that passes through a given point with the given slope. \\
+
+    \textbf{Local variables:} \\
+    $ a,b \gets $ variables used to store coordinates \\
+    $ \Delta x, \Delta y \gets $ the differences in $ x, y $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{DeriveEdge}{$ Slope, Point, Length, IsSideEdge $}
+
         \Switch{$Slope$}
             \Case{$0$}
-                \If{$ IsSideLine = True $}
+                \If{$ IsSideEdge = True $}
                     \State $ a.x \gets Point.x $
@@
 
-                \If{$ IsSideLine = True $}
+                \If{$ IsSideEdge = True $}
                     \State $ a.y \gets Point.y $
@@
             \Default
-                \State $ dx = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ dy = Slope \cdot dx $
-                \If{$ IsSideLine = True $}
+                \State $ \Delta x = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ \Delta y = Slope \cdot \Delta x $
+                \If{$ IsSideEdge = True $}
                     \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
                 \Else
-                    \State $ a.x \gets Point.x + dx $,~$ a.y \gets Point.y + dy $
+                    \State $ a.x \gets Point.x + \Delta x $,~$ a.y \gets Point.y + \Delta y $
                 \EndIf
 
-                \State $ b.x \gets Point.x - dx $,~$ b.y \gets Point.y - dy $
+                \State $ b.x \gets Point.x - \Delta x $,~$ b.y \gets Point.y - \Delta y $
             \EndDefault

### 2022-02-18 paper/section/related.tex
draft: update

--- 2022-01-16 paper/section/related.tex
@@
 
-This section introduces the definition of cartogram types, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
+This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
 
@@
 
-Dorling cartograms are non-contiguous and do not preserve geography and topology.
-A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size \cite{dorling2011Area}.
-In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
-Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.
-
-Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular}.
-Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.
-
-Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
-Some variants are able to preserve topological accuracy as well.
-
-\subsection{Peer-reviewed Applications}
-
-\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
-\citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms.
-\citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
-\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy.
-\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
-See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
-
 {
 \renewcommand{\arraystretch}{1.5}
-\begin{table*}[!tb]
+\begin{table*}[!htb]
 	\centering
@@
 	}
-	\caption{}
+	\caption{Related work with cartogram-based visualizations.
+\textbf{Cartogram type} is the type of cartogram used.
+\textbf{Geographic region} is the geographic region depicted by the cartogram.
+\textbf{Number of nodes} is the number of nodes depicted in the cartogram.
+}
 	\label{table:region vs node}
@@
 
+\thumbnail{dorling}{
+	Dorling cartograms are non-contiguous and do not preserve geography and topology.
+A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimension of interest is represented by the circle size \cite{dorling2011Area}.
+In a Demers Cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
+Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
+}
+
+\thumbnail{rectangular}{
+	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
+Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
+}
+
+\thumbnail{mosaic}{
+	Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+Some variants are able to preserve topological accuracy as well.
+}
+
+\subsection{Peer-reviewed Applications}
+
+\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
+\citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
+\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
+\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
+See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+
 \subsection{Other Applications}
 
-Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news.
-\citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
-The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
+Cartograms are a popular choice of representation in covering various topics by the media, in order to present an unbiased and complete picture of the news.
+The Washington Post uses cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
 The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
+\citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
 \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.
+
+% \subsection{Contiguous Applications}
+
+% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
+
+Compared to previous work on Demers Cartogram, we introduce a topological feature, a river, aiming to improve the interpretation, readability, and topological accuracy.

### 2022-02-23 src/router/index.js
chore: update router

@@ -8,17 +8,13 @@ const routes = [
   {
     path: "/",
     name: "Home",
-    component: Home,
+    component: () => import(/* webpackChunkName: "US" */ "../views/CCG.vue"),
+    // component: Home,
   },
   {
     path: "/ccg",
     name: "CCG",
     component: () => import(/* webpackChunkName: "US" */ "../views/CCG.vue"),
-  },
-  {
-    path: "/ccg2",
-    name: "CCG2",
-    component: () => import(/* webpackChunkName: "US" */ "../views/CCG2.vue"),
   },
 ];
 

### 2022-02-23 paper/section/main.tex
draft: update algorithm

@@
     $ \mathcal{D} \gets $ the distance to move nodes within the corridor \\
-    $ \mathcal{S} \gets $ the current size of all nodes \\
 
@@
             \ForEach{$ {n}_{in}^i $ inside $ corridor $}
+                \State $ {n}_{in}^i.dx \gets \Delta x $, $ {n}_{in}^i.dy \gets \Delta y $
                 \State $ e_{d} = $ \Call{DeriveEdge}{$ slope $, $ {n}_{in}^i $, $ \mathcal{D} $}
@@
                 \State $ a.y \gets Point.y $
+
+                \If{$ Point.dx < 0 $}
+                    \State $ a.x \gets Point.x + Length $
+                \Else
+                    \State $ a.x \gets Point.x - Length $
+                \EndIf
                 \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
@@
                     \State $ a.y \gets Point.y + Length $
+                \EndIf
+
+                \If{$ Point.dy < 0 $}
+                    \State $ a.y \gets Point.y + Length $
+                \Else
+                    \State $ a.y \gets Point.y - Length $
                 \EndIf
@@
 
+                \If{$ Point.dx > 0 $ and $ Slope = 1 $ and $ IsSideEdge = True $}
+                    \State $ b.x \gets Point.x + \Delta x $,~$ b.y \gets Point.y + \Delta y $
+                \Else
                     \State $ b.x \gets Point.x - \Delta x $,~$ b.y \gets Point.y - \Delta y $
+                \EndIf
+
             \EndDefault

### 2022-03-01 paper/section/data.tex
draft: update algorithm and review

@@
 
-Obtaining the data is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}.
+Obtaining the heterogenous data can be challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}.
 The data acquisition involves multiple steps.
 The first step is to obtain both geospatial boundaries and EHR data that match one another.
-The second step is to pre-process the EHR data for removing empty and erroneous values.
+The second step is to pre-process the EHR data to remove empty and erroneous values.
 The final step is to transform the data into a format that is suitable for visualization.
@@
     \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-    \caption{A map of 106 CCGs in England as of April 2021, obtained from the Open Geography portalx \cite{opengeographyportalxOpen}.}
+    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography portalx \cite{opengeographyportalxOpen}.}
     \label{fig:ccg}
@@
 
-\subsubsection{Clinical Commissioning Groups}
+\subsubsection{Clinical Commissioning Groups Shapefile}
 
 Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
-The number of CCGs changes each year due to the NHS re-organization.
+The number of CCGs changes over time due to the NHS re-organization.
 The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, there is no public EHR data published based on the latest CCG re-organization that took place in April 2021.
+We decided to use the CCG shapefile from 2020 at the time of writing, because there is no public EHR data published based on the latest CCG re-organization that took place in 2021.
 
-\subsubsection{Rivers}
+\subsubsection{River Features}
 
@@
         \begin{itemize}
-            \item Stoke
+            \item Stroke
             \item Alcohol-specific admission and readmission

### 2022-03-01 paper/section/main.tex
draft: update algorithm and review

--- 2022-02-16 paper/section/main.tex
@@
-\section{\software}
+\section{Cartograms with Dynamic Features}
 
-\algoref{alg:UpdateNodePosition} describes the overview of \software.
-We first load all processed data from \secref{sec:{Data Description}} and apply VPSC \cite{dwyer2006fast} recursively to remove overlaps.
+{
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=\columnwidth]{example-image-a}
+    \caption{An overview of \software. Figure TBA.}
+    \label{fig:overview}
+\end{figure}
+}
+
+\algoref{alg:UpdateNodePosition} provides an overview of our algorithm.
+We first load and render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a square node with $ s $ as its size.
+We then apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} repeatedly to remove overlaps.
 We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-During each recursion, we check node trajectories (See \algoref{alg:check river crossing}) and move back nodes that crossed a river.
-A recursion ends when 1) no overlaps are present; 2) no nodes crossed a river.
-We then increase the node size and repeat the recursion until the entire map canvas is filled.
+During each iteration, we compute node trajectories (See \algoref{alg:check river crossing}) and translate nodes to their new position.
+Nodes that cross a river are translated back to their previous position.
+This procedure ends when 1) no overlaps are present; 2) no nodes cross a river.
+We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to update node positions by removing overlaps and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
-    \textbf{Variables:} \\
-    $NodeList \gets$ a list of nodes representing CCGs \\
-    $Node_{size}\gets$ the current size of all nodes \\
-    $Node_{inc} \gets$ the increment node size of each iteration \\
-    $Node_{max} \gets$ the maximum size of a node \\
-    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
+    \caption{Procedure to update node positions by removing overlap and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
+    \textbf{Global variables:} \\
+    $ \nodeList \gets $ a list of nodes representing CCGs \\
+    $ \nodeSize \gets $ the current size of all nodes \\
+    $ \nodeSizeMax \gets $ the maximum size of a node \\
+    $ \stalemate \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \nodeListVPSC \gets $ a list of nodes representing CCGs after running VPSC \\
+    $ \crossRiver \gets $ boolean indicating a node crossed a river \\
+
     \begin{algorithmic}[1]
         \Procedure{UpdateNodePosition}{}
-        \While{$ Node_{size} < Node_{max} $}
+        \While{$ \nodeSize < \nodeSizeMax $}
 
-        \State $ nodeList_{ora} \gets $ RunVPSC ($ NodeList $)
+        \State $ \nodeListVPSC \gets $ RunVPSC ($ \nodeList $)
 
-            \For{$ node^i_{ora} \in nodeList_{ora} $}
+            \ForEach {$ \nodeListVPSCEach \in \nodeListVPSC $}
 
-                \If {MoveNode ($ Node^i, node^i_{ora} $)}
-                    \State $ cross \gets $ \Call{CheckRiverCrossing}{$ Node^i, node^i_{ora} $}
+                \If {TranslateNode ($ \nodeListEach, \nodeListVPSCEach $)}
+                    \State $ \crossRiver \gets $ \Call{TestRiverCross}{$ \nodeListEach, \nodeListVPSCEach $}
 
-                    \If{$ cross = True $}
-                        \State $ Node^i.stale \gets Node^i.stale + 1 $
+                    \If{$ \crossRiver = True $}
+                        \State $ \stalemate \gets \stalemate + 1 $
 
-                        \If{$ Node^i.stale < Node_{stale} $}
-                            \State \Call{MoveNode}{$ Node^i $}
+                        \If{$ \stalemate < \stalemate $}
+                            \State \Call{TranslateNode}{$ \nodeListEach $}
                         \Else
-                            \State \Call{DeriveCorridor}{$ Node^i, node^i_{ora} $}
+                            \State \Call{DeriveCorridor}{$ \nodeListEach, \nodeListVPSCEach $}
                         \EndIf
@@
 
-        \State $Node_{size} \gets Node_{size} + Node_{inc}$
+        \State $ \nodeSize \gets \nodeSize ++$
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to update a node's position.}\label{alg:move position}
+    \caption{Procedure to update a single node, $ \node $'s position.}\label{alg:move position}
+
+    \textbf{Input:} \\
+    $ \node \gets $ required, the node \\
+    $ \nodeVPSC \gets $ optional, the node after running VPSC \\
+
+    \textbf{Output:} \\
+    A boolean for whether the node's position is updated. \\
+
     \begin{algorithmic}[1]
-        \Procedure{MoveNode}{$ Node, Node_{new} $}
-        \If{$ Node.x = Node_{new}.x $ and $ Node.y = Node_{new}.y $}
+        \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
+        \If{$ \node.x = \nodeVPSC.x $ and $ \node.y = \nodeVPSC.y $}
+            \item[] \Comment{Node position remains unchanged}
             \State \Return{$ False $}
@@
 
-        \State init. $ x,y $
-
-        \If{$ Node_{new}$ is $ undefined $}
-            \State $previous \gets Node.history.pop() $
-            \State $ x \gets previous.x,~y \gets previous.y $
+        \If{$ \nodeVPSC$ is undefined } 
+        \item[] \Comment{When optional parameter $ \nodeVPSC $ is undefined}
+        \item[] \Comment{Move the node back to its previous position}
+            \State $ \node.x \gets \node.previous.x,~\node.y \gets \node.previous.y $
         \Else
-            \State $ x \gets Node_{new}.x,~y \gets Node_{new}.y $
-            \State $ Node.history.push(x, y) $
+            \State $ \node.x \gets \nodeVPSC.x,~\node.y \gets \nodeVPSC.y $
+            \State $ \node.history.push(\node.x, \node.y) $
         \EndIf
-
-        \State $ Node.x \gets x $
-        \State $ Node.y \gets y $
         \State \Return{$ True $}
@@
 
-\subsection{River Cross Checking}
+\subsection{River Cross Testing}
 
-We use rivers as topological boundaries to prevent nodes from crossing them.
-When a node's position changes, we check if the node's trajectory intersects any segment of a river, see \algoref{alg:check river crossing}.
-A boundary collision detection is performed to reduce the number of line intersection detections required.
+We use rivers as topological boundaries and prevent nodes from crossing them.
+When a node's position changes, we test if the node's trajectory intersects any segment of a river.
+See \algoref{alg:check river crossing}.
+A bounding box intersection test can be performed to reduce the number of edge intersection detections required.
 
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to check if a node crosses a river.}\label{alg:check river crossing}
-    \textbf{Variables:} \\
-    $RiverEdges \gets$ a list of river edges formed by two points \\
+    \caption{Procedure to test if a node crosses a river.}\label{alg:check river crossing}
+    \textbf{Input:} \\
+    $ \node \gets $ required, the node \\
+    $ \nodeVPSC \gets $ required, the node after running VPSC \\
+
+    \textbf{Output:} \\
+    A boolean indicating the node crosses a river. \\
+
+    \textbf{Global variables:} \\
+    $ \riverEdgeList \gets $ a list of river edges \\
+
+    \textbf{Local variables:} \\
+    $ \riverEdgeListEach \gets $ a river edge in $ \riverEdgeList $ \\ 
+    $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \node $ and $ \riverEdgeListEach $ \\
+    $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
+    $ \nodeEdge \gets $ the edge for $ \node $\\
+    $ \edgeBoxInt \gets $ boolean indicating $ \nodeEdge $ intersects $ \riverEdgeListEach $ \\
 
     \begin{algorithmic}[1]
-        \Procedure{CheckRiverCrossing}{$ Node, Node_{ora} $}
-            \For{$ edge \in RiverEdges $}
-                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
-                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
-                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
-                \If{$ collide = True $}
-                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
-                    \State $ edge_l \gets $ GenerateLine ($ edge $)
-                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
-                    \If{$ intersect = True $}
+        \Procedure{TestRiverCross}{$ \node, \nodeVPSC $}
+
+        \ForEach{$ \riverEdgeListEach \in \riverEdgeList $}
+            \State $ \nodeBoundingBox \gets $ GenerateBoundingBox ($ \node, \nodeVPSC $)
+            \State $ \riverEdgeBoundingBoxEach \gets $ GenerateBoundingBox ($ \riverEdgeListEach $)
+            \State $ \boundingBoxInt \gets $ TestBoundingIntersection ($ \nodeBoundingBox, \riverEdgeBoundingBoxEach $)
+
+            \If{$ \boundingBoxInt = True $}
+                \State $ e_{n} \gets $ GenerateEdge ($ \node, \nodeVPSC $)
+                \State $ \edgeBoxInt \gets $ TestEdgeIntersection ($ \nodeEdge, \riverEdgeListEach $)
+
+                \If{$ \edgeBoxInt = True $}
                     \State \Return{$ True $}
                 \EndIf
+
                 \EndIf
         \EndFor
+
         \State \Return{$ False $}
@@
 
-As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly traveling between two positions due to the lack of available space, creating a stalemate situation for recursions, as shown in \figref{fig:stalemate}.
-In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ Node_{stale} $, to break the stalemate: if a node is traveling between two positions for more than $ Node_{stale} $ iterations, it is considered to be in a stalemate.
-A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).
+As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node might be repeatedly translated between two positions due to river crossing and congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+In order to create space for VPSC to generate a new layout, we introduce a parameter, $ \stalemate $, to test for stalemate.
+If a node is translated between two positions for more than $ \stalemate $ iterations, it is considered to be in a stalemate.
+A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ d $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).
 
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node repeatedly travels between two positions (A and B) for more than $Node_{stale}$ iterations. }
+    \caption{A stalemate situation is when a node repeatedly translates between two positions (A and B) for more than $ stalemate $ iterations. }
     \label{fig:stalemate}
@@
     \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}.
-All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \vec{BA} $.
+All nodes within the corridor are moved by $ d $ in the direction of $ \Vector{B}{A} $.
 }
@@
 \begin{algorithm}[tb!]
-    \caption{The procedure to derive a corridor to move included nodes.
-We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
-    \textbf{Variables:} \\
-    $\mathcal{C}_l \gets$ the length of a corridor \\
-    $\mathcal{C}_w \gets$ the width of a corridor \\
-    $\mathcal{D} \gets$ the distance to move nodes within the corridor \\
-    $Node_{size}\gets$ the current size of all nodes \\
+    \caption{Procedure to derive a corridor to translate enclosed nodes.
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards.}\label{alg:derive corridor}
+
+    \textbf{Input:} \\
+    $ \node \gets $ the node used to derive the corridor \\
+
+    \textbf{Global variables:} \\
+    $ \nodeSize \gets $ the current size of all nodes \\
+    $ \CorridorLength \gets $ the length of a corridor \\
+    $ \CorridorWidth \gets $ the width of a corridor \\
+
+    \textbf{Local variables:} \\
+    $ \PointP \gets $ the point extending $ \Vector{\node.previous}{~\node} $ such that the length of $ \Vector{\node.previous}{~\PointP} $ is $ \CorridorLength $\\
+    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \Vector{\node}{\PointP} $ \\
+    $ corridor \gets $ a rectangular formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
+
     \begin{algorithmic}[1]
-        \Procedure{DeriveCorridor}{$ Node, Node_{old} $}
-        \State MoveNode ($ Node $)
+        \Procedure{DeriveCorridor}{$ \node $}
+            \State TranslateNode ($ \node $)
 
-        \State $ \Delta x \gets Node.x - Node_{old}.x,~ \Delta y \gets Node.y - Node_{old}.y $
+            \State $ \PointP \gets $ \Call{DerivePoint}{$ \Vector{\node.previous}{~\node} $, $ \CorridorLength $}
 
-        \State $ slope \gets \frac{\Delta x}{\Delta y}$
+            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \Vector{\node.previous}{~\PointP} $, $ \frac{\CorridorWidth}{2} $}
 
-        \State $ pLine = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ Node $, $ \mathcal{C}_w $}
-        \State $ sideLine_1 = $ \Call{DerivePoint}{$ slope $, $ pLine.a $, $ \mathcal{C}_l $, $ True $}
-        \State $ sideLine_2 = $ \Call{DerivePoint}{$ slope $, $ pLine.b $, $ \mathcal{C}_l $, $ True $}
+            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \Vector{\node.previous}{~\PointP} $, $ -\frac{\CorridorWidth}{2} $}
+
             \State $ corridor \gets
                 \begin{bmatrix}
-                pLine.a &
-                pLine.b \\
+                    \EdgeParallelA.start &
+                    \EdgeParallelA.end \\
 
-                sideLine_1.a &
-                sideLine_2.b \\
+                    \EdgeParallelB.start &
+                    \EdgeParallelB.end \\
                 \end{bmatrix} $
 
-        \For{$ node^i $ inside $ corridor $}
-            \State $ destination = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ node^i $, $ \mathcal{D} $}
+            \ForEach{$ \nodeInCorridorEach $ inside $ corridor $}
+                \State $ \nodeInCorridorEachDx \gets \nodeInCorridorEach.x - \nodeInCorridorEach.previous.x $
 
-            \State MoveNode ($ node^i, destination.b $)
+                \State $ \nodeInCorridorEachDy \gets \nodeInCorridorEach.y - \nodeInCorridorEach.previous.y $
+
+                \State $ hyp \gets \sqrt{(\nodeInCorridorEachDx)^2 +(\nodeInCorridorEachDy)^2} $
+
+                \State $ \nodeInCorridorEach.next \gets $ \Call{DerivePoint}{$ \Vector{\nodeInCorridorEach.previous}{~\nodeInCorridorEach} $, $ hyp + s $}
+
+                \State TranslateNode ($ \nodeInCorridorEach $, $ \nodeInCorridorEach.next $)
+
             \EndFor
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
 
-        \EndProcedure
+%\end{noindent}
 
-        \item[]
+% \begin{noindent}
 
-        \Procedure{DerivePoint}{$ Slope, Point, Length, IsSideLine $}
+    \begin{algorithm}[tb!]
+        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
+        \textbf{Input:} \\
+        $ \Edge \gets $ the edge used to derive the new point \\
+        $ \Distance \gets $ the distance between $ \PointP $ and $ \Edge.start $ \\
 
-        \State init. $ a $, $ b $
-        \Switch{$Slope$}
-            \Case{$0$}
-                \If{$ IsSideLine = True $}
-                    \State $ a.x \gets Point.x $
-                \Else
-                    \State $ a.x \gets Point.x + Length $
-                \EndIf
-                \State $ a.y \gets Point.y $
-                \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
-            \EndCase
-            \Case{$ \infty $ or $ -\infty $}
-                \State $ a.x \gets Point.x $
+        \textbf{Output:} \\
+        A point, $ \PointP $, that is distance $ \Distance $ away from $ \Edge.start $. \\
 
-                \If{$ IsSideLine = True $}
-                    \State $ a.y \gets Point.y $
-                \Else
-                    \State $ a.y \gets Point.y + Length $
-                \EndIf
+        \textbf{Local variables:} \\
+        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \Edge.start $ and $ \Edge.end $ \\
 
-                \State $ b.x \gets Point.x $,~$ b.y \gets Point.y - Length $
-            \EndCase
-            \Default
-                \State $ dx = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ dy = Slope \cdot dx $
-                \If{$ IsSideLine = True $}
-                    \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
-                \Else
-                    \State $ a.x \gets Point.x + dx $,~$ a.y \gets Point.y + dy $
-                \EndIf
+        \begin{algorithmic}[1]
+            \Procedure{DerivePoint}{$ \Edge $, $ \Distance $}
+                \State $ \dx \gets \Edge.start.x - \Edge.end.x $
 
-                \State $ b.x \gets Point.x - dx $,~$ b.y \gets Point.y - dy $
-            \EndDefault
+                \State $ \dy \gets \Edge.start.y - \Edge.end.y $
 
-        \EndSwitch
+                \State $ \PointP.x \gets \frac{\dx}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
 
-        \State \Return{$ a, b $}
+                \State $ \PointP.y \gets \frac{\dy}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
+
+            \State \Return{$ \PointP $}
 
@@
     %\end{noindent}
+
+% \begin{noindent}
+
+\begin{algorithm}[tb!]
+    \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}
+
+    \textbf{Input:} \\
+    $ \Edge \gets $ the edge used to derive the parallel edge $ \EdgeParallel $ \\
+    $ \Distance \gets $ the distance between $ \Edge $ and $ \EdgeParallel $ \\
+
+    \textbf{Output:} \\
+    An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $. \\
+
+    \textbf{Local variables:} \\
+    $ \dx, \dy \gets $ the differences in $ x, y $ for $ \Edge.start $ and $ \Edge.end $ \\
+    $ \Scale \gets $ the scale of $ \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{DeriveParallelEdge}{$ \Edge $, $ \Distance $}
+            \State $ \dx \gets \Edge.start.x - \Edge.end.x $
+
+            \State $ \dy \gets \Edge.start.y - \Edge.end.y $
+
+            \State $ \Scale \gets \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $
+
+            \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \Edge.start.x $
+
+            \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \Edge.start.y $
+
+            \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \Edge.end.x $
+
+            \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \Edge.end.y $
+
+        \State \Return{$ \EdgeParallel $}
+
+        \EndProcedure
+
+    \end{algorithmic}
+\end{algorithm}
+
+%\end{noindent}

### 2022-03-01 paper/section/related.tex
draft: update algorithm and review

@@
 
-This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.
+This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world implementations of cartogram-based visualizations.
 
@@
 
-Cartograms combine statistical and geographical information in thematic maps.
-We follow the categorization of cartograms by \citea{vankreveld2004Rectangular}: contiguous, non-contiguous, rectangular, and Dorling.
-\citea{nusrat2016State} summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
-Each cartogram design may make accuracy trade-offs between dimensions, we provide an overview in Table \ref{table:accuracy}.
+\new{Cartograms are representations of geographical space generated by the value-by-area mapping technique that combines statistical and geographical information.
+In this paper, we follow the categorization of cartograms by \citea{vankreveld2004Rectangular}: contiguous, non-contiguous, rectangular, and Dorling.
+While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent areas but have distortion in shape.
+Non-contiguous cartograms sacrifice the topological connectivity to enable growing or shrinking in size to maintain their original shape \cite{dent2009Cartography}.}
+
+\citea{nusrat2016State} defines summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+Each cartogram design may make accuracy trade-offs between dimensions, we provide an overview of these cartograms in Table \ref{table:accuracy}.
 
@@
 			\hline
-			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & Dorling, Mosaic, Contiguous & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
 			\hline
-			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Dorling, Non-contiguous, Contiguous & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
 			\hline
@@
 	}
-	\caption{Related work with cartogram-based visualizations.
+	\caption{Related work with non-contiguous cartogram-based visualizations.
 \textbf{Cartogram type} is the type of cartogram used.
@@
 	Dorling cartograms are non-contiguous and do not preserve geography and topology.
-A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimension of interest is represented by the circle size \cite{dorling2011Area}.
-In a Demers Cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
+A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
+In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
 Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
+\new{Here we focus on Demers cartograms as we use squares to depict regions.
+This style of cartogram offers the advantages that the comparisons between regions are more intuitive and the space utilization is more effcient.}
 }
@@
 Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
+\new{Statistical comparisons between regions are difficult due to the differing shapes.}
 }
@@
 \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
@@
 
-\subsection{Other Applications}
+\subsection{Implementations}
 
-Cartograms are a popular choice of representation in covering various topics by the media, in order to present an unbiased and complete picture of the news.
+Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
 The Washington Post uses cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
@@
 
-Compared to previous work on Demers Cartogram, we introduce a topological feature, a river, aiming to improve the interpretation, readability, and topological accuracy.
+Compared to previous work on Demers cartograms, we introduce dynamic topological feature, a river, aiming to improve the interpretation, readability, and topological accuracy.

### 2022-03-02 paper/section/introduction.tex
draft: introduction

\section{Introduction and Motivation}

\new{Cartograms are representations of geographical space generated by the value-by-area mapping technique that combines statistical and geographical information. Various styles of cartograms have been proposed and implemented in fields such as geography, social science, and visualization \cite{dent2009Cartography}.}

\new{In this paper, we follow the categorization of cartograms by \citea{nusrat2016State}: contiguous, non-contiguous, rectangular, and Dorling. Among these four types, trade-offs between accuracies are made (See \tableref{table:accuracy}).}

\new{
    Based on Demers cartograms, we include dynamic topological features, such as rivers and mountains, aiming to improve the topological accuracy without sacrificing statistical accuracy. We then introduce a cartographic layout algorithm that uses these topological features to update the layout of the nodes representing geographical regions. To minimize cartographic errors, the algorithm also updates the position of rivers to accommodate the node layout. We then visualize a real world Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
}

### 2022-03-02 paper/section/related.tex
draft: introduction

@@
 
-\new{Cartograms are representations of geographical space generated by the value-by-area mapping technique that combines statistical and geographical information.
-In this paper, we follow the categorization of cartograms by \citea{vankreveld2004Rectangular}: contiguous, non-contiguous, rectangular, and Dorling.
-While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent areas but have distortion in shape.
+\new{While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent areas but have distortion in shape.
 Non-contiguous cartograms sacrifice the topological connectivity to enable growing or shrinking in size to maintain their original shape \cite{dent2009Cartography}.}
@@
 \new{Here we focus on Demers cartograms as we use squares to depict regions.
-This style of cartogram offers the advantages that the comparisons between regions are more intuitive and the space utilization is more effcient.}
+This style of cartogram offers the advantages that the comparisons between regions are more intuitive and the space utilization is more effcient.
+Demers cartograms also have the advantage of preserving a higher level of topology.
+}
 }

### 2022-03-04 paper/section/data.tex
draft: introduction
draft: update and review

@@
 
-Obtaining the heterogenous data can be challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}.
+Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHRa}.
 The data acquisition involves multiple steps.
@@
 
-Geospatial boundaries, or shapefiles, were obtained from the following sources.
-After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile.
-Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size and improve the performance of our implementation.
-
+Geospatial boundaries, or shapefiles, were obtained from the sources described below.
 {
@@
 
-\subsubsection{Clinical Commissioning Groups Shapefile}
+\bobgraph{Clinical Commissioning Groups Shapefile}
 
@@
 The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, because there is no public EHR data published based on the latest CCG re-organization that took place in 2021.
+We decided to use the CCG shapefile from 2020 at the time of writing, because there is still no public EHR data published based on the latest CCG re-organization that took place in 2021.
 
-\subsubsection{River Features}
+\bobgraph{River Features}
 
@@
 \end{lstlisting}
+
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and scales, and merge them into one shapefile.
+Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size and improve the performance of our implementation.
 

### 2022-03-04 paper/section/introduction.tex
draft: update and review

@@
 
-\new{Cartograms are representations of geographical space generated by the value-by-area mapping technique that combines statistical and geographical information.
-Various styles of cartograms have been proposed and implemented in fields such as geography, social science, and visualization \cite{dent2009Cartography}.}
+Cartograms are representations of geographical space generated by a value-by-area mapping technique that combines statistical and geographical information \cite{dent2009Cartography}.
+\new{Various styles of cartograms have been proposed and implemented, covering topics on urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.}
 
-\new{In this paper, we follow the categorization of cartograms by \citea{nusrat2016State}: contiguous, non-contiguous, rectangular, and Dorling.
-Among these four types, trade-offs between accuracies are made (See \tableref{table:accuracy}).}
+In this paper, we follow the categorization of cartograms by \citea{nusrat2016State}: contiguous, non-contiguous, rectangular, and Dorling.
+Among these four types, trade-offs between accuracies are made (See \tableref{table:accuracy}).
+Building on Demers cartograms \cite{ian2002Cartogram}, we develop dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
+We incorporate a cartographic layout algorithm that includes these topological features to update the layout of the nodes representing geographical regions.
+To minimize geographical errors, the algorithm also updates the position of rivers to accommodate the node layout.
+We then apply the algorithm a real world Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
 
-\new{
-    Based on Demers cartograms, we include dynamic topological features, such as rivers and mountains, aiming to improve the topological accuracy without sacrificing statistical accuracy.
-We then introduce a cartographic layout algorithm that uses these topological features to update the layout of the nodes representing geographical regions.
-To minimize cartographic errors, the algorithm also updates the position of rivers to accommodate the node layout.
-We then visualize a real world Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
-}
+\new{Our contributions include:}
+
+\begin{itemize}
+    \item A new variant of Demers cartograms that incorporates dynamic topological features to improve readability and recognizability.
+    \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers.
+\end{itemize}

### 2022-03-04 paper/section/main.tex
draft: introduction
draft: update and review

@@
 
-\algoref{alg:UpdateNodePosition} provides an overview of our algorithm.
-We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a square node with $ s $ as its size.
-We then apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} repeatedly to remove overlaps.
+{
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/flowchart.png}
+    \caption{A flowchart abstraction of our layout algorithm.}
+    \label{fig:flowchart}
+\end{figure}
+}
+
+\algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of our algorithm.
+\textbf{Initialization:} We first load and render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a small square node with $ \nodeSize $ as the initial size.
+\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} repeatedly to remove overlaps.
 We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-During each iteration, we compute node trajectories (See \algoref{alg:check river crossing}) and translate nodes to their new position.
+During each iteration, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
 Nodes that cross a river are translated back to their previous position.
-This procedure ends when 1) no overlaps are present; 2) no nodes cross a river.
+This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
 We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
+\new{The size growing process provides a stability in the layout and minimizes geographical errors.}
 
@@
     $ \nodeSizeMax \gets $ the maximum size of a node \\
-    $ \stalemate \gets $ the maximum number of iterations indicating a stalemate \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
 
     \textbf{Local variables:} \\
-    $ \nodeListVPSC \gets $ a list of nodes representing CCGs after running VPSC \\
+    $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
     $ \crossRiver \gets $ boolean indicating a node crossed a river \\
+    $ \nodeListEach \gets $ a node in $ \nodeList $ \\
+    $ \nodeListEachPrevious \gets $ the previous position of $ \nodeListEach $\\
+    $ \nodeListVPSCEach \gets $ a node in $ \nodeListVPSC $, which is the next position of $ \nodeListEach $ after VPSC \\
 
@@
 
-                \If {TranslateNode ($ \nodeListEach, \nodeListVPSCEach $)}
-                    \State $ \crossRiver \gets $ \Call{TestRiverCross}{$ \nodeListEach, \nodeListVPSCEach $}
+                \If {\Call{TranslateNode}{$ \nodeListEach, \nodeListVPSCEach $}}
+                    \State $ \crossRiver \gets $ \Call{TestRiverIntersection}{$ \nodeListEach, \nodeListVPSCEach $}
 
                     \If{$ \crossRiver = True $}
-                        \State $ \stalemate \gets \stalemate + 1 $
+                        \State $ \nodeListEachStalemate \gets \nodeListEachStalemate + 1 $
 
-                        \If{$ \stalemate < \stalemate $}
-                            \State \Call{TranslateNode}{$ \nodeListEach $}
+                        \If{$ \nodeListEachStalemate < \stalemateMax $}
+                            \State $ \nodeListEach(x,y) \gets \nodeListEachPrevious(x,y) $
                         \Else
-                            \State \Call{DeriveCorridor}{$ \nodeListEach, \nodeListVPSCEach $}
+                            \State \Call{ProcessStalemate}{$ \nodeListEach, \nodeListVPSCEach $}
+                            \State $ \nodeListEachStalemate \gets 0 $
                         \EndIf
@@
 
+            \State $ \nodeSize \gets \nodeSize ++$
+
         \EndWhile
-
-        \State $ \nodeSize \gets \nodeSize ++$
 
@@
     \textbf{Input:} \\
-    $ \node \gets $ required, the node \\
-    $ \nodeVPSC \gets $ optional, the node after running VPSC \\
+    $ \node \gets $ the node \\
+    $ \nodeVPSC \gets $ the node after running VPSC \\
 
     \textbf{Output:} \\
-    A boolean for whether the node's position is updated. \\
+    Returns $ True $ if the node's position is updated. \\
 
@@
         \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
-        \If{$ \node.x = \nodeVPSC.x $ and $ \node.y = \nodeVPSC.y $}
+        \If{$ \node(x,y) = \nodeVPSC(x,y) $}
             \item[] \Comment{Node position remains unchanged}
@@
 
-        \If{$ \nodeVPSC$ is undefined } 
-        \item[] \Comment{When optional parameter $ \nodeVPSC $ is undefined}
-        \item[] \Comment{Move the node back to its previous position}
-            \State $ \node.x \gets \node.previous.x,~\node.y \gets \node.previous.y $
-        \Else
-            \State $ \node.x \gets \nodeVPSC.x,~\node.y \gets \nodeVPSC.y $
-            \State $ \node.history.push(\node.x, \node.y) $
-        \EndIf
+        \State $ \node(x,y) \gets \nodeVPSC(x,y) $
+
         \State \Return{$ True $}
@@
 
-\subsection{River Cross Testing}
+\subsection{River Intersection Testing}
 
 We use rivers as topological boundaries and prevent nodes from crossing them.
+\new{The resolution of rivers can be dynamically adjusted, as shown in \figref{fig:river resolution}}.
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
-See \algoref{alg:check river crossing}.
+See \algoref{alg:check river intersection}.
 A bounding box intersection test can be performed to reduce the number of edge intersection detections required.
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/river_resolution.png}
+    \caption{The resolution of rivers can be dynamically adjusted.
+The top figure shows River Thames at its original resolution.
+The bottom figure shows the river with a 60\% reduction in resolution.
+The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+    \label{fig:river resolution}
+\end{figure}
+}
 
@@
 \begin{algorithm}[tb!]
-    \caption{Procedure to test if a node crosses a river.}\label{alg:check river crossing}
+    \caption{Procedure to test if a node, $ \node $ crosses a river.}\label{alg:check river intersection}
     \textbf{Input:} \\
-    $ \node \gets $ required, the node \\
-    $ \nodeVPSC \gets $ required, the node after running VPSC \\
+    $ \node \gets $ the node \\
+    $ \nodeVPSC \gets $ the node after running VPSC \\
 
     \textbf{Output:} \\
-    A boolean indicating the node crosses a river. \\
+    Returns $ True $ if the node crosses a river. \\
 
@@
     \textbf{Local variables:} \\
-    $ \riverEdgeListEach \gets $ a river edge in $ \riverEdgeList $ \\ 
-    $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \node $ and $ \riverEdgeListEach $ \\
-    $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
-    $ \nodeEdge \gets $ the edge for $ \node $\\
-    $ \edgeBoxInt \gets $ boolean indicating $ \nodeEdge $ intersects $ \riverEdgeListEach $ \\
+    $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeListEachLine $ \\
+    $ \riverEdgeListEachLine \gets $ a river edge in $ \riverEdgeList $ \\ 
+    % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
+    % $ \nodeEdge \gets $ the edge for $ \node $\\
+    % $ \edgeBoxInt \gets $ boolean indicating $ \nodeEdge $ intersects $ \riverEdgeListEach $ \\
 
     \begin{algorithmic}[1]
-        \Procedure{TestRiverCross}{$ \node, \nodeVPSC $}
+        \Procedure{TestRiverIntersection}{$ \node, \nodeVPSC $}
 
-        \ForEach{$ \riverEdgeListEach \in \riverEdgeList $}
-            \State $ \nodeBoundingBox \gets $ GenerateBoundingBox ($ \node, \nodeVPSC $)
-            \State $ \riverEdgeBoundingBoxEach \gets $ GenerateBoundingBox ($ \riverEdgeListEach $)
-            \State $ \boundingBoxInt \gets $ TestBoundingIntersection ($ \nodeBoundingBox, \riverEdgeBoundingBoxEach $)
+        \ForEach{$ \riverEdgeListEachLine \in \riverEdgeList $}
+            \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
+            \State $ \riverEdgeBoundingBoxEach \gets $ GetBoundingBox ($ \riverEdgeListEachLine $)
+            % \State $ \boundingBoxInt \gets $ TestIntersection ($ \nodeBoundingBox, \riverEdgeBoundingBoxEach $)
+            % \If{$ \boundingBoxInt = True $}
 
-            \If{$ \boundingBoxInt = True $}
-                \State $ e_{n} \gets $ GenerateEdge ($ \node, \nodeVPSC $)
-                \State $ \edgeBoxInt \gets $ TestEdgeIntersection ($ \nodeEdge, \riverEdgeListEach $)
-
-                \If{$ \edgeBoxInt = True $}
+            \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBoxEach = True $}
+                \If{$ \nodeLineNV ~intersect~ \riverEdgeListEachLine = True $}
                     \State \Return{$ True $}
@@
 
-\subsection{Corridor Derivation}
+\subsection{Detecting Stalemates}
 
-As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node might be repeatedly translated between two positions due to river crossing and congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-In order to create space for VPSC to generate a new layout, we introduce a parameter, $ \stalemate $, to test for stalemate.
-If a node is translated between two positions for more than $ \stalemate $ iterations, it is considered to be in a stalemate.
-A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ d $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).
+As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node might be repeatedly translated between two positions due to river intersection and congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+In order to create space for VPSC to generate a new layout, we introduce a parameter, $ \stalemateMax $, to test for stalemate.
+If a node is translated between two positions for more than $ \stalemateMax $ iterations, it is considered to be in a stalemate.
+A stalemate triggers a corridor derivation, {a corridor is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, in the direction of the node $ \node $'s previous and current position, denoted $ \nodeVectorPC $ }, all enclosed nodes are then translated in the same direction and by the distance of that vector to alleviate the congestion.
 
@@
     \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}.
-All nodes within the corridor are moved by $ d $ in the direction of $ \Vector{B}{A} $.
+All nodes within the corridor are moved by $ d $ units in the direction of $ \Vector{B}{A} $, e.g, A is moved to C.
 }
@@
     \textbf{Local variables:} \\
-    $ \PointP \gets $ the point extending $ \Vector{\node.previous}{~\node} $ such that the length of $ \Vector{\node.previous}{~\PointP} $ is $ \CorridorLength $\\
-    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \Vector{\node}{\PointP} $ \\
-    $ corridor \gets $ a rectangular formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
+    $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    $ \PointP \gets $ the point extending $ \nodeVectorPC $ such that the length of $ \nodeLineNpP $ is $ \CorridorLength $\\
+    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNpP $ \\
+    $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
 
     \begin{algorithmic}[1]
-        \Procedure{DeriveCorridor}{$ \node $}
+        \Procedure{ProcessStalemate}{$ \node $}
             \State TranslateNode ($ \node $)
 
-            \State $ \PointP \gets $ \Call{DerivePoint}{$ \Vector{\node.previous}{~\node} $, $ \CorridorLength $}
+            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorPC $, $ \CorridorLength $}
 
-            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \Vector{\node.previous}{~\PointP} $, $ \frac{\CorridorWidth}{2} $}
+            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ \frac{\CorridorWidth}{2} $}
 
-            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \Vector{\node.previous}{~\PointP} $, $ -\frac{\CorridorWidth}{2} $}
+            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ -\frac{\CorridorWidth}{2} $}
 
@@
 
+                % TODO: calculate the vector once and use it for all nodes
             \ForEach{$ \nodeInCorridorEach $ inside $ corridor $}
@@
         $ \Edge \gets $ the edge used to derive the new point \\
-        $ \Distance \gets $ the distance between $ \PointP $ and $ \Edge.start $ \\
+        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\
 
         \textbf{Output:} \\
-        A point, $ \PointP $, that is distance $ \Distance $ away from $ \Edge.start $. \\
+        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\
 
         \textbf{Local variables:} \\
-        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \Edge.start $ and $ \Edge.end $ \\
+        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
 
@@
             \Procedure{DerivePoint}{$ \Edge $, $ \Distance $}
-                \State $ \dx \gets \Edge.start.x - \Edge.end.x $
+                \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
 
-                \State $ \dy \gets \Edge.start.y - \Edge.end.y $
+                \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
 
@@
     \textbf{Local variables:} \\
-    $ \dx, \dy \gets $ the differences in $ x, y $ for $ \Edge.start $ and $ \Edge.end $ \\
+    $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
     $ \Scale \gets $ the scale of $ \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $ \\
@@
         \Procedure{DeriveParallelEdge}{$ \Edge $, $ \Distance $}
-            \State $ \dx \gets \Edge.start.x - \Edge.end.x $
+            \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
 
-            \State $ \dy \gets \Edge.start.y - \Edge.end.y $
+            \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
 
@@
 
-            \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \Edge.start.x $
+            \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \EdgeStart.x $
 
-            \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \Edge.start.y $
+            \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \EdgeStart.y $
 
-            \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \Edge.end.x $
+            \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \EdgeEnd.x $
 
-            \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \Edge.end.y $
+            \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \EdgeEnd.y $
 

### 2022-03-04 paper/section/related.tex
draft: update and review

@@
 
-\new{While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent areas but have distortion in shape.
-Non-contiguous cartograms sacrifice the topological connectivity to enable growing or shrinking in size to maintain their original shape \cite{dent2009Cartography}.}
+While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but have distortion in shape.
+Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable growing or shrinking in size while maintaining their original shape \cite{dent2009Cartography}.
 
-\citea{nusrat2016State} defines summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
-Each cartogram design may make accuracy trade-offs between dimensions, we provide an overview of these cartograms in Table \ref{table:accuracy}.
+\citea{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+Each cartogram design may make accuracy trade-offs between dimensions.
+We provide an overview of these cartograms in Table \ref{table:accuracy}.
 
@@
 Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
-\new{Here we focus on Demers cartograms as we use squares to depict regions.
-This style of cartogram offers the advantages that the comparisons between regions are more intuitive and the space utilization is more effcient.
+Here we focus on Demers cartograms as we use squares to depict regions.
+This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
 Demers cartograms also have the advantage of preserving a higher level of topology.
-}
 }
@@
 Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
-\new{Statistical comparisons between regions are difficult due to the differing shapes.}
 }
@@
 The Washington Post uses cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
-The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
+National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
 \citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
@@
 
-Compared to previous work on Demers cartograms, we introduce dynamic topological feature, a river, aiming to improve the interpretation, readability, and topological accuracy.
+Our work extends Demers cartograms. We introduce dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy.

### 2022-03-04 paper/GHRVis.tex
draft: EHR data description
draft: add algorithms
draft: update main section
draft: update
draft: update algorithm and review
draft: update and review
draft: latex structure change

@@
 
-% code listing
-\usepackage{listings}
-\lstdefinestyle{customc}{
-  belowcaptionskip=1\baselineskip,
-  breaklines=true,
-  xleftmargin=\parindent,
-  language=C,
-  showstringspaces=false,
-  basicstyle=\footnotesize\ttfamily,
-  keywordstyle=\bfseries\color{green!40!black},
-  commentstyle=\itshape\color{purple!40!black},
-  identifierstyle=\color{blue},
-  stringstyle=\color{orange},
-}
-
-\lstdefinestyle{customasm}{
-  belowcaptionskip=1\baselineskip,
-  xleftmargin=\parindent,
-  language=[x86masm]Assembler,
-  basicstyle=\footnotesize\ttfamily,
-  commentstyle=\itshape\color{purple!40!black},
-}
-
-\lstset{escapechar=@,style=customc}
-
 \usepackage{dblfloatfix}
@@
 \usepackage[table, dvipsnames]{xcolor}
+
+\usepackage{multicol}
+
+\include{preamble/preamble.basic}
+\include{preamble/preamble.main}
+\include{preamble/preamble.math}
+\include{preamble/preamble.draft}
 
@@
 
-\title{GHRVis:}
+\title{Cartograms with Dynamic Features}
 
@@
 \pagestyle{plain}
-
-% auto labeling of sections
-\let\origsection=\section
-\renewcommand\section[1]{\origsection{#1}\label{sec:{#1}}}
-
-\let\origsubsection=\subsection
-\renewcommand\subsection[1]{\origsubsection{#1}\label{subsec:{#1}}}
-
-\let\origsubsubsection=\subsubsection
-\renewcommand\subsubsection[1]{\origsubsubsection{#1}\label{subsubsec:{#1}}}
-
-\newcommand{\figureref}[1]{Figure \ref{#1}}
-\newcommand{\tableref}[1]{Table \ref{#1}}
-\newcommand{\phantomref}[1]{$\hyperref[#1]{\ref{#1}}$}
-
-% Bob's dream paragraph style
-\newcommand{\bobgraph}[1]{ \noindent\textbf{#1} \phantomsection \label{bobgraph:{#1}}}
-
-\definecolor{Mycolor1}{HTML}{4C5866} %#4C5866
-\definecolor{Mycolor2}{HTML}{A6B6D2} %#A6B6D2
-\definecolor{Mycolor3}{HTML}{67daff} %#67daff
-\definecolor{Mycolor4}{HTML}{F7F7FE} %#F7F7FE
-\definecolor{Mycolor5}{HTML}{666ad1} %#666ad1
-\definecolor{Mycolor6}{HTML}{ffbb93} %#ffbb93
-\newcommand{\mycolorbox}[2]{\begingroup\setlength{\fboxsep}{2pt}\colorbox{#1}{#2}\endgroup}
-
-% table cite
-\newcommand{\citea}[1]{\citeauthor{#1}\cite{#1}}
-
-% customized table headers
-\newcommand{\theaderR}[1]{\multicolumn{1}{c|}{\cellcolor{Mycolor3} \textbf{\rotatebox{90}{#1 \phantom{-} }}}}
-\newcommand{\theaderP}[1]{\cellcolor{Mycolor3} \textbf{#1}}
-
-\newcommand{\note}[1]{\textcolor{blue}{#1}}
-
-\newcommand{\expertcomment}[1]{\textit{``#1''}}
-
-\newcommand{\task}[1]{\textbf{T#1}}
-\newcommand{\req}[1]{\textbf{R#1}}
 
@@
 
-\section{GHRVis}
-
-\subsection{VPSC}
+\input{section/main}
 

### 2022-03-04 paper/section/main.tex
draft: latex structure change

@@
     \includegraphics[width=\columnwidth]{figure/flowchart.png}
-    \caption{A flowchart abstraction of our layout algorithm.}
+    \caption{A flowchart showing the abstraction of our layout algorithm.}
     \label{fig:flowchart}
@@
     $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    $ \PointP \gets $ the point extending $ \nodeVectorPC $ such that the length of $ \nodeLineNpP $ is $ \CorridorLength $\\
+    $ \PointP \gets $ the point extending $ \nodeVectorPC $ such that $ \nodeLineWidthNpP = \CorridorLength $\\
     $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNpP $ \\

### 2022-03-04 paper/section/related.tex
draft: latex structure change

@@
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
-See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+See \tableref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
 

### 2022-03-04 paper/section/main.tex
draft: update figures and algorithms

@@
 {
-\begin{figure}[b!]
+\begin{figure}[tb!]
     \centering
@@
     $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
+    $ \solved \gets True $ if no node crossed a river \\
     $ \crossRiver \gets $ boolean indicating a node crossed a river \\
@@
     $ \nodeListEachPrevious \gets $ the previous position of $ \nodeListEach $\\
-    $ \nodeListVPSCEach \gets $ a node in $ \nodeListVPSC $, which is the next position of $ \nodeListEach $ after VPSC \\
+    $ \nodeListVPSCEach \gets $ a node in $ \nodeListVPSC $, which is the position of $ \nodeListEach $ after VPSC \\
 
@@
 
+            \While{$ \solved = False $}
+
+                \State $ \solved \gets True $
                 \State $ \nodeListVPSC \gets $ RunVPSC ($ \nodeList $)
@@
 
-                \If {\Call{TranslateNode}{$ \nodeListEach, \nodeListVPSCEach $}}
-                    \State $ \crossRiver \gets $ \Call{TestRiverIntersection}{$ \nodeListEach, \nodeListVPSCEach $}
+                    \If {$ \nodeListEach(x,y) \neq \nodeListVPSCEach(x,y) $}
+
+                        \State $ \crossRiver \gets $ \Call{TestRiverIntersection}{$ \nodeListVPSCEachLine $}
+                        \State $ \nodeListEach(x,y) \gets \nodeListVPSCEach(x,y) $
 
                         \If{$ \crossRiver = True $}
+
+                            \State $ \solved \gets False $
                             \State $ \nodeListEachStalemate \gets \nodeListEachStalemate + 1 $
@@
                             \If{$ \nodeListEachStalemate < \stalemateMax $}
+
                                 \State $ \nodeListEach(x,y) \gets \nodeListEachPrevious(x,y) $
+
                             \Else
+
                                 \State \Call{ProcessStalemate}{$ \nodeListEach, \nodeListVPSCEach $}
                                 \State $ \nodeListEachStalemate \gets 0 $
+
                             \EndIf
@@
                 \EndFor
+
+            \EndWhile
 
@@
 
-\begin{algorithm}[tb!]
-    \caption{Procedure to update a single node, $ \node $'s position.}\label{alg:move position}
+% \begin{algorithm}[tb!]
+%     \caption{Procedure to update the position of a single node $ \node $.}\label{alg:move position}
 
-    \textbf{Input:} \\
-    $ \node \gets $ the node \\
-    $ \nodeVPSC \gets $ the node after running VPSC \\
+%     \textbf{Input:} \\
+%     $ \node \gets $ the node \\
+%     $ \nodeVPSC \gets $ the node after running VPSC \\
 
-    \textbf{Output:} \\
-    Returns $ True $ if the node's position is updated. \\
+%     \textbf{Output:} \\
+%     Returns $ True $ if the node's position is updated. \\
 
-    \begin{algorithmic}[1]
-        \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
-        \If{$ \node(x,y) = \nodeVPSC(x,y) $}
-            \item[] \Comment{Node position remains unchanged}
-            \State \Return{$ False $}
-        \EndIf
+%     \begin{algorithmic}[1]
+%         \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
+%         \If{$ \node(x,y) = \nodeVPSC(x,y) $}
+%             \item[] \Comment{Node position remains unchanged}
+%             \State \Return{$ False $}
+%         \EndIf
 
-        \State $ \node(x,y) \gets \nodeVPSC(x,y) $
+%         \State $ \node(x,y) \gets \nodeVPSC(x,y) $
 
-        \State \Return{$ True $}
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
+%         \State \Return{$ True $}
+%         \EndProcedure
+%     \end{algorithmic}
+% \end{algorithm}
 
@@
 \begin{algorithm}[tb!]
-    \caption{Procedure to test if a node, $ \node $ crosses a river.}\label{alg:check river intersection}
+    \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
     \textbf{Input:} \\
-    $ \node \gets $ the node \\
-    $ \nodeVPSC \gets $ the node after running VPSC \\
+    $ \nodeLineNV \gets $ the node's translation path \\
 
@@
     \begin{algorithmic}[1]
-        \Procedure{TestRiverIntersection}{$ \node, \nodeVPSC $}
+        \Procedure{TestRiverIntersection}{$ \nodeLineNV $}
 
@@
             \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBoxEach = True $}
-                \If{$ \nodeLineNV ~intersect~ \riverEdgeListEachLine = True $}
-                    \State \Return{$ True $}
-                \EndIf
-
+                \State \Return{$ \nodeLineNV ~intersect~ \riverEdgeListEachLine $}
                 \EndIf
@@
 
-As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node might be repeatedly translated between two positions due to river intersection and congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-In order to create space for VPSC to generate a new layout, we introduce a parameter, $ \stalemateMax $, to test for stalemate.
-If a node is translated between two positions for more than $ \stalemateMax $ iterations, it is considered to be in a stalemate.
-A stalemate triggers a corridor derivation, {a corridor is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, in the direction of the node $ \node $'s previous and current position, denoted $ \nodeVectorPC $ }, all enclosed nodes are then translated in the same direction and by the distance of that vector to alleviate the congestion.
+As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is derived to improve the congestion.
+\new{A corridor is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ so that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (\figref{fig:corridor}C and D) .} All enclosed nodes are then translated based on $ \nodeVectorPC $ to alleviate the congestion (\figref{fig:corridor}D).
 
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node repeatedly translates between two positions (A and B) for more than $ stalemate $ iterations. }
+    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times.
+A stalemate often occurs when the area is congested and the node is unable to move to a position that is not intersecting a river.}
     \label{fig:stalemate}
@@
     \includegraphics[width=\columnwidth]{figure/corridor.png}
-    \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}.
-All nodes within the corridor are moved by $ d $ units in the direction of $ \Vector{B}{A} $, e.g, A is moved to C.
-}
+    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times, as shown in A.
+To address the issue, we derive a corridor (orange rectangle in D) based on $ \node $ and $ \nodePrevious $.
+All nodes within the corridor are translated based on $ \nodeVectorPC $, such that $ \nodeVectorPC = \nodeVectorNNn = \nodeVectorNinNinn $.
+In order to show a clear illustration, we place nodes sparsely in this figure.}
     \label{fig:corridor}
@@
     \textbf{Global variables:} \\
-    $ \nodeSize \gets $ the current size of all nodes \\
     $ \CorridorLength \gets $ the length of a corridor \\
@@
         \Procedure{ProcessStalemate}{$ \node $}
-            \State TranslateNode ($ \node $)
+            \State $ \node(x,y) \gets \nodePrevious(x,y) $
 
@@
 
-                % TODO: calculate the vector once and use it for all nodes
             \ForEach{$ \nodeInCorridorEach $ inside $ corridor $}
-                \State $ \nodeInCorridorEachDx \gets \nodeInCorridorEach.x - \nodeInCorridorEach.previous.x $
 
-                \State $ \nodeInCorridorEachDy \gets \nodeInCorridorEach.y - \nodeInCorridorEach.previous.y $
+                \State $ \Vector{\nodeInCorridorEach ~\nodeInCorridorEach.next} = \nodeVectorPC $
 
-                \State $ hyp \gets \sqrt{(\nodeInCorridorEachDx)^2 +(\nodeInCorridorEachDy)^2} $
-
-                \State $ \nodeInCorridorEach.next \gets $ \Call{DerivePoint}{$ \Vector{\nodeInCorridorEach.previous}{~\nodeInCorridorEach} $, $ hyp + s $}
-
-                \State TranslateNode ($ \nodeInCorridorEach $, $ \nodeInCorridorEach.next $)
+                \State $ \nodeInCorridorEach(x,y) \gets \nodeInCorridorEach.next(x,y) $
 
@@
     $ \Edge \gets $ the edge used to derive the parallel edge $ \EdgeParallel $ \\
-    $ \Distance \gets $ the distance between $ \Edge $ and $ \EdgeParallel $ \\
+    $ \Distance \gets $ the shortest distance between $ \Edge $ and $ \EdgeParallel $ \\
 

### 2022-03-10 paper/section/appendix.tex
draft: update and review

--- 2022-03-04 paper/section/main.tex
@@
-\section{Cartograms with Dynamic Features}
+\appendix
 
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{An overview of \software. Figure TBA.}
-    \label{fig:overview}
-\end{figure}
-}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth]{figure/flowchart.png}
-    \caption{A flowchart showing the abstraction of our layout algorithm.}
-    \label{fig:flowchart}
-\end{figure}
-}
-
-\algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of our algorithm.
-\textbf{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a small square node with $ \nodeSize $ as the initial size.
-\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} repeatedly to remove overlaps.
-We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-During each iteration, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
-Nodes that cross a river are translated back to their previous position.
-This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
-\new{The size growing process provides a stability in the layout and minimizes geographical errors.}
+\section{Procedure: DerivePoint}
 
@@
 
-\begin{algorithm}[tb!]
-    \caption{Procedure to update node positions by removing overlap and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
-    \textbf{Global variables:} \\
-    $ \nodeList \gets $ a list of nodes representing CCGs \\
-    $ \nodeSize \gets $ the current size of all nodes \\
-    $ \nodeSizeMax \gets $ the maximum size of a node \\
-    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
-
-    \textbf{Local variables:} \\
-    $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
-    $ \solved \gets True $ if no node crossed a river \\
-    $ \crossRiver \gets $ boolean indicating a node crossed a river \\
-    $ \nodeListEach \gets $ a node in $ \nodeList $ \\
-    $ \nodeListEachPrevious \gets $ the previous position of $ \nodeListEach $\\
-    $ \nodeListVPSCEach \gets $ a node in $ \nodeListVPSC $, which is the position of $ \nodeListEach $ after VPSC \\
-
-    \begin{algorithmic}[1]
-        \Procedure{UpdateNodePosition}{}
-        \While{$ \nodeSize < \nodeSizeMax $}
-
-            \While{$ \solved = False $}
-
-                \State $ \solved \gets True $
-                \State $ \nodeListVPSC \gets $ RunVPSC ($ \nodeList $)
-
-                \ForEach {$ \nodeListVPSCEach \in \nodeListVPSC $}
-
-                    \If {$ \nodeListEach(x,y) \neq \nodeListVPSCEach(x,y) $}
-
-                        \State $ \crossRiver \gets $ \Call{TestRiverIntersection}{$ \nodeListVPSCEachLine $}
-                        \State $ \nodeListEach(x,y) \gets \nodeListVPSCEach(x,y) $
-
-                        \If{$ \crossRiver = True $}
-
-                            \State $ \solved \gets False $
-                            \State $ \nodeListEachStalemate \gets \nodeListEachStalemate + 1 $
-
-                            \If{$ \nodeListEachStalemate < \stalemateMax $}
-
-                                \State $ \nodeListEach(x,y) \gets \nodeListEachPrevious(x,y) $
-
-                            \Else
-
-                                \State \Call{ProcessStalemate}{$ \nodeListEach, \nodeListVPSCEach $}
-                                \State $ \nodeListEachStalemate \gets 0 $
-
-                            \EndIf
-
-                        \EndIf
-
-                    \EndIf
-
-                \EndFor
-
-            \EndWhile
-
-            \State $ \nodeSize \gets \nodeSize ++$
-
-        \EndWhile
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
-% \begin{algorithm}[tb!]
-%     \caption{Procedure to update the position of a single node $ \node $.}\label{alg:move position}
-
-%     \textbf{Input:} \\
-%     $ \node \gets $ the node \\
-%     $ \nodeVPSC \gets $ the node after running VPSC \\
-
-%     \textbf{Output:} \\
-%     Returns $ True $ if the node's position is updated. \\
-
-%     \begin{algorithmic}[1]
-%         \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
-%         \If{$ \node(x,y) = \nodeVPSC(x,y) $}
-%             \item[] \Comment{Node position remains unchanged}
-%             \State \Return{$ False $}
-%         \EndIf
-
-%         \State $ \node(x,y) \gets \nodeVPSC(x,y) $
-
-%         \State \Return{$ True $}
-%         \EndProcedure
-%     \end{algorithmic}
-% \end{algorithm}
-
-%\end{noindent}
-
-\subsection{River Intersection Testing}
-
-We use rivers as topological boundaries and prevent nodes from crossing them.
-\new{The resolution of rivers can be dynamically adjusted, as shown in \figref{fig:river resolution}}.
-When a node's position changes, we test if the node's trajectory intersects any segment of a river.
-See \algoref{alg:check river intersection}.
-A bounding box intersection test can be performed to reduce the number of edge intersection detections required.
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth]{figure/river_resolution.png}
-    \caption{The resolution of rivers can be dynamically adjusted.
-The top figure shows River Thames at its original resolution.
-The bottom figure shows the river with a 60\% reduction in resolution.
-The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
-    \label{fig:river resolution}
-\end{figure}
-}
-
-% \begin{noindent}
-\begin{algorithm}[tb!]
-    \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
-    \textbf{Input:} \\
-    $ \nodeLineNV \gets $ the node's translation path \\
-
-    \textbf{Output:} \\
-    Returns $ True $ if the node crosses a river. \\
-
-    \textbf{Global variables:} \\
-    $ \riverEdgeList \gets $ a list of river edges \\
-
-    \textbf{Local variables:} \\
-    $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeListEachLine $ \\
-    $ \riverEdgeListEachLine \gets $ a river edge in $ \riverEdgeList $ \\ 
-    % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
-    % $ \nodeEdge \gets $ the edge for $ \node $\\
-    % $ \edgeBoxInt \gets $ boolean indicating $ \nodeEdge $ intersects $ \riverEdgeListEach $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TestRiverIntersection}{$ \nodeLineNV $}
-
-        \ForEach{$ \riverEdgeListEachLine \in \riverEdgeList $}
-            \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
-            \State $ \riverEdgeBoundingBoxEach \gets $ GetBoundingBox ($ \riverEdgeListEachLine $)
-            % \State $ \boundingBoxInt \gets $ TestIntersection ($ \nodeBoundingBox, \riverEdgeBoundingBoxEach $)
-            % \If{$ \boundingBoxInt = True $}
-
-            \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBoxEach = True $}
-                \State \Return{$ \nodeLineNV ~intersect~ \riverEdgeListEachLine $}
-                \EndIf
-        \EndFor
-
-        \State \Return{$ False $}
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-%\end{noindent}
-
-\subsection{Detecting Stalemates}
-
-As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is derived to improve the congestion.
-\new{A corridor is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ so that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (\figref{fig:corridor}C and D) .} All enclosed nodes are then translated based on $ \nodeVectorPC $ to alleviate the congestion (\figref{fig:corridor}D).
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times.
-A stalemate often occurs when the area is congested and the node is unable to move to a position that is not intersecting a river.}
-    \label{fig:stalemate}
-\end{figure}
-}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth]{figure/corridor.png}
-    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times, as shown in A.
-To address the issue, we derive a corridor (orange rectangle in D) based on $ \node $ and $ \nodePrevious $.
-All nodes within the corridor are translated based on $ \nodeVectorPC $, such that $ \nodeVectorPC = \nodeVectorNNn = \nodeVectorNinNinn $.
-In order to show a clear illustration, we place nodes sparsely in this figure.}
-    \label{fig:corridor}
-\end{figure}
-}
-
-% \begin{noindent}
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to derive a corridor to translate enclosed nodes.
-We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards.}\label{alg:derive corridor}
-
-    \textbf{Input:} \\
-    $ \node \gets $ the node used to derive the corridor \\
-
-    \textbf{Global variables:} \\
-    $ \CorridorLength \gets $ the length of a corridor \\
-    $ \CorridorWidth \gets $ the width of a corridor \\
-
-    \textbf{Local variables:} \\
-    $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    $ \PointP \gets $ the point extending $ \nodeVectorPC $ such that $ \nodeLineWidthNpP = \CorridorLength $\\
-    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNpP $ \\
-    $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{ProcessStalemate}{$ \node $}
-            \State $ \node(x,y) \gets \nodePrevious(x,y) $
-
-            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorPC $, $ \CorridorLength $}
-
-            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ \frac{\CorridorWidth}{2} $}
-
-            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ -\frac{\CorridorWidth}{2} $}
-
-            \State $ corridor \gets
-                \begin{bmatrix}
-                    \EdgeParallelA.start &
-                    \EdgeParallelA.end \\
-
-                    \EdgeParallelB.start &
-                    \EdgeParallelB.end \\
-                \end{bmatrix} $
-
-            \ForEach{$ \nodeInCorridorEach $ inside $ corridor $}
-
-                \State $ \Vector{\nodeInCorridorEach ~\nodeInCorridorEach.next} = \nodeVectorPC $
-
-                \State $ \nodeInCorridorEach(x,y) \gets \nodeInCorridorEach.next(x,y) $
-
-            \EndFor
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
-%\end{noindent}
-
-% \begin{noindent}
-
-    \begin{algorithm}[tb!]
+    \begin{algorithm}[h!]
         \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
@@
 
+\section{Procedure: DeriveParallelEdge}
+
 % \begin{noindent}
 
-\begin{algorithm}[tb!]
+\begin{algorithm}[h!]
     \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

### 2022-03-10 paper/section/main.tex
draft: update and review

@@
     \includegraphics[width=\columnwidth]{example-image-a}
-    \caption{An overview of \software. Figure TBA.}
+    \caption{An screenshot of \software. Figure TBA.}
     \label{fig:overview}
@@
 \textbf{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a small square node with $ \nodeSize $ as the initial size.
-\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} repeatedly to remove overlaps.
+For each CCG we compute the centroid and represent it using a small unit size square node with $ \nodeSize $ as the initial size.
+\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} to remove overlaps.
 We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-During each iteration, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
+During overlap removal, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
 Nodes that cross a river are translated back to their previous position.
@@
     $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
-    $ \solved \gets True $ if no node crossed a river \\
-    $ \crossRiver \gets $ boolean indicating a node crossed a river \\
+    $ \nodeListCross \gets True $ if any node in the list crossed a river \\
+    $ \nodeCross \gets True $ if the node crossed a river \\
     $ \nodeListEach \gets $ a node in $ \nodeList $ \\
@@
 
-            \While{$ \solved = False $}
+            \While{$ \nodeListCross = True $}
 
-                \State $ \solved \gets True $
-                \State $ \nodeListVPSC \gets $ RunVPSC ($ \nodeList $)
+                \State $ \nodeListCross \gets False $
+                \State $ \nodeListVPSC \gets $ RemoveOverlap ($ \nodeList $)
 
@@
 
-                        \State $ \crossRiver \gets $ \Call{TestRiverIntersection}{$ \nodeListVPSCEachLine $}
+                        \State $ \nodeCross \gets $ \Call{TestRiverIntersection}{$ \nodeListVPSCEachLine $}
                         \State $ \nodeListEach(x,y) \gets \nodeListVPSCEach(x,y) $
 
-                        \If{$ \crossRiver = True $}
+                        \If{$ \nodeCross = True $}
 
-                            \State $ \solved \gets False $
+                            \State $ \nodeListCross \gets True $
                             \State $ \nodeListEachStalemate \gets \nodeListEachStalemate + 1 $
@@
                                 \State \Call{ProcessStalemate}{$ \nodeListEach, \nodeListVPSCEach $}
-                                \State $ \nodeListEachStalemate \gets 0 $
+                                \State $ \nodeListEachStalemate \gets 0 $ \Comment{Reset counter}
 
@@
 We use rivers as topological boundaries and prevent nodes from crossing them.
-\new{The resolution of rivers can be dynamically adjusted, as shown in \figref{fig:river resolution}}.
+The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from 10,170 to 65.
+Enabling river simplification reduces the number of edge intersection tests that need to be performed.
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
 See \algoref{alg:check river intersection}.
-A bounding box intersection test can be performed to reduce the number of edge intersection detections required.
+A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection detections required.
 
@@
     \includegraphics[width=\columnwidth]{figure/river_resolution.png}
-    \caption{The resolution of rivers can be dynamically adjusted.
-The top figure shows River Thames at its original resolution.
-The bottom figure shows the river with a 60\% reduction in resolution.
+    \caption{The resolution of rivers can be dynamically adjusted by the user.
+The top figure shows River Thames at its original resolution with 10,170 edges.
+The bottom figure shows the river at a  reduced resolution of 65 edges.
 The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
@@
     \textbf{Local variables:} \\
+    $ \riverEdgeListEachLine \gets $ a river edge in $ \riverEdgeList $ \\ 
     $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeListEachLine $ \\
-    $ \riverEdgeListEachLine \gets $ a river edge in $ \riverEdgeList $ \\ 
     % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
@@
 As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is derived to improve the congestion.
-\new{A corridor is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ so that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (\figref{fig:corridor}C and D) .} All enclosed nodes are then translated based on $ \nodeVectorPC $ to alleviate the congestion (\figref{fig:corridor}D).
+If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is derived to break up the congestion.
+A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (See \figref{fig:corridor}C and D).
+All enclosed nodes are then translated by $ \nodeVectorPC $ to alleviate the congestion (See \figref{fig:corridor}D).
 
@@
     \caption{Procedure to derive a corridor to translate enclosed nodes.
-We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards.}\label{alg:derive corridor}
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \figref{fig:stalemate}).}\label{alg:derive corridor}
 
@@
     \textbf{Local variables:} \\
+    $ \Corridor \gets $ the corridor \\
     $ \nodePrevious \gets $ the previous position of $ \node $ \\
@@
 
-            \State $ corridor \gets
+            \State $ \Corridor \gets
                 \begin{bmatrix}
@@
 
-            \ForEach{$ \nodeInCorridorEach $ inside $ corridor $}
+            \ForEach{$ \nodeInCorridorEach $ inside $ \Corridor $}
 
-                \State $ \Vector{\nodeInCorridorEach ~\nodeInCorridorEach.next} = \nodeVectorPC $
+                \State $ \Vector{\nodeInCorridorEach ~\nodeInCorridorEachT} = \nodeVectorPC $
 
-                \State $ \nodeInCorridorEach(x,y) \gets \nodeInCorridorEach.next(x,y) $
+                \State $ \nodeInCorridorEach(x,y) \gets \nodeInCorridorEachT(x,y) $
 
@@
 %\end{noindent}
-
-% \begin{noindent}
-
-    \begin{algorithm}[tb!]
-        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
-        \textbf{Input:} \\
-        $ \Edge \gets $ the edge used to derive the new point \\
-        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\
-
-        \textbf{Output:} \\
-        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\
-
-        \textbf{Local variables:} \\
-        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
-
-        \begin{algorithmic}[1]
-            \Procedure{DerivePoint}{$ \Edge $, $ \Distance $}
-                \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
-
-                \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
-
-                \State $ \PointP.x \gets \frac{\dx}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
-
-                \State $ \PointP.y \gets \frac{\dy}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
-
-            \State \Return{$ \PointP $}
-
-            \EndProcedure
-
-        \end{algorithmic}
-    \end{algorithm}
-
-    %\end{noindent}
-
-% \begin{noindent}
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}
-
-    \textbf{Input:} \\
-    $ \Edge \gets $ the edge used to derive the parallel edge $ \EdgeParallel $ \\
-    $ \Distance \gets $ the shortest distance between $ \Edge $ and $ \EdgeParallel $ \\
-
-    \textbf{Output:} \\
-    An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $. \\
-
-    \textbf{Local variables:} \\
-    $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
-    $ \Scale \gets $ the scale of $ \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{DeriveParallelEdge}{$ \Edge $, $ \Distance $}
-            \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
-
-            \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
-
-            \State $ \Scale \gets \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $
-
-            \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \EdgeStart.x $
-
-            \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \EdgeStart.y $
-
-            \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \EdgeEnd.x $
-
-            \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \EdgeEnd.y $
-
-        \State \Return{$ \EdgeParallel $}
-
-        \EndProcedure
-
-    \end{algorithmic}
-\end{algorithm}
-
-%\end{noindent}

### 2022-03-11 paper/section/data.tex

@@
 
-Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
+Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
 The number of CCGs changes over time due to the NHS re-organization.
@@
 We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
-These rivers were chosen as they pass through regions with dense population, and provide informative geographical and topological cues.
+These rivers were chosen as they are well-known rivers and pass through regions with dense population, and provide informative geographical and topological cues.
 
@@
 After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and scales, and merge them into one shapefile.
-Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size and improve the performance of our implementation.
+Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file.
+TopoJSON eliminates redundant coordinates in the data, this improves the rendering speed of our implementation.
 
@@
 \end{itemize}
+
+Each CCG has a unique ONS code, which is used to link the CCG shapefile with the statistical data.

### 2022-03-11 paper/section/introduction.tex

@@
 
-Cartograms are representations of geographical space generated by a value-by-area mapping technique that combines statistical and geographical information \cite{dent2009Cartography}.
-\new{Various styles of cartograms have been proposed and implemented, covering topics on urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.}
+Cartograms are representations of hybrid geographical and data space based on a value-by-area mapping technique that combines statistical and geographical information \cite{dent2009Cartography}.
+Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.
 
 In this paper, we follow the categorization of cartograms by \citea{nusrat2016State}: contiguous, non-contiguous, rectangular, and Dorling.
-Among these four types, trade-offs between accuracies are made (See \tableref{table:accuracy}).
-Building on Demers cartograms \cite{ian2002Cartogram}, we develop dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-We incorporate a cartographic layout algorithm that includes these topological features to update the layout of the nodes representing geographical regions.
-To minimize geographical errors, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm a real world Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
+Among these four types, a trade-off between accuracy is made (See \tableref{table:accuracy}).
+For this project we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
+Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
+We implement a cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions.
+To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
+We then apply the algorithm a real world case-study using an Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
 
@@
     \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers.
+    \item A user study evaluation of the technique with an application to EHRs
 \end{itemize}

### 2022-03-11 paper/section/related.tex
draft: update figures and algorithms

@@
 
-While we focus on rectangular and Dorling cartograms in this paper, we provide brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but have distortion in shape.
-Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable growing or shrinking in size while maintaining their original shape \cite{dent2009Cartography}.
+While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
+Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
 
@@
 \renewcommand{\arraystretch}{1.5}
-\begin{table}[!b]
+\begin{table}[!tb]
 	\centering
@@
 \renewcommand{\arraystretch}{1.5}
-\begin{table*}[!htb]
+\begin{table*}[!tb]
 	\centering
@@
 
-Our work extends Demers cartograms. We introduce dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy.
+One of the disadvantages of Dorling and Demers cartograms is the legibility.
+The layout algorithms may displace regions far from their original position and making them more difficult to interpret.
+Our work extends Demers cartograms.
+We introduce a new type of dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.

### 2022-03-18 paper/section/introduction.tex
draft: review

@@
 
-Cartograms are representations of hybrid geographical and data space based on a value-by-area mapping technique that combines statistical and geographical information \cite{dent2009Cartography}.
+Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information \cite{dent2009Cartography}.
 Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.
 
-In this paper, we follow the categorization of cartograms by \citea{nusrat2016State}: contiguous, non-contiguous, rectangular, and Dorling.
-Among these four types, a trade-off between accuracy is made (See \tableref{table:accuracy}).
+Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \tableref{table:accuracy}).
 For this project we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-We implement a cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions.
+We implement a new cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions.
 To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm a real world case-study using an Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
+One of the major challenges involves is how to develop a layout algorithm that handles different types of nodes.
+In other words, the layout algorithm is novel because it handles different types of nodes: rectangular representing regions and polylines representing rivers.
+We then apply the algorithm to a real world case-study using an Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
+We also present a user study demonstrating its effectiveness.
 

### 2022-03-18 paper/section/main.tex
draft: review

@@
     \includegraphics[width=\columnwidth]{figure/flowchart.png}
-    \caption{A flowchart showing the abstraction of our layout algorithm.}
+    \caption{A flowchart showing our layout algorithm.}
     \label{fig:flowchart}
@@
 \textbf{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a small unit size square node with $ \nodeSize $ as the initial size.
+For each CCG we compute the centroid and represent it using a small unit size square node with initial size, $ \nodeSize $.
+We start with small unit size squares because they do not overlap initially.
 \textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} to remove overlaps.
-We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+We chose VPSC over other node overlap removal algorithms because VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+We gradually increase the node size by 1 unit at a time to ensure smooth transitions.
+An increase in size can cause the nodes to overlap.
 During overlap removal, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
@@
 We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
-\new{The size growing process provides a stability in the layout and minimizes geographical errors.}
+The gradual size increase process provides stability to the layout and helps minimize geographical error.
 
@@
     $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
-    $ \nodeListCross \gets True $ if any node in the list crossed a river \\
-    $ \nodeCross \gets True $ if the node crossed a river \\
-    $ \nodeListEach \gets $ a node in $ \nodeList $ \\
+    $ \nodeListCross \gets True $ if any node in the list crosses a river \\
+    $ \nodeCross \gets True $ if node crosses a river \\
+    % $ \nodeListEach \gets $ a node in $ \nodeList $ \\
     $ \nodeListEachPrevious \gets $ the previous position of $ \nodeListEach $\\
@@
 
+                \State \Comment{We should merge two lists into one}
                 \State $ \nodeListCross \gets False $
@@
 We use rivers as topological boundaries and prevent nodes from crossing them.
-The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from 10,170 to 65.
+The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from its original 10,170 to 65.
+We do not require a pixel-level river resolution as Demers cartograms do not represent geography at this level od detail, instead a resolution that matches the nodes is used.
 Enabling river simplification reduces the number of edge intersection tests that need to be performed.
@@
 
-As the VPSC always tries to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is derived to break up the congestion.
+As the VPSC always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is constructed to break up the congestion.
 A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (See \figref{fig:corridor}C and D).
-All enclosed nodes are then translated by $ \nodeVectorPC $ to alleviate the congestion (See \figref{fig:corridor}D).
+All enclosed nodes are then translated by $ \nodeVectorNC $ to alleviate the congestion (See \figref{fig:corridor}D).
 
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times.
+    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCN $ intersects a river for $ \stalemateMax $ times.
 A stalemate often occurs when the area is congested and the node is unable to move to a position that is not intersecting a river.}
@@
     \includegraphics[width=\columnwidth]{figure/corridor.png}
-    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCP $ intersects a river for $ \stalemateMax $ times, as shown in A.
-To address the issue, we derive a corridor (orange rectangle in D) based on $ \node $ and $ \nodePrevious $.
-All nodes within the corridor are translated based on $ \nodeVectorPC $, such that $ \nodeVectorPC = \nodeVectorNNn = \nodeVectorNinNinn $.
+    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCN $ intersects a river for $ \stalemateMax $ times, as shown in A.
+To address the issue, we derive a corridor (orange rectangle in D) based on $ \node $ and $ \nodeVPSC $.
+All nodes within the corridor are translated based on $ \nodeVectorNC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
 In order to show a clear illustration, we place nodes sparsely in this figure.}
@@
     $ \Corridor \gets $ the corridor \\
-    $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    $ \PointP \gets $ the point extending $ \nodeVectorPC $ such that $ \nodeLineWidthNpP = \CorridorLength $\\
+    $ \nodePrevious \gets $ the previous position of $ \node $ (this here is confusing because $ \nodePrevious $ is actually $ \nodeVPSC $ in the previous iteration)\\
+    $ \PointP \gets $ the point extending $ \nodeVectorNC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
     $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNpP $ \\

### 2022-03-18 paper/section/related.tex
draft: review

@@
 
-While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms for interested readers: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
+While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
@@
 \citea{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
-Each cartogram design may make accuracy trade-offs between dimensions.
+Each cartogram design may make various types of accuracy trade-offs between dimensions.
 We provide an overview of these cartograms in Table \ref{table:accuracy}.
@@
 This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
+This is important in our use case scenario involving EHRs.
 Demers cartograms also have the advantage of preserving a higher level of topology.
@@
 Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
-The Washington Post uses cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
+The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
 National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.

### 2022-03-25 paper/section/data.tex
draft: review

@@
 The second step is to pre-process the EHR data to remove empty and erroneous values.
-The final step is to transform the data into a format that is suitable for visualization.
+The final step is to transform the data into a format that is suitable for cartograms.
 
@@
 
-\bobgraph{Clinical Commissioning Groups Shapefile}
-
-Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
+\bobgraph{Clinical Commissioning Groups Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
 The number of CCGs changes over time due to the NHS re-organization.
@@
 
-\bobgraph{River Features}
-
-We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
+\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
 These rivers were chosen as they are well-known rivers and pass through regions with dense population, and provide informative geographical and topological cues.
@@
 
-After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and scales, and merge them into one shapefile.
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, translations and scales, and merge them into one shapefile.
 Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file.

### 2022-03-25 paper/section/introduction.tex
draft: review

@@
 To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-One of the major challenges involves is how to develop a layout algorithm that handles different types of nodes.
-In other words, the layout algorithm is novel because it handles different types of nodes: rectangular representing regions and polylines representing rivers.
-We then apply the algorithm to a real world case-study using an Electronic Health Records (EHR) dataset to assess the accuracy of our algorithm.
+We then apply the algorithm to a real world case-study using an Electronic Health Records (EHR) dataset to evaluate of our algorithm.
 We also present a user study demonstrating its effectiveness.
@@
 \end{itemize}
+
+\new{One of the major challenges involves is how to develop a layout algorithm that handles different shapes.
+In other words, the layout algorithm is novel because it handles different types of nodes: rectangular representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to revolve stalemate situations introduced by dynamic topological features while minimizing the geographical error.}

### 2022-03-25 paper/section/main.tex
draft: review

@@
         \includegraphics[width=\columnwidth]{figure/flowchart.png}
-    \caption{A flowchart showing our layout algorithm.}
+        \caption{A flowchart illustration of our layout algorithm. See also \algoref{alg:UpdateNodePosition} for more detail.}
         \label{fig:flowchart}
+    \end{figure}
+
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/flowchart2.png}
+        \caption{An awesome flowchart illustration of our layout algorithm. See also \algoref{alg:UpdateNodePosition} for more detail.}
+        \label{fig:flowchart2}
     \end{figure}
@@
 
-\algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of our algorithm.
+\algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of the process.
 \textbf{Initialization:} We first load and render the CCG geospatial boundaries.
@@
 We start with small unit size squares because they do not overlap initially.
-\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} to remove overlaps.
-We chose VPSC over other node overlap removal algorithms because VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
+We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
 We gradually increase the node size by 1 unit at a time to ensure smooth transitions.
-An increase in size can cause the nodes to overlap.
+An increase in $ \nodeSize $ can cause the nodes to overlap.
 During overlap removal, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
@@
     \textbf{Global variables:} \\
-    $ \nodeList \gets $ a list of nodes representing CCGs \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs with the following properties: \\
+    \-\hspace{1em} $ \nodeListCross \gets True $ if any $ \node $ in $ \nodeList $ crosses a river \\
     $ \nodeSize \gets $ the current size of all nodes \\
@@
     \textbf{Local variables:} \\
-    $ \nodeListVPSC \gets $ a list of nodes representing CCGs after VPSC \\
-    $ \nodeListCross \gets True $ if any node in the list crosses a river \\
-    $ \nodeCross \gets True $ if node crosses a river \\
-    % $ \nodeListEach \gets $ a node in $ \nodeList $ \\
-    $ \nodeListEachPrevious \gets $ the previous position of $ \nodeListEach $\\
-    $ \nodeListVPSCEach \gets $ a node in $ \nodeListVPSC $, which is the position of $ \nodeListEach $ after VPSC \\
+    $ \node \gets $ a node is an object with the following property: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets True $ if $ \node $ crosses a river \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
 
@@
 
-                \State \Comment{We should merge two lists into one}
                 \State $ \nodeListCross \gets False $
-                \State $ \nodeListVPSC \gets $ RemoveOverlap ($ \nodeList $)
+                \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
 
-                \ForEach {$ \nodeListVPSCEach \in \nodeListVPSC $}
+                \ForEach {$ \node \in \nodeList $}
 
-                    \If {$ \nodeListEach(x,y) \neq \nodeListVPSCEach(x,y) $}
+                    \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
 
-                        \State $ \nodeCross \gets $ \Call{TestRiverIntersection}{$ \nodeListVPSCEachLine $}
-                        \State $ \nodeListEach(x,y) \gets \nodeListVPSCEach(x,y) $
+                        \State $ \node(x,y) \gets \nodeFNOR(x,y) $
+                        \State $ \nodeCross \gets $ \Call{TestRiverIntersection}{$ \nodeFNORLine $}
 
@@
                             \State $ \nodeListCross \gets True $
-                            \State $ \nodeListEachStalemate \gets \nodeListEachStalemate + 1 $
+                            \State $ \nodeStalemate \gets \nodeStalemate + 1 $
 
-                            \If{$ \nodeListEachStalemate < \stalemateMax $}
-
-                                \State $ \nodeListEach(x,y) \gets \nodeListEachPrevious(x,y) $
+                            \If{$ \nodeStalemate < \stalemateMax $}
+                                \State \hskip-2em \Comment{Transalte back to previous position}
+                                \State $ \node(x,y) \gets \nodePrevious(x,y) $
 
@@
 
-                                \State \Call{ProcessStalemate}{$ \nodeListEach, \nodeListVPSCEach $}
-                                \State $ \nodeListEachStalemate \gets 0 $ \Comment{Reset counter}
+                                \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
+                                \State $ \nodeStalemate \gets 0 $ \Comment{Reset counter}
 
@@
 
-% \begin{algorithm}[tb!]
-%     \caption{Procedure to update the position of a single node $ \node $.}\label{alg:move position}
-
-%     \textbf{Input:} \\
-%     $ \node \gets $ the node \\
-%     $ \nodeVPSC \gets $ the node after running VPSC \\
-
-%     \textbf{Output:} \\
-%     Returns $ True $ if the node's position is updated. \\
-
-%     \begin{algorithmic}[1]
-%         \Procedure{TranslateNode}{$ \node, \nodeVPSC $}
-%         \If{$ \node(x,y) = \nodeVPSC(x,y) $}
-%             \item[] \Comment{Node position remains unchanged}
-%             \State \Return{$ False $}
-%         \EndIf
-
-%         \State $ \node(x,y) \gets \nodeVPSC(x,y) $
-
-%         \State \Return{$ True $}
-%         \EndProcedure
-%     \end{algorithmic}
-% \end{algorithm}
-
 %\end{noindent}
@@
 The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from its original 10,170 to 65.
-We do not require a pixel-level river resolution as Demers cartograms do not represent geography at this level od detail, instead a resolution that matches the nodes is used.
+We do not require a pixel-level river resolution as Demers cartograms do not represent geography at this level od detail.
+Instead a resolution that matches the node size is used.
 Enabling river simplification reduces the number of edge intersection tests that need to be performed.
@@
 See \algoref{alg:check river intersection}.
-A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection detections required.
+A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
 
@@
     \textbf{Local variables:} \\
-    $ \riverEdgeListEachLine \gets $ a river edge in $ \riverEdgeList $ \\ 
-    $ \nodeBoundingBox, \riverEdgeBoundingBoxEach \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeListEachLine $ \\
-    % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBoxEach $ \\
+    $ \riverEdgeLine \gets $ a river edge in $ \riverEdgeList $ \\ 
+    $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeLine $ \\
+    % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBox $ \\
     % $ \nodeEdge \gets $ the edge for $ \node $\\
@@
 
-        \ForEach{$ \riverEdgeListEachLine \in \riverEdgeList $}
+        \ForEach{$ \riverEdgeLine \in \riverEdgeList $}
             \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
-            \State $ \riverEdgeBoundingBoxEach \gets $ GetBoundingBox ($ \riverEdgeListEachLine $)
-            % \State $ \boundingBoxInt \gets $ TestIntersection ($ \nodeBoundingBox, \riverEdgeBoundingBoxEach $)
-            % \If{$ \boundingBoxInt = True $}
+            \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \riverEdgeLine $)
 
-            \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBoxEach = True $}
-                \State \Return{$ \nodeLineNV ~intersect~ \riverEdgeListEachLine $}
+            \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
+                \State \Return{$ \nodeLineNV ~intersect~ \riverEdgeLine $}
                 \EndIf
@@
 
-As the VPSC always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might be repeatedly intersecting a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions for $ \stalemateMax $ iterations, a corridor is constructed to break up the congestion.
-A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNpP $ (See \figref{fig:corridor}C and D).
-All enclosed nodes are then translated by $ \nodeVectorNC $ to alleviate the congestion (See \figref{fig:corridor}D).
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations, we introduce a heuristic solution: constructing a corridor to break up the congestion.
+A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \figref{fig:corridor}C and D).
+All enclosed nodes are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \figref{fig:corridor}D).
 
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCN $ intersects a river for $ \stalemateMax $ times.
-A stalemate often occurs when the area is congested and the node is unable to move to a position that is not intersecting a river.}
+    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (in iteration 1) intersects a river for $ \stalemateMax $ times.
+The node is translated back to its previous position (iteration 2).
+A stalemate often occurs when the area is congested and the node is unable to translate to a position without intersecting a river.}
     \label{fig:stalemate}
@@
     \includegraphics[width=\columnwidth]{figure/corridor.png}
-    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCN $ intersects a river for $ \stalemateMax $ times, as shown in A.
-To address the issue, we derive a corridor (orange rectangle in D) based on $ \node $ and $ \nodeVPSC $.
-All nodes within the corridor are translated based on $ \nodeVectorNC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
+    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in A.
+To address the issue, we derive a corridor (orange rectangle in E) based on $ \node $ and $ \nodeFNOR $.
+All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
 In order to show a clear illustration, we place nodes sparsely in this figure.}
@@
     $ \Corridor \gets $ the corridor \\
-    $ \nodePrevious \gets $ the previous position of $ \node $ (this here is confusing because $ \nodePrevious $ is actually $ \nodeVPSC $ in the previous iteration)\\
-    $ \PointP \gets $ the point extending $ \nodeVectorNC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
-    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNpP $ \\
+    $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
+    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
     $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
@@
 
-            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorPC $, $ \CorridorLength $}
+            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}
 
-            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ \frac{\CorridorWidth}{2} $}
+            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}
 
-            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNpP $, $ -\frac{\CorridorWidth}{2} $}
+            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}
 
@@
 
-            \ForEach{$ \nodeInCorridorEach $ inside $ \Corridor $}
+            \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}
 
-                \State $ \Vector{\nodeInCorridorEach ~\nodeInCorridorEachT} = \nodeVectorPC $
+                \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $
 
-                \State $ \nodeInCorridorEach(x,y) \gets \nodeInCorridorEachT(x,y) $
+                \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $
 

### 2022-03-25 paper/section/related.tex
draft: review

@@
 
-\subsection{Implementations}
+\new{Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates static topological features into Demers cartograms.
+Making topological features dynamic further improves the geographical accuracy and legibility.}
+
+\subsection{Implementation in Popular Media}
 
@@
 One of the disadvantages of Dorling and Demers cartograms is the legibility.
-The layout algorithms may displace regions far from their original position and making them more difficult to interpret.
+The layout algorithms may displace regions far from their original position and making the maps more difficult to interpret.
 Our work extends Demers cartograms.

### 2022-04-02 paper/GHRVis.tex
draft: review

@@
 \begin{abstract}
-
+Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information.
+Based on Dorling cartogram, Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
+One of the disadvantages of Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.
+To address that, we propose a novel layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
+We evaluate the proposed layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
+The results demonstrate that the proposed layout algorithm is able to preserve the original layout of the cartogram, while improving the legibility, readability, and geographical accuracy of the cartogram.
 \end{abstract}
@@
 
+\input{section/appendix}
+
 \end{document}

### 2022-04-02 paper/section/appendix.tex
draft: review

@@
 \appendix
+
+\section{Reproject Coordinate Reference System (CRS) of Shapefiles}
+Shapefiles obtained use different CRS for projections.
+The river shapefiles use EPSG:4326 (WGS-84), whereas the CCG shapefile uses EPSG:27700 (OSGB36 - British National Grid).
+We first load all three river shapefiles into QGIS, followed by the CCG shapefile.
+QGIS automatically prompts to transform the latter into WGS-84, see \figref{fig:reproject}.
+The reprojection result is shown in \figref{fig:all_rivers}.
+
+{
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/qgis/reproject.png}
+    \caption{QGIS interface. A window prompt to transform the CCG shapefile's CRS into WGS-84.}
+    \label{fig:reproject}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{figure/qgis/all_rivers.png}
+    \caption{The reprojection result produced by \figref{fig:reproject}, which includes the River Trent (red), River Great Ouse (green), and River Thames (orange).}
+    \label{fig:all_rivers}
+\end{figure}
+}
 

### 2022-04-02 paper/section/data.tex
draft: review

@@
 Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHRa}.
-The data acquisition involves multiple steps.
 The first step is to obtain both geospatial boundaries and EHR data that match one another.
@@
     \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography portalx \cite{opengeographyportalxOpen}.}
+    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography portalx \cite{opengeographyportalxOpen} with EPSG:27700 (OSGB36 - British National Grid) as the Coordinate Reference System (CRS).}
     \label{fig:ccg}
@@
     relation(2263653);>>;
+    // River Great Ouse: 2798097
+    // River Trent: 2863468
     out skel;

### 2022-04-02 paper/section/introduction.tex
draft: review

@@
 
-\new{Our contributions include:}
+Our contributions include:
 
 \begin{itemize}
-    \item A new variant of Demers cartograms that incorporates dynamic topological features to improve readability and recognizability.
-    \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers.
-    \item A user study evaluation of the technique with an application to EHRs
+    \item A new variant of Demers cartograms that incorporates dynamic topological features to improve readability and recognizability,
+    \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers,
+    \item A user study evaluation of the technique with an application to EHRs.
 \end{itemize}
 
-\new{One of the major challenges involves is how to develop a layout algorithm that handles different shapes.
-In other words, the layout algorithm is novel because it handles different types of nodes: rectangular representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to revolve stalemate situations introduced by dynamic topological features while minimizing the geographical error.}
+One of the major challenges involved is how to develop a layout algorithm that handles different shapes.
+In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing the geographical error.

### 2022-04-02 paper/section/main.tex
draft: review

@@
     \end{figure}
-
-    \begin{figure}[tb!]
-        \centering
-        \includegraphics[width=\columnwidth]{figure/flowchart2.png}
-        \caption{An awesome flowchart illustration of our layout algorithm. See also \algoref{alg:UpdateNodePosition} for more detail.}
-        \label{fig:flowchart2}
-    \end{figure}
 }
@@
 Nodes that cross a river are translated back to their previous position.
+If a node oscillates across a river, we process this as a stalemate.
 This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
@@
     \textbf{Local variables:} \\
-    $ \node \gets $ a node is an object with the following property: \\
+    $ \node \gets $ a node is an object with the following properties: \\
     \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets True $ if $ \node $ crosses a river \\
+    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
     \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
@@
         \While{$ \nodeSize < \nodeSizeMax $}
-
+            \State $ \nodeListCross \gets True $
             \While{$ \nodeListCross = True $}
@@
                         \State $ \node(x,y) \gets \nodeFNOR(x,y) $
-                        \State $ \nodeCross \gets $ \Call{TestRiverIntersection}{$ \nodeFNORLine $}
 
-                        \If{$ \nodeCross = True $}
+                        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
+                            \State $ \nodeCross \gets \nodeCross ++ $
+                            \State $ \nodeListCross \gets True $
 
-                            \State $ \nodeListCross \gets True $
-                            \State $ \nodeStalemate \gets \nodeStalemate + 1 $
-
-                            \If{$ \nodeStalemate < \stalemateMax $}
-                                \State \hskip-2em \Comment{Transalte back to previous position}
+                            \If{$ \nodeCross < \stalemateMax $}
+                                \State \hskip-2em \Comment{Translate back to previous position}
                                 \State $ \node(x,y) \gets \nodePrevious(x,y) $
@@
                                 \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
-                                \State $ \nodeStalemate \gets 0 $ \Comment{Reset counter}
+                                \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
 
@@
 The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from its original 10,170 to 65.
-We do not require a pixel-level river resolution as Demers cartograms do not represent geography at this level od detail.
+We do not require a pixel-level river resolution as Demers cartograms do not need to represent geography at this level od detail.
 Instead a resolution that matches the node size is used.
@@
     \begin{algorithmic}[1]
-        \Procedure{TestRiverIntersection}{$ \nodeLineNV $}
+        \Procedure{TestIntersection}{$ \nodeLineNV $}
 
@@
 As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations, we introduce a heuristic solution: constructing a corridor to break up the congestion.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to break up the congestion.
 A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \figref{fig:corridor}C and D).
@@
     \includegraphics[width=\columnwidth]{figure/stalemate.png}
-    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (in iteration 1) intersects a river for $ \stalemateMax $ times.
+    \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (in iteration 1) intersects a river $ \stalemateMax $ times.
 The node is translated back to its previous position (iteration 2).
-A stalemate often occurs when the area is congested and the node is unable to translate to a position without intersecting a river.}
+A stalemate often occurs when the area is congested and the node is unable to translate to a new position without intersecting a river.}
     \label{fig:stalemate}

### 2022-04-02 paper/section/related.tex
draft: review

@@
 
-\new{Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates static topological features into Demers cartograms.
-Making topological features dynamic further improves the geographical accuracy and legibility.}
+Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms.
+We introduce multiple features (rivers) into the layout (as opposed to a single).
+Making topological features dynamic further improves the legibility and geographical accuracy.
+
+The way we evaluate the cartograms is also different.
+\citeauthor{tong2018cartograms} count river crossings to evaluate error.
+Here our focus is on readability, thus we include a preliminary user study.
 
@@
 
-One of the disadvantages of Dorling and Demers cartograms is the legibility.
+One of the disadvantages of Dorling and Demers cartograms is legibility.
 The layout algorithms may displace regions far from their original position and making the maps more difficult to interpret.

### 2022-04-02 src/helper/river_vector.js
feat: use river sections to derive corridor

@@ -1,0 +1,18 @@
+const bounding_box = (topLeft, bottomRight, point) => {
+  return (
+    topLeft[0] <= point[0] &&
+    point[0] <= bottomRight[0] &&
+    topLeft[1] <= point[1] &&
+    point[1] <= bottomRight[1]
+  );
+};
+
+const cal_slope = (start, end) => {
+  return (start[1] - end[1]) / (start[0] - end[0]);
+};
+
+const is_upper_side = (node, color) => {
+  return node.attr("original_fill") === color;
+};
+
+export { bounding_box, cal_slope, is_upper_side };

### 2022-04-23 paper/GHRVis.tex
draft: review

--- 2022-03-04 paper/GHRVis.tex
@@
 \begin{abstract}
-
+Cartograms are hybrid representations of geographical and abstract data based on a value-by-area mapping technique.
+Based on the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
+One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.
+To address that, we propose a novel layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
+We evaluate the proposed layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
+The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and geographical accuracy of the cartogram.
 \end{abstract}

### 2022-04-23 paper/section/data.tex
draft: review

@@
 The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, because there is still no public EHR data published based on the latest CCG re-organization that took place in 2021.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organization that took place in 2021.
 
@@
 Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file.
-TopoJSON eliminates redundant coordinates in the data, this improves the rendering speed of our implementation.
+TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
 

### 2022-04-23 paper/section/main.tex
draft: review

@@
 \textbf{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a small unit size square node with initial size, $ \nodeSize $.
+For each CCG we compute the centroid and represent it using a small unit size square node, $ \node $, with initial size, $ \nodeSize $.
 We start with small unit size squares because they do not overlap initially.
@@
 The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from its original 10,170 to 65.
-We do not require a pixel-level river resolution as Demers cartograms do not need to represent geography at this level od detail.
+We do not require a pixel-level river resolution as Demers cartograms do not need to represent geography at this level of detail.
 Instead a resolution that matches the node size is used.
@@
 
-As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path might repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
-If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to break up the congestion.
-A corridor, $ c $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \figref{fig:corridor}C and D).
-All enclosed nodes are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \figref{fig:corridor}D).
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate the congestion.
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \figref{fig:corridor}C and D).
+All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \figref{fig:corridor}E).
 

### 2022-04-23 paper/section/related.tex
draft: review

@@
 Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms.
-We introduce multiple features (rivers) into the layout (as opposed to a single).
-Making topological features dynamic further improves the legibility and geographical accuracy.
-
+Our work enhances that of \citeauthor{tong2018cartograms} in three ways.
+We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic further improves the legibility and geographical accuracy.
+We improve the algorithm for resolving stalemates.
 The way we evaluate the cartograms is also different.

### 2022-04-23 src/helper/river_vector.js
feat: switch to Point object

@@ -1,9 +1,9 @@
 const bounding_box = (topLeft, bottomRight, point) => {
   return (
-    topLeft[0] <= point[0] &&
-    point[0] <= bottomRight[0] &&
-    topLeft[1] <= point[1] &&
-    point[1] <= bottomRight[1]
+    topLeft.x <= point.x &&
+    point.x <= bottomRight.x &&
+    topLeft.y <= point.y &&
+    point.y <= bottomRight.y
   );
 };
 

### 2022-04-23 src/model/LinkedList.js

@@ -1,0 +1,70 @@
+export default class LinkedList {
+  constructor() {
+    this.nodes = [];
+  }
+
+  get size() {
+    return this.nodes.length;
+  }
+
+  get first() {
+    return this.size ? this.nodes[0] : null;
+  }
+
+  get last() {
+    return this.size ? this.nodes[this.size - 1] : null;
+  }
+
+  get secondLast() {
+    return this.size ? this.nodes[this.size - 2] : null;
+  }
+
+  insertAt(index, value) {
+    const previousNode = this.nodes[index - 1] || null;
+    const nextNode = this.nodes[index] || null;
+    const node = { value, next: nextNode };
+
+    if (previousNode) previousNode.next = node;
+    this.nodes.splice(index, 0, node);
+
+    if (this.size > 4) {
+      this.removeAt(0);
+    }
+  }
+
+  insertFirst(value) {
+    this.insertAt(0, value);
+  }
+
+  insertLast(value) {
+    this.insertAt(this.size, value);
+  }
+
+  getAt(index) {
+    return this.nodes[index];
+  }
+
+  removeAt(index) {
+    const previousNode = this.nodes[index - 1];
+    const nextNode = this.nodes[index + 1] || null;
+
+    if (previousNode) previousNode.next = nextNode;
+
+    return this.nodes.splice(index, 1);
+  }
+
+  clear() {
+    this.nodes = [];
+  }
+
+  reverse() {
+    this.nodes = this.nodes.reduce(
+      (acc, { value }) => [{ value, next: acc[0] || null }, ...acc],
+      []
+    );
+  }
+
+  *[Symbol.iterator]() {
+    yield* this.nodes;
+  }
+}

### 2022-04-23 src/model/Point.js

@@ -1,0 +1,27 @@
+export default class Point {
+  constructor(x, y, size, shiftXY = false) {
+    this.size = size && size > 0 ? size : 0;
+    this.halfSize = this.size > 0 ? this.size / 2 : 0;
+
+    // shiftXY is used to shift the point to the center of the rectangle
+    if (shiftXY) {
+      this.x = x + this.halfSize;
+      this.y = y + this.halfSize;
+      this.xRect = x;
+      this.yRect = y;
+    } else {
+      this.x = x;
+      this.y = y;
+      this.xRect = x - this.halfSize;
+      this.yRect = y - this.halfSize;
+    }
+  }
+
+  get coord() {
+    return [this.x, this.y];
+  }
+
+  get coordRect() {
+    return [this.xRect, this.yRect];
+  }
+}

### 2022-04-23 src/router/index.js

@@ -1,6 +1,5 @@
 import Vue from "vue";
 import VueRouter from "vue-router";
-import Home from "../views/Home.vue";
 
 Vue.use(VueRouter);
 
@@ -10,11 +9,6 @@ const routes = [
     name: "Home",
     component: () => import(/* webpackChunkName: "US" */ "../views/CCG.vue"),
     // component: Home,
-  },
-  {
-    path: "/ccg",
-    name: "CCG",
-    component: () => import(/* webpackChunkName: "US" */ "../views/CCG.vue"),
   },
 ];
 

### 2022-04-25 src/helper/flattener.js
fix: wrong corridor direction

@@ -5,7 +5,7 @@ const flatten_path = (pathData, translate) => {
     result += pair.type;
     result += pair.values[0] + translate[0];
     result += ",";
-    result += pair.values[1] += translate[1];
+    result += pair.values[1] + translate[1];
   });
   return result;
 };

### 2022-04-30 paper/section/main.tex
draft: review

@@
         \centering
-        \includegraphics[width=\columnwidth]{figure/flowchart.png}
-        \caption{A flowchart illustration of our layout algorithm. See also \algoref{alg:UpdateNodePosition} for more detail.}
+        \includegraphics[width=0.8\columnwidth]{figure/flowchart.png}
+        \caption{A flowchart illustration of our layout algorithm.See \figref{fig:flowchart-stalemate} for the flow of processing stalemate. See also \algoref{alg:UpdateNodePosition} for more detail.}
         \label{fig:flowchart}
+    \end{figure}
+
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=0.9\columnwidth]{figure/flowchart stalemate.png}
+        \caption{A flowchart illustration of stalemate processing.}
+        \label{fig:flowchart-stalemate}
     \end{figure}
@@
 \algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of the process.
-\textbf{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a small unit size square node, $ \node $, with initial size, $ \nodeSize $.
-We start with small unit size squares because they do not overlap initially.
-\textbf{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
+
+\bobgraph{Initialization:} We first load and render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $.
+\new{We then load the river shapefiles and render the rivers.
+Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next closest vertex.
+This rendering approach enables us to adjust the river resolution as shown in \figref{fig:river resolution}.
+We further apply a smoothing logic by removing vertices that are too close to each other.}
+
+\bobgraph{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
 We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
@@
 We use rivers as topological boundaries and prevent nodes from crossing them.
-The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of edges is reduced from its original 10,170 to 65.
-We do not require a pixel-level river resolution as Demers cartograms do not need to represent geography at this level of detail.
+The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of vertices is reduced from 10,170 to 30.
+A pixel-level river resolution is not required as Demers cartograms do not need to represent geography at this level of detail.
 Instead a resolution that matches the node size is used.
@@
     \caption{The resolution of rivers can be dynamically adjusted by the user.
-The top figure shows River Thames at its original resolution with 10,170 edges.
-The bottom figure shows the river at a  reduced resolution of 65 edges.
-The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+A shows River Thames at its original resolution with 10,170 edges.
+B shows the river at a reduced resolution of 49 edges.
+\new{We further smooth the river by removing vertices in dense areas, as shown in C.} The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
     \label{fig:river resolution}

### 2022-04-30 src/main.js
chore: remove unused dependencies

@@ -1,7 +1,6 @@
 import Vue from "vue";
 import App from "./App.vue";
 import router from "./router";
-import store from "./store";
 import "path-data-polyfill";
 
 import { BootstrapVue, IconsPlugin } from "bootstrap-vue";
@@ -15,6 +14,5 @@ Vue.config.productionTip = false;
 
 new Vue({
   router,
-  store,
   render: (h) => h(App),
 }).$mount("#app");

### 2022-05-15 src/model/LinkedList.js
feat: move river based on average node crossing distance

@@ -40,6 +40,10 @@ export default class LinkedList {
     this.insertAt(this.size, value);
   }
 
+  removeLast(value) {
+    this.removeAt(this.size);
+  }
+
   getAt(index) {
     return this.nodes[index];
   }

### 2022-05-18 paper/section/main.tex
draft: add river translation in algo

@@
 For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $.
-\new{We then load the river shapefiles and render the rivers.
+We then load the river shapefiles and render the rivers.
 Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next closest vertex.
 This rendering approach enables us to adjust the river resolution as shown in \figref{fig:river resolution}.
-We further apply a smoothing logic by removing vertices that are too close to each other.}
+We further apply a smoothing logic by removing vertices that are too close to each other.
 
@@
 
+% \begin{noindent}
+
+    \begin{algorithm}[tb!]
+        \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers.}\label{alg:UpdateLayout}
+        \textbf{Global variables:} \\
+        $ \nodeList \gets $ a list of $ \node $ representing CCGs with the following properties: \\
+        \-\hspace{1em}  \textcolor{blue}{$\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river} \\
+        $ \nodeSize \gets $ the current size of all nodes \\
+        $ \nodeSizeMax \gets $ the maximum size of a node \\
+        \textcolor{blue}{ $ \riverList \gets $ a list of $ \river $ representing river features} \\
+        $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+        \textbf{Local variables:} \\
+        $ \node \gets $ a node is an object with the following properties: \\
+        \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+        \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+        \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+        \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{UpdateLayout}{}
+            \While{$ \nodeSize < \nodeSizeMax $}
+                \State \textcolor{blue}{$ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}}
+                \While{$ \nodeListCross > 0 $}
+
+                    \State $ \nodeListCross \gets 0 $
+                    \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
+                    \color{blue}
+                    \ForEach {$ \river \in \riverList $}
+                        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
+                        \ForEach {$ \node \in \nodeList $}
+
+                            \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+                                \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
+                                \State $ \riverVector \gets \riverVector + \nodeVectorCT$
+                                \State $ \nodeListCross \gets \nodeListCross ++ $
+                                \EndIf
+
+                            \EndIf
+
+                        \EndFor
+
+                        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeListCross} $
+                        \State $ \nodeListCross \gets 0 $ \Comment{Clear the counter}
+                    \EndFor
+
+                    \color{black}
+                    \ForEach {$ \node \in \nodeList $}
+
+                        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+
+                            \State $ \node(x,y) \gets \nodeFNOR(x,y) $
+
+                            \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
+                                \State $ \nodeCross \gets \nodeCross ++ $
+                                \State $ \nodeListCross \gets \nodeListCross++ $
+
+                                \If{$ \nodeCross < \stalemateMax $}
+                                    \State \hskip-2em \Comment{Translate back to previous position}
+                                    \State $ \node(x,y) \gets \nodePrevious(x,y) $
+
+                                \Else
+
+                                    \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
+                                    \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
+
+                                \EndIf
+
+                            \EndIf
+
+                        \EndIf
+
+                    \EndFor
+
+                \EndWhile
+
+                \State $ \nodeSize \gets \nodeSize ++$
+
+            \EndWhile
+
+            \EndProcedure
+        \end{algorithmic}
+    \end{algorithm}
+
+%\end{noindent}
+
 \subsection{River Intersection Testing}
@@
 B shows the river at a reduced resolution of 49 edges.
-\new{We further smooth the river by removing vertices in dense areas, as shown in C.} The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+We further smooth the river by removing vertices in dense areas, as shown in C.
+The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
     \label{fig:river resolution}
@@
     $ \nodeLineNV \gets $ the node's translation path \\
+    $ \river \gets $ a river feature \\
 
@@
 
-    \textbf{Global variables:} \\
-    $ \riverEdgeList \gets $ a list of river edges \\
-
     \textbf{Local variables:} \\
-    $ \riverEdgeLine \gets $ a river edge in $ \riverEdgeList $ \\ 
-    $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \riverEdgeLine $ \\
-    % $ \boundingBoxInt \gets $ boolean indicating $ \nodeBoundingBox $ intersects $ \riverEdgeBoundingBox $ \\
-    % $ \nodeEdge \gets $ the edge for $ \node $\\
-    % $ \edgeBoxInt \gets $ boolean indicating $ \nodeEdge $ intersects $ \riverEdgeListEach $ \\
+    $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \Edge $ \\
+    $ \Edge \gets $ an edge of $ \river $ \\
 
     \begin{algorithmic}[1]
-        \Procedure{TestIntersection}{$ \nodeLineNV $}
+        \Procedure{TestIntersection}{$ \nodeLineNV $, $ \river $}
 
-        \ForEach{$ \riverEdgeLine \in \riverEdgeList $}
+        \ForEach{$ \Edge \in \river $}
             \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
-            \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \riverEdgeLine $)
+            \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \Edge $)
 
             \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
-                \State \Return{$ \nodeLineNV ~intersect~ \riverEdgeLine $}
+                \State \Return{$ \nodeLineNV ~intersect~ \Edge $}
                 \EndIf

### 2022-05-25 paper/section/main.tex
draft: update algorithm

@@
 
-% \begin{noindent}
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to update node positions by removing overlap and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
-    \textbf{Global variables:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs with the following properties: \\
-    \-\hspace{1em} $ \nodeListCross \gets True $ if any $ \node $ in $ \nodeList $ crosses a river \\
-    $ \nodeSize \gets $ the current size of all nodes \\
-    $ \nodeSizeMax \gets $ the maximum size of a node \\
-    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
-
-    \textbf{Local variables:} \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{UpdateNodePosition}{}
-        \While{$ \nodeSize < \nodeSizeMax $}
-            \State $ \nodeListCross \gets True $
-            \While{$ \nodeListCross = True $}
-
-                \State $ \nodeListCross \gets False $
-                \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
-
-                \ForEach {$ \node \in \nodeList $}
-
-                    \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
-
-                        \State $ \node(x,y) \gets \nodeFNOR(x,y) $
-
-                        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
-                            \State $ \nodeCross \gets \nodeCross ++ $
-                            \State $ \nodeListCross \gets True $
-
-                            \If{$ \nodeCross < \stalemateMax $}
-                                \State \hskip-2em \Comment{Translate back to previous position}
-                                \State $ \node(x,y) \gets \nodePrevious(x,y) $
-
-                            \Else
-
-                                \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
-                                \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
-
-                            \EndIf
-
-                        \EndIf
-
-                    \EndIf
-
-                \EndFor
-
-            \EndWhile
-
-            \State $ \nodeSize \gets \nodeSize ++$
-
-        \EndWhile
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
-%\end{noindent}
-
-% \begin{noindent}
-
 \begin{algorithm}[tb!]
@@
                 \color{blue}
+                \State \Call{TranslateRiver}{$ \nodeList $}
+                \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}
+
+                \color{black}
+
+            \EndWhile
+
+            \State $ \nodeSize \gets \nodeSize ++$
+
+        \EndWhile
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[tb!]
+    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
+    \color{blue}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+
+    \textbf{Local variables:} \\
+    $ count \gets $ the number of river crossing nodes \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateRiver}{$ \nodeList $}
         \ForEach {$ \river \in \riverList $}
+        \State $ count \gets 0 $
         \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
@@
         \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-                                \State $ \nodeListCross \gets \nodeListCross ++ $
+        \State $ count \gets count ++ $
         \EndIf
@@
 
-                        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeListCross} $
-                        \State $ \nodeListCross \gets 0 $ \Comment{Clear the counter}
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
         \EndFor
 
-                    \color{black}
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[tb!]
+    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
+
+    \color{blue}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+
+    \textbf{Output:} \\
+    $ count \gets $ the number of river crossing nodes in the input \\
+
+    \textbf{Global variables:} \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateNode}{$ \nodeList $}
+        \State $ count \gets 0 $
         \ForEach {$ \node \in \nodeList $}
@@
                             \State $ \nodeCross \gets \nodeCross ++ $
-                                \State $ \nodeListCross \gets \nodeListCross++ $
+                            \State $ count \gets count ++ $
 
@@
                 \EndFor
-
-                \EndWhile
-
-                \State $ \nodeSize \gets \nodeSize ++$
-
-            \EndWhile
-
+        \State \Return{$ count $}
         \EndProcedure
@@
 
-%\end{noindent}
+\color{black}
 

### 2022-05-30 paper/section/main.tex
draft: update flowchart

@@
       \centering
-      \includegraphics[width=\columnwidth]{example-image-a}
+      \includegraphics[width=\columnwidth,keepaspectratio]{example-image-a}
       \caption{An screenshot of \software. Figure TBA.}
@@
  {
-  \begin{figure}[tb!]
+  \begin{figure*}[tb!]
       \centering
-      \includegraphics[width=0.8\columnwidth]{figure/flowchart.png}
-      \caption{A flowchart illustration of our layout algorithm.See \figref{fig:flowchart-stalemate} for the flow of processing stalemate. See also \algoref{alg:UpdateNodePosition} for more detail.}
+      \includegraphics[width=\textwidth,height=0.95\textheight,keepaspectratio]{figure/flowchart.png}
+      \caption{A flowchart illustration of our layout algorithm.See \figref{fig:flowchart-stalemate} for the flow of processing stalemate. See also \algoref{alg:UpdateLayout} for more detail.}
       \label{fig:flowchart}
-  \end{figure}
+  \end{figure*}
 
@@
       \centering
-      \includegraphics[width=0.9\columnwidth]{figure/flowchart stalemate.png}
+      \includegraphics[width=0.9\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
       \caption{A flowchart illustration of stalemate processing.}
@@
 
-\algoref{alg:UpdateNodePosition} and \figref{fig:flowchart} provide an overview of the process.
+\algoref{alg:UpdateLayout} and \figref{fig:flowchart} provide an overview of the process.
 
@@
             \centering
-            \includegraphics[width=\columnwidth]{figure/river_resolution.png}
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/river_resolution.png}
             \caption{The resolution of rivers can be dynamically adjusted by the user.
@@
             \centering
-            \includegraphics[width=\columnwidth]{figure/stalemate.png}
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
             \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (in iteration 1) intersects a river $ \stalemateMax $ times.
@@
             \centering
-            \includegraphics[width=\columnwidth]{figure/corridor.png}
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/corridor.png}
             \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in A.

### 2022-06-04 paper/section/main.tex
drat: update flowchart

@@
       \includegraphics[width=\textwidth,height=0.95\textheight,keepaspectratio]{figure/flowchart.png}
-      \caption{A flowchart illustration of our layout algorithm.See \figref{fig:flowchart-stalemate} for the flow of processing stalemate. See also \algoref{alg:UpdateLayout} for more detail.}
+      \caption{A flowchart illustration of our layout algorithm.
+See \figref{fig:flowchart-stalemate} for the flow of processing stalemate.
+See also \algoref{alg:UpdateLayout} for more detail.
+For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
       \label{fig:flowchart}

### 2022-06-04 src/model/Point.js
feat: river cross nodes

@@ -24,4 +24,8 @@ export default class Point {
   get coordRect() {
     return [this.xRect, this.yRect];
   }
+
+  set crossedRiver(river) {
+    this.crossedRiver = river;
+  }
 }

### 2022-06-04 vue.config.js

@@ -4,4 +4,7 @@ module.exports = {
       config.output.filename("[name].[hash].js").end();
     }
   },
+  configureWebpack: {
+    devtool: "source-map",
+  },
 };

### 2022-06-05 src/model/LinkedList.js
feat: dynamic thames

@@ -40,7 +40,7 @@ export default class LinkedList {
     this.insertAt(this.size, value);
   }
 
-  removeLast(value) {
+  removeLast() {
     this.removeAt(this.size);
   }
 

### 2022-07-16 paper/section/appendix.tex
draft: appendix on QGIS pre-processing

--- 2022-03-10 paper/section/appendix.tex
@@
 \appendix
+
+\section{Pre-processing Shapefiles}
+
+Shapefiles from different sources are likely to be incompatible.
+In our case, the NHS CCG shapefile is incompatible with the river shapefiles.
+The major reason for the incompatibility is the coordinate reference system (CRS).
+The CRS of the CCG shapefile is EPSG:27700 (OSGB36 - British National Grid).
+The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System).
+Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility issue and reduce shapefile size to improve performance.
+
+\subsection{Import Shapefiles into QGIS}
+We first load all three river shapefiles into QGIS \autoref{fig:import_rivers}, followed by the CCG shapefile \autoref{fig:import_ccgs}.
+
+{
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/import_rivers.png}
+        \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
+        \label{fig:import_rivers}
+    \end{figure}
+
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/import_ccgs.png}
+        \caption{QGIS interface, with all NHS CCGs imported.}
+        \label{fig:import_ccgs}
+    \end{figure}
+
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/unify_crs.png}
+        \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
+        \label{fig:unify_crs}
+    \end{figure}
+}
+
+\subsection{Export Shapefiles in GeoJSON and Unify the Coordinate Reference System (CRS)}
+
+We then use QGIS to unify the CRS, and export both layers in the GeoJSON format. See \autoref{fig:export_rivers} and \autoref{fig:export_ccgs}. The unified layer is shown in \autoref{fig:unify_crs}.
+
+{
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/export_rivers.png}
+        \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
+        \label{fig:export_rivers}
+    \end{figure}
+
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/export_ccgs.png}
+        \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
+        \label{fig:export_ccgs}
+    \end{figure}
+
+}
+
+\subsection{Merge Shapefiles and Reduce File Size}
+
+We then merge two layers into one layer, and export it in the TopoJSON format using Mapshaper \cite{blochMapshaper}.
+Mapshaper also supports the simplification of GeoJSON shapefiles.
+See \autoref{fig:export_topojson}.
+
+{
+    \begin{figure}[tbh!]
+        \centering
+        \includegraphics[width=\columnwidth]{figure/qgis/export_topojson.png}
+        \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
+        \label{fig:export_topojson}
+    \end{figure}
+}
+
+\subsection{Pre-processing Result}
+
+\autoref{tbale:pre-processing_result} shows the pre-processing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+
+{
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[!tb]
+	\centering
+	\resizebox{\columnwidth}{!}{
+		\begin{tabulary}{\columnwidth}{|*{4}{l|}}
+			\hhline{~|*{3}{-}}
+			\multicolumn{1}{c|}{\textbf{Shapefile}} &
+			\cellcolor{Mycolor2}\textbf{Original} &
+			\cellcolor{Mycolor2}\textbf{GeoJSON} &
+			\cellcolor{Mycolor2}\textbf{TopoJSON} \\
+			\hline
+			Rivers & 2.0 MB (GeoJSON) & 1.4 MB & -  \\
+			\hline
+			NHS CCGs & 46.6 MB (.shp, Esri vector shapefile) & 140.2 MB & -  \\
+			\hline
+			Merged & - & - & 16.3 MB \\
+			\hline
+		\end{tabulary}
+	}
+	\caption{The file size is reduced by 88.5\% from the original size.}
+	\label{table:pre-processing_result}
+\end{table}
+}
+
+\clearpage
 
@@
 
-    \begin{algorithm}[h!]
+    \begin{algorithm}[tbh!]
         \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
@@
 
-\begin{algorithm}[h!]
+\begin{algorithm}[tbh!]
     \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

### 2022-07-16 paper/section/data.tex
draft: appendix on QGIS pre-processing

@@
     \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography portalx \cite{opengeographyportalxOpen} with EPSG:27700 (OSGB36 - British National Grid) as the Coordinate Reference System (CRS).}
+    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
     \label{fig:ccg}
@@
 The number of CCGs changes over time due to the NHS re-organization.
-The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organization that took place in 2021.
+The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations took place in 2021 and 2022.
 

### 2022-07-16 paper/section/introduction.tex
draft: appendix on QGIS pre-processing

@@
 
-Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \tableref{table:accuracy}).
+Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \autoref{table:accuracy}).
 For this project we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.

### 2022-07-16 paper/section/main.tex
draft: appendix on QGIS pre-processing

@@
       \caption{A flowchart illustration of our layout algorithm.
-See \figref{fig:flowchart-stalemate} for the flow of processing stalemate.
-See also \algoref{alg:UpdateLayout} for more detail.
+See \autoref{fig:flowchart-stalemate} for the flow of processing stalemate.
+See also \autoref{alg:UpdateLayout} for more detail.
 For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
@@
 
-\algoref{alg:UpdateLayout} and \figref{fig:flowchart} provide an overview of the process.
+\autoref{alg:UpdateLayout} and \autoref{fig:flowchart} provide an overview of the process.
 
@@
 Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next closest vertex.
-This rendering approach enables us to adjust the river resolution as shown in \figref{fig:river resolution}.
+This rendering approach enables us to adjust the river resolution as shown in \autoref{fig:river resolution}.
 We further apply a smoothing logic by removing vertices that are too close to each other.
@@
 An increase in $ \nodeSize $ can cause the nodes to overlap.
-During overlap removal, we compute node trajectories (See \algoref{alg:check river intersection}) and translate nodes to their new position.
+During overlap removal, we compute node trajectories (See \autoref{alg:check river intersection}) and translate nodes to their new position.
 Nodes that cross a river are translated back to their previous position.
@@
 We use rivers as topological boundaries and prevent nodes from crossing them.
-The resolution of rivers can be adjusted by the user, as shown in \figref{fig:river resolution}, the number of vertices is reduced from 10,170 to 30.
+The resolution of rivers can be adjusted by the user, as shown in \autoref{fig:river resolution}, the number of vertices is reduced from 10,170 to 30.
 A pixel-level river resolution is not required as Demers cartograms do not need to represent geography at this level of detail.
@@
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
-See \algoref{alg:check river intersection}.
+See \autoref{alg:check river intersection}.
 A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
@@
 
-As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \figref{fig:stalemate}.
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \autoref{fig:stalemate}.
 If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate the congestion.
-A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \figref{fig:corridor}C and D).
-All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \figref{fig:corridor}E).
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \autoref{fig:corridor}C and D).
+All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \autoref{fig:corridor}E).
 
@@
     \caption{Procedure to derive a corridor to translate enclosed nodes.
-We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \figref{fig:stalemate}).}\label{alg:derive corridor}
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \autoref{fig:stalemate}).}\label{alg:derive corridor}
 

### 2022-07-16 paper/section/related.tex
draft: appendix on QGIS pre-processing

@@
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
-See \tableref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+See \autoref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
 

### 2022-07-16 paper/section/data.tex
draft: autoref algorithm

@@
 
-After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, translations and scales, and merge them into one shapefile.
-Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file.
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files.
+Finally, mapshaper \cite{blochMapshaper} is used to merge adn convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file.
 TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
+See \autoref{table:pre-processing_result} for the pre-processing result.
+
+We describe the above steps in detail in \autoref{app:pre-processing}.
 

### 2022-08-11 paper/section/evaluation.tex
draft: evaluation

\color{blue}

\section{Evaluation}

\subsection{Hypotheses Formulation}

We formulate the following hypotheses to evaluate the effectiveness of our approach:

\textbf{H1:} The introduction of dynamic topological features can effectively improve the recognizability of CCGs.

\textbf{H2:} \dots

\subsection{Participants}

We recruited X participants, [add demographics here].

\subsection{Datasets}

We used the following datasets for our evaluation:

\begin{itemize}
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol related liver disease
\end{itemize}

Datasets are projected on a map that consists of 135 CGGs in England.

\subsection{Tasks}

We conducted tasks to evaluate the effectiveness of our approach.

\textbf{T1:} The participants were asked to familiarize with the design, including the user options and the interaction methods.
\textbf{T2:} The participants

\subsection{Questionnaire}

We asked participants to complete a questionnaire to collect their feedback. The questionnaire consisted of the following questions:

\begin{itemize}
    \item \dots
    \item \dots
\end{itemize}

\color{black}

### 2022-08-11 paper/section/main.tex
draft: evaluation

@@
     $ \nodeList \gets $ a list of $ \node $ representing CCGs with the following properties: \\
-    \-\hspace{1em}  \textcolor{blue}{$\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river} \\
+    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
     $ \nodeSize \gets $ the current size of all nodes \\
     $ \nodeSizeMax \gets $ the maximum size of a node \\
-    \textcolor{blue}{ $ \riverList \gets $ a list of $ \river $ representing river features} \\
+    $ \riverList \gets $ a list of $ \river $ representing river features \\
     $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
@@
         \While{$ \nodeSize < \nodeSizeMax $}
-            \State \textcolor{blue}{$ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}}
+            \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
             \While{$ \nodeListCross > 0 $}
@@
                 \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
-                \color{blue}
                 \State \Call{TranslateRiver}{$ \nodeList $}
                 \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}
-
-                \color{black}
 
@@
     \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
-    \color{blue}
     \textbf{Input:} \\
@@
 
-    \color{blue}
     \textbf{Input:} \\

### 2022-09-15 paper/section/data.tex
fix: typos

@@
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations took place in 2021 and 2022.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
 
-\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England.
-These rivers were chosen as they are well-known rivers and pass through regions with dense population, and provide informative geographical and topological cues.
+\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England.
+These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
 
@@
 After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files.
-Finally, mapshaper \cite{blochMapshaper} is used to merge adn convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file.
+Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file.
 TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.

### 2022-09-15 paper/section/evaluation.tex
fix: typos

@@
     \item Under 75 mortality from cardiovascular disease
-    \item Emergency admissions for alcohol related liver disease
+    \item Emergency admissions for alcohol-related liver disease
 \end{itemize}
@@
 
-\textbf{T1:} The participants were asked to familiarize with the design, including the user options and the interaction methods.
+\textbf{T1:} The participants were asked to familiarize themselves with the design, including the user options and the interaction methods.
 \textbf{T2:} The participants

### 2022-09-15 paper/section/introduction.tex
fix: typos

@@
 Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information \cite{dent2009Cartography}.
-Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision making \cite{gao2020Visualising, sack2021Visualizing}.
+Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision-making \cite{gao2020Visualising, sack2021Visualizing}.
 
 Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \autoref{table:accuracy}).
-For this project we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
+For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
@@
 To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm to a real world case-study using an Electronic Health Records (EHR) dataset to evaluate of our algorithm.
+We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate our algorithm.
 We also present a user study demonstrating its effectiveness.
@@
 In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing the geographical error.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing geographical error.

### 2022-09-15 paper/section/related.tex
fix: typos

@@
 	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
-Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
+Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
 }
@@
 \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
@@
 Our work enhances that of \citeauthor{tong2018cartograms} in three ways.
-We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic further improves the legibility and geographical accuracy.
+We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic and further improving the legibility and geographical accuracy.
 We improve the algorithm for resolving stalemates.
@@
 One of the disadvantages of Dorling and Demers cartograms is legibility.
-The layout algorithms may displace regions far from their original position and making the maps more difficult to interpret.
+The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret.
 Our work extends Demers cartograms.

### 2022-10-11 paper/section/evaluation.tex
feat: add evaluation draft

@@
 
-We recruited X participants, [add demographics here].
+We recruited X participants, [add demographic analysis here].
 
@@
 
-Datasets are projected on a map that consists of 135 CGGs in England.
+135 CGGs are projected on the screen with a choropleth, we then generated another view with our visual design. The color in both visualizations are mapped to population. See \autoref{fig:task}.
 
-\subsection{Tasks}
+{
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task.png}
+        \caption{A typical task for participants.
+The left shows the choropleth map, and the right shows the cartogram.
+Both visualizations show the three longest rivers in England, and the color is mapped to CCG population.
+The target CGG will be blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.}
+        \label{fig:task}
+    \end{figure}
+}
 
-We conducted tasks to evaluate the effectiveness of our approach.
+\subsection{User Experiment}
 
-\textbf{T1:} The participants were asked to familiarize themselves with the design, including the user options and the interaction methods.
-\textbf{T2:} The participants
+We conducted a user experiment to evaluate the effectiveness of our approach.
+The experiment was designed to be conducted remotely, and participants were asked to complete the tasks on their own computers.
+The experiment includes four parts:
 
-\subsection{Questionnaire}
+\textbf{P1:} The participants were asked read instructions and trainings provided in both text and videos.
+The instructions were designed to help participants understand the concepts used in the tasks.
+Videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
 
-We asked participants to complete a questionnaire to collect their feedback. The questionnaire consisted of the following questions:
+\textbf{P2:} The participants were given three practice tasks to familiarize themselves with the experiment. A demonstration of the task is also included in the instructional video.
 
-\begin{itemize}
-    \item \dots
-    \item \dots
-\end{itemize}
+\textbf{P3:} The participants were asked to complete 16 location tasks.
+The response and reaction time are recorded.
+These 16 CCGs were carefully selected to avoid extreme cases (thus biases the result), in terms of size, color, and location.
+
+\textbf{P4:} The participants were asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
+
+The user experiment is available at \url{https://osf.io/q39w7}.
 

### 2022-10-11 paper/section/main.tex
fix: typos
feat: add evaluation draft

@@
       \centering
-      \includegraphics[width=\columnwidth,keepaspectratio]{example-image-a}
-      \caption{An screenshot of \software. Figure TBA.}
+      \includegraphics[width=\columnwidth,keepaspectratio]{figure/UI.png}
+      \caption{An screenshot of \software.
+User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
+Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained.}
       \label{fig:overview}
@@
       \centering
-      \includegraphics[width=\textwidth,height=0.95\textheight,keepaspectratio]{figure/flowchart.png}
+      \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
       \caption{A flowchart illustration of our layout algorithm.
@@
       \centering
-      \includegraphics[width=0.9\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
+      \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
       \caption{A flowchart illustration of stalemate processing.}
@@
 A pixel-level river resolution is not required as Demers cartograms do not need to represent geography at this level of detail.
-Instead a resolution that matches the node size is used.
+Instead, a resolution that matches the node size is used.
 Enabling river simplification reduces the number of edge intersection tests that need to be performed.
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
-            \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (in iteration 1) intersects a river $ \stalemateMax $ times.
+            \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
 The node is translated back to its previous position (iteration 2).

### 2022-10-26 paper/GHRVis.tex
draft: appendix on QGIS pre-processing
draft: autoref algorithm
draft: evaluation
fix: typos
feat: revised paper

@@
 \documentclass{egpubl}
-\usepackage{eurovis2021}
+\usepackage{eurovis2023}
 
@@
 
+\newcommand{\algorithmautorefname}{Algorithm}
+
 \usepackage{dblfloatfix}
@@
 
-\author[Q. Wang \& R. S. Laramee]
-{\parbox{\textwidth}{\centering Q. Wang
+\author[Q. Wang, K. Xu \& R. S. Laramee]
+{\parbox{\textwidth}{\centering Q. Wang, K. Xu
   and R. S. Laramee
@@
 
-\title{Cartograms with Dynamic Features}
+\title{Cartograms with Rivers}
 
@@
 One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.
-To address that, we propose a novel layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
+To address this, we propose a novel cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
+Rivers affect both the layout and appearance of cartograms.
 We evaluate the proposed layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
@@
 
-\section{Evaluation}
+\input{section/evaluation}
 
@@
 
+\clearpage
 % reset \section command for section title "References"
@@
 
+\cleardoublepage
+
 \input{section/appendix}

### 2022-10-26 paper/section/data.tex
feat: revised paper

@@
 \section{Data Description}
+
+{
+\begin{figure}[htb!]
+    \centering
+    \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
+    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
+    \label{fig:ccg}
+\end{figure}
+}
 
@@
 
-\subsection{Geospatial Data}
+\bobgraph{Geospatial Data}
 
-Geospatial boundaries, or shapefiles, were obtained from the sources described below.
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-    \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
-    \label{fig:ccg}
-\end{figure}
-}
+Geospatial boundaries, or shapefiles, were obtained from the sources described here.
 
@@
 These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
+Other rivers cloud be chose, including additional and smaller rivers remains future work.
 
@@
 TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
-See \autoref{table:pre-processing_result} for the pre-processing result.
+See \autoref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result.
+We describe the above steps in more detail in \autoref{app:pre-processing}.
 
-We describe the above steps in detail in \autoref{app:pre-processing}.
-
-\subsection{EHR Data}
+\bobgraph{EHR Data}
 
@@
 \begin{itemize}
-    \item Under 75 mortality
-    \begin{multicols}{2}
-        \begin{itemize}
-            \item Cardiovascular disease
-            \item Respiratory disease
-            \item Liver disease
-            \item Cancer
-        \end{itemize}
-    \end{multicols}
-    \item Emergency hospital admission
-        \begin{itemize}
-            \item Stroke
-            \item Alcohol-specific admission and readmission
-            \item Coronary heart disease
-            \item Readmissions within 30 days of discharge
-            \item Children with lower respiratory tract infections
-        \end{itemize}
+    \item Under 75 mortality: cardiovascular disease, respiratory disease, liver disease, and cancer
+    \item Emergency hospital admission: stroke, alcohol-specific admission and readmission, coronary heart disease, re-admissions within 30 days of discharge, children with lower respiratory tract infections
 \end{itemize}

### 2022-10-26 paper/section/introduction.tex
feat: revised paper

@@
+{
+    \begin{figure*}[ht!]
+        \centering
+        \includegraphics[width=0.8\textwidth,keepaspectratio]{figure/UI.png}
+        \caption{A screenshot of the software.
+User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
+Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained.}
+        \label{fig:overview}
+    \end{figure*}
+}
+
 \section{Introduction and Motivation}
 
-Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping technique combining statistical and geographical information \cite{dent2009Cartography}.
+Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
 Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision-making \cite{gao2020Visualising, sack2021Visualizing}.
@@
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-We implement a new cartographic layout algorithm that includes these topological features into the layout of the nodes representing geographical regions.
-To minimize geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate our algorithm.
+Standard cartograms are composed of uniform shape nodes.
+As such, this can reduce their legibility.
+We implement a new cartographic layout algorithm that includes rivers into the layout of the nodes representing geographical regions.
+To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
+We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate the result.
 We also present a user study demonstrating its effectiveness.
@@
 
+The results of the user study indicate that rivers can improve the legibility of cartograms.
 One of the major challenges involved is how to develop a layout algorithm that handles different shapes.

### 2022-10-26 paper/section/main.tex
feat: revised paper

@@
-\section{Cartograms with Dynamic Features}
-
+\section{Cartograms with Rivers}
 {
-  \begin{figure}[tb!]
+    \begin{figure}[htb!]
         \centering
-      \includegraphics[width=\columnwidth,keepaspectratio]{figure/UI.png}
-      \caption{An screenshot of \software.
-User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
-Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained.}
-      \label{fig:overview}
+        \includegraphics[width=0.7\columnwidth,keepaspectratio]{figure/river_resolution.png}
+        \caption{The resolution of rivers can be dynamically adjusted by the user.
+A shows River Thames at its original resolution with 10,170 edges.
+B shows the river at a reduced resolution of 49 edges.
+We further smooth the river by removing vertices in dense areas, as shown in C.
+The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+        \label{fig:river resolution}
     \end{figure}
@@
 {
-  \begin{figure*}[tb!]
+  \begin{figure*}[htb!]
       \centering
@@
       \caption{A flowchart illustration of our layout algorithm.
-See \autoref{fig:flowchart-stalemate} for the flow of processing stalemate.
-See also \autoref{alg:UpdateLayout} for more detail.
+See \autoref{fig:flowchart-stalemate} for the logic of processing a stalemate.
+See also \autoref{alg:UpdateLayout} in \autoref{sec:{Cartograms with Rivers}} for more detail.
 For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
@@
 
-  \begin{figure}[tb!]
+  \begin{figure}[htb!]
       \centering
       \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
-      \caption{A flowchart illustration of stalemate processing.}
+      \caption{A flowchart illustration of stalemate processing. See \autoref{subsec:{Detecting Stalemates}}, \autoref{fig:stalemate}, and \autoref{fig:corridor} for more detail.}
       \label{fig:flowchart-stalemate}
@@
 
-\autoref{alg:UpdateLayout} and \autoref{fig:flowchart} provide an overview of the process.
-
-\bobgraph{Initialization:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $.
-We then load the river shapefiles and render the rivers.
-Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next closest vertex.
-This rendering approach enables us to adjust the river resolution as shown in \autoref{fig:river resolution}.
-We further apply a smoothing logic by removing vertices that are too close to each other.
-
-\bobgraph{Node layout:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
-We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-We gradually increase the node size by 1 unit at a time to ensure smooth transitions.
-An increase in $ \nodeSize $ can cause the nodes to overlap.
-During overlap removal, we compute node trajectories (See \autoref{alg:check river intersection}) and translate nodes to their new position.
-Nodes that cross a river are translated back to their previous position.
-If a node oscillates across a river, we process this as a stalemate.
-This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
-The gradual size increase process provides stability to the layout and helps minimize geographical error.
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers.}\label{alg:UpdateLayout}
+\begin{algorithm}[htb!]
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \autoref{sec:{Cartograms with Rivers}}.}\label{alg:UpdateLayout}
     \textbf{Global variables:} \\
@@
 
-\begin{algorithm}[tb!]
+\autoref{alg:UpdateLayout} and \autoref{fig:flowchart} provide an overview of the cartogram layout process that includes rivers.
+
+\bobgraph{Initialization with Rivers:} We first load and render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $.
+We then load the river shapefiles and render the rivers.
+Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex.
+We do not need the original river resolution to incorporate them into cartograms.
+We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests.
+This rendering approach enables us to adjust the river resolution as shown in \autoref{fig:river resolution}.
+We further apply smoothing by removing vertices that are too close to each other.
+
+\bobgraph{Node Layout and Overlap Removal:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
+We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+We gradually increase the node size by 1 unit at a time to ensure smooth transitions.
+An increase in $ \nodeSize $ can cause the nodes to overlap.
+During overlap removal, we compute node trajectories (See \autoref{alg:check river intersection}) and translate nodes to their new position.
+Nodes that cross a river are translated back to their previous position.
+If a node oscillates across a river, we process this as a stalemate.
+This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
+We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
+The gradual size increase process provides stability to the layout and helps minimize geographical error.
+
+\begin{algorithm}[htb!]
     \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
@@
 
-\begin{algorithm}[tb!]
+\begin{algorithm}[htb!]
     \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
@@
 
-We use rivers as topological boundaries and prevent nodes from crossing them.
-The resolution of rivers can be adjusted by the user, as shown in \autoref{fig:river resolution}, the number of vertices is reduced from 10,170 to 30.
+% We use rivers as topological boundaries and prevent nodes from crossing them.
+The resolution of rivers can be adjusted by the user, as shown in \autoref{fig:river resolution}.
+In this example, the number of vertices is reduced from 10,170 to 30.
 A pixel-level river resolution is not required as Demers cartograms do not need to represent geography at this level of detail.
@@
 Enabling river simplification reduces the number of edge intersection tests that need to be performed.
+%
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
@@
 A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
+Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
 
-    {
-        \begin{figure}[tb!]
-            \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{figure/river_resolution.png}
-            \caption{The resolution of rivers can be dynamically adjusted by the user.
-A shows River Thames at its original resolution with 10,170 edges.
-B shows the river at a reduced resolution of 49 edges.
-We further smooth the river by removing vertices in dense areas, as shown in C.
-The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
-            \label{fig:river resolution}
-        \end{figure}
-    }
+\subsection{Translating Rivers}
+
+For all nodes that cross a river, we compute an average vector $ \frac{\riverVector}{count} $ used to translate the corresponding river.
+See \autoref{alg:TranslateRiver}.
+This step intends to create room for the next iteration of node translation without crossing a river.
+
+During river translations, some nodes might be crossed by the trajectory of river translations. We move these nodes by the same vector.
 
 % \begin{noindent}
-\begin{algorithm}[tb!]
+\begin{algorithm}[htb!]
     \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
@@
     {
-        \begin{figure}[tb!]
+        \begin{figure}[htb!]
             \centering
@@
 
-\begin{algorithm}[tb!]
+\begin{algorithm}[htb!]
     \caption{Procedure to derive a corridor to translate enclosed nodes.

### 2022-11-01 paper/section/evaluation.tex
feat: revised paper
draft: evaluation

@@
 
-\subsection{Hypotheses Formulation}
+We conducted a pilot user study to evaluate the effectiveness of our approach.
 
-We formulate the following hypotheses to evaluate the effectiveness of our approach:
+\subsection{User Study Tasks}
 
-\textbf{H1:} The introduction of dynamic topological features can effectively improve the recognizability of CCGs.
+We formulate the following hypotheses to design our user study:
 
-\textbf{H2:} \dots
+\textbf{H1:} The introduction of rivers can improve the accuracy of locating a CCG in a cartogram.
 
-\subsection{Participants}
+\textbf{H2:} The introduction of rivers can reduce the time needed to locate a CCG in a cartogram.
 
-We recruited X participants, [add demographic analysis here].
+\subsection{User Study Variables}
 
-\subsection{Datasets}
+We discuss the variables of our user study in this section.
+
+\subsubsection{Independant Variables}
+
+The primary independent variable is the river-crossing behavior of the nodes, which directly decides the final layout of our cartograms.
+We set the parameter to either allow or disallow the nodes from crossing a river.
+The presence of the rivers on the screen is another independent variable.
+\autoref{table:Independant Variables} shows the combinations we used in our user study.
+
+{
+\renewcommand{\arraystretch}{1.5}
+\begin{table}[!tb]
+	\centering
+		\begin{tabulary}{\columnwidth}{|*{2}{c|}}
+			\hline
+			\cellcolor{Mycolor2}\textbf{River Crossing} &
+			\cellcolor{Mycolor2}\textbf{River Presence} \\
+			\hline
+			Allow & Present \\
+			\hline
+			Allow & Absent \\
+			\hline
+			Disallow & Present \\
+			\hline
+			Disallow & Absent \\
+			\hline
+		\end{tabulary}
+	\caption{The combinations of independent variables.}
+	\label{table:Independant Variables}
+\end{table}
+}
+
+\subsubsection{Dependant Variables}
+
+\bobgraph{Accuracy:} The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
+
+\bobgraph{Response time:} Another dependant variable is the time taken by the participants to complete each task.
+
+\subsubsection{Control Variables}
+\bobgraph{Choice of the color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+
+\bobgraph{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find.
+The target CCG is shown in the form of a non-stop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
+
+\subsection{User Study}
+
+\subsubsection{Participants}
+
+We recruited X participants, ...
+
+\subsubsection{Datasets}
 
@@
 \begin{itemize}
+    \item Population
     \item Under 75 mortality from cardiovascular disease
     \item Emergency admissions for alcohol-related liver disease
+    \item Alcohol-specific admission and readmission
 \end{itemize}
@@
 {
-    \begin{figure}[tb!]
+    \begin{figure}[htb!]
         \centering
@@
 
-\subsection{User Experiment}
+\subsubsection{Procedure}
 
-We conducted a user experiment to evaluate the effectiveness of our approach.
-The experiment was designed to be conducted remotely, and participants were asked to complete the tasks on their own computers.
-The experiment includes four parts:
+The user study was designed to be conducted remotely, and participants were asked to complete the tasks on their own computers. The user study includes four parts:
 
@@
 
-\textbf{P2:} The participants were given three practice tasks to familiarize themselves with the experiment. A demonstration of the task is also included in the instructional video.
+\textbf{P2:} The participants were given three practice tasks to familiarize themselves with the user study. A demonstration of the task is also included in the instructional video.
 
@@
 
+\subsection{User Study Results}
+
 \color{black}

### 2022-11-01 user-study/index.js
feat: add user study analysis

@@ -1,0 +1,208 @@
+const fs = require("fs");
+const { parse } = require("csv-parse");
+const likertscale = () => {
+    return {
+        "Strongly agree": 0,
+        "Agree": 0,
+        "Neither agree nor disagree": 0,
+        "Disagree": 0,
+        "Strongly disagree": 0,
+    }
+}
+
+const taskList = [
+    { ccg: "E38000243", withRiver: true, allowCrossing: false, correct: 0 },
+    { ccg: "E38000136", withRiver: true, allowCrossing: false, correct: 0 },
+    { ccg: "E38000242", withRiver: true, allowCrossing: false, correct: 0 },
+    { ccg: "E38000244", withRiver: true, allowCrossing: false, correct: 0 },
+    { ccg: "E38000243", withRiver: false, allowCrossing: true, correct: 0 },
+    { ccg: "E38000136", withRiver: false, allowCrossing: true, correct: 0 },
+    { ccg: "E38000242", withRiver: false, allowCrossing: true, correct: 0 },
+    { ccg: "E38000244", withRiver: false, allowCrossing: true, correct: 0 },
+    { ccg: "E38000243", withRiver: true, allowCrossing: true, correct: 0 },
+    { ccg: "E38000136", withRiver: true, allowCrossing: true, correct: 0 },
+    { ccg: "E38000242", withRiver: true, allowCrossing: true, correct: 0 },
+    { ccg: "E38000244", withRiver: true, allowCrossing: true, correct: 0 },
+    { ccg: "E38000243", withRiver: false, allowCrossing: false, correct: 0 },
+    { ccg: "E38000136", withRiver: false, allowCrossing: false, correct: 0 },
+    { ccg: "E38000242", withRiver: false, allowCrossing: false, correct: 0 },
+    { ccg: "E38000244", withRiver: false, allowCrossing: false, correct: 0 },
+];
+
+const resultPerCCG = {
+    E38000243: { withRiver: 0, withoutRiver: 0 },
+    E38000136: { withRiver: 0, withoutRiver: 0 },
+    E38000242: { withRiver: 0, withoutRiver: 0 },
+    E38000244: { withRiver: 0, withoutRiver: 0 },
+};
+
+const resultPerQuestion = [
+    { withRiver: 0, withoutRiver: 0, responseTime: 0 },
+    { withRiver: 0, withoutRiver: 0, responseTime: 0 },
[…]

### 2022-11-08 paper/GHRVis.tex
draft

@@
 
+\usepackage[nameinlink]{cleveref}
+
 \author[Q. Wang, K. Xu \& R. S. Laramee]
@@
 Cartograms are hybrid representations of geographical and abstract data based on a value-by-area mapping technique.
-Based on the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
+A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
 One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.

### 2022-11-08 paper/section/appendix.tex
draft: autoref algorithm
draft

@@
 \section{Pre-processing Shapefiles}
-
+\label{app:pre-processing}
 Shapefiles from different sources are likely to be incompatible.
@@
 \subsection{Import Shapefiles into QGIS}
-We first load all three river shapefiles into QGIS \autoref{fig:import_rivers}, followed by the CCG shapefile \autoref{fig:import_ccgs}.
+We first load all three river shapefiles into QGIS \Cref{fig:import_rivers}, followed by the CCG shapefile \Cref{fig:import_ccgs}.
 
@@
 
-We then use QGIS to unify the CRS, and export both layers in the GeoJSON format. See \autoref{fig:export_rivers} and \autoref{fig:export_ccgs}. The unified layer is shown in \autoref{fig:unify_crs}.
+We then use QGIS to unify the CRS, and export both layers in the GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.
 
@@
 Mapshaper also supports the simplification of GeoJSON shapefiles.
-See \autoref{fig:export_topojson}.
+See \Cref{fig:export_topojson}.
 
@@
 
-\autoref{tbale:pre-processing_result} shows the pre-processing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+\Cref{table:pre-processing_result} shows the pre-processing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
 
@@
 \clearpage
+
+\section{Procedure: TestIntersection}
+% \begin{noindent}
+    \begin{algorithm}[htb!]
+        \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
+        \textbf{Input:} \\
+        $ \nodeLineNV \gets $ the node's translation path \\
+        $ \river \gets $ a river feature \\
+
+        \textbf{Output:} \\
+        Returns $ True $ if the node crosses a river. \\
+
+        \textbf{Local variables:} \\
+        $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \Edge $ \\
+        $ \Edge \gets $ an edge of $ \river $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{TestIntersection}{$ \nodeLineNV $, $ \river $}
+
+            \ForEach{$ \Edge \in \river $}
+                \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
+                \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \Edge $)
+
+                \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
+                    \State \Return{$ \nodeLineNV ~intersect~ \Edge $}
+                    \EndIf
+            \EndFor
+
+            \State \Return{$ False $}
+            \EndProcedure
+        \end{algorithmic}
+    \end{algorithm}
+    %\end{noindent}
 

### 2022-11-08 paper/section/data.tex
draft

@@
 Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHRa}.
-The first step is to obtain both geospatial boundaries and EHR data that match one another.
+The first step is to obtain both geospatial boundaries and EHR data.
 The second step is to pre-process the EHR data to remove empty and erroneous values.
 The final step is to transform the data into a format that is suitable for cartograms.
-
-\bobgraph{Geospatial Data}
 
@@
 \bobgraph{Clinical Commissioning Groups Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
-The number of CCGs changes over time due to the NHS re-organization.
+The number of CCGs changes over time due to NHS re-organization.
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
@@
 These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
-Other rivers cloud be chose, including additional and smaller rivers remains future work.
+Including additional and smaller rivers remains future work.
 
@@
 TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
-See \autoref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result.
-We describe the above steps in more detail in \autoref{app:pre-processing}.
+See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result.
+We describe the above steps in more detail in \Cref{app:pre-processing}.
 
-\bobgraph{EHR Data}
-
-We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
+\bobgraph{EHR Data: }We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
 The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.

### 2022-11-08 paper/section/evaluation.tex
draft

--- 2022-09-15 paper/section/evaluation.tex
@@
 
-\section{Evaluation}
+\newcommand{\pCount}{14 }
+\section{User-Centered Evaluation}
 
-\subsection{Hypotheses Formulation}
+We conduct a preliminary user study to evaluate the effectiveness of our approach.
 
-We formulate the following hypotheses to evaluate the effectiveness of our approach:
+\subsection{User Study Tasks}
 
-\textbf{H1:} The introduction of dynamic topological features can effectively improve the recognizability of CCGs.
+We formulate the following hypotheses to design our user study:
 
-\textbf{H2:} \dots
+\textbf{H1:} The introduction of rivers can improve the legibility and recognizability of cartograms.
 
-\subsection{Participants}
+\textbf{H2:} The introduction of rivers can reduce the time needed to locate a CCG in a cartogram.
 
-We recruited X participants, [add demographics here].
+We test these hypotheses using location based tasks.
+Given a standard choropleth on the left side of the screen and a cartogram on the right side, we ask user study participants to locate the given region in the choropleth in the cartogram.
 
-\subsection{Datasets}
+\subsection{User Study Variables}
+
+We discuss the variables of our user study in this section.
+
+\subsubsection{Independant Variables}
+
+The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. We set this parameter to either render rivers or not.
+
+% {
+% \renewcommand{\arraystretch}{1.5}
+% \begin{table}[!tb]
+% 	\centering
+% 		\begin{tabulary}{\columnwidth}{|*{2}{c|}}
+% 			\hline
+% 			\cellcolor{Mycolor2}\textbf{River Crossing} &
+% 			\cellcolor{Mycolor2}\textbf{River Presence} \\
+% 			\hline
+% 			Allow & Present \\
+% 			\hline
+% 			Allow & Absent \\
+% 			\hline
+% 			Disallow & Present \\
+% 			\hline
+% 			Disallow & Absent \\
+% 			\hline
+% 		\end{tabulary}
+% 	\caption{The combinations of independent variables.}
+% 	\label{table:Independant Variables}
+% \end{table}
+% }
+
+\subsubsection{Dependant Variables}
+
+\bobgraph{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
+
+\bobgraph{Response time:} Another dependant variable is the time taken by the participants to complete each task.
+
+\subsubsection{Control Variables}
+\bobgraph{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+
+\bobgraph{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find.
+The target CCG is shown in the form of a non-stop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
+
+\subsection{User Study}
+
+\subsubsection{Participants}
+
+We recruited \pCount participants, ...
+
+\subsubsection{Datasets}
 
@@
 \begin{itemize}
+    \item Population
     \item Under 75 mortality from cardiovascular disease
     \item Emergency admissions for alcohol-related liver disease
+    \item Alcohol-specific admission and readmission
 \end{itemize}
 
-Datasets are projected on a map that consists of 135 CGGs in England.
+135 CGGs are rendered on the screen as a choropleth on the left.
+We then generated another view on the right using cartograms.
+The color in both visual designs is mapped to the dataset.
+See \Cref{fig:task}.
 
-\subsection{Tasks}
+{
+    \begin{figure}[htb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task.png}
+        \caption{A sample task for participants.
+The left shows the choropleth map, and the right shows the corresponding cartogram.
+Both images show the three longest rivers in England, and the color is mapped to CCG population.
+The target CGG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.}
+        \label{fig:task}
+    \end{figure}
+}
 
-We conducted tasks to evaluate the effectiveness of our approach.
+\subsubsection{Procedure}
 
-\textbf{T1:} The participants were asked to familiarize themselves with the design, including the user options and the interaction methods.
-\textbf{T2:} The participants
+The user study is designed to be conducted online, and participants are asked to complete the tasks using their own computer. The user study includes four parts:
 
-\subsection{Questionnaire}
+\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos.
+The instructions are designed to help participants understand the concepts used in the tasks.
+Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
 
-We asked participants to complete a questionnaire to collect their feedback. The questionnaire consisted of the following questions:
+\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design. A demonstration of three sample tasks is also included in the instructional video.
 
-\begin{itemize}
-    \item \dots
-    \item \dots
-\end{itemize}
+\textbf{P3:} The participants are asked to complete 16 location tasks.
+The response and reaction time are recorded.
+These 16 CCGs were carefully selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+
+\textbf{P4:} The participants were asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
+
+These are provided in the appendix, and available at \url{https://osf.io/q39w7}.
+
+\subsection{User Study Results}
+
+In this section we analyze the results of our user study. We recruited \pCount participants to perform 16 location tasks.
+
+{
+    \begin{figure}[htb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task-rt.png}
+        \caption{The scatter plot shows the response time (in seconds) and answers for each task, filtering out [add the exact number in the end] responses that took over 20 seconds.}
+        \label{fig:task-rt}
+    \end{figure}
+}
 

### 2022-11-08 paper/section/introduction.tex
draft

@@
 User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
-Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained.}
+Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained [why too complicated?
+this is what we are doing in the software].}
         \label{fig:overview}
@@
 
-Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \autoref{table:accuracy}).
+Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
 For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-Standard cartograms are composed of uniform shape nodes.
+Standard Demers cartograms are composed of uniform shape nodes.
 As such, this can reduce their legibility.
 We implement a new cartographic layout algorithm that includes rivers into the layout of the nodes representing geographical regions.
+Introducing rivers into the cartogram improves their legibility.
 To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
 We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate the result.
-We also present a user study demonstrating its effectiveness.
+We present a user study demonstrating its effectiveness.
 
@@
 \begin{itemize}
-    \item A new variant of Demers cartograms that incorporates dynamic topological features to improve readability and recognizability,
+    \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
     \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers,
@@
 In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic topological features while minimizing geographical error.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic rivers while minimizing geographical error.

### 2022-11-08 paper/section/main.tex
draft

@@
       \caption{A flowchart illustration of our layout algorithm.
-See \autoref{fig:flowchart-stalemate} for the logic of processing a stalemate.
-See also \autoref{alg:UpdateLayout} in \autoref{sec:{Cartograms with Rivers}} for more detail.
-For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
+See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
+See also \Cref{alg:UpdateLayout} in \Cref{sec:{Cartograms with Rivers}} for more detail.
+For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.
+\new{the notations here do not match those in Algorithms.
+Which ones to use?}}
       \label{fig:flowchart}
@@
       \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
-      \caption{A flowchart illustration of stalemate processing. See \autoref{subsec:{Detecting Stalemates}}, \autoref{fig:stalemate}, and \autoref{fig:corridor} for more detail.}
+      \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Detecting Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
       \label{fig:flowchart-stalemate}
@@
 \begin{algorithm}[htb!]
-    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \autoref{sec:{Cartograms with Rivers}}.}\label{alg:UpdateLayout}
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Cartograms with Rivers}}.}\label{alg:UpdateLayout}
     \textbf{Global variables:} \\
@@
     \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
-    $ \nodeSize \gets $ the current size of all nodes \\
+    $ \nodeSizeAvg \gets $ the average size of all nodes \\
     $ \nodeSizeMax \gets $ the maximum size of a node \\
@@
         \Procedure{UpdateLayout}{}
-        \While{$ \nodeSize < \nodeSizeMax $}
+        \While{$ \nodeSizeAvg < \nodeSizeMax $}
             \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
@@
 
-            \State $ \nodeSize \gets \nodeSize ++$
+            \State $ \nodeSize ++ $
 
@@
 
-\autoref{alg:UpdateLayout} and \autoref{fig:flowchart} provide an overview of the cartogram layout process that includes rivers.
+\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the cartogram layout process that includes rivers.
 
@@
 We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests.
-This rendering approach enables us to adjust the river resolution as shown in \autoref{fig:river resolution}.
+This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}.
 We further apply smoothing by removing vertices that are too close to each other.
 
-\bobgraph{Node Layout and Overlap Removal:} The basic algorithm is as follows: we first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlaps.
+\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlap.
 We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-We gradually increase the node size by 1 unit at a time to ensure smooth transitions.
+Initialized with a pixel size of unity, we gradually increase the node size by one unit (proportional to its data) at a time to ensure smooth transitions.
 An increase in $ \nodeSize $ can cause the nodes to overlap.
-During overlap removal, we compute node trajectories (See \autoref{alg:check river intersection}) and translate nodes to their new position.
-Nodes that cross a river are translated back to their previous position.
-If a node oscillates across a river, we process this as a stalemate.
+During overlap removal, we compute node trajectories (See \Cref{alg:check river intersection}) and translate nodes to their new position.
+Nodes that cross a river, denoted $ \nodeCross $, are translated back to their previous position.
+If a node oscillates across a river, we identify \new{(or label?)} this as a stalemate.
 This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one pixel and repeat the algorithm until the max node size $ \nodeSizeMax $ is reached.
+We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
 The gradual size increase process provides stability to the layout and helps minimize geographical error.
-
-\begin{algorithm}[htb!]
-    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
-    \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
-
-    \textbf{Local variables:} \\
-    $ count \gets $ the number of river crossing nodes \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TranslateRiver}{$ \nodeList $}
-        \ForEach {$ \river \in \riverList $}
-        \State $ count \gets 0 $
-        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
-        \ForEach {$ \node \in \nodeList $}
-
-        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
-        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
-        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-        \State $ count \gets count ++ $
-        \EndIf
-
-        \EndIf
-
-        \EndFor
-
-        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
-        \EndFor
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
 
@@
                         \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
-                            \State $ \nodeCross \gets \nodeCross ++ $
-                            \State $ count \gets count ++ $
+                            \State $ \nodeCross ++ $
+                            \State $ count ++ $
 
@@
 
-\color{black}
+\begin{algorithm}[htb!]
+    \caption{Procedure to translate rivers. \new{I think this algorithm is nolonger necessary.}}\label{alg:TranslateRiver}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+
+    \textbf{Local variables:} \\
+    $ count \gets $ the number of river crossing nodes \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateRiver}{$ \nodeList $}
+        \ForEach {$ \river \in \riverList $}
+        \State $ count \gets 0 $
+        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
+        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
+        \State $ count ++ $
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
+        \EndFor
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
 
@@
 
-% We use rivers as topological boundaries and prevent nodes from crossing them.
-The resolution of rivers can be adjusted by the user, as shown in \autoref{fig:river resolution}.
-In this example, the number of vertices is reduced from 10,170 to 30.
-A pixel-level river resolution is not required as Demers cartograms do not need to represent geography at this level of detail.
-Instead, a resolution that matches the node size is used.
-Enabling river simplification reduces the number of edge intersection tests that need to be performed.
-%
+The logic for translating a node's position is detailed in \Cref{alg:TranslateNode}.
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
-See \autoref{alg:check river intersection}.
+See \Cref{alg:check river intersection}.
 A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
@@
 
-For all nodes that cross a river, we compute an average vector $ \frac{\riverVector}{count} $ used to translate the corresponding river.
-See \autoref{alg:TranslateRiver}.
-This step intends to create room for the next iteration of node translation without crossing a river.
+For all nodes that cross a river, we compute an average vector used to translate the corresponding river.
+Whenever a node, $ \node $, crosses a river , we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
+We then use a heuristic to translate the corresponding river using the average vector of node intersection $ \vectorAvg = \sum_{i=1}^j \frac{\Vector{\node_i\node_{i_t}}}{j} $, where $ j $ is the number of nodes intersecting the river.
+%  See \Cref{alg:TranslateRiver}.
+This step intends to create space for the next iteration of node translation without crossing a river.
 
-During river translations, some nodes might be crossed by the trajectory of river translations. We move these nodes by the same vector.
-
-% \begin{noindent}
-\begin{algorithm}[htb!]
-    \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
-    \textbf{Input:} \\
-    $ \nodeLineNV \gets $ the node's translation path \\
-    $ \river \gets $ a river feature \\
-
-    \textbf{Output:} \\
-    Returns $ True $ if the node crosses a river. \\
-
-    \textbf{Local variables:} \\
-    $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \Edge $ \\
-    $ \Edge \gets $ an edge of $ \river $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TestIntersection}{$ \nodeLineNV $, $ \river $}
-
-        \ForEach{$ \Edge \in \river $}
-            \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
-            \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \Edge $)
-
-            \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
-                \State \Return{$ \nodeLineNV ~intersect~ \Edge $}
-                \EndIf
-        \EndFor
-
-        \State \Return{$ False $}
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-%\end{noindent}
+When a river is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $.
+As a heuristic we also translate these nodes by $ \vectorAvg $.
+The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated in order to create space for nodes that are too crowded together.
 
@@
 
-As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \autoref{fig:stalemate}.
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}.
 If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate the congestion.
-A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \autoref{fig:corridor}C and D).
-All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \autoref{fig:corridor}E).
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
+All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
 
@@
             \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in A.
-To address the issue, we derive a corridor (orange rectangle in E) based on $ \node $ and $ \nodeFNOR $.
+To address this, we derive a corridor (orange rectangle in E) based on $ \node $ and $ \nodeFNOR $.
 All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
-In order to show a clear illustration, we place nodes sparsely in this figure.}
+For clarity in the illustration, we place nodes sparsely in this figure.}
             \label{fig:corridor}
@@
     \caption{Procedure to derive a corridor to translate enclosed nodes.
-We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \autoref{fig:stalemate}).}\label{alg:derive corridor}
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}
 
@@
 %\end{noindent}
+
+\subsection{Terminating the Algorithm}
+
+\new{The algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
+At this point, the node layout is optimal where no node has crossed, or crossed by a river.}

### 2022-11-08 paper/section/related.tex
feat: add evaluation draft
feat: revised paper
draft

@@
 
-\subsection{Definitions}
-
-While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
+\bobgraph{Definitions: }While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
@@
 	}
-	\caption{\textcolor{Mycolor3}{\textbf{Trade-off between dimensions}}.
-\textcolor{Mycolor6}{\textbf{Dimension sacrificed}} in order to improve \textcolor{Mycolor5}{\textbf{target dimension}}'s accuracy.}
+	\caption{\coloredsquare[Mycolor3]{0.25cm}{0.25cm} Trade-off between dimensions.
+\coloredsquare[Mycolor6]{0.25cm}{0.25cm} Dimension sacrificed in order to improve \coloredsquare[Mycolor5]{0.25cm}{0.25cm} target dimension's accuracy.}
 	\label{table:accuracy}
@@
 This is important in our use case scenario involving EHRs.
-Demers cartograms also have the advantage of preserving a higher level of topology.
+Demers cartograms also have the advantage of preserving a higher level of topology at the cost of geographical accuracy.
 }
@@
 
-\subsection{Peer-reviewed Applications}
-
-\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States.
+\bobgraph{Peer-reviewed Applications: }\citea{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
 \citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
@@
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
-See \autoref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
 
@@
 
-\subsection{Implementation in Popular Media}
-
-Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
+\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
 The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.

### 2022-11-08 user-study/index.js
feat: user study analysis

@@ -1,208 +1,0 @@
-const fs = require("fs");
-const { parse } = require("csv-parse");
-const likertscale = () => {
-    return {
-        "Strongly agree": 0,
-        "Agree": 0,
-        "Neither agree nor disagree": 0,
-        "Disagree": 0,
-        "Strongly disagree": 0,
-    }
-}
-
-const taskList = [
-    { ccg: "E38000243", withRiver: true, allowCrossing: false, correct: 0 },
-    { ccg: "E38000136", withRiver: true, allowCrossing: false, correct: 0 },
-    { ccg: "E38000242", withRiver: true, allowCrossing: false, correct: 0 },
-    { ccg: "E38000244", withRiver: true, allowCrossing: false, correct: 0 },
-    { ccg: "E38000243", withRiver: false, allowCrossing: true, correct: 0 },
-    { ccg: "E38000136", withRiver: false, allowCrossing: true, correct: 0 },
-    { ccg: "E38000242", withRiver: false, allowCrossing: true, correct: 0 },
-    { ccg: "E38000244", withRiver: false, allowCrossing: true, correct: 0 },
-    { ccg: "E38000243", withRiver: true, allowCrossing: true, correct: 0 },
-    { ccg: "E38000136", withRiver: true, allowCrossing: true, correct: 0 },
-    { ccg: "E38000242", withRiver: true, allowCrossing: true, correct: 0 },
-    { ccg: "E38000244", withRiver: true, allowCrossing: true, correct: 0 },
-    { ccg: "E38000243", withRiver: false, allowCrossing: false, correct: 0 },
-    { ccg: "E38000136", withRiver: false, allowCrossing: false, correct: 0 },
-    { ccg: "E38000242", withRiver: false, allowCrossing: false, correct: 0 },
-    { ccg: "E38000244", withRiver: false, allowCrossing: false, correct: 0 },
-];
-
-const resultPerCCG = {
-    E38000243: { withRiver: 0, withoutRiver: 0 },
-    E38000136: { withRiver: 0, withoutRiver: 0 },
-    E38000242: { withRiver: 0, withoutRiver: 0 },
-    E38000244: { withRiver: 0, withoutRiver: 0 },
-};
-
-const resultPerQuestion = [
-    { withRiver: 0, withoutRiver: 0, responseTime: 0 },
-    { withRiver: 0, withoutRiver: 0, responseTime: 0 },
[…]

### 2022-11-08 user-study/likert.js

@@ -1,0 +1,80 @@
+const fs = require("fs");
+const { parse } = require("csv-parse");
+const { stringify } = require("csv-stringify");
+const likertscale = () => {
+    return {
+        "Strongly agree": 0,
+        "Agree": 0,
+        "Neither agree nor disagree": 0,
+        "Disagree": 0,
+        "Strongly disagree": 0,
+    }
+}
+
+
+
+const choropleth = likertscale();
+const cartogram = likertscale();
+const likertScales = {
+    1: likertscale(),
+    2: likertscale(),
+    3: likertscale(),
+    4: likertscale(),
+    5: likertscale(),
+}
+
+function readCSV() {
+    fs.createReadStream("./result.csv")
+        .pipe(parse({ delimiter: ", ", relax_quotes: true, from_line: 1 }))
+        .on("data", function (row) {
+            addFamiliarity(row);
+            addLikert(row);
+
+        })
+        .on("end", function () {
+            // printFamiliarity();
+            // printLikert();
+        });
+}
+readCSV();
+
+function writeCSV(name, columns, data) {
+    const writableStream = fs.createWriteStream(`./${name}.csv`);
+    const stringifier = stringify({ header: true, columns: columns });
+    data.forEach((row) => {
+        stringifier.write(row);
+    });
+    stringifier.pipe(writableStream);
+}
+
+function addFamiliarity(row) {
+    choropleth[row[0].trim()]++;
+    cartogram[row[1].trim()]++;
+}
+
+function addLikert(row) {
+    for (let i = 1; i <= 5; i++) {
+        likertScales[i][row[49 + i].trim()]++;
+    }
+}
+function printLikert() {
+    for (let i = 1; i <= 5; i++) {
+        for (const key in likertScales[i]) {
+            console.log(`${likertScales[i][key]}`)
+        }
+        console.log("");
+    }
+}
+
+function printFamiliarity() {
+    console.log("choropleth familiarity");
+    for (const key in choropleth) {
+        console.log(`${choropleth[key]}`)
+    }
+
+    console.log("cartogram familiarity");
+    for (const key in cartogram) {
+        console.log(`${cartogram[key]}`)
+    }
+}
+

### 2022-11-08 user-study/splitCSV.js

@@ -1,0 +1,218 @@
+const fs = require("fs");
+const { parse } = require("csv-parse");
+const { stringify } = require("csv-stringify");
+const ttest = require('ttest');
+
+const taskList = [
+    { id: 1, ccg: "E38000243", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 2, ccg: "E38000136", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 3, ccg: "E38000242", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 4, ccg: "E38000244", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 5, ccg: "E38000243", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 6, ccg: "E38000136", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 7, ccg: "E38000242", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 8, ccg: "E38000244", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 9, ccg: "E38000243", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 10, ccg: "E38000136", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 11, ccg: "E38000242", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 12, ccg: "E38000244", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 13, ccg: "E38000243", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 14, ccg: "E38000136", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 15, ccg: "E38000242", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 16, ccg: "E38000244", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+];
+
+// with or without river
+// always disallow crossing
[…]

### 2022-11-08 user-study/ttest.js

@@ -1,0 +1,126 @@
+const fs = require("fs");
+const { parse } = require("csv-parse");
+const { stringify } = require("csv-stringify");
+const ttest = require('ttest');
+
+const taskList = [
+    { id: 1, ccg: "E38000243", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 2, ccg: "E38000136", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 3, ccg: "E38000242", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 4, ccg: "E38000244", withRiver: true, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 5, ccg: "E38000243", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 6, ccg: "E38000136", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 7, ccg: "E38000242", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 8, ccg: "E38000244", withRiver: false, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 9, ccg: "E38000243", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 10, ccg: "E38000136", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 11, ccg: "E38000242", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 12, ccg: "E38000244", withRiver: true, allowCrossing: true, correct: 0, time: 0, validCount: 0 },
+    { id: 13, ccg: "E38000243", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 14, ccg: "E38000136", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 15, ccg: "E38000242", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+    { id: 16, ccg: "E38000244", withRiver: false, allowCrossing: false, correct: 0, time: 0, validCount: 0 },
+];
+
+// with or without river
+// always disallow crossing
[…]

### 2022-11-09 paper/GHRVis.tex
draft: limitations and conclusions

@@
 
+% For cover's captionof
+\usepackage[font=normalfont,labelfont=bf]{caption}
+
 \include{preamble/preamble.basic}
@@
 
+\begin{document}
+
 \author[Q. Wang, K. Xu \& R. S. Laramee]
-{\parbox{\textwidth}{\centering Q. Wang, K. Xu
-  and R. S. Laramee
+{
+\parbox{\textwidth}{\centering Q. Wang, K. Xu and R. S. Laramee}
+\\
+\parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
+    \centering
+    \includegraphics[height=3cm,keepaspectratio]{figure/cover.png}
+    \captionof{figure}{The River Thames passing through NHS Clinical Commissioning Groups in London and surrounding areas.}
+\end{center}
 }
-    \\
-\parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK}
 }
@@
 \title{Cartograms with Rivers}
-
-\begin{document}
 
@@
 
-\section{Limitations and Future Work}
+\input{section/limitation}
 
-\section{Conclusions}
+\input{section/conclusion}
 
-\section{Acknowledgments}
-This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC).
+% \section{Acknowledgments}
+% This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC).
 EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.
@@
 
-\cleardoublepage
+\clearpage
 

### 2022-11-09 paper/section/conclusion.tex
draft: limitations and conclusions

\section{Conclusions}

In this paper, we present a novel approach to generate cartograms with rivers. We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers. We evaluate our approach using a user study, and the results partially support our hypotheses: the introduction of rivers improves the legibility but doesn't reduce the time needed to locate a region. We also discuss the limitations of our work and the future directions.

### 2022-11-09 paper/section/evaluation.tex
draft: limitations and conclusions

@@
-\color{blue}
-
-\newcommand{\pCount}{14 }
+\newcommand{\pCount}{15 }
 \section{User-Centered Evaluation}
@@
 \subsection{User Study}
+In this section we describe the user study participants, datasets, and the experiment procedure.
 
-\subsubsection{Participants}
+\bobgraph{Participants:} We recruited \pCount participants.
 
-We recruited \pCount participants, ...
+\begin{itemize}
+    \item Gender:  8 females and 7 males
+    \item Age Group:   18-24 (7), 25-29 (8)
+    \item Education:   3 Bachelor's, 11 Master's, 1 PhD
+\end{itemize}
 
-\subsubsection{Datasets}
-
-We used the following datasets for our evaluation:
+\bobgraph{Datasets:} We used the following datasets for our evaluation:
 
@@
 
-In this section we analyze the results of our user study. We recruited \pCount participants to perform 16 location tasks.
+In this section we analyze the results of our user study.
+We recruited \pCount participants to perform 16 location tasks each, in total we collected 240 responses, 59 responses that took over 20 seconds are excluded from the analysis.
 
@@
         \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task-rt.png}
-        \caption{The scatter plot shows the response time (in seconds) and answers for each task, filtering out [add the exact number in the end] responses that took over 20 seconds.}
+        \caption{The scatter plot shows the response time (in seconds) and answers for each task.
+Correct answers are shown in green circle, and incorrect answers are shown in red triangle.
+Solid glyphs represent the task without rivers, and hollow glyphs represent the task with rivers.}
+        \label{fig:task-all}
+    \end{figure}
+}
+
+\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
+The chart shows that the introduction of rivers has improved the accuracy for two CCGs.
+The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(176.43) = 2.92, $p$ = 0.0039.
+This supports our hypothesis H1 that the introduction of rivers improves the legibility of cartogram.
+
+{
+    \begin{figure}[htb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
+        \caption{The bar chart shows the accuracy of locating CCGs.}
+        \label{fig:task-acc}
+    \end{figure}
+}
+
+\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average response time measured for four CCGs.
+The chart shows that the introduction of rivers has slightly reduced the time needed to locate CCGs.
+The two groups show no significant differences according to the two-sample unequal variance $t$-test performed, $t$(168.03) = 1.29, $p$ = 0.20.
+The statistical analysis does not support our hypothesis H2, that the introduction of rivers does not reduce the response time significantly.
+
+{
+    \begin{figure}[htb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/rt.png}
+        \caption{The bar chart shows the average response time for locating CCGs.}
         \label{fig:task-rt}
@@
 }
-
-\color{black}

### 2022-11-09 paper/section/limitation.tex
draft: limitations and conclusions

\section{Limitations and Future Work}

In this section we discuss some limitations of our work and the future directions. The first limitation is the color map we use to depict the data in our cartograms. We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green. However, we believe that the choice of color map can have a significant impact on the legibility of cartogram. In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target, see \Cref{fig:extreme}. We plan to explore the impact of different color maps on the legibility of cartogram in future work.

{
    \begin{figure}[htb!]
        \centering
        \includegraphics[width=\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
        \caption{Due to color and relative location, the CCGs in the black circle are extremely easy to locate.}
        \label{fig:extreme}
    \end{figure}
}

Another limitation is the overlap removal algorithm (FNOR) we use. We believe that developing a new algorithm with built-in constraint support can significantly reduce the time needed to generate cartograms with rivers.

### 2022-11-09 paper/section/main.tex
draft: limitations and conclusions

@@
 
-\new{The algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
-At this point, the node layout is optimal where no node has crossed, or crossed by a river.}
+The algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
+At this point, the node layout is optimal where no node has crossed, or crossed by a river.

### 2022-11-09 user-study/splitCSV.js

@@ -50,7 +50,6 @@ function readCSV() {
     fs.createReadStream("./result.csv")
         .pipe(parse({ delimiter: ",", from_line: 1 }))
         .on("data", function (row) {
-            rowCount++;
             for (let i = 0; i < taskList.length; i++) {
                 const time = parseInt(row[i * 2 + 1]) / 1000;
 
@@ -59,6 +58,7 @@ function readCSV() {
                     continue;
                 }
 
+                rowCount++;
                 const correctAnswer = taskList[i].ccg;
                 const answer = row[i * 2];
 
@@ -95,7 +95,7 @@ function readCSV() {
 
         })
         .on("end", function () {
-
+            console.log(rowCount);
             let w = result.filter((row) => !row.allowCrossing && row.withRiver).map((row) => row.rt);
             let wo = result.filter((row) => !row.allowCrossing && !row.withRiver).map((row) => row.rt);
             tTest(w, wo, "group1");
@@ -112,6 +112,7 @@ function readCSV() {
             wo = result.filter((row) => !row.allowCrossing && !row.withRiver).map((row) => row.rt);
             tTest(w, wo, "group4");
 
+            //E38000136
             w = result.filter((row) => row.withRiver).map((row) => row.rt);
             wo = result.filter((row) => !row.withRiver).map((row) => row.rt);
             tTest(w, wo, "group5");
@@ -203,9 +204,16 @@ function computeSingleGroup() {
 }
 
 function tTest(w, wo, group) {
-    let t = ttest(w, wo)
-    // console.log(group);
-    // console.log(t.testValue());
+    let t = ttest(wo, w)
+    console.log(w);
+    console.log(wo);
+    console.log(group);
+    console.log("df:", t.freedom());
+    console.log("t:", t.testValue());
+    console.log("confidence:", t.confidence());
+    console.log("p:", t.pValue());
+    console.log("t:", t.valid());
+    console.log(t);
 }
 
 function writeCSV(name, columns, data) {

### 2022-11-15 paper/section/appendix.tex
draft: review

@@
 
+\section{Procedure: TranslateRiver}
+
+\begin{algorithm}[tb!]
+    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+
+    \textbf{Local variables:} \\
+    $ count \gets $ the number of river crossing nodes \\
+    $ \riverList \gets $ a list of $ \river $ representing river features \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateRiver}{$ \nodeList $}
+        \ForEach {$ \river \in \riverList $}
+        \State $ count \gets 0 $
+        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
+        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
+        \State $ count ++ $
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
+        \EndFor
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
 \section{Procedure: TestIntersection}
 % \begin{noindent}
-    \begin{algorithm}[htb!]
+    \begin{algorithm}[tb!]
         \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}

### 2022-11-15 paper/section/conclusion.tex
draft: review

@@
 
-In this paper, we present a novel approach to generate cartograms with rivers.
+In this paper, we present a novel approach to generate non-contiguous cartograms with rivers.
 We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers.
-We evaluate our approach using a user study, and the results partially support our hypotheses: the introduction of rivers improves the legibility but doesn't reduce the time needed to locate a region.
-We also discuss the limitations of our work and the future directions.
+We evaluate our approach using a user study, and the results  support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
+We also discuss the limitations of our work and future research directions.

### 2022-11-15 paper/section/data.tex
draft: review

@@
  {
-\begin{figure}[htb!]
+  \begin{figure}[tb!]
       \centering
@@
 The final step is to transform the data into a format that is suitable for cartograms.
-
 Geospatial boundaries, or shapefiles, were obtained from the sources described here.
@@
 
-\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
+\begin{lstlisting}[float=tp,caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
     relation(2263653);>>;
@@
 
-For all datasets, a spreadsheet including the following data is provided:
+For all datasets, a spreadsheet including the following is provided:
 

### 2022-11-15 paper/section/evaluation.tex
draft: review

@@
-\newcommand{\pCount}{15 }
+\newcommand{\pCount}{24 }
 \section{User-Centered Evaluation}
@@
 
-\textbf{H2:} The introduction of rivers can reduce the time needed to locate a CCG in a cartogram.
+\textbf{H1.1:} The introduction of rivers can increase target location accuracy.
+
+\textbf{H1.2:} The introduction of rivers can reduce the time needed to locate a CCG in a cartogram.
 
@@
 Given a standard choropleth on the left side of the screen and a cartogram on the right side, we ask user study participants to locate the given region in the choropleth in the cartogram.
+See \Cref{fig:task}.
 
@@
 
-\subsubsection{Independant Variables}
-
-The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. We set this parameter to either render rivers or not.
+\bobgraph{Independant Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram.
+We set this parameter to either render rivers or not.
 
@@
 
-\subsubsection{Dependant Variables}
-
-\bobgraph{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+\bobgraph{Dependant Variables: }\bobgraph{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
 The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
@@
 
-\subsubsection{Control Variables}
-\bobgraph{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+\bobgraph{Control Variables: }\bobgraph{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
 
@@
 
-\subsection{User Study}
+\subsection{User Study Design}
 In this section we describe the user study participants, datasets, and the experiment procedure.
@@
 {
-    \begin{figure}[htb!]
+\begin{figure}[tb!]
     \centering
@@
 
-The user study is designed to be conducted online, and participants are asked to complete the tasks using their own computer. The user study includes four parts:
+The user study is designed to be conducted online, and participants are asked to complete the tasks using their own hardware. The user study includes four parts:
 
@@
 The instructions are designed to help participants understand the concepts used in the tasks.
-Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+Instructional videos are available at \URL{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
 
-\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design. A demonstration of three sample tasks is also included in the instructional video.
+\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
+These tasks also serve as a quality test to see if the users are taking the study seriously.
+A demonstration of three sample tasks is also included in the instructional video.
 
@@
 The response and reaction time are recorded.
-These 16 CCGs were carefully selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+These 16 CCGs were selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
 
@@
 
-These are provided in the appendix, and available at \url{https://osf.io/q39w7}.
+The entire study is available at \URL{https://osf.io/q39w7}.
 
@@
 
-{
-    \begin{figure}[htb!]
-        \centering
-        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task-rt.png}
-        \caption{The scatter plot shows the response time (in seconds) and answers for each task.
-Correct answers are shown in green circle, and incorrect answers are shown in red triangle.
-Solid glyphs represent the task without rivers, and hollow glyphs represent the task with rivers.}
-        \label{fig:task-all}
-    \end{figure}
-}
-
 \bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
@@
 The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(176.43) = 2.92, $p$ = 0.0039.
-This supports our hypothesis H1 that the introduction of rivers improves the legibility of cartogram.
+This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.
 
     {
-    \begin{figure}[htb!]
+        \begin{figure}[tb!]
             \centering
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
-        \caption{The bar chart shows the accuracy of locating CCGs.}
+            \caption{The bar chart shows the accuracy of locating CCGs with and without rivers.}
             \label{fig:task-acc}
@@
 The two groups show no significant differences according to the two-sample unequal variance $t$-test performed, $t$(168.03) = 1.29, $p$ = 0.20.
-The statistical analysis does not support our hypothesis H2, that the introduction of rivers does not reduce the response time significantly.
+The statistical analysis does not support our hypothesis H1.2, that the introduction of rivers reduces the response time significantly.
 
     {
-    \begin{figure}[htb!]
+        \begin{figure}[tb!]
             \centering
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/rt.png}
-        \caption{The bar chart shows the average response time for locating CCGs.}
+            \caption{The bar chart shows the average response time for locating CCGs with and without rivers.}
             \label{fig:task-rt}
@@
     }
+
+\bobgraph{Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
+\begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
+    \item 87.50\% of participants agree that including rivers is useful.
+    \item 83.33\% of participants agree that rivers increase the legibility of a cartogram.
+    \item 95.83\% of participants agree that including rivers makes cartograms easier to understand.
+    \item 62.50\% of participants agree that including rivers makes CCGs easier to locate.
+    \item 87.50\% of participants agree that including rivers adds value to the standard cartogram.
+\end{enumerate}
+
+    {
+        \begin{figure}[b!]
+            \centering
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/likert.png}
+            \caption{The stacked bar chart shows the responses of Likert Scale questions.}
+            \label{fig:likert}
+        \end{figure}
+    }

### 2022-11-15 paper/section/introduction.tex
draft: review

@@
-{
-    \begin{figure*}[ht!]
-        \centering
-        \includegraphics[width=0.8\textwidth,keepaspectratio]{figure/UI.png}
-        \caption{A screenshot of the software.
-User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
-Other options include the ability to control the overlap removal behavior: rivers can be static, dynamic, or constrained [why too complicated?
-this is what we are doing in the software].}
-        \label{fig:overview}
-    \end{figure*}
-}
-
 \section{Introduction and Motivation}
@@
 Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
-Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and public health decision-making \cite{gao2020Visualising, sack2021Visualizing}.
+Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.
 
@@
 As such, this can reduce their legibility.
-We implement a new cartographic layout algorithm that includes rivers into the layout of the nodes representing geographical regions.
+We implement a new cartographic layout algorithm that includes rivers into the placement of the nodes representing geographical regions.
 Introducing rivers into the cartogram improves their legibility.
 To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm to a real-world case study using an Electronic Health Records (EHR) dataset to evaluate the result.
+We then apply the algorithm to a real-world case study using Electronic Health Records (EHR) data to evaluate the result.
 We present a user study demonstrating its effectiveness.

### 2022-11-15 paper/section/limitation.tex
draft: review

@@
 
-In this section we discuss some limitations of our work and the future directions.
+In this section we discuss some limitations of our work and future research directions.
 The first limitation is the color map we use to depict the data in our cartograms.
 We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
-However, we believe that the choice of color map can have a significant impact on the legibility of cartogram.
+However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
 In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target, see \Cref{fig:extreme}.
@@
 {
-    \begin{figure}[htb!]
+    \begin{figure}[tb!]
         \centering
         \includegraphics[width=\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
-        \caption{Due to color and relative location, the CCGs in the black circle are extremely easy to locate.}
+        \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
         \label{fig:extreme}

### 2022-11-15 paper/section/main.tex
draft: review

@@
  {
-    \begin{figure}[htb!]
-        \centering
-        \includegraphics[width=0.7\columnwidth,keepaspectratio]{figure/river_resolution.png}
-        \caption{The resolution of rivers can be dynamically adjusted by the user.
-A shows River Thames at its original resolution with 10,170 edges.
-B shows the river at a reduced resolution of 49 edges.
-We further smooth the river by removing vertices in dense areas, as shown in C.
-The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
-        \label{fig:river resolution}
-    \end{figure}
-}
-
-{
-  \begin{figure*}[htb!]
+  \begin{figure*}[tb!]
       \centering
@@
   \end{figure*}
+ }
 
-  \begin{figure}[htb!]
+ {
+  \begin{figure}[tb!]
+      \centering
+      \includegraphics[width=0.7\columnwidth,keepaspectratio]{figure/river_resolution.png}
+      \caption{The resolution of rivers can be dynamically adjusted by the user.
+(A) shows River Thames at its original resolution with 10,170 edges.
+(B) shows the river at a reduced resolution of 49 edges.
+We further smooth the river by removing 19 vertices in dense areas, as shown in (C).
+The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+      \label{fig:river resolution}
+  \end{figure}
+ }
+
+ {
+  \begin{figure}[tb!]
       \centering
@@
 
-\begin{algorithm}[htb!]
+\begin{algorithm}[tb!]
     \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Cartograms with Rivers}}.}\label{alg:UpdateLayout}
@@
     $ \nodeSizeMax \gets $ the maximum size of a node \\
-    $ \riverList \gets $ a list of $ \river $ representing river features \\
     $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
@@
 \bobgraph{Initialization with Rivers:} We first load and render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ of 1 pixel.
 We then load the river shapefiles and render the rivers.
@@
 This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}.
-We further apply smoothing by removing vertices that are too close to each other.
+We further apply simplification by removing vertices that are too close to each other.
 
@@
 Nodes that cross a river, denoted $ \nodeCross $, are translated back to their previous position.
-If a node oscillates across a river, we identify \new{(or label?)} this as a stalemate.
+If a node oscillates across a river, we identify this as a stalemate.
 This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
@@
 
-\begin{algorithm}[htb!]
+\begin{algorithm}[tb!]
     \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
@@
 
-\begin{algorithm}[htb!]
-    \caption{Procedure to translate rivers. \new{I think this algorithm is nolonger necessary.}}\label{alg:TranslateRiver}
-    \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
-
-    \textbf{Local variables:} \\
-    $ count \gets $ the number of river crossing nodes \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TranslateRiver}{$ \nodeList $}
-        \ForEach {$ \river \in \riverList $}
-        \State $ count \gets 0 $
-        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
-        \ForEach {$ \node \in \nodeList $}
-
-        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
-        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
-        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-        \State $ count ++ $
-        \EndIf
-
-        \EndIf
-
-        \EndFor
-
-        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
-        \EndFor
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
 \subsection{River Intersection Testing}
@@
 Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
+We label these nodes, $ \NxR $
 
@@
 
-For all nodes that cross a river, we compute an average vector used to translate the corresponding river.
+For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $.
 Whenever a node, $ \node $, crosses a river , we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
-We then use a heuristic to translate the corresponding river using the average vector of node intersection $ \vectorAvg = \sum_{i=1}^j \frac{\Vector{\node_i\node_{i_t}}}{j} $, where $ j $ is the number of nodes intersecting the river.
-%  See \Cref{alg:TranslateRiver}.
+We then use a heuristic to translate $ \river $ using the average vector of node intersection $ \vectorAvg = \sum_{i=1}^j \frac{\Vector{\node_i\node_{i_t}}}{j} $, where $ j $ is the number of nodes intersecting the river.
 This step intends to create space for the next iteration of node translation without crossing a river.
+The procedure for translating riers is provided by \Cref{alg:TranslateRiver} in the Appendix.
 
@@
 All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
+The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.
 
     {
-        \begin{figure}[htb!]
+        \begin{figure}[tb!]
             \centering
@@
     {
-        \begin{figure}[htb!]
+        \begin{figure}[tb!]
             \centering
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/corridor.png}
-            \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in A.
-To address this, we derive a corridor (orange rectangle in E) based on $ \node $ and $ \nodeFNOR $.
+            \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A).
+To address this, we derive a corridor (orange rectangle in (E)) based on $ \node $ and $ \nodeFNOR $.
 All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
@@
 
-\begin{algorithm}[htb!]
+\begin{algorithm}[tb!]
     \caption{Procedure to derive a corridor to translate enclosed nodes.
@@
 At this point, the node layout is optimal where no node has crossed, or crossed by a river.
+
+\subsection{User Options}
+
+{
+\begin{figure*}[htb!]
+    \centering
+    \includegraphics[width=\textwidth,keepaspectratio]{figure/UI.png}
+    \caption{A screenshot of the software.
+User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
+Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.}
+    \label{fig:overview}
+\end{figure*}
+}

### 2022-11-15 paper/section/related.tex
draft: review

@@
 
-\bobgraph{Definitions: }While we focus on rectangular and Dorling cartograms in this paper, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
+\bobgraph{Definitions: }While we focus on rectangular and Dorling cartograms, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.

### 2022-11-15 paper/section/main.tex
draft: User Options

@@
 
+\Cref{fig:overview} presents an overview of the implementation, including the user options. The user can adjust the following parameters:
+
+\bobgraph{Visibility:} The visibility of various elements, including the choropleth, rivers, nodes, and node centroids, can be toggled on and off.
+
+\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR datasets or set to uniform.
+
+\bobgraph{Overlap Removal Behavior:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.
+The behavior of rivers during the process can also be adjusted by the user: 1) Disallow river crossing: this option forbids nodes from crossing rivers; 2) Disable river translation: this option disables the translation of rivers.
+Both options are useful for generating different layouts for observations.
+
+\bobgraph{Corridor Length:} The user can define the length of a corridor, which is used to resolve stalemate situations.
+A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less legible layout.
+
+\bobgraph{River Thickess and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
+Similarity, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
+
 {

### 2022-11-17 paper/GHRVis.tex
draft: review

@@
     \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
-    \centering
-    \includegraphics[height=3cm,keepaspectratio]{figure/cover.png}
-    \captionof{figure}{The River Thames passing through NHS Clinical Commissioning Groups in London and surrounding areas.}
+            \vspace{2em}
+            \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
+            \captionof{figure}{The River Thames passing through a cartographic representation of the NHS Clinical Commissioning Groups in London and surrounding areas.}
         \end{center}
@@
 \begin{abstract}
-Cartograms are hybrid representations of geographical and abstract data based on a value-by-area mapping technique.
+    Cartograms are representations of geographical and abstract data based on a value-by-area mapping technique.
 A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
 One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.
-To address this, we propose a novel cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
+To address this, we propose a novel hybrid cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
 Rivers affect both the layout and appearance of cartograms.
-We evaluate the proposed layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
+We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
 The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and geographical accuracy of the cartogram.
@@
 
-\clearpage
 % reset \section command for section title "References"

### 2022-11-17 paper/section/appendix.tex
draft: User Options
draft: review

@@
 {
-\begin{figure}[tbh!]
+\begin{figure}[h!]
     \centering
@@
 
-\begin{figure}[tbh!]
+\begin{figure}[h!]
     \centering
@@
 
-\begin{algorithm}[tb!]
+\begin{algorithm}[h!]
     \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
@@
     \textbf{Local variables:} \\
-    $ count \gets $ the number of river crossing nodes \\
     $ \riverList \gets $ a list of $ \river $ representing river features \\
@@
     \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
     \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
     \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+    $ j \gets $ the number of nodes intersecting $ \river $  \\
 
@@
         \ForEach {$ \river \in \riverList $}
-        \State $ count \gets 0 $
+        \State $ j \gets 0 $
         \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
@@
         \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-        \State $ count ++ $
+        \State $ j ++ $
         \EndIf
@@
 
-        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{count} $
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{j} $
         \EndFor
@@
 % \begin{noindent}
-    \begin{algorithm}[tb!]
+    \begin{algorithm}[h!]
         \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
@@
 
-    \begin{algorithm}[tbh!]
+    \begin{algorithm}[h!]
         \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
@@
 
-\begin{algorithm}[tbh!]
+\begin{algorithm}[h!]
     \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

### 2022-11-17 paper/section/conclusion.tex
draft: review

@@
 We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers.
-We evaluate our approach using a user study, and the results  support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
+We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
 We also discuss the limitations of our work and future research directions.

### 2022-11-17 paper/section/data.tex
draft: review

@@
 
-\bobgraph{Clinical Commissioning Groups Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
+\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
 The number of CCGs changes over time due to NHS re-organization.
@@
 
-\bobgraph{River Features:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England.
+\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England.
 These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.

### 2022-11-17 paper/section/evaluation.tex
draft: User Options
draft: review

@@
 
-\textbf{H1.2:} The introduction of rivers can reduce the time needed to locate a CCG in a cartogram.
+\textbf{H1.2:} The introduction of rivers can reduce the time needed to locate a target region in a cartogram.
 
@@
 
-\bobgraph{Dependant Variables: }\bobgraph{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+\bobgraph{Dependant Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
 The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
+\textit{Response time:} Another dependant variable is the time taken by the participants to complete each task.
 
-\bobgraph{Response time:} Another dependant variable is the time taken by the participants to complete each task.
-
-\bobgraph{Control Variables: }\bobgraph{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
-
-\bobgraph{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find.
+\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find.
 The target CCG is shown in the form of a non-stop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
@@
 \subsection{User Study Design}
-In this section we describe the user study participants, datasets, and the experiment procedure.
+In this section we describe the user study participants, datasets, and the experimental procedure.
 
@@
 \begin{itemize}
-    \item Gender:  8 females and 7 males
-    \item Age Group:   18-24 (7), 25-29 (8)
-    \item Education:   3 Bachelor's, 11 Master's, 1 PhD
+    \item Gender: 10 females and 14 males
+    \item Age Group: 18-24 (16), 25-29 (8)
+    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
 \end{itemize}
@@
 The color in both visual designs is mapped to the dataset.
-See \Cref{fig:task}.
+See \Cref{fig:task} for an example.
 
@@
     \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/task.png}
-    \caption{A sample task for participants.
+    \caption{A sample location task for participants.
 The left shows the choropleth map, and the right shows the corresponding cartogram.
@@
 
-The entire study is available at \URL{https://osf.io/q39w7}.
+The materials for P1-P4 are available at \URL{https://osf.io/q39w7}.
 
@@
 
-In this section we analyze the results of our user study.
-We recruited \pCount participants to perform 16 location tasks each, in total we collected 240 responses, 59 responses that took over 20 seconds are excluded from the analysis.
+In this section we analyze the results of our preliminary user study.
+We recruited \pCount participants to perform 16 location tasks each, in total we collected 384 responses, 24 responses that took over 60 seconds are excluded from the analysis.
+We hypothesize that participants taking longer than 60 seconds to complete a task were distracted from the study.
 
@@
 The chart shows that the introduction of rivers has improved the accuracy for two CCGs.
-The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(176.43) = 2.92, $p$ = 0.0039.
+The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002.
 This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
-            \caption{The bar chart shows the accuracy of locating CCGs with and without rivers.}
+            \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers, the x-axis shows the target CCGs.}
             \label{fig:task-acc}
@@
 \bobgraph{Response Time:} \Cref{fig:task-rt} shows the average response time measured for four CCGs.
-The chart shows that the introduction of rivers has slightly reduced the time needed to locate CCGs.
-The two groups show no significant differences according to the two-sample unequal variance $t$-test performed, $t$(168.03) = 1.29, $p$ = 0.20.
-The statistical analysis does not support our hypothesis H1.2, that the introduction of rivers reduces the response time significantly.
+The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.
+The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38.
 
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/rt.png}
-            \caption{The bar chart shows the average response time for locating CCGs with and without rivers.}
+            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers, the x-axis shows the target CCGs.}
             \label{fig:task-rt}
@@
 
-\bobgraph{Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
+\bobgraph{Participant Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
 \begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
@@
     }
+
+The qualitative results clearly support our hypothesis H1 that rivers generally increase the legibility of cartograms.

### 2022-11-17 paper/section/introduction.tex
draft: review

@@
 
-Cartograms are representations of hybrid geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
+Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
 Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.
@@
 For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
-Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel dynamic topological features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
+Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
 Standard Demers cartograms are composed of uniform shape nodes.
 As such, this can reduce their legibility.
-We implement a new cartographic layout algorithm that includes rivers into the placement of the nodes representing geographical regions.
+We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geographical regions.
 Introducing rivers into the cartogram improves their legibility.
@@
     \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
-    \item A novel layout algorithm that preserves node positions relative to dynamic topological features such as rivers,
+    \item A novel hybrid layout algorithm that combines node positions with features such as rivers,
     \item A user study evaluation of the technique with an application to EHRs.
@@
 One of the major challenges involved is how to develop a layout algorithm that handles different shapes.
-In other words, the layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by dynamic rivers while minimizing geographical error.
+In other words, the hybrid layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing geographical error.

### 2022-11-17 paper/section/limitation.tex
draft: review

@@
 However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
-In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target, see \Cref{fig:extreme}.
-We plan to explore the impact of different color maps on the legibility of cartogram in future work.
+In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
+See \Cref{fig:extreme} for an example.
+We plan to explore the impact of different color maps on the legibility of cartograms in future work.
 

### 2022-11-17 paper/section/main.tex
draft: review

--- 2022-11-15 paper/section/main.tex
@@
       \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
-      \caption{A flowchart illustration of our layout algorithm.
+      \caption{An overview of our hybrid layout algorithm.
 See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
 See also \Cref{alg:UpdateLayout} in \Cref{sec:{Cartograms with Rivers}} for more detail.
-For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.
-\new{the notations here do not match those in Algorithms.
-Which ones to use?}}
+For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
       \label{fig:flowchart}
@@
     \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
     \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
@@
 
-\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the cartogram layout process that includes rivers.
+\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the hybrid cartogram layout process that includes rivers.
 
@@
 An increase in $ \nodeSize $ can cause the nodes to overlap.
-During overlap removal, we compute node trajectories (See \Cref{alg:check river intersection}) and translate nodes to their new position.
-Nodes that cross a river, denoted $ \nodeCross $, are translated back to their previous position.
+During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position.
+Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position.
 If a node oscillates across a river, we identify this as a stalemate.
@@
     \textbf{Output:} \\
-    $ count \gets $ the number of river crossing nodes in the input \\
+    $ j \gets $ the number of river crossing nodes in the input \\
 
@@
     \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the counter for river crosses for $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
     \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
@@
         \Procedure{TranslateNode}{$ \nodeList $}
-        \State $ count \gets 0 $
+        \State $ j \gets 0 $
         \ForEach {$ \node \in \nodeList $}
@@
         \State $ \nodeCross ++ $
-        \State $ count ++ $
+        \State $ j ++ $
 
@@
         \EndFor
-        \State \Return{$ count $}
+        \State \Return{$ j $}
         \EndProcedure
@@
 Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
-We label these nodes, $ \NxR $
+We label these nodes, $ \NxR $.
 
@@
 The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated in order to create space for nodes that are too crowded together.
+In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
+Thus, we can influence how far a river, $ \river $, is translated in each iteration of the algorithm.
+We can use this to ensure smooth transitions between iterations.
 
-\subsection{Detecting Stalemates}
+\subsection{Process Stalemates}
 
@@
 
+\Cref{fig:overview} presents an overview of the implementation, including the user options. The user can adjust the following parameters:
+
+\bobgraph{Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
+
+\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR attributes or set to uniform.
+
+\bobgraph{Overlap Removal Speed:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.
+
+\bobgraph{Max Node Size:} The user can adjust the max node size, which is used to terminate the algorithm.
+
+\bobgraph{RIver Translation Behavior:} The behavior of rivers during the process can also be adjusted by the user: 1) Prevent river crossing: this option forbids nodes from crossing rivers; 2) Disable river translation: this option disables the translation of rivers.
+Both options are useful for generating different layouts and for observing the behavior of the hybrid layout algorithm.
+
+\bobgraph{Corridor Length:} The user can define the length of a corridor, which is used to resolve stalemate situations.
+A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.
+
+\bobgraph{River Thickess and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
+Similarity, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
+
 {

### 2022-11-22 paper/section/evaluation.tex
fix: typo

@@
 
-\bobgraph{Independant Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram.
+\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram.
 We set this parameter to either render rivers or not.
@@
 % 	\caption{The combinations of independent variables.}
-% 	\label{table:Independant Variables}
+% 	\label{table:Independent Variables}
 % \end{table}
@@
 
-\bobgraph{Dependant Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
 The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
-\textit{Response time:} Another dependant variable is the time taken by the participants to complete each task.
+\textit{Response time:} Another dependent variable is the time taken by the participants to complete each task.
 

### 2022-11-25 paper/section/evaluation.tex
draft: review

@@
 
-\subsection{User Study Tasks}
+\subsection{Study Hypothesis}
 
-We formulate the following hypotheses to design our user study:
+We formulate the following hypotheses to motivate our user study:
 
@@
 We test these hypotheses using location based tasks.
-Given a standard choropleth on the left side of the screen and a cartogram on the right side, we ask user study participants to locate the given region in the choropleth in the cartogram.
+Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
 See \Cref{fig:task}.
@@
 
-\bobgraph{Datasets:} We used the following datasets for our evaluation:
+\bobgraph{Datasets:} We used the following EHR datasets for our evaluation:
 
@@
 The response and reaction time are recorded.
-These 16 CCGs were selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+These 16 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
 
-\textbf{P4:} The participants were asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
+\textbf{P4:} The participants are asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
 
-The materials for P1-P4 are available at \URL{https://osf.io/q39w7}.
+The materials for P1-P4 are available on the Open Science Foundation's website at \URL{https://osf.io/q39w7}.
 
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
-            \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers, the x-axis shows the target CCGs.}
+            \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
             \label{fig:task-acc}
@@
 
-\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average response time measured for four CCGs.
+\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to located a CCG measured for four CCGs.
 The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.
 The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38.
+However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.
 
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/rt.png}
-            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers, the x-axis shows the target CCGs.}
+            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
             \label{fig:task-rt}
@@
 {
-        \begin{figure}[b!]
+\begin{figure}[tb!]
     \centering
     \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/likert.png}
-            \caption{The stacked bar chart shows the responses of Likert Scale questions.}
+    \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
     \label{fig:likert}

### 2022-11-25 paper/section/introduction.tex
draft: review

@@
 Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
-For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
+For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space, and comparison of regions is useful when studying EHR data.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.

### 2022-11-25 paper/section/main.tex
draft: review

@@
       \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
-      \caption{An overview of our hybrid layout algorithm.
+      \caption{An overview of our hybrid layout algorithm incorporating rivers.
+See also \Cref{alg:UpdateLayout} in \Cref{sec:{Cartograms with Rivers}} for more detail.
 See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
-See also \Cref{alg:UpdateLayout} in \Cref{sec:{Cartograms with Rivers}} for more detail.
 For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
@@
       \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
-      \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Detecting Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+      \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
       \label{fig:flowchart-stalemate}
@@
 This step intends to create space for the next iteration of node translation without crossing a river.
-The procedure for translating riers is provided by \Cref{alg:TranslateRiver} in the Appendix.
+The procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
 
-When a river is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $.
+\begin{algorithm}[tbh!]
+    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+
+    \textbf{Local variables:} \\
+    $ \riverList \gets $ a list of $ \river $ representing river features \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+    $ j \gets $ the number of nodes intersecting $ \river $  \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateRiver}{$ \nodeList $}
+        \ForEach {$ \river \in \riverList $}
+        \State $ j \gets 0 $
+        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
+        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
+        \State $ j ++ $
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{j} $
+        \EndFor
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $.
 As a heuristic we also translate these nodes by $ \vectorAvg $.
@@
 In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
-Thus, we can influence how far a river, $ \river $, is translated in each iteration of the algorithm.
+Thus, we can influence how far $ \river $ is translated in each iteration of the algorithm.
 We can use this to ensure smooth transitions between iterations.
@@
 
-\bobgraph{Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
+\bobgraph{Rendering Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
 
@@
 
-\bobgraph{RIver Translation Behavior:} The behavior of rivers during the process can also be adjusted by the user: 1) Prevent river crossing: this option forbids nodes from crossing rivers; 2) Disable river translation: this option disables the translation of rivers.
+\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option forbids nodes from crossing rivers.
+Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
 Both options are useful for generating different layouts and for observing the behavior of the hybrid layout algorithm.
@@
 A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.
+The default corridor length is three times the max node size.
 
@@
 {
-\begin{figure*}[htb!]
+\begin{figure*}[tb!]
     \centering
@@
 User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
-Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.}
+Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
+See \Cref{subsec:{User Options}} for more on user options.}
     \label{fig:overview}

### 2022-11-25 paper/section/related.tex
draft: review

@@
 Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms.
-Our work enhances that of \citeauthor{tong2018cartograms} in three ways.
-We introduce multiple features (rivers) into the layout (as opposed to a single), making topological features dynamic and further improving the legibility and geographical accuracy.
-We improve the algorithm for resolving stalemates.
-The way we evaluate the cartograms is also different.
+Our work enhances that of \citeauthor{tong2018cartograms} in multiple ways.
+First we introduce multiple features (rivers) into the layout, as opposed to a single river.
+Second, we make topological features dynamic and further improve the legibility and geographical accuracy.
+By the term \textit{dynamic}, we mean that the position of the rivers are updated as part of the layout algorithm.
+In previous work they are static and merely serve as boundaries.
+Third, we improve the algorithm for resolving stalemates.
+Finally, the way we evaluate the cartograms is also different.
 \citeauthor{tong2018cartograms} count river crossings to evaluate error.

### 2022-12-01 paper/section/appendix.tex
draft: review
submitted: EuroVis 2023

--- 2022-11-08 paper/section/appendix.tex
@@
 
+\let\origappendixsection=\section
+\renewcommand\section[1]{\origappendixsection{\textbf{#1}}}
 \section{Pre-processing Shapefiles}
@@
 The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System).
-Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility issue and reduce shapefile size to improve performance.
+Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.
 
@@
 {
-    \begin{figure}[tbh!]
+\begin{figure}[h!]
     \centering
@@
 
-    \begin{figure}[tbh!]
+\begin{figure}[h!]
     \centering
@@
 % \begin{noindent}
-    \begin{algorithm}[htb!]
+    \begin{algorithm}[h!]
         \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
@@
 %\end{noindent}
-
+\vfill\eject
 \section{Procedure: DeriveParallelEdge}

### 2022-12-01 paper/section/data.tex
submitted: EuroVis 2023

@@
  {
-  \begin{figure}[tb!]
+  \begin{figure}[b!]
       \centering
@@
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
+We decided to use the CCG shapefile from 2020 at the time of writing (See \Cref{fig:ccg}), due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
 
@@
 See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result.
-We describe the above steps in more detail in \Cref{app:pre-processing}.
+We describe the one-time pre-processing steps in more detail in \Cref{app:pre-processing}.
 

### 2022-12-01 paper/section/evaluation.tex
submitted: EuroVis 2023

@@
 We conduct a preliminary user study to evaluate the effectiveness of our approach.
+We chose this type of evaluation because legibility is a human-centred characteristic.
+Many different types of statistical error metrics have been evaluated in previous work \cite{nusrat2016State}, however, our focus is more user-centric in nature.
 
@@
 
-\textbf{P3:} The participants are asked to complete 16 location tasks.
+\textbf{P3:} The participants are asked to complete 16 location tasks involving 4 CCGs.
 The response and reaction time are recorded.
-These 16 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
 
@@
 \subsection{User Study Results}
+
+{
+    \begin{figure}[b!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
+        \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
+        \label{fig:task-acc}
+    \end{figure}
+}
 
@@
 
-    {
-        \begin{figure}[tb!]
-            \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{figure/evaluation/accuracy.png}
-            \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
-            \label{fig:task-acc}
-        \end{figure}
-    }
-
 \bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to located a CCG measured for four CCGs.
@@
     {
-        \begin{figure}[tb!]
+        \begin{figure}[t!]
             \centering
@@
 \begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
-    \item 87.50\% of participants agree that including rivers is useful.
-    \item 83.33\% of participants agree that rivers increase the legibility of a cartogram.
-    \item 95.83\% of participants agree that including rivers makes cartograms easier to understand.
-    \item 62.50\% of participants agree that including rivers makes CCGs easier to locate.
-    \item 87.50\% of participants agree that including rivers adds value to the standard cartogram.
+    \item 87.5\% of participants agree that including rivers is useful.
+    \item 83.3\% of participants agree that rivers increase the legibility of a cartogram.
+    \item 95.8\% of participants agree that including rivers makes cartograms easier to understand.
+    \item 62.5\% of participants agree that including rivers makes CCGs easier to locate.
+    \item 87.5\% of participants agree that including rivers adds value to the standard cartogram.
 \end{enumerate}

### 2022-12-01 paper/section/introduction.tex
submitted: EuroVis 2023

--- 2022-11-17 paper/section/introduction.tex
@@
 Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
-Various styles of cartograms have been proposed and implemented, covering applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.
+Various styles of cartograms have been proposed and implemented for applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.
 
-Among the four types of cartograms categorized by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
-For this project, we focus on non-contiguous cartograms like Demers cartograms, because they facilitate statistical comparison between regions and they can make good use of screen space.
+Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
+For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying EHR data.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
@@
 We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geographical regions.
-Introducing rivers into the cartogram improves their legibility.
+We hypothesize that introducing rivers into the cartogram improves their legibility.
 To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.

### 2022-12-01 paper/section/main.tex
submitted: EuroVis 2023

@@
       \label{fig:river resolution}
-  \end{figure}
- }
-
- {
-  \begin{figure}[tb!]
-      \centering
-      \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
-      \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
-      \label{fig:flowchart-stalemate}
   \end{figure}
@@
 
-\bobgraph{Initialization with Rivers:} We first load and render the CCG geospatial boundaries.
+\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
 For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ of 1 pixel.
@@
 We further apply simplification by removing vertices that are too close to each other.
+The initialization procedure is a one-time process that can be saved for re-use.
 
@@
 If a node oscillates across a river, we identify this as a stalemate.
-This procedure ends when 1) no node overlap is present; 2) no nodes cross a river.
+One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river.
 We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
@@
 For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $.
-Whenever a node, $ \node $, crosses a river , we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
+Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
 We then use a heuristic to translate $ \river $ using the average vector of node intersection $ \vectorAvg = \sum_{i=1}^j \frac{\Vector{\node_i\node_{i_t}}}{j} $, where $ j $ is the number of nodes intersecting the river.
 This step intends to create space for the next iteration of node translation without crossing a river.
-The procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
+The detailed procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
 
@@
 In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
-Thus, we can influence how far $ \river $ is translated in each iteration of the algorithm.
+Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm.
 We can use this to ensure smooth transitions between iterations.
@@
         \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
+        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+        \label{fig:flowchart-stalemate}
+    \end{figure}
+   }
+
+    {
+        \begin{figure}[tb!]
+            \centering
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
@@
 The node is translated back to its previous position (iteration 2).
-A stalemate often occurs when the area is congested and the node is unable to translate to a new position without intersecting a river.}
+A stalemate can occur when the area is congested and the node is unable to translate to a new position without intersecting a river.}
             \label{fig:stalemate}
@@
 
-The algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
+The overall algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
 At this point, the node layout is optimal where no node has crossed, or crossed by a river.
@@
 
-\Cref{fig:overview} presents an overview of the implementation, including the user options. The user can adjust the following parameters:
+\Cref{fig:overview} presents an overview of the application, including the user options. The user can adjust the following parameters:
 

### 2022-12-01 paper/section/related.tex
submitted: EuroVis 2023

@@
 By the term \textit{dynamic}, we mean that the position of the rivers are updated as part of the layout algorithm.
-In previous work they are static and merely serve as boundaries.
+In previous work the river is static and merely serves as a boundary.
 Third, we improve the algorithm for resolving stalemates.
 Finally, the way we evaluate the cartograms is also different.
-\citeauthor{tong2018cartograms} count river crossings to evaluate error.
+\citeauthor{tong2018cartograms} count river crossings to evaluate error (a statistical metric).
 Here our focus is on readability, thus we include a preliminary user study.
@@
 Our work extends Demers cartograms.
-We introduce a new type of dynamic topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.
+We introduce a new type of topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.

### 2023-03-12 paper/GHRVis.tex
submitted: EuroVis 2023
draft: review 3

@@
 
-\addbibresource{GeoVis.bib}
+\addbibresource{references.bib}
 
@@
 
-\author[Q. Wang, K. Xu \& R. S. Laramee]
+% \author[Q. Wang, K. Xu \& R. S. Laramee]
+\author[Authors]
 {
-    \parbox{\textwidth}{\centering Q. Wang, K. Xu and R. S. Laramee}
+% \parbox{\textwidth}{\centering Q. Wang, K. Xu and R. S. Laramee}
+% \\
+% \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
+\parbox{\textwidth}{\centering Authors }
 \\
-    \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
+\parbox{\textwidth}{\centering Affiliation \\ \begin{center}
         \vspace{2em}
         \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
-            \captionof{figure}{The River Thames passing through a cartographic representation of the NHS Clinical Commissioning Groups in London and surrounding areas.}
+        \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
     \end{center}

### 2023-03-12 paper/section/data.tex
draft: review 3

@@
 
-Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHRa}.
+Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHR}.
 The first step is to obtain both geospatial boundaries and EHR data.
@@
 These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
-Including additional and smaller rivers remains future work.
+Including additional and smaller rivers remains future work, as they may create complicated stalemate situations for the algorithm to resolve.
 

### 2023-03-12 paper/section/evaluation.tex
draft: review 3

@@
 \bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
-The is the primary dependent variable measured by the number of correct CCGs chosen by the participants.
+Accuracy as the primary dependent variable is measured by the number of correct CCGs chosen by the participants.
 \textit{Response time:} Another dependent variable is the time taken by the participants to complete each task.
@@
 
-135 CGGs are rendered on the screen as a choropleth on the left.
+135 CCGs are rendered on the screen as a choropleth on the left.
 We then generated another view on the right using cartograms.
@@
 Both images show the three longest rivers in England, and the color is mapped to CCG population.
-The target CGG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.}
+The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.}
             \label{fig:task}

### 2023-03-12 paper/section/introduction.tex
draft: review 3

@@
 Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
-For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying EHR data.
+For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-Standard Demers cartograms are composed of uniform shape nodes.
-As such, this can reduce their legibility.
+Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
 We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geographical regions. 
-We hypothesize that introducing rivers into the cartogram improves their legibility.
+We hypothesize that introducing rivers improves the overall legibility of cartogram. 
+By \textit{legibility} we mean readability and ability to interpret the cartogram.  
+We test this by asking user-study participants to perform correspondence and location tasks.
 To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
-We then apply the algorithm to a real-world case study using Electronic Health Records (EHR) data to evaluate the result.
+We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
 We present a user study demonstrating its effectiveness.
@@
 One of the major challenges involved is how to develop a layout algorithm that handles different shapes.
-In other words, the hybrid layout algorithm is novel because it handles different types of nodes -- rectangular representing regions and polylines representing rivers.
+In other words, the hybrid layout algorithm is novel because it handles different types of elements -- rectangular representing regions and polylines representing rivers.
 Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing geographical error.

### 2023-03-12 paper/section/limitation.tex
draft: review
submitted: EuroVis 2023
draft: review 3

@@
 In this section we discuss some limitations of our work and future research directions.
+Because this is a new concept it opens the door for many future research directions.
 The first limitation is the color map we use to depict the data in our cartograms. 
@@
 We believe that developing a new algorithm with built-in constraint support can significantly reduce the time needed to generate cartograms with rivers.
+
+Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, coastlines etc.
+We also consider if increasing the length of the rivers as the size of the nodes increases would be a useful option.
+We would like to explore the cast of river-river intersections and testing out more geographies such as the U.S. and Europe which have more complex rivers.

### 2023-03-12 paper/section/main.tex
draft: review 3

@@
     \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
+
+    $ \nodeSize \gets $ the initial size of all nodes \\
     $ \nodeSizeAvg \gets $ the average size of all nodes \\
@@
         \Procedure{UpdateLayout}{}
+
+        \State $ \nodeSize \gets 1 $
         \While{$ \nodeSizeAvg < \nodeSizeMax $}
@@
 
-\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006fast} in order to remove overlap.
+\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap of nodes.
 We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
@@
 At this point, the node layout is optimal where no node has crossed, or crossed by a river.
+Every node remains on the same side of the river as its centroid.
 

### 2023-03-12 paper/section/related.tex
draft: review 3

@@
 
-\bobgraph{Definitions: }While we focus on rectangular and Dorling cartograms, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
+\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
@@
 Each cartogram design may make various types of accuracy trade-offs between dimensions.
-We provide an overview of these cartograms in Table \ref{table:accuracy}.
+We provide a comparison of these trade-offs as introduced by \citea{nusrat2016State} in Table \ref{table:accuracy}.
+In addition, we include the Demers as it is one of the focuses in our work.
 
@@
 			\hhline{~|*{4}{-}}
-			\multicolumn{1}{c|}{\textbf{Catogram Type}} &
+			\multicolumn{1}{c|}{\textbf{Cartogram Variant}} &
 			\textbf{Statistical} &
@@
 			\hline
-			Dorling & Accurate & Inaccurate & Inaccurate & No  \\
+			Dorling & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
+			\hline
+			Demers & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
 			\hline
@@
 			\hline
-			\citea{tong2018cartograms} & \citetitlea{tong2018cartograms} & Demers & England & 209 & \citeyear{tong2018cartograms} \\
+			\citea{tong2018Cartograms} & \citetitlea{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
 			\hline
@@
 \thumbnail{dorling}{
-	Dorling cartograms are non-contiguous and do not preserve geography and topology.
+	Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
 	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
-In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}.
-Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
-Here we focus on Demers cartograms as we use squares to depict regions.
+	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citea{cano2015Mosaic} in their work.
+	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
 	This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
 	This is important in our use case scenario involving EHRs.
-Demers cartograms also have the advantage of preserving a higher level of topology at the cost of geographical accuracy.
+	Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}. 
 }
@@
 \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citea{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
 \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
@@
 
-Our work extends the algorithm described by \citea{tong2018cartograms} which incorporates a static topological feature into Demers cartograms.
-Our work enhances that of \citeauthor{tong2018cartograms} in multiple ways.
+Our work extends the algorithm described by \citea{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms.
+Our work enhances that of \citeauthor{tong2018Cartograms} in multiple ways.
 First we introduce multiple features (rivers) into the layout, as opposed to a single river.
@@
 Finally, the way we evaluate the cartograms is also different.
-\citeauthor{tong2018cartograms} count river crossings to evaluate error (a statistical metric).
+\citeauthor{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
 Here our focus is on readability, thus we include a preliminary user study.

### 2023-03-23 paper/section/evaluation.tex
draft: review 3

@@
 Both images show the three longest rivers in England, and the color is mapped to CCG population.
-The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.}
+The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.
+In this figure, $ \nodeErrCrossing $ is reduced from 9 to 0.
+}
             \label{fig:task}

### 2023-03-23 paper/section/introduction.tex
draft: review 3

@@
 
-Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography}.
+Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography,inoue2011New}.
 Various styles of cartograms have been proposed and implemented for applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.

### 2023-03-23 paper/section/main.tex
draft: review 3

@@
 
-The overall algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
-At this point, the node layout is optimal where no node has crossed, or crossed by a river.
+\new{There are two conditions that can terminate the algorithm.
+The first condition is when the average node size $ \nodeSizeAvg $ reaches the max node size, $ \nodeSizeMax $, set by the user.
+The second condition is when the displacement error, $ \nodeErrDisplacement $, reaches the max displacement error, $ \nodeErrDisplacementMax $ , allowed by the user.}
+
+When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river \new{(denoted by $ \nodeErrCrossing $ = 0)}.
 Every node remains on the same side of the river as its centroid.
@@
 
+\bobgraph{Error Metrics:} \new{The displacement error metric is used to evaluate the accuracy of the layout. The user can adjust the error threshold, which is used to terminate the algorithm.}
+
 {
@@
     \caption{A screenshot of the software.
-User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
+User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers.
 Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
-See \Cref{subsec:{User Options}} for more on user options.}
+See \Cref{subsec:{User Options}} for more on user options.
+\new{In this figure, $ \nodeErrCrossing $ is reduced from 13 to 0, and $ \nodeErrDisplacement $ for X and Y axes are 59.5 and 56.56, respectively.
+}}
     \label{fig:overview}

### 2023-03-23 paper/section/related.tex
draft: review 3

@@
 The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret.
-Our work extends Demers cartograms.
-We introduce a new type of topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.
+\new{\citea{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with zero overlapping nodes.}
+
+In this paper, we introduce a new type of topological feature, a river, \new{as a constraint to compute the final layout}, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.

### 2023-03-23 paper/section/main.tex
draft: review 3

@@
     $ \nodeSizeMax \gets $ the maximum size of a node \\
+    $ \nodeErrDisplacement \gets $ the current displacement error \\
+    $ \nodeErrDisplacementMax \gets $ the maximum displacement error allowed \\
     $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
@@
         \State $ \nodeSize \gets 1 $
-        \While{$ \nodeSizeAvg < \nodeSizeMax $}
+        \While{$ \nodeSizeAvg < \nodeSizeMax $ and $ \nodeErrDisplacement < \nodeErrDisplacementMax $}
         \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}

### 2023-04-12 paper/GHRVis.tex
draft: review 3

@@
 
-\title{Cartograms with Rivers}
+\title{Demers Cartogram with Rivers}
 
@@
 We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
-The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and geographical accuracy of the cartogram.
+The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and accuracy of the cartogram.
 \end{abstract}

### 2023-04-12 paper/section/data.tex
draft: review 3

--- 2022-12-01 paper/section/data.tex
@@
 
-Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved \cite{wang2021EHRa}.
+Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
 The first step is to obtain both geospatial boundaries and EHR data.
@@
 These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
-Including additional and smaller rivers remains future work.
+While including smaller rivers is technically feasible it may not increase the legibility of the cartogram.
 

### 2023-04-12 paper/section/evaluation.tex
draft: review 3

@@
 
-We conduct a preliminary user study to evaluate the effectiveness of our approach.
+We conduct a user study to evaluate the effectiveness of our approach.
 We chose this type of evaluation because legibility is a human-centred characteristic.
@@
 \bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
-Accuracy as the primary dependent variable is measured by the number of correct CCGs chosen by the participants.
+Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants.
 \textit{Response time:} Another dependent variable is the time taken by the participants to complete each task.
@@
 The left shows the choropleth map, and the right shows the corresponding cartogram.
-Both images show the three longest rivers in England, and the color is mapped to CCG population.
+Both images show the three longest rivers in England, and color is mapped to CCG population.
 The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.
-In this figure, $ \nodeErrCrossing $ is reduced from 9 to 0.
+In this figure, $ \nodeTopologicalError $ is reduced from 9 to 0.
 }

### 2023-04-12 paper/section/introduction.tex
draft: review 3

@@
 
-Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
+Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
 For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
+Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
+Also comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
 Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
-We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geographical regions. 
-We hypothesize that introducing rivers improves the overall legibility of cartogram. 
+We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions. 
+We hypothesize that introducing rivers improves the overall legibility of a cartogram. 
 By \textit{legibility} we mean readability and ability to interpret the cartogram.  
 We test this by asking user-study participants to perform correspondence and location tasks.
-To reduce geographical errors and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
+To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
 We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
@@
 One of the major challenges involved is how to develop a layout algorithm that handles different shapes.
-In other words, the hybrid layout algorithm is novel because it handles different types of elements -- rectangular representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing geographical error.
+In other words, the hybrid layout algorithm is novel because it handles different types of elements -- square representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.

### 2023-04-12 paper/section/limitation.tex
draft: review 3

@@
 We also consider if increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the cast of river-river intersections and testing out more geographies such as the U.S. and Europe which have more complex rivers.
+We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe which have more complex rivers.
+We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
+Furture work also includes adding rivers to dorling-style cartograms.

### 2023-04-12 paper/section/main.tex
draft: review 3

--- 2023-03-12 paper/section/main.tex
@@
-\section{Cartograms with Rivers}
+\section{Demers Cartogram with Rivers}
  {
@@
       \caption{An overview of our hybrid layout algorithm incorporating rivers.
-See also \Cref{alg:UpdateLayout} in \Cref{sec:{Cartograms with Rivers}} for more detail.
+See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail.
 See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
@@
 \begin{algorithm}[tb!]
-    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Cartograms with Rivers}}.}\label{alg:UpdateLayout}
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Demers Cartogram with Rivers}}.}\label{alg:UpdateLayout}
     \textbf{Global variables:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs with the following properties: \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
     \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
@@
     $ \nodeSize \gets $ the initial size of all nodes \\
-    $ \nodeSizeAvg \gets $ the average size of all nodes \\
-    $ \nodeSizeMax \gets $ the maximum size of a node \\
+    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
+    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
     $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
@@
         \State $ \nodeSize \gets 1 $
-        \While{$ \nodeSizeAvg < \nodeSizeMax $}
+        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
         \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
@@
 \bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ of 1 pixel.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel.
 We then load the river shapefiles and render the rivers.
@@
 
-\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap of nodes.
+\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
 We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
-Initialized with a pixel size of unity, we gradually increase the node size by one unit (proportional to its data) at a time to ensure smooth transitions.
+Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions.
 An increase in $ \nodeSize $ can cause the nodes to overlap.
@@
 One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
+We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ nodeCartographicError $, a measure set by the user, reaches its maximum value $ nodeCartographicErrorMax $.
 The gradual size increase process provides stability to the layout and helps minimize geographical error.
@@
     \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
 
     \textbf{Output:} \\
-    $ j \gets $ the number of river crossing nodes in the input \\
+    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\
 
@@
         \Procedure{TranslateNode}{$ \nodeList $}
-        \State $ j \gets 0 $
+        \State $ \nodeTopologicalError \gets 0 $
         \ForEach {$ \node \in \nodeList $}
@@
         \State $ \nodeCross ++ $
-        \State $ j ++ $
+        \State $ \nodeTopologicalError ++ $
 
@@
         \EndFor
-        \State \Return{$ j $}
+        \State \Return{$ \nodeTopologicalError $}
         \EndProcedure
@@
 Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
-We then use a heuristic to translate $ \river $ using the average vector of node intersection $ \vectorAvg = \sum_{i=1}^j \frac{\Vector{\node_i\node_{i_t}}}{j} $, where $ j $ is the number of nodes intersecting the river.
+We then use a heuristic to translate $ \river $ using the average vector of node intersection $$ \vectorAvg = \sum_{i=1}^{\nodeTopologicalError} \frac{\Vector{\node_i\node_{i_t}}}{\nodeTopologicalError} $$, where $ \nodeTopologicalError $ is the number of nodes intersecting the river.
 This step intends to create space for the next iteration of node translation without crossing a river.
@@
     \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
 
@@
     \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-    $ j \gets $ the number of nodes intersecting $ \river $  \\
+    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\
 
@@
         \ForEach {$ \river \in \riverList $}
-        \State $ j \gets 0 $
+        \State $ \nodeTopologicalError \gets 0 $
         \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
@@
         \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-        \State $ j ++ $
+        \State $ \nodeTopologicalError ++ $
         \EndIf
@@
 
-        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{j} $
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeTopologicalError} $
         \EndFor
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
-            \caption{A stalemate situation is when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
+            \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
 The node is translated back to its previous position (iteration 2).
@@
 
-The overall algorithm terminates when the average node size $ \nodeSizeAvg $ reaches the max node size $ \nodeSizeMax $.
-At this point, the node layout is optimal where no node has crossed, or crossed by a river.
+The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
+We adopt the maximum cartographic error from \citea{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in L} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
+$$
+
+where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, it is a normalized value that we express as a percentage.
+Because the algorithm precesses node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
+We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers.
+We express $ \nodeTopologicalError $ in the normalized range $ \nodeTopologicalError \in [0,1] $, where $ \nodeTopologicalErrorMax $ is the case where all nodes cross a river.
+
+When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river (denoted by $ \nodeTopologicalError $ = 0).
 Every node remains on the same side of the river as its centroid.
@@
 
-\bobgraph{Max Node Size:} The user can adjust the max node size, which is used to terminate the algorithm.
+\bobgraph{Maximum Cartographic Error:} The user can adjust the maximum cartographic error, $ \nodeCartographicErrorMax $, which is used to terminate the algorithm.
 
@@
 
-\bobgraph{River Thickess and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
+\bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
 Similarity, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
@@
     \caption{A screenshot of the software.
-User options are provided to adjust the size, color mapping, visibility of nodes and rivers.
+User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers.
 Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
-See \Cref{subsec:{User Options}} for more on user options.}
+See \Cref{subsec:{User Options}} for more on user options.
+\new{In this figure, $ \nodeTopologicalError $ is reduced from 13 to 0, and $ \nodeCartographicError $ for X and Y axes are 59.5 and 56.56, respectively.
+}}
     \label{fig:overview}

### 2023-04-12 paper/section/related.tex
draft: review 3

--- 2023-03-12 paper/section/related.tex
@@
 We provide a comparison of these trade-offs as introduced by \citea{nusrat2016State} in Table \ref{table:accuracy}.
-In addition, we include the Demers as it is one of the focuses in our work.
+In addition, we include the Demers as it is the focus of our work.
 
@@
 			\hline
+			\citea{nickel2022Multicriteriaa} & \citetitlea{nickel2022Multicriteriaa} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteriaa} \\
+			\hline
 
@@
 	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
-	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citea{cano2015Mosaic} in their work.
+	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citea{cano2015Mosaic} in their related work section.
 	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
@@
 Second, we make topological features dynamic and further improve the legibility and geographical accuracy.
-By the term \textit{dynamic}, we mean that the position of the rivers are updated as part of the layout algorithm.
+By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm.
 In previous work the river is static and merely serves as a boundary.
@@
 \citeauthor{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
-Here our focus is on readability, thus we include a preliminary user study.
+Here our focus is on readability, thus we include a user study.
 
@@
 The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret.
-Our work extends Demers cartograms.
-We introduce a new type of topological feature, a river, aiming to improve the interpretation, readability, and geographical accuracy of this class of cartograms.
+\citea{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.
+
+In this paper, we introduce a new type of topological feature, a river, as a constraint to compute the final layout, aiming to improve the interpretation, readability, and accuracy of this class of cartograms.

### 2023-04-20 paper/section/main.tex
draft: e_t percentage

@@
 One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ nodeCartographicError $, a measure set by the user, reaches its maximum value $ nodeCartographicErrorMax $.
-The gradual size increase process provides stability to the layout and helps minimize geographical error.
+We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
+The gradual size increase process provides stability to the layout.
 
@@
     \includegraphics[width=\textwidth,keepaspectratio]{figure/UI.png}
-    \caption{A screenshot of the software.
+    \caption{A screenshot of the user interface.
 User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers.
@@
 See \Cref{subsec:{User Options}} for more on user options.
-\new{In this figure, $ \nodeTopologicalError $ is reduced from 13 to 0, and $ \nodeCartographicError $ for X and Y axes are 59.5 and 56.56, respectively.
-}}
+In this figure, an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
     \label{fig:overview}

### 2023-05-03 paper/section/main.tex
draft: review

@@
 We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
-The gradual size increase process provides stability to the layout.
+The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.
 
@@
 The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
-We adopt the maximum cartographic error from \citea{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in L} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
+We adopt the maximum cartographic error from \citea{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
 $$
 
-where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, it is a normalized value that we express as a percentage.
-Because the algorithm precesses node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
+where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
+Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
 We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers.
@@
 
-When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river (denoted by $ \nodeTopologicalError $ = 0).
+When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river (denoted by $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
 Every node remains on the same side of the river as its centroid.
@@
 
-\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option forbids nodes from crossing rivers.
+\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers.
 Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
@@
 See \Cref{subsec:{User Options}} for more on user options.
-In this figure, an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
+In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
     \label{fig:overview}

### 2023-05-03 paper/section/related.tex
draft: review

@@
 			\hline
-			\citea{nickel2022Multicriteriaa} & \citetitlea{nickel2022Multicriteriaa} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteriaa} \\
+			\citea{nickel2022Multicriteria} & \citetitlea{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
 			\hline

### 2023-05-11 paper/section/appendix.tex
draft: VMV 10 page

--- 2022-11-17 paper/section/appendix.tex
@@
 
+\let\origappendixsection=\section
+\renewcommand\section[1]{\origappendixsection{\textbf{#1}}}
+
+\section{Obtaining Shapefile}
+\label{obtaining-shapefile}
+\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
+    relation(2263653);>>;
+    // River Great Ouse: 2798097
+    // River Trent: 2863468
+    out skel;
+\end{lstlisting}
 \section{Pre-processing Shapefiles}
@@
 {
-\begin{figure}[h!]
+\begin{figure}[H]
     \centering
@@
 
-\begin{figure}[h!]
+\begin{figure}[H]
     \centering
@@
         \renewcommand{\arraystretch}{1.5}
-        \begin{table}[!tb]
+        \begin{table}[H]
             \centering
@@
 
+\section{Procedure: TranslateNode}
+
+\begin{algorithm}[H]
+    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
+
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
+
+    \textbf{Output:} \\
+    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\
+
+    \textbf{Global variables:} \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateNode}{$ \nodeList $}
+        \State $ \nodeTopologicalError \gets 0 $
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+
+        \State $ \node(x,y) \gets \nodeFNOR(x,y) $
+
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
+        \State $ \nodeCross ++ $
+        \State $ \nodeTopologicalError ++ $
+
+        \If{$ \nodeCross < \stalemateMax $}
+        \State \hskip-2em \Comment{Translate back to previous position}
+        \State $ \node(x,y) \gets \nodePrevious(x,y) $
+
+        \Else
+
+        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
+        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
+
+        \EndIf
+
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+        \State \Return{$ \nodeTopologicalError $}
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Procedure: UpdateLayout}
+
+\begin{algorithm}[H]
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Demers Cartogram with Rivers}}.}\label{alg:UpdateLayout}
+    \textbf{Global variables:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
+    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
+
+    $ \nodeSize \gets $ the initial size of all nodes \\
+    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
+    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{UpdateLayout}{}
+
+        \State $ \nodeSize \gets 1 $
+        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
+        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
+        \While{$ \nodeListCross > 0 $}
+
+        \State $ \nodeListCross \gets 0 $
+        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
+        \State \Call{TranslateRiver}{$ \nodeList $}
+        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}
+
+        \EndWhile
+
+        \State $ \nodeSize ++ $
+
+        \EndWhile
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
 \section{Procedure: TranslateRiver}
 
-\begin{algorithm}[h!]
+\begin{algorithm}[H]
     \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
     \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing CCGs \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
 
@@
 
+\section{Procedure: ProcessStalemate}
+
+% \begin{noindent}
+
+    \begin{algorithm}[H]
+        \caption{Procedure to derive a corridor to resolve stalemate.
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}
+
+        \textbf{Input:} \\
+        $ \node \gets $ the node used to derive the corridor \\
+
+        \textbf{Global variables:} \\
+        $ \CorridorLength \gets $ the length of a corridor \\
+        $ \CorridorWidth \gets $ the width of a corridor \\
+
+        \textbf{Local variables:} \\
+        $ \Corridor \gets $ the corridor \\
+        $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
+        $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
+        $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{ProcessStalemate}{$ \node $}
+                \State $ \node(x,y) \gets \nodePrevious(x,y) $
+
+                \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}
+
+                \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}
+
+                \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}
+
+                \State $ \Corridor \gets
+                    \begin{bmatrix}
+                        \EdgeParallelA.start &
+                        \EdgeParallelA.end \\
+
+                        \EdgeParallelB.start &
+                        \EdgeParallelB.end \\
+                    \end{bmatrix} $
+
+                \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}
+
+                    \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $
+
+                    \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $
+
+                \EndFor
+            \EndProcedure
+        \end{algorithmic}
+    \end{algorithm}
+
+    %\end{noindent}
 \section{Procedure: TestIntersection}
 % \begin{noindent}
-    \begin{algorithm}[h!]
+    \begin{algorithm}[H]
         \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
@@
 
-    \begin{algorithm}[h!]
+    \begin{algorithm}[H]
         \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
@@
 
-\begin{algorithm}[h!]
+\begin{algorithm}[H]
     \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

### 2023-05-11 paper/section/data.tex
draft: review
draft: VMV 10 page

@@
 
- {
-  \begin{figure}[b!]
-      \centering
-      \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-      \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
-      \label{fig:ccg}
-  \end{figure}
- }
+%  {
+%   \begin{figure}[b!]
+%       \centering
+%       \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
+%       \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
+%       \label{fig:ccg}
+%   \end{figure}
+%  }
 
@@
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing (See \Cref{fig:ccg}), due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
 
@@
 While including smaller rivers is technically feasible it may not increase the legibility of the cartogram.
+This is an open question for future work.
 
@@
 River Thames, on OpenStreetMap.
-The relation ID is used to construct a query (See Listing~\ref{overpass}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
-
-\begin{lstlisting}[float=tp,caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
-    relation(2263653);>>;
-    // River Great Ouse: 2798097
-    // River Trent: 2863468
-    out skel;
-\end{lstlisting}
+The relation ID is used to construct a query (See \cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
 

### 2023-05-11 paper/section/main.tex
draft: VMV 10 page

@@
 See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail.
-See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
+% See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
        For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
@@
  }
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:{Demers Cartogram with Rivers}}.}\label{alg:UpdateLayout}
-    \textbf{Global variables:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
-    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
-
-    $ \nodeSize \gets $ the initial size of all nodes \\
-    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
-    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
-    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
-
-    \textbf{Local variables:} \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{UpdateLayout}{}
-
-        \State $ \nodeSize \gets 1 $
-        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
-        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
-        \While{$ \nodeListCross > 0 $}
-
-        \State $ \nodeListCross \gets 0 $
-        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
-        \State \Call{TranslateRiver}{$ \nodeList $}
-        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}
-
-        \EndWhile
-
-        \State $ \nodeSize ++ $
-
-        \EndWhile
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
 
@@
 
-\begin{algorithm}[tb!]
-    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
-
-    \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing regions \\
-
-    \textbf{Output:} \\
-    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\
-
-    \textbf{Global variables:} \\
-    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
-
-    \textbf{Local variables:} \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TranslateNode}{$ \nodeList $}
-        \State $ \nodeTopologicalError \gets 0 $
-        \ForEach {$ \node \in \nodeList $}
-
-        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
-
-        \State $ \node(x,y) \gets \nodeFNOR(x,y) $
-
-        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
-        \State $ \nodeCross ++ $
-        \State $ \nodeTopologicalError ++ $
-
-        \If{$ \nodeCross < \stalemateMax $}
-        \State \hskip-2em \Comment{Translate back to previous position}
-        \State $ \node(x,y) \gets \nodePrevious(x,y) $
-
-        \Else
-
-        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
-        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
-
-        \EndIf
-
-        \EndIf
-
-        \EndIf
-
-        \EndFor
-        \State \Return{$ \nodeTopologicalError $}
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
 \subsection{River Intersection Testing}
@@
 
-\begin{algorithm}[tbh!]
-    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
-    \textbf{Input:} \\
-    $ \nodeList \gets $ a list of $ \node $ representing regions \\
-
-    \textbf{Local variables:} \\
-    $ \riverList \gets $ a list of $ \river $ representing river features \\
-    $ \node \gets $ a node is an object with the following properties: \\
-    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
-    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
-    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
-    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\
-
-    \begin{algorithmic}[1]
-        \Procedure{TranslateRiver}{$ \nodeList $}
-        \ForEach {$ \river \in \riverList $}
-        \State $ \nodeTopologicalError \gets 0 $
-        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
-        \ForEach {$ \node \in \nodeList $}
-
-        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
-        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
-        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
-        \State $ \nodeTopologicalError ++ $
-        \EndIf
-
-        \EndIf
-
-        \EndFor
-
-        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeTopologicalError} $
-        \EndFor
-
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
 When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $.
@@
 
-{
-    \begin{figure}[tb!]
-        \centering
-        \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
-        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
-        \label{fig:flowchart-stalemate}
-    \end{figure}
-   }
+% {
+%     \begin{figure}[tb!]
+%         \centering
+%         \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
+%         \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+%         \label{fig:flowchart-stalemate}
+%     \end{figure}
+%    }
 
@@
     }
-
-% \begin{noindent}
-
-\begin{algorithm}[tb!]
-    \caption{Procedure to derive a corridor to translate enclosed nodes.
-We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}
-
-    \textbf{Input:} \\
-    $ \node \gets $ the node used to derive the corridor \\
-
-    \textbf{Global variables:} \\
-    $ \CorridorLength \gets $ the length of a corridor \\
-    $ \CorridorWidth \gets $ the width of a corridor \\
-
-    \textbf{Local variables:} \\
-    $ \Corridor \gets $ the corridor \\
-    $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
-    $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
-    $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
-
-    \begin{algorithmic}[1]
-        \Procedure{ProcessStalemate}{$ \node $}
-            \State $ \node(x,y) \gets \nodePrevious(x,y) $
-
-            \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}
-
-            \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}
-
-            \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}
-
-            \State $ \Corridor \gets
-                \begin{bmatrix}
-                    \EdgeParallelA.start &
-                    \EdgeParallelA.end \\
-
-                    \EdgeParallelB.start &
-                    \EdgeParallelB.end \\
-                \end{bmatrix} $
-
-            \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}
-
-                \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $
-
-                \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $
-
-            \EndFor
-        \EndProcedure
-    \end{algorithmic}
-\end{algorithm}
-
-%\end{noindent}
 
@@
     \centering
-    \includegraphics[width=\textwidth,keepaspectratio]{figure/UI.png}
+    \includegraphics[width=0.7\textwidth,keepaspectratio]{figure/UI.png}
     \caption{A screenshot of the user interface.

### 2023-05-11 paper/GHRVis.tex
draft: VMV 10 page
fix: grammars

@@
 \documentclass{egpubl}
-\usepackage{eurovis2023}
+\usepackage{VMV2023}
 
@@
 % \SpecialIssueSubmission    % uncomment for submission to , special issue
-\SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
+% \SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
 %                          % EuroVis, SGP, Rendering, PG
 % --- for  EG Workshop Proceedings
-% \WsSubmission      % uncomment for submission to EG Workshop
+\WsSubmission      % uncomment for submission to EG Workshop
 % \WsPaper           % uncomment for final version of EG Workshop contribution
@@
 A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
-One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing the legibility, readability, and geographical accuracy.
+One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing legibility, readability, and geographical accuracy.
 To address this, we propose a novel hybrid cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
@@
 We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
-The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram, while improving the legibility, readability, and accuracy of the cartogram.
+The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram while improving the legibility, readability, and accuracy of the cartogram.
 \end{abstract}

### 2023-05-11 paper/section/data.tex
fix: grammars

@@
 
-Obtaining the heterogenous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
+Obtaining heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
 The first step is to obtain both geospatial boundaries and EHR data.
-The second step is to pre-process the EHR data to remove empty and erroneous values.
+The second step is to preprocess the EHR data to remove empty and erroneous values.
 The final step is to transform the data into a format that is suitable for cartograms.
@@
 \bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
-The number of CCGs changes over time due to NHS re-organization.
+The number of CCGs changes over time due to NHS reorganization.
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG re-organizations that took place in 2021 and 2022.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG reorganizations that took place in 2021 and 2022.
 
-\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent, and River Great Ouse in England.
-These rivers were chosen as they are well-known rivers and pass through regions with dense populations, and provide informative geographical and topological cues.
-While including smaller rivers is technically feasible it may not increase the legibility of the cartogram.
+\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England.
+These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues.
+Although including smaller rivers is technically feasible, it may not increase the legibility of the cartogram.
 This is an open question for future work.
@@
 
-After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections, and convert them into GeoJSON files.
-Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON \cite{TopoJSON} file.
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust the projections and convert them into GeoJSON files.
+Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON file \cite{TopoJSON}.
 TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
-See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the pre-processing result.
-We describe the one-time pre-processing steps in more detail in \Cref{app:pre-processing}.
+See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the preprocessing result.
+We describe the one-time preprocessing steps in more detail in \Cref{app:pre-processing}.
 

### 2023-05-11 paper/section/evaluation.tex
draft: e_t percentage
draft: review
fix: grammars

@@
 
-\textbf{H1.1:} The introduction of rivers can increase target location accuracy.
+\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.
 
-\textbf{H1.2:} The introduction of rivers can reduce the time needed to locate a target region in a cartogram.
+\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.
 
-We test these hypotheses using location based tasks.
+We test these hypotheses using location-based tasks.
 Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
@@
 
-We discuss the variables of our user study in this section.
+In this section, we discuss the variables of our user study.
 
 \bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram.
-We set this parameter to either render rivers or not.
+We set this parameter to render rivers or not.
 
@@
 Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants.
-\textit{Response time:} Another dependent variable is the time taken by the participants to complete each task.
+\textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.
 
 \bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
-\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG that they need to find.
-The target CCG is shown in the form of a non-stop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
+\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find.
+The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
 
@@
 We then generated another view on the right using cartograms.
-The color in both visual designs is mapped to the dataset.
+The color of both visual designs is mapped to the dataset.
 See \Cref{fig:task} for an example.
@@
 Both images show the three longest rivers in England, and color is mapped to CCG population.
-The target CCG is blinking on the choropleth (shown in black), participants are asked to identify this CCG on the cartogram.
-In this figure, $ \nodeTopologicalError $ is reduced from 9 to 0.
+The target CCG is blinking on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
+In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated.
 }
@@
 \textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
-These tasks also serve as a quality test to see if the users are taking the study seriously.
+These tasks also serve as a quality test to see whether users are taking the study seriously.
 A demonstration of three sample tasks is also included in the instructional video.
 
-\textbf{P3:} The participants are asked to complete 16 location tasks involving 4 CCGs.
-The response and reaction time are recorded.
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 CCGs.
+Response and reaction time are recorded.
 These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
@@
 
-The materials for P1-P4 are available on the Open Science Foundation's website at \URL{https://osf.io/q39w7}.
+The materials for P1-P4 are available on the Open Science Foundation website at \URL{https://osf.io/q39w7}.
 
@@
 
-In this section we analyze the results of our preliminary user study.
+In this section, we analyze the results of our preliminary user study.
 We recruited \pCount participants to perform 16 location tasks each, in total we collected 384 responses, 24 responses that took over 60 seconds are excluded from the analysis.
-We hypothesize that participants taking longer than 60 seconds to complete a task were distracted from the study.
+We hypothesize that participants who took longer than 60 seconds to complete a task were distracted from the study.
 
 \bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
-The chart shows that the introduction of rivers has improved the accuracy for two CCGs.
+The chart shows that the introduction of rivers has improved the accuracy of two CCGs.
 The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002.
@@
 
-\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to located a CCG measured for four CCGs.
+\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs.
 The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.

### 2023-05-11 paper/section/introduction.tex
fix: grammars

@@
 
-Among the four types of cartograms categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
+Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
 For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
 Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
-Also comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
+Also, comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
@@
 We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
-We present a user study demonstrating its effectiveness.
+We present a user study that demonstrates its effectiveness.
 
@@
 The results of the user study indicate that rivers can improve the legibility of cartograms.
-One of the major challenges involved is how to develop a layout algorithm that handles different shapes.
-In other words, the hybrid layout algorithm is novel because it handles different types of elements -- square representing regions and polylines representing rivers.
+One of the major challenges involved is developing a layout algorithm that handles different shapes.
+In other words, the hybrid layout algorithm is novel because it handles different types of elements: square representing regions and polylines representing rivers.
 Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.

### 2023-05-11 paper/section/limitation.tex
draft: review
draft: VMV 10 page
fix: grammars

--- 2022-11-17 paper/section/limitation.tex
@@
 
-In this section we discuss some limitations of our work and future research directions.
-The first limitation is the color map we use to depict the data in our cartograms.
+In this section, we discuss some limitations of our work and future research directions.
+Because this is a new concept, it opens the door for many future research directions.
+The first limitation is the color map that we use to depict the data in our cartograms. 
 We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
@@
         \centering
-        \includegraphics[width=\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
+        \includegraphics[width=0.7\columnwidth,keepaspectratio]{figure/limitations/extreme.png}
         \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
@@
 Another limitation is the overlap removal algorithm (FNOR) we use.
-We believe that developing a new algorithm with built-in constraint support can significantly reduce the time needed to generate cartograms with rivers.
+We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
+Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
+
+Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, coastlines, etc.
+We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
+We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
+We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
+Furture work also includes adding rivers to dorling-style cartograms.

### 2023-05-11 paper/section/main.tex
fix: grammars

@@
 We further apply simplification by removing vertices that are too close to each other.
-The initialization procedure is a one-time process that can be saved for re-use.
+The initialization procedure is a one-time process that can be saved for reuse.
 
 \bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
-We chose FNOR over other node overlap removal algorithms because FNOR is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}.
+We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}.
 Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions.
@@
 
-The logic for translating a node's position is detailed in \Cref{alg:TranslateNode}.
+The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}.
 When a node's position changes, we test if the node's trajectory intersects any segment of a river.
@@
 
-When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where the nodes are crossed by a translated river, denoted $ \RxN $.
-As a heuristic we also translate these nodes by $ \vectorAvg $.
-The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated in order to create space for nodes that are too crowded together.
+When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $.
+As a heuristic, we also translate these nodes by $ \vectorAvg $.
+The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together.
 In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
@@
 As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}.
-If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate the congestion.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion.
 A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
@@
 The node is translated back to its previous position (iteration 2).
-A stalemate can occur when the area is congested and the node is unable to translate to a new position without intersecting a river.}
+A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
             \label{fig:stalemate}
@@
 
-When the algorithm terminates, the node layout is considered optimal where no node has crossed, or crossed by a river (denoted by $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
+When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
 Every node remains on the same side of the river as its centroid.
@@
 
-\Cref{fig:overview} presents an overview of the application, including the user options. The user can adjust the following parameters:
+\Cref{fig:overview} presents an overview of the application, including user options. The user can adjust the following parameters:
 
@@
 Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
-Both options are useful for generating different layouts and for observing the behavior of the hybrid layout algorithm.
+Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.
 
-\bobgraph{Corridor Length:} The user can define the length of a corridor, which is used to resolve stalemate situations.
+\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations.
 A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.

### 2023-05-11 paper/section/related.tex
fix: grammars

@@
 
-\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving that maintain connectivity with their adjacent neighbors but are also subject to distortion in shape.
-Non-contiguous cartograms sacrifice the topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
+\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
+Non-contiguous cartograms sacrifice topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
 
@@
 \citea{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
-\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US.
+\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US.
 See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
@@
 Our work enhances that of \citeauthor{tong2018Cartograms} in multiple ways.
-First we introduce multiple features (rivers) into the layout, as opposed to a single river.
-Second, we make topological features dynamic and further improve the legibility and geographical accuracy.
+First, we introduce multiple features (rivers) into the layout, as opposed to a single river.
+Second, we make topological features dynamic and further improve legibility and geographical accuracy.
 By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm.
-In previous work the river is static and merely serves as a boundary.
-Third, we improve the algorithm for resolving stalemates.
+In previous work, the river is static and merely serves as a boundary.
+Third, we improve the algorithm to resolve stalemates.
 Finally, the way we evaluate the cartograms is also different.
 \citeauthor{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
-Here our focus is on readability, thus we include a user study.
+Here our focus is on readability, and thus we include a user study.
 
@@
 
-In this paper, we introduce a new type of topological feature, a river, as a constraint to compute the final layout, aiming to improve the interpretation, readability, and accuracy of this class of cartograms.
+In this paper, we introduce a new type of topological feature, a river, as a constraint to compute the final layout, with the aim of improving the interpretation, readability, and accuracy of this class of cartograms.

### 2023-05-12 paper/GHRVis.tex
fix: grammar with writefull

@@
 \begin{abstract}
-    Cartograms are representations of geographical and abstract data based on a value-by-area mapping technique.
-A relative of the Dorling cartogram, the Demers cartogram uses squares instead of circles to represent regions, enabling the comparison of regions in a more intuitive way with a higher screen space utilization.
-One of the disadvantages of the Dorling cartogram and its variants is that regions may be displaced from their original positions, reducing legibility, readability, and geographical accuracy.
-To address this, we propose a novel hybrid cartogram layout algorithm that integrates dynamic topological features such as rivers into Demers cartograms.
-Rivers affect both the layout and appearance of cartograms.
-We evaluate the proposed hybrid layout algorithm through a user study with an application to an Electronic Health Records (EHR) dataset.
-The results demonstrate that the proposed layout algorithm is able to preserve some of the original properties of the cartogram while improving the legibility, readability, and accuracy of the cartogram.
+    Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique.
+A relative of the Dorling cartogram, the Demers cartogram, utilizing squares instead of circles to represent regions.
+This alternative approach allows for a more intuitive comparison of regions, utilizing screen space more efficiently.
+However, a drawback of the Dorling cartogram and its variants lies in the potential displacement of regions from their original positions, ultimately compromising legibility, readability, and accuracy.
+To tackle this limitation, we propose a novel hybrid cartogram layout algorithm that incorporates dynamic topological elements, such as rivers, into Demers cartograms.
+The presence of rivers significantly impacts both the layout and visual appearance of the cartograms.
+Through a user study conducted on an Electronic Health Records (EHR) dataset, we evaluate the efficacy of the proposed hybrid layout algorithm.
+The obtained results illustrate that this approach successfully retains key aspects of the original cartogram while enhancing legibility, readability, and overall accuracy.
 \end{abstract}

### 2023-05-12 paper/section/appendix.tex
fix: grammar with writefull

@@
 The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System).
-Here we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.
+Here, we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.
 
@@
 
-We then use QGIS to unify the CRS, and export both layers in the GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.
+We then use QGIS to unify the CRS, and export both layers in GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.
 
@@
 
-\Cref{table:pre-processing_result} shows the pre-processing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+\Cref{table:pre-processing_result} shows the preprocessing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
 
@@
     \textbf{Output:} \\
-    $ \nodeTopologicalError \gets $ the number of river crossing nodes in the input \\
+    $ \nodeTopologicalError \gets $ the number of nodes crossing the river in the input \\
 

### 2023-05-12 paper/section/data.tex
fix: grammar with writefull

@@
 The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
-We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of public EHR data published based on the latest CCG reorganizations that took place in 2021 and 2022.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of published public EHR data based on the latest CCG reorganizations that took place in 2021 and 2022.
 

### 2023-05-12 paper/section/introduction.tex
fix: grammar with writefull

@@
 
-Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off between types of accuracy is made (See \Cref{table:accuracy}).
+Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
 For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
 Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
-Also, comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
+Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.

### 2023-05-12 paper/section/limitation.tex
fix: grammar with writefull

@@
 
-Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, coastlines, etc.
+Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.

### 2023-05-12 paper/section/main.tex
fix: grammar with writefull

@@
       \centering
-      \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
+      \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
       \caption{An overview of our hybrid layout algorithm incorporating rivers.
@@
 \bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
-Similarity, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
+Similarly, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
 

### 2023-05-12 paper/section/related.tex
fix: grammar with writefull

@@
 
-\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms are topology-preserving, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
+\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms preserve topology, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
@@
 By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm.
-In previous work, the river is static and merely serves as a boundary.
+In previous work, the river is static and serves merely as a boundary.
 Third, we improve the algorithm to resolve stalemates.

### 2023-05-13 paper/section/evaluation.tex
fix: grammar with writefull
draft: adjust layout

@@
 Both images show the three longest rivers in England, and color is mapped to CCG population.
-The target CCG is blinking on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
+The target CCG blinks on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
 In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated.
@@
 
-In this section, we analyze the results of our preliminary user study.
-We recruited \pCount participants to perform 16 location tasks each, in total we collected 384 responses, 24 responses that took over 60 seconds are excluded from the analysis.
-We hypothesize that participants who took longer than 60 seconds to complete a task were distracted from the study.
+In this section, we delve into the analysis of the preliminary user study conducted.
+Our study involved the recruitment of \pCount participants, each tasked with completing 16 location-related tasks.
+As a result, we collected a total of 384 responses.
+However, we excluded 24 responses from our analysis as they took over 60 seconds to complete.
+We hypothesize that these prolonged response times indicate potential distractions experienced by participants during the study.
 

### 2023-07-06 paper/GHRVis.tex
draft: IV

@@
-\documentclass{egpubl}
-\usepackage{VMV2023}
+\documentclass[Afour,sagev,times]{sagej}
 
-% --- for  Annual CONFERENCE
-% \ConferenceSubmission   % uncomment for Conference submission
-% \ConferencePaper        % uncomment for (final) Conference Paper
-% \STAR                   % uncomment for STAR contribution
-% \Tutorial               % uncomment for Tutorial contribution
-% \ShortPresentation      % uncomment for (final) Short Conference Presentation
-% \Areas                  % uncomment for Areas contribution
-% \MedicalPrize           % uncomment for Medical Prize contribution
-% \Education              % uncomment for Education contribution
-% \Poster                 % uncomment for Poster contribution
-% \DC                     % uncomment for Doctoral Consortium
-%
-% --- for  CGF Journal
-% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
-% \JournalPaper         % uncomment for final version of Journal Paper
-%
-% --- for  CGF Journal: special issue
-% \SpecialIssueSubmission    % uncomment for submission to , special issue
-% \SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
-%                          % EuroVis, SGP, Rendering, PG
-% --- for  EG Workshop Proceedings
-\WsSubmission      % uncomment for submission to EG Workshop
-% \WsPaper           % uncomment for final version of EG Workshop contribution
-% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
-% \WsPaperJoint      % for joint events, for example ICAT-EGVE
-% \Expressive        % for SBIM, CAe, NPAR
-% \DigitalHeritagePaper
-% \PaperL2P          % for events EG only asks for License to Publish
+\usepackage{moreverb,url}
 
-% --- for EuroVis 
-% for full papers use \SpecialIssuePaper
-% \STAREurovis   % for EuroVis additional material 
-% \EuroVisPoster % for EuroVis additional material 
-% \EuroVisShort  % for EuroVis additional material
-
-% !! *please* don't change anything above
-% !! unless you REALLY know what you are doing
-% ------------------------------------------------------------------------
-
-\usepackage[T1]{fontenc}
-\usepackage{dfadobe}
-
-% \usepackage{cite}  % comment out for biblatex with backend=biber
-% ---------------------------
-\biberVersion
-\BibtexOrBiblatex
-\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true,url=false]{biblatex}
-%\addbibresource{egbibsample.bib}
-
-\renewbibmacro*{doi+eprint+url}{%
-    \printfield{doi}%
-    \newunit\newblock%
-    \iftoggle{bbx:eprint}{%
-        \usebibmacro{eprint}%
-    }{}%
-    \newunit\newblock%
-    \iffieldundef{doi}{%
-        \usebibmacro{url+urldate}}%
-        {}%
-    }
-
-\addbibresource{references.bib}
-
-\newcommand{\citetitlea}[1]{\citefield{#1}{title}}
-
-% ---------------------------  
-\electronicVersion
-\PrintedOrElectronic
-% for including postscript figures
-% mind: package option 'draft' will replace PS figure by a filename within a frame
-\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
-\else \usepackage[dvips]{graphicx} \fi
-
-\usepackage{egweblnk}
-% end of prologue
-
-\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}
+\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow, multicol}
 
@@
 
-\usepackage{dblfloatfix}
-
 \usepackage[table, dvipsnames]{xcolor}
 
-\usepackage{multicol}
-
-% For cover's captionof
-\usepackage[font=normalfont,labelfont=bf]{caption}
-
-\include{preamble/preamble.basic}
 \include{preamble/preamble.main}
 \include{preamble/preamble.math}
-\include{preamble/preamble.draft}
+
+\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}
+
+\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
+T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
+\setcounter{secnumdepth}{3}
+
 \begin{document}
-
-% \author[Q. Wang, K. Xu \& R. S. Laramee]
-\author[Authors]
-{
-% \parbox{\textwidth}{\centering Q. Wang, K. Xu and R. S. Laramee}
-% \\
-% \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
-\parbox{\textwidth}{\centering Authors }
-\\
-\parbox{\textwidth}{\centering Affiliation \\ \begin{center}
-        \vspace{2em}
-        \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
-        \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
-    \end{center}
-}
-}
 
@@
 
-\pagestyle{plain}
-
-\maketitle
+% \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
+%         \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
 
@@
 \end{abstract}
+
+\keywords{Cartogram, Demers Cartogram}
+
+\maketitle
 
@@
 
-% \section{Acknowledgments}
-% This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC).
-EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.
+% \bibliographystyle{mslapa}
+\urlstyle{same}
+\bibliographystyle{plainnat}
+\bibliography{references}
 
-% reset \section command for section title "References"
-\let\section=\origsection
-\printbibliography
-
-\clearpage
+\cleardoublepage
 

### 2023-07-06 paper/section/evaluation.tex
draft: IV

@@
 \newcommand{\pCount}{24 }
-\section{User-Centered Evaluation}
+\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}
 
@@
 
-\subsection{Study Hypothesis}
+\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}
 
@@
 
-\subsection{User Study Variables}
+\subsection{User Study Variables}\label{subsec:{User Study Variables}}
 
@@
 
-\subsection{User Study Design}
+\subsection{User Study Design}\label{subsec:{User Study Design}}
+
 In this section we describe the user study participants, datasets, and the experimental procedure.
@@
 \begin{itemize}
+    \setlength\itemsep{0px}
     \item Gender: 10 females and 14 males
@@
 \begin{itemize}
+    \setlength\itemsep{0px}
     \item Population
@@
 \subsubsection{Procedure}
+\label{subsec:{Procedure}}
 
@@
 The instructions are designed to help participants understand the concepts used in the tasks.
-Instructional videos are available at \URL{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
 
@@
 
-The materials for P1-P4 are available on the Open Science Foundation website at \URL{https://osf.io/q39w7}.
+The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.
 
 \subsection{User Study Results}
+\label{subsec:{User Study Results}}
 

### 2023-07-06 paper/section/introduction.tex
draft: adjust layout
draft: IV

@@
 
-Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
+Among the four types of cartogram categorized in a survey by \citet{nusrat2016State} (See \Cref{sec:RelatedWork} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
 For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
@@
 By \textit{legibility} we mean readability and ability to interpret the cartogram.  
-We test this by asking user-study participants to perform correspondence and location tasks.
+To assess this hypothesis, we designed an experimental setup where participants engaged in correspondence and location tasks as part of a user study.
 To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
@@
 \begin{itemize}
+    \setlength\itemsep{0px}
     \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,

### 2023-07-06 paper/section/main.tex
draft: IV

@@
 \section{Demers Cartogram with Rivers}
+\label{sec:main}
+
  {
@@
       \caption{An overview of our hybrid layout algorithm incorporating rivers.
-See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail.
+See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail.
 % See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
@@
 \subsection{River Intersection Testing}
+\label{subsec:{River Intersection Testing}}
 
@@
 \subsection{Translating Rivers}
+\label{subsec:{Translating Rivers}}
 
@@
 \subsection{Process Stalemates}
+\label{subsec:{Process Stalemates}}
 
@@
 \subsection{Terminating the Algorithm}
+\label{subsec:{Terminating the Algorithm}}
 
 The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
-We adopt the maximum cartographic error from \citea{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
+We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
 $$
@@
 
-\subsection{User Options}
+\subsection{User Options}\label{subsec:{UserOptions}}
 
@@
 Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
-See \Cref{subsec:{User Options}} for more on user options.
+See \Cref{subsec:{UserOptions}} for more on user options.
 In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}

### 2023-07-06 paper/section/related.tex
draft: adjust layout
draft: IV

@@
 \section{Related Work}
+\label{sec:RelatedWork}
 
@@
 
-\citea{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+\citet{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
 Each cartogram design may make various types of accuracy trade-offs between dimensions.
-We provide a comparison of these trade-offs as introduced by \citea{nusrat2016State} in Table \ref{table:accuracy}.
+We provide a comparison of these trade-offs as introduced by \citet{nusrat2016State} in \Cref{table:accuracy}.
 In addition, we include the Demers as it is the focus of our work.
+
+{
+\renewcommand{\arraystretch}{1.5}
+\begin{table*}[!tb]
+	\centering
+	\resizebox{\textwidth}{!}{
+		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
+			\hhline{~|*{4}{-}|~}
+			\multicolumn{1}{c|}{\textbf{Literature}} &
+			\textbf{Title} &
+			\textbf{Cartogram Type(s)} &
+			\textbf{Geographic Region(s)} &
+			\textbf{Number of Nodes} &
+			\multicolumn{1}{c}{\textbf{Year}} \\
+			\hline
+
+			% \citet{auber2007Geographical} & \citep{auber2007Geographical} & US & 19 & 2007 \\
+			% \hline
+			\citet{warf2008Geography} & \citep{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
+			\hline
+			\citet{sun2010Effectiveness} & \citep{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+			\hline
+			\citet{cruz2017Adapted} & \citep{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+			\hline
+			\citet{tong2018Cartograms} & \citep{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
+			\hline
+			\citet{gao2020Visualising} & \citep{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
+			\hline
+			\citet{nusrat2020Recognition} & \citep{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
+			\hline
+			\citet{nickel2022Multicriteria} & \citep{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
+			\hline
+
+			% \multicolumn{1}{c|}{\textbf{Total unique papers: 51}} & 24&15&12& \multicolumn{1}{c}{}  \\
+			% \hhline{~|*{3}{-}|~}
+
+		\end{tabulary}
+	}
+	\caption{Related work with non-contiguous cartogram-based visualizations.
+\textbf{Cartogram type} is the type of cartogram used.
+\textbf{Geographic region} is the geographic region depicted by the cartogram.
+\textbf{Number of nodes} is the number of nodes depicted in the cartogram.
+}
+	\label{table:region vs node}
+\end{table*}
+}
+
+\thumbnail{dorling}{
+	Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
+	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
+	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
+	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
+	This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
+	This is important in our use case scenario involving EHRs.
+	Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}. 
+}
+
+\thumbnail{rectangular}{
+	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
+Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
+}
+
+\thumbnail{mosaic}{
+	Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+Some variants are able to preserve topological accuracy as well.
+}
 
@@
 
-{
-\renewcommand{\arraystretch}{1.5}
-\begin{table*}[!tb]
-	\centering
-	\resizebox{\textwidth}{!}{
-		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
-			\hhline{~|*{4}{-}|~}
-			\multicolumn{1}{c|}{\textbf{Literature}} &
-			\textbf{Title} &
-			\textbf{Cartogram Type(s)} &
-			\textbf{Geographic Region(s)} &
-			\textbf{Number of Nodes} &
-			\multicolumn{1}{c}{\textbf{Year}} \\
-			\hline
-
-			% \citea{auber2007Geographical} & \citetitlea{auber2007Geographical} & US & 19 & 2007 \\
-			% \hline
-			\citea{warf2008Geography} & \citetitlea{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
-			\hline
-			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
-			\hline
-			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
-			\hline
-			\citea{tong2018Cartograms} & \citetitlea{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
-			\hline
-			\citea{gao2020Visualising} & \citetitlea{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
-			\hline
-			\citea{nusrat2020Recognition} & \citetitlea{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
-			\hline
-			\citea{nickel2022Multicriteria} & \citetitlea{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
-			\hline
-
-			% \multicolumn{1}{c|}{\textbf{Total unique papers: 51}} & 24&15&12& \multicolumn{1}{c}{}  \\
-			% \hhline{~|*{3}{-}|~}
-
-		\end{tabulary}
-	}
-	\caption{Related work with non-contiguous cartogram-based visualizations.
-\textbf{Cartogram type} is the type of cartogram used.
-\textbf{Geographic region} is the geographic region depicted by the cartogram.
-\textbf{Number of nodes} is the number of nodes depicted in the cartogram.
-}
-	\label{table:region vs node}
-\end{table*}
-}
-
-\thumbnail{dorling}{
-	Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
-	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
-	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citea{cano2015Mosaic} in their related work section.
-	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
-	This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
-	This is important in our use case scenario involving EHRs.
-	Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}. 
-}
-
-\thumbnail{rectangular}{
-	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
-Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
-}
-
-\thumbnail{mosaic}{
-	Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
-Some variants are able to preserve topological accuracy as well.
-}
-
-\bobgraph{Peer-reviewed Applications: }\citea{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
-\citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
-\citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citea{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
-\citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US.
+\bobgraph{Peer-reviewed Applications: }\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
+\citet{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
+\citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citet{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US.
 See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
 
-Our work extends the algorithm described by \citea{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms.
-Our work enhances that of \citeauthor{tong2018Cartograms} in multiple ways.
+Our work extends the algorithm described by \citet{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms.
+Our work enhances that of \citet{tong2018Cartograms} in multiple ways.
 First, we introduce multiple features (rivers) into the layout, as opposed to a single river.
@@
 Finally, the way we evaluate the cartograms is also different.
-\citeauthor{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
+\citet{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
 Here our focus is on readability, and thus we include a user study.
@@
 National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
-\citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
-\citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.
-
-% \subsection{Contiguous Applications}
-
-% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.
+\citet{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
+\citet{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.
 
@@
 The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret.
-\citea{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.
+\citet{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.
 

### 2023-07-14 paper/section/related.tex
draft: increase thumbnail size

@@
 			\multicolumn{1}{c|}{\textbf{Literature}} &
-			\textbf{Title} &
 			\textbf{Cartogram Type(s)} &
@@
 			% \hline
-			\citet{warf2008Geography} & \citep{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
+			\citet{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
 			\hline
-			\citet{sun2010Effectiveness} & \citep{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+			\citet{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
 			\hline
-			\citet{cruz2017Adapted} & \citep{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+			\citet{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
 			\hline
-			\citet{tong2018Cartograms} & \citep{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
+			\citet{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
 			\hline
-			\citet{gao2020Visualising} & \citep{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
+			\citet{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
 			\hline
-			\citet{nusrat2020Recognition} & \citep{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
+			\citet{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
 			\hline
-			\citet{nickel2022Multicriteria} & \citep{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
+			\citet{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
 			\hline

### 2023-07-15 paper/main.tex
draft: IV submission

\documentclass[Afour,sagev,times]{sagej}

\usepackage{moreverb,url}
\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow, multicol}

\newcommand{\algorithmautorefname}{Algorithm}

\usepackage[table, dvipsnames]{xcolor}

% \include{preamble/preamble.main}

\definecolor{Mycolor1}{HTML}{4C5866} %#4C5866
\definecolor{Mycolor2}{HTML}{A6B6D2} %#A6B6D2
\definecolor{Mycolor3}{HTML}{67daff} %#67daff
\definecolor{Mycolor4}{HTML}{F7F7FE} %#F7F7FE
\definecolor{Mycolor5}{HTML}{666ad1} %#666ad1
\definecolor{Mycolor6}{HTML}{ffbb93} %#ffbb93
\newcommand{\mycolorbox}[2]{\begingroup\setlength{\fboxsep}{2pt}\colorbox{#1}{#2}\endgroup}

\newcommand{\bobgraph}[1]{\noindent\textbf{#1}}

% cite with author names
\newcommand{\citea}[1]{\citeauthor{#1}~\cite{#1}}

% draw a colored square
% color, width, height
\newcommand\coloredsquare[3][black]{\textcolor{#1}{\rule{#2}{#3}}}

% thumbnail image
\usepackage{wrapfig}
\newcommand{\thumbnail}[2]{
    \begin{samepage}

      \setlength{\intextsep}{3pt}
      \setlength{\columnsep}{3pt}
      \begin{wrapfigure}{r}{0.4\columnwidth}
        \vspace{-15pt}
        \begin{center}
          \includegraphics[width=0.38\columnwidth]{"#1".png}
        \end{center}
      \end{wrapfigure}
      #2

    \end{samepage}
}

\usepackage{algpseudocode}
\usepackage{algorithm}
% switch for algorithmic
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicdefault{\textbf{default}}

\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\textbf{:}}{\algorithmicend\ \algorithmicswitch}%
\algtext*{EndSwitch}%

\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1\textbf{:}}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndCase}%

\algdef{SE}[DEFUALT]{Default}{EndDefault}{\algorithmicdefault\textbf{:}}{\algorithmicend\ \algorithmicdefault}%
\algtext*{EndDefault}%

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

% code listing
\usepackage{listings}
\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  xleftmargin=\parindent,
  language=C,
  showstringspaces=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
}

\lstdefinestyle{customasm}{
  belowcaptionskip=1\baselineskip,
  xleftmargin=\parindent,
  language=[x86masm]Assembler,
  basicstyle=\footnotesize\ttfamily,
  commentstyle=\itshape\color{purple!40!black},
}

\lstset{escapechar=@,style=customc}

% list with styles
\usepackage{enumitem}
% \include{preamble/preamble.main}

% \include{preamble/preamble.math}
\usepackage{amsmath,bm}
\usepackage{amssymb}

% bold \mathcal{}
\DeclareMathAlphabet{\mathbcal}{OMS}{cmsy}{b}{n}

% wrapper for math symbols
\newcommand{\Vector}[1]{\protect\overrightarrow{#1}}
\newcommand{\Line}[1]{\protect\overline{#1}}
\newcommand{\dx}{\Delta x}
\newcommand{\dy}{\Delta y}
\newcommand{\Width}[1]{\lvert #1 \rvert}

% UpdateNodePosition
\newcommand{\nodeList}{\bm{L}}

\newcommand{\nodeFNORLine}{\Line{\node\nodeFNOR}}
\newcommand{\nodeSize}{s}
\newcommand{\nodeSizeAvg}{\nodeSize_{avg}}
\newcommand{\nodeSizeMax}{\nodeSize_{max}}
\newcommand{\nodeError}{\epsilon}
\newcommand{\nodeCartographicError}{\nodeError_{c}}
\newcommand{\nodeCartographicErrorMax}{\nodeError_{c_{max}}}
\newcommand{\nodeTopologicalError}{\nodeError_{t}}
\newcommand{\nodeTopologicalErrorMax}{\nodeError_{t_{max}}}
\newcommand{\stalemateMax}{w}
\newcommand{\nodeCross}{\node.cross}
\newcommand{\nodeListCross}{\nodeList.cross}
\newcommand{\nodeListCrossCount}{\nodeList.count}

% move position

\newcommand{\node}{\bm{n}}
\newcommand{\nodePrevious}{\node_{p}}
\newcommand{\nodeFNOR}{\node_t}

\newcommand{\nodeVectorTC}{\Vector{\nodeFNOR\node}}
\newcommand{\nodeVectorCT}{\Vector{\node\nodeFNOR}}

% check river crossing

\newcommand{\river}{\bm{r}}
\newcommand{\riverVector}{\Vector{v_{\river}}}
\newcommand{\riverList}{\bm{R}}
\newcommand{\nodeBoundingBox}{b_{\Line{n}}}
\newcommand{\riverEdgeBoundingBox}{b_{\Edge}}
\newcommand{\boundingBoxInt}{b_{int.}}
\newcommand{\nodeEdge}{e_n}
\newcommand{\edgeBoxInt}{e_{int.}}

% derive corridor
\newcommand{\Corridor}{c}
\newcommand{\CorridorLength}{\Corridor_l}
\newcommand{\CorridorWidth}{\Corridor_w}
\newcommand{\PointP}{\node_c}
\newcommand{\Edge}{\Line{e}}
\newcommand{\EdgeStart}{\Edge.start}
\newcommand{\EdgeEnd}{\Edge.end}
\newcommand{\EdgeParallel}{\Line{e_p}}
\newcommand{\EdgeParallelA}{\Line{e_{p^1}}}
\newcommand{\EdgeParallelB}{\Line{e_{p^2}}}
\newcommand{\Distance}{d}
\newcommand{\nodeVectorNP}{\Vector{\node\PointP}}
\newcommand{\nodeVectorNV}{\Vector{\node\nodeFNOR}}
\newcommand{\nodeLineNV}{\Line{\node\nodeFNOR}}
\newcommand{\nodeLinePN}{\Line{\nodePrevious\node}}
\newcommand{\nodeVectorNtP}{\Vector{\nodeFNOR\PointP}}
\newcommand{\nodeLineNtNc}{\Line{\nodeFNOR\PointP}}
\newcommand{\nodeLineWidthNtP}{\Width{\nodeFNOR\PointP}}
\newcommand{\nodeInCorridor}{\node_{in}}
\newcommand{\nodeInCorridorT}{\node_{in_t}}

% derive corridor point
\newcommand{\Scale}{scale}

% figure fig:corridor

\newcommand{\nodeVectorNNn}{\Vector{\node\node_{t}}}
\newcommand{\nodeVectorNinNinn}{\Vector{\nodeInCorridor\nodeInCorridorT}}

\newcommand{\vectorAvg}{\bm{v}_{avg}}
\newcommand{\NxR}{\node_{n_{x}r}}
\newcommand{\RxN}{\node_{r_{x}n}}
% \include{preamble/preamble.math}

\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage[nameinlink]{cleveref}

\setcounter{secnumdepth}{3}

\begin{document}

\title{Demers Cartogram with Rivers}

\begin{abstract}
    Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique. A relative of the Dorling cartogram, the Demers cartogram, utilizing squares instead of circles to represent regions. This alternative approach allows for a more intuitive comparison of regions, utilizing screen space more efficiently. However, a drawback of the Dorling cartogram and its variants lies in the potential displacement of regions from their original positions, ultimately compromising legibility, readability, and accuracy. To tackle this limitation, we propose a novel hybrid cartogram layout algorithm that incorporates dynamic topological elements, such as rivers, into Demers cartograms. The presence of rivers significantly impacts both the layout and visual appearance of the cartograms. Through a user study conducted on an Electronic Health Records (EHR) dataset, we evaluate the efficacy of the proposed hybrid layout algorithm. The obtained results illustrate that this approach successfully retains key aspects of the original cartogram while enhancing legibility, readability, and overall accuracy.
\end{abstract}

\keywords{Cartogram, Demers Cartogram}

\maketitle

\section{Introduction and Motivation}

\begin{samepage}

    Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography,inoue2011New}.
    Various styles of cartograms have been proposed and implemented for applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth,keepaspectratio]{cover.png}
        \caption{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
        \label{fig:cover}
    \end{figure}

\end{samepage}

Among the four types of cartogram categorized in a survey by \citet{nusrat2016State} (See \Cref{sec:RelatedWork} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes. Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions.
We hypothesize that introducing rivers improves the overall legibility of a cartogram.
By \textit{legibility} we mean readability and ability to interpret the cartogram.
To assess this hypothesis, we designed an experimental setup where participants engaged in correspondence and location tasks as part of a user study.
To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
We present a user study that demonstrates its effectiveness.

Our contributions include:

\begin{itemize}
    \setlength\itemsep{0px}
    \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
    \item A novel hybrid layout algorithm that combines node positions with features such as rivers,
    \item A user study evaluation of the technique with an application to EHRs.
\end{itemize}

The results of the user study indicate that rivers can improve the legibility of cartograms.
One of the major challenges involved is developing a layout algorithm that handles different shapes.
In other words, the hybrid layout algorithm is novel because it handles different types of elements: square representing regions and polylines representing rivers.
Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.

\section{Related Work}
\label{sec:RelatedWork}

This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world implementations of cartogram-based visualizations.

\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms preserve topology, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape. Non-contiguous cartograms sacrifice topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.

\citet{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological. Each cartogram design may make various types of accuracy trade-offs between dimensions. We provide a comparison of these trade-offs as introduced by \citet{nusrat2016State} in \Cref{table:accuracy}. In addition, we include the Demers as it is the focus of our work.

    {
        \renewcommand{\arraystretch}{1.5}
        \begin{table*}[!tb]
            \centering
            \resizebox{\textwidth}{!}{
                \begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
                    \hhline{~|*{4}{-}|~}
                    \multicolumn{1}{c|}{\textbf{Literature}} &
                    \textbf{Cartogram Type(s)} &
                    \textbf{Geographic Region(s)} &
                    \textbf{Number of Nodes} &
                    \multicolumn{1}{c}{\textbf{Year}} \\
                    \hline

                    % \citet{auber2007Geographical} & \citep{auber2007Geographical} & US & 19 & 2007 \\
                    % \hline
                    \citet{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
                    \hline
                    \citet{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
                    \hline
                    \citet{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
                    \hline
                    \citet{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
                    \hline
                    \citet{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
                    \hline
                    \citet{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
                    \hline
                    \citet{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
                    \hline

                    % \multicolumn{1}{c|}{\textbf{Total unique papers: 51}} & 24&15&12& \multicolumn{1}{c}{}  \\
                    % \hhline{~|*{3}{-}|~}

                \end{tabulary}
            }
            \caption{Related work with non-contiguous cartogram-based visualizations. \textbf{Cartogram type} is the type of cartogram used. \textbf{Geographic region} is the geographic region depicted by the cartogram. \textbf{Number of nodes} is the number of nodes depicted in the cartogram. }
            \label{table:region vs node}
        \end{table*}
    }

\thumbnail{dorling}{
    Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
    A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
    In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
    Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
    This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
    This is important in our use case scenario involving EHRs.
    Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}.
}

\thumbnail{rectangular}{
    Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}. Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
}

\thumbnail{mosaic}{
    Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}. Some variants are able to preserve topological accuracy as well.
}

{
    \renewcommand{\arraystretch}{1.5}
    \begin{table}[!tb]
        \centering
        \resizebox{\columnwidth}{!}{
            \begin{tabulary}{\columnwidth}{|*{5}{l|}}
                \hhline{~|*{3}{-}|~}
                \multicolumn{1}{c|}{} & \multicolumn{3}{c|}{\cellcolor{Mycolor2} \textbf{Accuracy}} \\
                \hhline{~|*{4}{-}}
                \multicolumn{1}{c|}{\textbf{Cartogram Variant}} &
                \textbf{Statistical} &
                \textbf{Geographical} &
                \textbf{Topological} &
                \textbf{Contiguity} \\
                \hline
                Contiguous & \cellcolor{Mycolor3}Variable & \cellcolor{Mycolor3}Variable & Accurate & Yes \\
                \hline
                Non-contiguous & \cellcolor{Mycolor5}\textcolor{white}{Accurate} & \cellcolor{Mycolor5}\textcolor{white}{Shape is accurate} &\cellcolor{Mycolor6}Inaccurate & No   \\
                \hline
                Rectangular & \cellcolor{Mycolor3}Variable & Shape is inaccurate & \cellcolor{Mycolor3}Variable & Yes   \\
                \hline
                Dorling & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
                \hline
                Demers & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
                \hline
            \end{tabulary}
        }
        \caption{\coloredsquare[Mycolor3]{0.25cm}{0.25cm} Trade-off between dimensions. \coloredsquare[Mycolor6]{0.25cm}{0.25cm} Dimension sacrificed in order to improve \coloredsquare[Mycolor5]{0.25cm}{0.25cm} target dimension's accuracy.}
        \label{table:accuracy}
    \end{table}
}

\bobgraph{Peer-reviewed Applications: }\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US. \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citet{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citet{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US. See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

Our work extends the algorithm described by \citet{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms. Our work enhances that of \citet{tong2018Cartograms} in multiple ways. First, we introduce multiple features (rivers) into the layout, as opposed to a single river. Second, we make topological features dynamic and further improve legibility and geographical accuracy. By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm. In previous work, the river is static and serves merely as a boundary. Third, we improve the algorithm to resolve stalemates. Finally, the way we evaluate the cartograms is also different. \citet{tong2018Cartograms} count river crossings to evaluate error (a statistical metric). Here our focus is on readability, and thus we include a user study.

\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news. The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citet{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citet{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

One of the disadvantages of Dorling and Demers cartograms is legibility. The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret. \citet{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.

In this paper, we introduce a new type of topological feature, a river, as a constraint to compute the final layout, with the aim of improving the interpretation, readability, and accuracy of this class of cartograms.

\section{Data Description}

 {
  \begin{figure}[b!]
      \centering
      \includegraphics[width=0.6\columnwidth]{ccg.png}
      \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
      \label{fig:ccg}
  \end{figure}
 }

Obtaining heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}. The first step is to obtain both geospatial boundaries and EHR data. The second step is to preprocess the EHR data to remove empty and erroneous values. The final step is to transform the data into a format that is suitable for cartograms. Geospatial boundaries, or shapefiles, were obtained from the sources described here.

\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes over time due to NHS reorganization. The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of published public EHR data based on the latest CCG reorganizations that took place in 2021 and 2022.

\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England. These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues. Although including smaller rivers is technically feasible, it may not increase the legibility of the cartogram. This is an open question for future work.

We first obtain a relation ID by searching for a river, e.g. River Thames, on OpenStreetMap. The relation ID is used to construct a query (See \Cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.

After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust the projections and convert them into GeoJSON files. Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON file \cite{TopoJSON}. TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation. See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the preprocessing result. We describe the one-time preprocessing steps in more detail in \Cref{app:pre-processing}.

\bobgraph{EHR Data: }We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}. The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS. Some datasets include:
\begin{itemize}
    \setlength\itemsep{0px}
    \item Under 75 mortality: cardiovascular disease, respiratory disease, liver disease, and cancer
    \item Emergency hospital admission: stroke, alcohol-specific admission and readmission, coronary heart disease, re-admissions within 30 days of discharge, children with lower respiratory tract infections
\end{itemize}

For all datasets, a spreadsheet including the following is provided:

\begin{itemize}
    \setlength\itemsep{0px}
    \item Reporting period: Calendar year of registration
    \item Period of coverage: Start and end date or reporting period
    \item Breakdown: Organization type
    \item ONS code: UK Office for National Statistics CCG code
    \item Level: CCG Code
    \item Level description: CCG Name
    \item Gender
    \item Indicator value: Directly standardized mortality rate
    \item CI lower: lower 95\% confidence interval
    \item CI upper: upper 95\% confidence interval
    \item Denominator: The count of registered patients
    \item Numerator: Number of deaths
\end{itemize}

Each CCG has a unique ONS code, which is used to link the CCG shapefile with the statistical data.

\section{Demers Cartogram with Rivers}
\label{sec:main}

{
    \begin{figure*}[tb!]
        \centering
        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{flowchart.png}
        \caption{An overview of our hybrid layout algorithm incorporating rivers. See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail. See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
        \label{fig:flowchart}
    \end{figure*}
}

{
    \begin{figure}[tb!]
        \centering
        \includegraphics[width=0.7\columnwidth,keepaspectratio]{river_resolution.png}
        \caption{The resolution of rivers can be dynamically adjusted by the user. (A) shows River Thames at its original resolution with 10,170 edges. (B) shows the river at a reduced resolution of 49 edges. We further smooth the river by removing 19 vertices in dense areas, as shown in (C). The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
        \label{fig:river resolution}
    \end{figure}
}

\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the hybrid cartogram layout process that includes rivers.

\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries. For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel. We then load the river shapefiles and render the rivers. Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex. We do not need the original river resolution to incorporate them into cartograms. We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests. This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}. We further apply simplification by removing vertices that are too close to each other. The initialization procedure is a one-time process that can be saved for reuse.

\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes. We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}. Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions. An increase in $ \nodeSize $ can cause the nodes to overlap. During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position. Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position. If a node oscillates across a river, we identify this as a stalemate. One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river. We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $. The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.

\subsection{River Intersection Testing}
\label{subsec:{River Intersection Testing}}

The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}. When a node's position changes, we test if the node's trajectory intersects any segment of a river. See \Cref{alg:check river intersection}. A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required. Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm. We label these nodes, $ \NxR $.

\subsection{Translating Rivers}
\label{subsec:{Translating Rivers}}

For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $. Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation. We then use a heuristic to translate $ \river $ using the average vector of node intersection $$ \vectorAvg = \sum_{i=1}^{\nodeTopologicalError} \frac{\Vector{\node_i\node_{i_t}}}{\nodeTopologicalError} $$, where $ \nodeTopologicalError $ is the number of nodes intersecting the river. This step intends to create space for the next iteration of node translation without crossing a river. The detailed procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.

When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $. As a heuristic, we also translate these nodes by $ \vectorAvg $. The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together. In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $. Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm. We can use this to ensure smooth transitions between iterations.

\subsection{Process Stalemates}
\label{subsec:{Process Stalemates}}

As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}. If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion. A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D). All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E). The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.

{
    \begin{figure}[tb!]
        \centering
        \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
        \label{fig:flowchart-stalemate}
    \end{figure}
   }

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{stalemate.png}
    \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times. The node is translated back to its previous position (iteration 2). A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
    \label{fig:stalemate}
\end{figure}
}

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{corridor.png}
    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A). To address this, we derive a corridor (orange rectangle in (E)) based on $ \node $ and $ \nodeFNOR $. All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $. For clarity in the illustration, we place nodes sparsely in this figure.}
    \label{fig:corridor}
\end{figure}
}

\subsection{Terminating the Algorithm}
\label{subsec:{Terminating the Algorithm}}

The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user. We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
$$

where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $. We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers. We express $ \nodeTopologicalError $ in the normalized range $ \nodeTopologicalError \in [0,1] $, where $ \nodeTopologicalErrorMax $ is the case where all nodes cross a river.

When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $). Every node remains on the same side of the river as its centroid.

\subsection{User Options}\label{subsec:{UserOptions}}

\Cref{fig:overview} presents an overview of the application, including user options. The user can adjust the following parameters:

\bobgraph{Rendering Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.

\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR attributes or set to uniform.

\bobgraph{Overlap Removal Speed:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.

\bobgraph{Maximum Cartographic Error:} The user can adjust the maximum cartographic error, $ \nodeCartographicErrorMax $, which is used to terminate the algorithm.

\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers. Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default. Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.

\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations. A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout. The default corridor length is three times the max node size.

\bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram. Similarly, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.

    {
        \begin{figure*}[tb!]
            \centering
            \includegraphics[width=0.7\textwidth,keepaspectratio]{UI.png}
            \caption{A screenshot of the user interface. User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers. Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic. See \Cref{subsec:{UserOptions}} for more on user options. In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
            \label{fig:overview}
        \end{figure*}
    }

\newcommand{\pCount}{24 }
\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}

We conduct a user study to evaluate the effectiveness of our approach. We chose this type of evaluation because legibility is a human-centred characteristic. Many different types of statistical error metrics have been evaluated in previous work \cite{nusrat2016State}, however, our focus is more user-centric in nature.

\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}

We formulate the following hypotheses to motivate our user study:

\textbf{H1:} The introduction of rivers can improve the legibility and recognizability of cartograms.

\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.

\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.

We test these hypotheses using location-based tasks. Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right). See \Cref{fig:task}.

\subsection{User Study Variables}\label{subsec:{User Study Variables}}

In this section, we discuss the variables of our user study.

\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. We set this parameter to render rivers or not.

\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram. Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants. \textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.

\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms. \textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find. The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.

\subsection{User Study Design}\label{subsec:{User Study Design}}

In this section we describe the user study participants, datasets, and the experimental procedure.

\bobgraph{Participants:} We recruited \pCount participants.

\begin{itemize}
    \setlength\itemsep{0px}
    \item Gender: 10 females and 14 males
    \item Age Group: 18-24 (16), 25-29 (8)
    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
\end{itemize}

\bobgraph{Datasets:} We used the following EHR datasets for our evaluation:

\begin{itemize}
    \setlength\itemsep{0px}
    \item Population
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol-related liver disease
    \item Alcohol-specific admission and readmission
\end{itemize}

135 CCGs are rendered on the screen as a choropleth on the left. We then generated another view on the right using cartograms. The color of both visual designs is mapped to the dataset. See \Cref{fig:task} for an example.

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{task.png}
            \caption{A sample location task for participants. The left shows the choropleth map, and the right shows the corresponding cartogram. Both images show the three longest rivers in England, and color is mapped to CCG population. The target CCG blinks on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram. In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated. }
            \label{fig:task}
        \end{figure}
    }

\subsubsection{Procedure}
\label{subsec:{Procedure}}

The user study is designed to be conducted online, and participants are asked to complete the tasks using their own hardware. The user study includes four parts:

\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos. The instructions are designed to help participants understand the concepts used in the tasks. Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.

\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design. These tasks also serve as a quality test to see whether users are taking the study seriously. A demonstration of three sample tasks is also included in the instructional video.

\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 CCGs. Response and reaction time are recorded. These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.

\textbf{P4:} The participants are asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.

The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.

\subsection{User Study Results}
\label{subsec:{User Study Results}}

{
\begin{figure}[b!]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
    \label{fig:task-acc}
\end{figure}
}

In this section, we delve into the analysis of the preliminary user study conducted. Our study involved the recruitment of \pCount participants, each tasked with completing 16 location-related tasks. As a result, we collected a total of 384 responses. However, we excluded 24 responses from our analysis as they took over 60 seconds to complete. We hypothesize that these prolonged response times indicate potential distractions experienced by participants during the study.

\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants. The chart shows that the introduction of rivers has improved the accuracy of two CCGs. The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002. This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.

\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs. The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs. The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38. However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.

    {
        \begin{figure}[t!]
            \centering
            \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
            \label{fig:task-rt}
        \end{figure}
    }

\bobgraph{Participant Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
\begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
    \item 87.5\% of participants agree that including rivers is useful.
    \item 83.3\% of participants agree that rivers increase the legibility of a cartogram.
    \item 95.8\% of participants agree that including rivers makes cartograms easier to understand.
    \item 62.5\% of participants agree that including rivers makes CCGs easier to locate.
    \item 87.5\% of participants agree that including rivers adds value to the standard cartogram.
\end{enumerate}

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{likert.png}
    \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
    \label{fig:likert}
\end{figure}
}

The qualitative results clearly support our hypothesis H1 that rivers generally increase the legibility of cartograms.

\section{Limitations and Future Work}
\label{sec:{Limitations and Future Work}}

In this section, we discuss some limitations of our work and future research directions.
Because this is a new concept, it opens the door for many future research directions.
The first limitation is the color map that we use to depict the data in our cartograms.
We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
See \Cref{fig:extreme} for an example.
We plan to explore the impact of different color maps on the legibility of cartograms in future work.

    {
        \begin{figure}[tb!]
            \centering
            \includegraphics[width=0.7\columnwidth,keepaspectratio]{extreme.png}
            \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
            \label{fig:extreme}
        \end{figure}
    }

Another limitation is the overlap removal algorithm (FNOR) we use.
We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers. Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.

Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
Furture work also includes adding rivers to dorling-style cartograms.

\section{Conclusions}\label{sec:{Conclusions}}

In this paper, we present a novel hybrid layout approach to generate non-contiguous cartograms with rivers. We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers. We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions. We also discuss the limitations of our work and future research directions.

\urlstyle{same}
\bibliographystyle{plainnat}
\bibliography{references}

\cleardoublepage

\section{Appendix}
\appendix

\section{Obtaining Shapefile}
\label{obtaining-shapefile}
\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    // River Great Ouse: 2798097
    // River Trent: 2863468
    out skel;
\end{lstlisting}
\section{Pre-processing Shapefiles}
\label{app:pre-processing}
Shapefiles from different sources are likely to be incompatible. In our case, the NHS CCG shapefile is incompatible with the river shapefiles. The major reason for the incompatibility is the coordinate reference system (CRS). The CRS of the CCG shapefile is EPSG:27700 (OSGB36 - British National Grid). The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System). Here, we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.

\subsection{Import Shapefiles into QGIS}
We first load all three river shapefiles into QGIS \Cref{fig:import_rivers}, followed by the CCG shapefile \Cref{fig:import_ccgs}.

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{import_rivers.png}
    \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
    \label{fig:import_rivers}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{import_ccgs.png}
    \caption{QGIS interface, with all NHS CCGs imported.}
    \label{fig:import_ccgs}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{unify_crs.png}
    \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
    \label{fig:unify_crs}
\end{figure}
}

\subsection{Export Shapefiles in GeoJSON and Unify the Coordinate Reference System (CRS)}

We then use QGIS to unify the CRS, and export both layers in GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{export_rivers.png}
    \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_rivers}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{export_ccgs.png}
    \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
    \label{fig:export_ccgs}
\end{figure}

}

\subsection{Merge Shapefiles and Reduce File Size}

We then merge two layers into one layer, and export it in the TopoJSON format using Mapshaper \cite{blochMapshaper}. Mapshaper also supports the simplification of GeoJSON shapefiles. See \Cref{fig:export_topojson}.

{
\begin{figure}[H]
    \centering
    \includegraphics[width=\columnwidth]{export_topojson.png}
    \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
    \label{fig:export_topojson}
\end{figure}
}

\subsection{Pre-processing Result}

\Cref{table:pre-processing_result} shows the preprocessing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.

    {
        \renewcommand{\arraystretch}{1.5}
        \begin{table}[H]
            \centering
            \resizebox{\columnwidth}{!}{
                \begin{tabulary}{\columnwidth}{|*{4}{l|}}
                    \hhline{~|*{3}{-}}
                    \multicolumn{1}{c|}{\textbf{Shapefile}} &
                    \cellcolor{Mycolor2}\textbf{Original} &
                    \cellcolor{Mycolor2}\textbf{GeoJSON} &
                    \cellcolor{Mycolor2}\textbf{TopoJSON} \\
                    \hline
                    Rivers & 2.0 MB (GeoJSON) & 1.4 MB & -  \\
                    \hline
                    NHS CCGs & 46.6 MB (.shp, Esri vector shapefile) & 140.2 MB & -  \\
                    \hline
                    Merged & - & - & 16.3 MB \\
                    \hline
                \end{tabulary}
            }
            \caption{The file size is reduced by 88.5\% from the original size.}
            \label{table:pre-processing_result}
        \end{table}
    }

% \clearpage
\section{Procedure: DeriveParallelEdge}
% \begin{noindent}

    \begin{algorithm}[H]
        \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}

        \textbf{Input:} \\
        $ \Edge \gets $ the edge used to derive the parallel edge $ \EdgeParallel $ \\
        $ \Distance \gets $ the shortest distance between $ \Edge $ and $ \EdgeParallel $ \\

        \textbf{Output:} \\
        An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $. \\

        \textbf{Local variables:} \\
        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
        $ \Scale \gets $ the scale of $ \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $ \\

        \begin{algorithmic}[1]
            \Procedure{DeriveParallelEdge}{$ \Edge $, $ \Distance $}
                \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $

                \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $

                \State $ \Scale \gets \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $

                \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \EdgeStart.x $

                \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \EdgeStart.y $

                \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \EdgeEnd.x $

                \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \EdgeEnd.y $

            \State \Return{$ \EdgeParallel $}

            \EndProcedure

        \end{algorithmic}
    \end{algorithm}
%\end{noindent}

\section{Procedure: TranslateNode}

\begin{algorithm}[H]
    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}

    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Output:} \\
    $ \nodeTopologicalError \gets $ the number of nodes crossing the river in the input \\

    \textbf{Global variables:} \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{TranslateNode}{$ \nodeList $}
        \State $ \nodeTopologicalError \gets 0 $
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}

        \State $ \node(x,y) \gets \nodeFNOR(x,y) $

        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
        \State $ \nodeCross ++ $
        \State $ \nodeTopologicalError ++ $

        \If{$ \nodeCross < \stalemateMax $}
        \State \hskip-2em \Comment{Translate back to previous position}
        \State $ \node(x,y) \gets \nodePrevious(x,y) $

        \Else

        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}

        \EndIf

        \EndIf

        \EndIf

        \EndFor
        \State \Return{$ \nodeTopologicalError $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}
\vfill\eject

\section{Procedure: UpdateLayout}

\begin{algorithm}[H]
    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:main}.}\label{alg:UpdateLayout}
    \textbf{Global variables:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\

    $ \nodeSize \gets $ the initial size of all nodes \\
    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\

    \textbf{Local variables:} \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\

    \begin{algorithmic}[1]
        \Procedure{UpdateLayout}{}

        \State $ \nodeSize \gets 1 $
        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
        \While{$ \nodeListCross > 0 $}

        \State $ \nodeListCross \gets 0 $
        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
        \State \Call{TranslateRiver}{$ \nodeList $}
        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}

        \EndWhile

        \State $ \nodeSize ++ $

        \EndWhile

        \EndProcedure
    \end{algorithmic}
\end{algorithm}
\vfill\eject

\section{Procedure: TranslateRiver}

\begin{algorithm}[H]
    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
    \textbf{Input:} \\
    $ \nodeList \gets $ a list of $ \node $ representing regions \\

    \textbf{Local variables:} \\
    $ \riverList \gets $ a list of $ \river $ representing river features \\
    $ \node \gets $ a node is an object with the following properties: \\
    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\

    \begin{algorithmic}[1]
        \Procedure{TranslateRiver}{$ \nodeList $}
        \ForEach {$ \river \in \riverList $}
        \State $ \nodeTopologicalError \gets 0 $
        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
        \ForEach {$ \node \in \nodeList $}

        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
        \State $ \nodeTopologicalError ++ $
        \EndIf

        \EndIf

        \EndFor

        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeTopologicalError} $
        \EndFor

        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\section{Procedure: DerivePoint}
% \begin{noindent}

    \begin{algorithm}[H]
        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
        \textbf{Input:} \\
        $ \Edge \gets $ the edge used to derive the new point \\
        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\

        \textbf{Output:} \\
        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\

        \textbf{Local variables:} \\
        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\

        \begin{algorithmic}[1]
            \Procedure{DerivePoint}{$ \Edge $, $ \Distance $}
                \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $

                \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $

                \State $ \PointP.x \gets \frac{\dx}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $

                \State $ \PointP.y \gets \frac{\dy}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $

            \State \Return{$ \PointP $}

            \EndProcedure

        \end{algorithmic}
    \end{algorithm}

%\end{noindent}
% \vfill\eject

\section{Procedure: ProcessStalemate}

% \begin{noindent}

    \begin{algorithm}[H]
        \caption{Procedure to derive a corridor to resolve stalemate. We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}

        \textbf{Input:} \\
        $ \node \gets $ the node used to derive the corridor \\

        \textbf{Global variables:} \\
        $ \CorridorLength \gets $ the length of a corridor \\
        $ \CorridorWidth \gets $ the width of a corridor \\

        \textbf{Local variables:} \\
        $ \Corridor \gets $ the corridor \\
        $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
        $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
        $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\

        \begin{algorithmic}[1]
            \Procedure{ProcessStalemate}{$ \node $}
                \State $ \node(x,y) \gets \nodePrevious(x,y) $

                \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}

                \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}

                \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}

                \State $ \Corridor \gets
                    \begin{bmatrix}
                        \EdgeParallelA.start &
                        \EdgeParallelA.end \\

                        \EdgeParallelB.start &
                        \EdgeParallelB.end \\
                    \end{bmatrix} $

                \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}

                    \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $

                    \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $

                \EndFor
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}

    %\end{noindent}
% \vfill\eject

\section{Procedure: TestIntersection}
% \begin{noindent}
    \begin{algorithm}[H]
        \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
        \textbf{Input:} \\
        $ \nodeLineNV \gets $ the node's translation path \\
        $ \river \gets $ a river feature \\

        \textbf{Output:} \\
        Returns $ True $ if the node crosses a river. \\

        \textbf{Local variables:} \\
        $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \Edge $ \\
        $ \Edge \gets $ an edge of $ \river $ \\

        \begin{algorithmic}[1]
            \Procedure{TestIntersection}{$ \nodeLineNV $, $ \river $}

            \ForEach{$ \Edge \in \river $}
                \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
                \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \Edge $)

                \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
                    \State \Return{$ \nodeLineNV ~intersect~ \Edge $}
                    \EndIf
            \EndFor

            \State \Return{$ False $}
            \EndProcedure
        \end{algorithmic}
    \end{algorithm}
    %\end{noindent}
    % \vfill\eject

\end{document}

### 2023-07-15 paper/section/data.tex
draft: IV
draft: IV submission

@@
 
-%  {
-%   \begin{figure}[b!]
-%       \centering
-%       \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
-%       \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
-%       \label{fig:ccg}
-%   \end{figure}
-%  }
+ {
+  \begin{figure}[b!]
+      \centering
+      \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
+      \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
+      \label{fig:ccg}
+  \end{figure}
+ }
 
@@
 River Thames, on OpenStreetMap.
-The relation ID is used to construct a query (See \cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
+The relation ID is used to construct a query (See \Cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
 
@@
 \begin{itemize}
+    \setlength\itemsep{0px}
     \item Under 75 mortality: cardiovascular disease, respiratory disease, liver disease, and cancer
@@
 \begin{itemize}
+    \setlength\itemsep{0px}
     \item Reporting period: Calendar year of registration

### 2023-08-18 user-study/splitCSV.js
draft: IV reviews

@@ -54,7 +54,7 @@ function readCSV() {
                 const time = parseInt(row[i * 2 + 1]) / 1000;
 
                 // exclude invalid data, RT >= 20s
-                if (time >= 20) {
+                if (time >= 60) {
                     continue;
                 }
 

### 2023-08-18 user-study/ttest.js

@@ -47,7 +47,7 @@ function readCSV() {
             for (let i = 0; i < taskList.length; i++) {
                 const time = parseInt(row[i * 2 + 1]) / 1000;
 
-                if (time >= 20) {
+                if (time >= 60) {
                     continue;
                 }
 

### 2023-08-18 paper/GHRVis.tex
draft: increase thumbnail size

@@
 
-\include{preamble/preamble.main}
-\include{preamble/preamble.math}
+% \include{preamble/preamble.main}
+
+\definecolor{Mycolor1}{HTML}{4C5866} %#4C5866
+\definecolor{Mycolor2}{HTML}{A6B6D2} %#A6B6D2
+\definecolor{Mycolor3}{HTML}{67daff} %#67daff
+\definecolor{Mycolor4}{HTML}{F7F7FE} %#F7F7FE
+\definecolor{Mycolor5}{HTML}{666ad1} %#666ad1
+\definecolor{Mycolor6}{HTML}{ffbb93} %#ffbb93
+\newcommand{\mycolorbox}[2]{\begingroup\setlength{\fboxsep}{2pt}\colorbox{#1}{#2}\endgroup}
+
+\newcommand{\bobgraph}[1]{\noindent\textbf{#1}}
+
+% cite with author names
+\newcommand{\citea}[1]{\citeauthor{#1}~\cite{#1}}
+
+% draw a colored square
+% color, width, height
+\newcommand\coloredsquare[3][black]{\textcolor{#1}{\rule{#2}{#3}}}
+
+% thumbnail image
+\usepackage{wrapfig}
+\newcommand{\thumbnail}[2]{
+    \begin{samepage}
+
+      \setlength{\intextsep}{3pt}
+      \setlength{\columnsep}{3pt}
+      \begin{wrapfigure}{r}{0.4\columnwidth}
+        \vspace{-15pt}
+        \begin{center}
+          \includegraphics[width=0.38\columnwidth]{"#1".png}
+        \end{center}
+      \end{wrapfigure}
+      #2
+
+    \end{samepage}
+}
+
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+% switch for algorithmic
+% New definitions
+\algnewcommand\algorithmicswitch{\textbf{switch}}
+\algnewcommand\algorithmiccase{\textbf{case}}
+\algnewcommand\algorithmicdefault{\textbf{default}}
+
+\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\textbf{:}}{\algorithmicend\ \algorithmicswitch}%
+\algtext*{EndSwitch}%
+
+\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1\textbf{:}}{\algorithmicend\ \algorithmiccase}%
+\algtext*{EndCase}%
+
+\algdef{SE}[DEFUALT]{Default}{EndDefault}{\algorithmicdefault\textbf{:}}{\algorithmicend\ \algorithmicdefault}%
+\algtext*{EndDefault}%
+
+\algnewcommand\algorithmicforeach{\textbf{for each}}
+\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
+
+% code listing
+\usepackage{listings}
+\lstdefinestyle{customc}{
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  xleftmargin=\parindent,
+  language=C,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  keywordstyle=\bfseries\color{green!40!black},
+  commentstyle=\itshape\color{purple!40!black},
+  identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+
+\lstdefinestyle{customasm}{
+  belowcaptionskip=1\baselineskip,
+  xleftmargin=\parindent,
+  language=[x86masm]Assembler,
+  basicstyle=\footnotesize\ttfamily,
+  commentstyle=\itshape\color{purple!40!black},
+}
+
+\lstset{escapechar=@,style=customc}
+
+% list with styles
+\usepackage{enumitem}
+% \include{preamble/preamble.main}
+
+% \include{preamble/preamble.math}
+\usepackage{amsmath,bm}
+\usepackage{amssymb}
+
+% bold \mathcal{}
+\DeclareMathAlphabet{\mathbcal}{OMS}{cmsy}{b}{n}
+
+% wrapper for math symbols
+\newcommand{\Vector}[1]{\protect\overrightarrow{#1}}
+\newcommand{\Line}[1]{\protect\overline{#1}}
+\newcommand{\dx}{\Delta x}
+\newcommand{\dy}{\Delta y}
+\newcommand{\Width}[1]{\lvert #1 \rvert}
+
+% UpdateNodePosition
+\newcommand{\nodeList}{\bm{L}}
+
+\newcommand{\nodeFNORLine}{\Line{\node\nodeFNOR}}
+\newcommand{\nodeSize}{s}
+\newcommand{\nodeSizeAvg}{\nodeSize_{avg}}
+\newcommand{\nodeSizeMax}{\nodeSize_{max}}
+\newcommand{\nodeError}{\epsilon}
+\newcommand{\nodeCartographicError}{\nodeError_{c}}
+\newcommand{\nodeCartographicErrorMax}{\nodeError_{c_{max}}}
+\newcommand{\nodeTopologicalError}{\nodeError_{t}}
+\newcommand{\nodeTopologicalErrorMax}{\nodeError_{t_{max}}}
+\newcommand{\stalemateMax}{w}
+\newcommand{\nodeCross}{\node.cross}
+\newcommand{\nodeListCross}{\nodeList.cross}
+\newcommand{\nodeListCrossCount}{\nodeList.count}
+
+% move position
+
+\newcommand{\node}{\bm{n}}
+\newcommand{\nodePrevious}{\node_{p}}
+\newcommand{\nodeFNOR}{\node_t}
+
+\newcommand{\nodeVectorTC}{\Vector{\nodeFNOR\node}}
+\newcommand{\nodeVectorCT}{\Vector{\node\nodeFNOR}}
+
+% check river crossing
+
+\newcommand{\river}{\bm{r}}
+\newcommand{\riverVector}{\Vector{v_{\river}}}
+\newcommand{\riverList}{\bm{R}}
+\newcommand{\nodeBoundingBox}{b_{\Line{n}}}
+\newcommand{\riverEdgeBoundingBox}{b_{\Edge}}
+\newcommand{\boundingBoxInt}{b_{int.}}
+\newcommand{\nodeEdge}{e_n}
+\newcommand{\edgeBoxInt}{e_{int.}}
+
+% derive corridor
+\newcommand{\Corridor}{c}
+\newcommand{\CorridorLength}{\Corridor_l}
+\newcommand{\CorridorWidth}{\Corridor_w}
+\newcommand{\PointP}{\node_c}
+\newcommand{\Edge}{\Line{e}}
+\newcommand{\EdgeStart}{\Edge.start}
+\newcommand{\EdgeEnd}{\Edge.end}
+\newcommand{\EdgeParallel}{\Line{e_p}}
+\newcommand{\EdgeParallelA}{\Line{e_{p^1}}}
+\newcommand{\EdgeParallelB}{\Line{e_{p^2}}}
+\newcommand{\Distance}{d}
+\newcommand{\nodeVectorNP}{\Vector{\node\PointP}}
+\newcommand{\nodeVectorNV}{\Vector{\node\nodeFNOR}}
+\newcommand{\nodeLineNV}{\Line{\node\nodeFNOR}}
+\newcommand{\nodeLinePN}{\Line{\nodePrevious\node}}
+\newcommand{\nodeVectorNtP}{\Vector{\nodeFNOR\PointP}}
+\newcommand{\nodeLineNtNc}{\Line{\nodeFNOR\PointP}}
+\newcommand{\nodeLineWidthNtP}{\Width{\nodeFNOR\PointP}}
+\newcommand{\nodeInCorridor}{\node_{in}}
+\newcommand{\nodeInCorridorT}{\node_{in_t}}
+
+% derive corridor point
+\newcommand{\Scale}{scale}
+
+% figure fig:corridor
+
+\newcommand{\nodeVectorNNn}{\Vector{\node\node_{t}}}
+\newcommand{\nodeVectorNinNinn}{\Vector{\nodeInCorridor\nodeInCorridorT}}
+
+\newcommand{\vectorAvg}{\bm{v}_{avg}}
+\newcommand{\NxR}{\node_{n_{x}r}}
+\newcommand{\RxN}{\node_{r_{x}n}}
+% \include{preamble/preamble.math}
 
@@
 \title{Demers Cartogram with Rivers}
-
-% \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
-%         \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
 
@@
 
-\input{section/introduction}
+\section{Introduction and Motivation}
 
-\input{section/related}
+\begin{samepage}
 
-\input{section/data}
+    Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography,inoue2011New}.
+    Various styles of cartograms have been proposed and implemented for applications such as urban planning \cite{harris2018Mapping, arranz-lopez2021Enduser}, natural hazard forecasting \cite{pappenberger2019Cartograms, park2020Flood}, conservation and environmental planning \cite{galluzzi2018Mapping, rocchini2019Cartogramming}, political and social demographics \cite{breitzman2018Using, alieva2021How}, and decision-making for public health \cite{gao2020Visualising, sack2021Visualizing}.
 
-\input{section/main}
+    \begin{figure}[b!]
+        \centering
+        \includegraphics[width=\linewidth,keepaspectratio]{cover.png}
+        \caption{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
+        \label{fig:cover}
+    \end{figure}
 
-\input{section/evaluation}
+\end{samepage}
 
-\input{section/limitation}
+Among the four types of cartogram categorized in a survey by \citet{nusrat2016State} (See \Cref{sec:RelatedWork} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
+For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
+Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
+Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
+Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
+Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
+We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions.
+We hypothesize that introducing rivers improves the overall legibility of a cartogram.
+By \textit{legibility} we mean readability and ability to interpret the cartogram.
+To assess this hypothesis, we designed an experimental setup where participants engaged in correspondence and location tasks as part of a user study.
+To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
+We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
+We present a user study that demonstrates its effectiveness.
 
-\input{section/conclusion}
+Our contributions include:
 
-% \bibliographystyle{mslapa}
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
+    \item A novel hybrid layout algorithm that combines node positions with features such as rivers,
+    \item A user study evaluation of the technique with an application to EHRs.
+\end{itemize}
+
+The results of the user study indicate that rivers can improve the legibility of cartograms.
+One of the major challenges involved is developing a layout algorithm that handles different shapes.
+In other words, the hybrid layout algorithm is novel because it handles different types of elements: square representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.
+
+\section{Related Work}
+\label{sec:RelatedWork}
+
+This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world implementations of cartogram-based visualizations.
+
+\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms preserve topology, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
+Non-contiguous cartograms sacrifice topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
+
+\citet{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological.
+Each cartogram design may make various types of accuracy trade-offs between dimensions.
+We provide a comparison of these trade-offs as introduced by \citet{nusrat2016State} in \Cref{table:accuracy}.
+In addition, we include the Demers as it is the focus of our work.
+
+    {
+        \renewcommand{\arraystretch}{1.5}
+        \begin{table*}[!tb]
+            \centering
+            \resizebox{\textwidth}{!}{
+                \begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
+                    \hhline{~|*{4}{-}|~}
+                    \multicolumn{1}{c|}{\textbf{Literature}} &
+                    \textbf{Cartogram Type(s)} &
+                    \textbf{Geographic Region(s)} &
+                    \textbf{Number of Nodes} &
+                    \multicolumn{1}{c}{\textbf{Year}} \\
+                    \hline
+
+                    % \citet{auber2007Geographical} & \citep{auber2007Geographical} & US & 19 & 2007 \\
+                    % \hline
+                    \citet{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
+                    \hline
+                    \citet{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
+                    \hline
+                    \citet{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
+                    \hline
+                    \citet{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
+                    \hline
+                    \citet{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
+                    \hline
+                    \citet{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
+                    \hline
+                    \citet{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
+                    \hline
+
+                    % \multicolumn{1}{c|}{\textbf{Total unique papers: 51}} & 24&15&12& \multicolumn{1}{c}{}  \\
+                    % \hhline{~|*{3}{-}|~}
+
+                \end{tabulary}
+            }
+            \caption{Related work with non-contiguous cartogram-based visualizations.
+\textbf{Cartogram type} is the type of cartogram used.
+\textbf{Geographic region} is the geographic region depicted by the cartogram.
+\textbf{Number of nodes} is the number of nodes depicted in the cartogram.
+}
+            \label{table:region vs node}
+        \end{table*}
+    }
+
+\thumbnail{dorling}{
+    Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
+    A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
+    In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
+    Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
+    This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
+    This is important in our use case scenario involving EHRs.
+    Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}.
+}
+
+\thumbnail{rectangular}{
+    Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
+Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
+}
+
+\thumbnail{mosaic}{
+    Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+Some variants are able to preserve topological accuracy as well.
+}
+
+{
+    \renewcommand{\arraystretch}{1.5}
+    \begin{table}[!tb]
+        \centering
+        \resizebox{\columnwidth}{!}{
+            \begin{tabulary}{\columnwidth}{|*{5}{l|}}
+                \hhline{~|*{3}{-}|~}
+                \multicolumn{1}{c|}{} & \multicolumn{3}{c|}{\cellcolor{Mycolor2} \textbf{Accuracy}} \\
+                \hhline{~|*{4}{-}}
+                \multicolumn{1}{c|}{\textbf{Cartogram Variant}} &
+                \textbf{Statistical} &
+                \textbf{Geographical} &
+                \textbf{Topological} &
+                \textbf{Contiguity} \\
+                \hline
+                Contiguous & \cellcolor{Mycolor3}Variable & \cellcolor{Mycolor3}Variable & Accurate & Yes \\
+                \hline
+                Non-contiguous & \cellcolor{Mycolor5}\textcolor{white}{Accurate} & \cellcolor{Mycolor5}\textcolor{white}{Shape is accurate} &\cellcolor{Mycolor6}Inaccurate & No   \\
+                \hline
+                Rectangular & \cellcolor{Mycolor3}Variable & Shape is inaccurate & \cellcolor{Mycolor3}Variable & Yes   \\
+                \hline
+                Dorling & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
+                \hline
+                Demers & Accurate & \cellcolor{Mycolor6}Inaccurate & \cellcolor{Mycolor3}Variable & No  \\
+                \hline
+            \end{tabulary}
+        }
+        \caption{\coloredsquare[Mycolor3]{0.25cm}{0.25cm} Trade-off between dimensions.
+\coloredsquare[Mycolor6]{0.25cm}{0.25cm} Dimension sacrificed in order to improve \coloredsquare[Mycolor5]{0.25cm}{0.25cm} target dimension's accuracy.}
+        \label{table:accuracy}
+    \end{table}
+}
+
+\bobgraph{Peer-reviewed Applications: }\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
+\citet{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
+\citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citet{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US.
+See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.
+
+Our work extends the algorithm described by \citet{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms.
+Our work enhances that of \citet{tong2018Cartograms} in multiple ways.
+First, we introduce multiple features (rivers) into the layout, as opposed to a single river.
+Second, we make topological features dynamic and further improve legibility and geographical accuracy.
+By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm.
+In previous work, the river is static and serves merely as a boundary.
+Third, we improve the algorithm to resolve stalemates.
+Finally, the way we evaluate the cartograms is also different.
+\citet{tong2018Cartograms} count river crossings to evaluate error (a statistical metric).
+Here our focus is on readability, and thus we include a user study.
+
+\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
+The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
+National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
+\citet{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal.
+\citet{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.
+
+One of the disadvantages of Dorling and Demers cartograms is legibility.
+The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret.
+\citet{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.
+
+In this paper, we introduce a new type of topological feature, a river, as a constraint to compute the final layout, with the aim of improving the interpretation, readability, and accuracy of this class of cartograms.
+
+\section{Data Description}
+
+ {
+  \begin{figure}[b!]
+      \centering
+      \includegraphics[width=0.6\columnwidth]{ccg.png}
+      \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
+      \label{fig:ccg}
+  \end{figure}
+ }
+
+Obtaining heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
+The first step is to obtain both geospatial boundaries and EHR data.
+The second step is to preprocess the EHR data to remove empty and erroneous values.
+The final step is to transform the data into a format that is suitable for cartograms.
+Geospatial boundaries, or shapefiles, were obtained from the sources described here.
+
+\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
+The number of CCGs changes over time due to NHS reorganization.
+The most up-to-date shapefile is available from the Open Geography Portalx \cite{opengeographyportalxOpen}.
+We decided to use the CCG shapefile from 2020 at the time of writing, due to the absence of published public EHR data based on the latest CCG reorganizations that took place in 2021 and 2022.
+
+\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England.
+These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues.
+Although including smaller rivers is technically feasible, it may not increase the legibility of the cartogram.
+This is an open question for future work.
+
+We first obtain a relation ID by searching for a river, e.g.
+River Thames, on OpenStreetMap.
+The relation ID is used to construct a query (See \cref{obtaining-shapefile}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.
+
+After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust the projections and convert them into GeoJSON files.
+Finally, mapshaper \cite{blochMapshaper} is used to merge and convert the GeoJSON files into a TopoJSON file \cite{TopoJSON}.
+TopoJSON eliminates redundant coordinates in the data, improving the rendering speed of our implementation.
+See \Cref{table:pre-processing_result} on page \pageref{table:pre-processing_result} for the preprocessing result.
+We describe the one-time preprocessing steps in more detail in \Cref{app:pre-processing}.
+
+\bobgraph{EHR Data: }We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
+The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.
+Some datasets include:
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Under 75 mortality: cardiovascular disease, respiratory disease, liver disease, and cancer
+    \item Emergency hospital admission: stroke, alcohol-specific admission and readmission, coronary heart disease, re-admissions within 30 days of discharge, children with lower respiratory tract infections
+\end{itemize}
+
+For all datasets, a spreadsheet including the following is provided:
+
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Reporting period: Calendar year of registration
+    \item Period of coverage: Start and end date or reporting period
+    \item Breakdown: Organization type
+    \item ONS code: UK Office for National Statistics CCG code
+    \item Level: CCG Code
+    \item Level description: CCG Name
+    \item Gender
+    \item Indicator value: Directly standardized mortality rate
+    \item CI lower: lower 95\% confidence interval
+    \item CI upper: upper 95\% confidence interval
+    \item Denominator: The count of registered patients
+    \item Numerator: Number of deaths
+\end{itemize}
+
+Each CCG has a unique ONS code, which is used to link the CCG shapefile with the statistical data.
+
+\section{Demers Cartogram with Rivers}
+\label{sec:main}
+
+{
+    \begin{figure*}[tb!]
+        \centering
+        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{flowchart.png}
+        \caption{An overview of our hybrid layout algorithm incorporating rivers.
+See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail.
+See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
+            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
+        \label{fig:flowchart}
+    \end{figure*}
+}
+
+{
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=0.7\columnwidth,keepaspectratio]{river_resolution.png}
+        \caption{The resolution of rivers can be dynamically adjusted by the user.
+(A) shows River Thames at its original resolution with 10,170 edges.
+(B) shows the river at a reduced resolution of 49 edges.
+We further smooth the river by removing 19 vertices in dense areas, as shown in (C).
+The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+        \label{fig:river resolution}
+    \end{figure}
+}
+
+\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the hybrid cartogram layout process that includes rivers.
+
+\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel.
+We then load the river shapefiles and render the rivers.
+Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex.
+We do not need the original river resolution to incorporate them into cartograms.
+We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests.
+This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}.
+We further apply simplification by removing vertices that are too close to each other.
+The initialization procedure is a one-time process that can be saved for reuse.
+
+\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
+We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}.
+Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions.
+An increase in $ \nodeSize $ can cause the nodes to overlap.
+During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position.
+Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position.
+If a node oscillates across a river, we identify this as a stalemate.
+One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river.
+We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
+The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.
+
+\subsection{River Intersection Testing}
+\label{subsec:{River Intersection Testing}}
+
+The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}.
+When a node's position changes, we test if the node's trajectory intersects any segment of a river.
+See \Cref{alg:check river intersection}.
+A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
+Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
+We label these nodes, $ \NxR $.
+
+\subsection{Translating Rivers}
+\label{subsec:{Translating Rivers}}
+
+For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $.
+Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
+We then use a heuristic to translate $ \river $ using the average vector of node intersection $$ \vectorAvg = \sum_{i=1}^{\nodeTopologicalError} \frac{\Vector{\node_i\node_{i_t}}}{\nodeTopologicalError} $$, where $ \nodeTopologicalError $ is the number of nodes intersecting the river.
+This step intends to create space for the next iteration of node translation without crossing a river.
+The detailed procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
+
+When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $.
+As a heuristic, we also translate these nodes by $ \vectorAvg $.
+The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together.
+In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
+Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm.
+We can use this to ensure smooth transitions between iterations.
+
+\subsection{Process Stalemates}
+\label{subsec:{Process Stalemates}}
+
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion.
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
+All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
+The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.
+
+{
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
+        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+        \label{fig:flowchart-stalemate}
+    \end{figure}
+   }
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{stalemate.png}
+    \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
+The node is translated back to its previous position (iteration 2).
+A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
+    \label{fig:stalemate}
+\end{figure}
+}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{corridor.png}
+    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A).
+To address this, we derive a corridor (orange rectangle in (E)) based on $ \node $ and $ \nodeFNOR $.
+All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
+For clarity in the illustration, we place nodes sparsely in this figure.}
+    \label{fig:corridor}
+\end{figure}
+}
+
+\subsection{Terminating the Algorithm}
+\label{subsec:{Terminating the Algorithm}}
+
+The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
+We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
+$$
+
+where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
+Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
+We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers.
+We express $ \nodeTopologicalError $ in the normalized range $ \nodeTopologicalError \in [0,1] $, where $ \nodeTopologicalErrorMax $ is the case where all nodes cross a river.
+
+When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
+Every node remains on the same side of the river as its centroid.
+
+\subsection{User Options}\label{subsec:{UserOptions}}
+
+\Cref{fig:overview} presents an overview of the application, including user options. The user can adjust the following parameters:
+
+\bobgraph{Rendering Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
+
+\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR attributes or set to uniform.
+
+\bobgraph{Overlap Removal Speed:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.
+
+\bobgraph{Maximum Cartographic Error:} The user can adjust the maximum cartographic error, $ \nodeCartographicErrorMax $, which is used to terminate the algorithm.
+
+\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers.
+Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
+Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.
+
+\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations.
+A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.
+The default corridor length is three times the max node size.
+
+\bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
+Similarly, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
+
+    {
+        \begin{figure*}[tb!]
+            \centering
+            \includegraphics[width=0.7\textwidth,keepaspectratio]{UI.png}
+            \caption{A screenshot of the user interface.
+User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers.
+Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
+See \Cref{subsec:{UserOptions}} for more on user options.
+In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
+            \label{fig:overview}
+        \end{figure*}
+    }
+
+\newcommand{\pCount}{24 }
+\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}
+
+We conduct a user study to evaluate the effectiveness of our approach.
+We chose this type of evaluation because legibility is a human-centred characteristic.
+Many different types of statistical error metrics have been evaluated in previous work \cite{nusrat2016State}, however, our focus is more user-centric in nature.
+
+\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}
+
+We formulate the following hypotheses to motivate our user study:
+
+\textbf{H1:} The introduction of rivers can improve the legibility and recognizability of cartograms.
+
+\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.
+
+\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.
+
+We test these hypotheses using location-based tasks.
+Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
+See \Cref{fig:task}.
+
+\subsection{User Study Variables}\label{subsec:{User Study Variables}}
+
+In this section, we discuss the variables of our user study.
+
+\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram.
+We set this parameter to render rivers or not.
+
+\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants.
+\textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.
+
+\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find.
+The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
+
+\subsection{User Study Design}\label{subsec:{User Study Design}}
+
+In this section we describe the user study participants, datasets, and the experimental procedure.
+
+\bobgraph{Participants:} We recruited \pCount participants.
+
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Gender: 10 females and 14 males
+    \item Age Group: 18-24 (16), 25-29 (8)
+    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
+\end{itemize}
+
+\bobgraph{Datasets:} We used the following EHR datasets for our evaluation:
+
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Population
+    \item Under 75 mortality from cardiovascular disease
+    \item Emergency admissions for alcohol-related liver disease
+    \item Alcohol-specific admission and readmission
+\end{itemize}
+
+135 CCGs are rendered on the screen as a choropleth on the left.
+We then generated another view on the right using cartograms.
+The color of both visual designs is mapped to the dataset.
+See \Cref{fig:task} for an example.
+
+    {
+        \begin{figure}[tb!]
+            \centering
+            \includegraphics[width=\columnwidth,keepaspectratio]{task.png}
+            \caption{A sample location task for participants.
+The left shows the choropleth map, and the right shows the corresponding cartogram.
+Both images show the three longest rivers in England, and color is mapped to CCG population.
+The target CCG blinks on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
+In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated.
+}
+            \label{fig:task}
+        \end{figure}
+    }
+
+\subsubsection{Procedure}
+\label{subsec:{Procedure}}
+
+The user study is designed to be conducted online, and participants are asked to complete the tasks using their own hardware. The user study includes four parts:
+
+\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos.
+The instructions are designed to help participants understand the concepts used in the tasks.
+Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+
+\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
+These tasks also serve as a quality test to see whether users are taking the study seriously.
+A demonstration of three sample tasks is also included in the instructional video.
+
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 CCGs.
+Response and reaction time are recorded.
+These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+
+\textbf{P4:} The participants are asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
+
+The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.
+
+\subsection{User Study Results}
+\label{subsec:{User Study Results}}
+
+{
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
+    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
+    \label{fig:task-acc}
+\end{figure}
+}
+
+In this section, we delve into the analysis of the preliminary user study conducted.
+Our study involved the recruitment of \pCount participants, each tasked with completing 16 location-related tasks.
+As a result, we collected a total of 384 responses.
+However, we excluded 24 responses from our analysis as they took over 60 seconds to complete.
+We hypothesize that these prolonged response times indicate potential distractions experienced by participants during the study.
+
+\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
+The chart shows that the introduction of rivers has improved the accuracy of two CCGs.
+The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002.
+This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.
+
+\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs.
+The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.
+The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38.
+However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.
+
+    {
+        \begin{figure}[t!]
+            \centering
+            \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
+            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
+            \label{fig:task-rt}
+        \end{figure}
+    }
+
+\bobgraph{Participant Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
+\begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
+    \item 87.5\% of participants agree that including rivers is useful.
+    \item 83.3\% of participants agree that rivers increase the legibility of a cartogram.
+    \item 95.8\% of participants agree that including rivers makes cartograms easier to understand.
+    \item 62.5\% of participants agree that including rivers makes CCGs easier to locate.
+    \item 87.5\% of participants agree that including rivers adds value to the standard cartogram.
+\end{enumerate}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{likert.png}
+    \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
+    \label{fig:likert}
+\end{figure}
+}
+
+The qualitative results clearly support our hypothesis H1 that rivers generally increase the legibility of cartograms.
+
+\section{Limitations and Future Work}
+\label{sec:{Limitations and Future Work}}
+
+In this section, we discuss some limitations of our work and future research directions.
+Because this is a new concept, it opens the door for many future research directions.
+The first limitation is the color map that we use to depict the data in our cartograms.
+We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
+However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
+In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
+See \Cref{fig:extreme} for an example.
+We plan to explore the impact of different color maps on the legibility of cartograms in future work.
+
+    {
+        \begin{figure}[tb!]
+            \centering
+            \includegraphics[width=0.7\columnwidth,keepaspectratio]{extreme.png}
+            \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
+            \label{fig:extreme}
+        \end{figure}
+    }
+
+Another limitation is the overlap removal algorithm (FNOR) we use.
+We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
+Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
+
+Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
+We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
+We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
+We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
+Furture work also includes adding rivers to dorling-style cartograms.
+
+\section{Conclusions}\label{sec:{Conclusions}}
+
+In this paper, we present a novel hybrid layout approach to generate non-contiguous cartograms with rivers.
+We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers.
+We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
+We also discuss the limitations of our work and future research directions.
+
 \urlstyle{same}
@@
 
-\input{section/appendix}
+\section{Appendix}
+\appendix
+
+\section{Obtaining Shapefile}
+\label{obtaining-shapefile}
+\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
+    relation(2263653);>>;
+    // River Great Ouse: 2798097
+    // River Trent: 2863468
+    out skel;
+\end{lstlisting}
+\section{Pre-processing Shapefiles}
+\label{app:pre-processing}
+Shapefiles from different sources are likely to be incompatible.
+In our case, the NHS CCG shapefile is incompatible with the river shapefiles.
+The major reason for the incompatibility is the coordinate reference system (CRS).
+The CRS of the CCG shapefile is EPSG:27700 (OSGB36 - British National Grid).
+The CRS of the river shapefiles is EPSG:4326 (WGS84 - World Geodetic System).
+Here, we provide some pre-processing steps using QGIS (version: 3.26.0-Buenos Aires) \cite{qgisWelcome} to handle the incompatibility and reduce shapefile size to improve performance.
+
+\subsection{Import Shapefiles into QGIS}
+We first load all three river shapefiles into QGIS \Cref{fig:import_rivers}, followed by the CCG shapefile \Cref{fig:import_ccgs}.
+
+{
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{import_rivers.png}
+    \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
+    \label{fig:import_rivers}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{import_ccgs.png}
+    \caption{QGIS interface, with all NHS CCGs imported.}
+    \label{fig:import_ccgs}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{unify_crs.png}
+    \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
+    \label{fig:unify_crs}
+\end{figure}
+}
+
+\subsection{Export Shapefiles in GeoJSON and Unify the Coordinate Reference System (CRS)}
+
+We then use QGIS to unify the CRS, and export both layers in GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.
+
+{
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{export_rivers.png}
+    \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
+    \label{fig:export_rivers}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{export_ccgs.png}
+    \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
+    \label{fig:export_ccgs}
+\end{figure}
+
+}
+
+\subsection{Merge Shapefiles and Reduce File Size}
+
+We then merge two layers into one layer, and export it in the TopoJSON format using Mapshaper \cite{blochMapshaper}.
+Mapshaper also supports the simplification of GeoJSON shapefiles.
+See \Cref{fig:export_topojson}.
+
+{
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\columnwidth]{export_topojson.png}
+    \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
+    \label{fig:export_topojson}
+\end{figure}
+}
+
+\subsection{Pre-processing Result}
+
+\Cref{table:pre-processing_result} shows the preprocessing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+
+    {
+        \renewcommand{\arraystretch}{1.5}
+        \begin{table}[H]
+            \centering
+            \resizebox{\columnwidth}{!}{
+                \begin{tabulary}{\columnwidth}{|*{4}{l|}}
+                    \hhline{~|*{3}{-}}
+                    \multicolumn{1}{c|}{\textbf{Shapefile}} &
+                    \cellcolor{Mycolor2}\textbf{Original} &
+                    \cellcolor{Mycolor2}\textbf{GeoJSON} &
+                    \cellcolor{Mycolor2}\textbf{TopoJSON} \\
+                    \hline
+                    Rivers & 2.0 MB (GeoJSON) & 1.4 MB & -  \\
+                    \hline
+                    NHS CCGs & 46.6 MB (.shp, Esri vector shapefile) & 140.2 MB & -  \\
+                    \hline
+                    Merged & - & - & 16.3 MB \\
+                    \hline
+                \end{tabulary}
+            }
+            \caption{The file size is reduced by 88.5\% from the original size.}
+            \label{table:pre-processing_result}
+        \end{table}
+    }
+
+\clearpage
+
+\section{Procedure: TranslateNode}
+
+\begin{algorithm}[H]
+    \caption{Procedure to translate node positions.}\label{alg:TranslateNode}
+
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
+
+    \textbf{Output:} \\
+    $ \nodeTopologicalError \gets $ the number of nodes crossing the river in the input \\
+
+    \textbf{Global variables:} \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river  \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateNode}{$ \nodeList $}
+        \State $ \nodeTopologicalError \gets 0 $
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+
+        \State $ \node(x,y) \gets \nodeFNOR(x,y) $
+
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $} $ = True $}
+        \State $ \nodeCross ++ $
+        \State $ \nodeTopologicalError ++ $
+
+        \If{$ \nodeCross < \stalemateMax $}
+        \State \hskip-2em \Comment{Translate back to previous position}
+        \State $ \node(x,y) \gets \nodePrevious(x,y) $
+
+        \Else
+
+        \State \Call{ProcessStalemate}{$ \node, \nodeFNOR $}
+        \State $ \nodeCross \gets 0 $ \Comment{Reset counter}
+
+        \EndIf
+
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+        \State \Return{$ \nodeTopologicalError $}
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Procedure: UpdateLayout}
+
+\begin{algorithm}[H]
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:main}.}\label{alg:UpdateLayout}
+    \textbf{Global variables:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions with the following properties: \\
+    \-\hspace{1em}  $\nodeListCross \gets $ the number of $ \node $ in $ \nodeList $ that crosses a river \\
+
+    $ \nodeSize \gets $ the initial size of all nodes \\
+    $ \nodeCartographicError \gets $ the average cartographic error of all nodes \\
+    $ \nodeCartographicErrorMax \gets $ the maximum cartographic error of all nodes \\
+    $ \stalemateMax \gets $ the maximum number of iterations indicating a stalemate \\
+
+    \textbf{Local variables:} \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodeCross \gets $ the number of times that $ \node $ crosses a river \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{UpdateLayout}{}
+
+        \State $ \nodeSize \gets 1 $
+        \While{$ \nodeCartographicError < \nodeCartographicErrorMax $}
+        \State $ \nodeListCross \gets 1 $ \Comment{Trigger the while loop}
+        \While{$ \nodeListCross > 0 $}
+
+        \State $ \nodeListCross \gets 0 $
+        \State $ \nodeList \gets $ RemoveOverlap ($ \nodeList $)
+        \State \Call{TranslateRiver}{$ \nodeList $}
+        \State $\nodeListCross \gets $ \Call{TranslateNode}{$ \nodeList $}
+
+        \EndWhile
+
+        \State $ \nodeSize ++ $
+
+        \EndWhile
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Procedure: TranslateRiver}
+
+\begin{algorithm}[H]
+    \caption{Procedure to translate rivers.}\label{alg:TranslateRiver}
+    \textbf{Input:} \\
+    $ \nodeList \gets $ a list of $ \node $ representing regions \\
+
+    \textbf{Local variables:} \\
+    $ \riverList \gets $ a list of $ \river $ representing river features \\
+    $ \node \gets $ a node is an object with the following properties: \\
+    \-\hspace{1em} $ \node.x, \node.y, $ or $ \node(x,y) \gets $ the x and y coordinates of $ \node $ \\
+    \-\hspace{1em} $ \nodePrevious \gets $ the previous position of $ \node $ \\
+    \-\hspace{1em} $ \nodeFNOR \gets $ the translated position of $ \node $ \\
+    $ \nodeTopologicalError \gets $ the number of nodes intersecting $ \river $  \\
+
+    \begin{algorithmic}[1]
+        \Procedure{TranslateRiver}{$ \nodeList $}
+        \ForEach {$ \river \in \riverList $}
+        \State $ \nodeTopologicalError \gets 0 $
+        \State $ \riverVector \gets (0,0) $ \Comment{Hold the sum of vectors $ \nodeVectorCT $}
+        \ForEach {$ \node \in \nodeList $}
+
+        \If {$ \node(x,y) \neq \nodeFNOR(x,y) $}
+        \If{ \Call{TestIntersection}{$ \nodeFNORLine $, $ \river $} $ = True $}
+        \State $ \riverVector \gets \riverVector + \nodeVectorCT$
+        \State $ \nodeTopologicalError ++ $
+        \EndIf
+
+        \EndIf
+
+        \EndFor
+
+        \State Translate river $ \river $ by the average vector $ \frac{\riverVector}{\nodeTopologicalError} $
+        \EndFor
+
+        \EndProcedure
+    \end{algorithmic}
+\end{algorithm}
+
+\section{Procedure: ProcessStalemate}
+
+% \begin{noindent}
+
+    \begin{algorithm}[H]
+        \caption{Procedure to derive a corridor to resolve stalemate.
+We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}
+
+        \textbf{Input:} \\
+        $ \node \gets $ the node used to derive the corridor \\
+
+        \textbf{Global variables:} \\
+        $ \CorridorLength \gets $ the length of a corridor \\
+        $ \CorridorWidth \gets $ the width of a corridor \\
+
+        \textbf{Local variables:} \\
+        $ \Corridor \gets $ the corridor \\
+        $ \PointP \gets $ the point extending $ \nodeVectorTC $ such that $ \nodeLineWidthNtP = \CorridorLength $\\
+        $ \EdgeParallelA, \EdgeParallelB \gets $ the edges parallel to $ \nodeLineNtNc $ \\
+        $ corridor \gets $ a rectangle formed by $ \EdgeParallelA $ and $ \EdgeParallelB $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{ProcessStalemate}{$ \node $}
+                \State $ \node(x,y) \gets \nodePrevious(x,y) $
+
+                \State $ \PointP \gets $ \Call{DerivePoint}{$ \nodeVectorTC $, $ \CorridorLength $}
+
+                \State $ \EdgeParallelA \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ \frac{\CorridorWidth}{2} $}
+
+                \State $ \EdgeParallelB \gets $ \Call{DeriveParallelEdge}{$ \nodeVectorNtP $, $ -\frac{\CorridorWidth}{2} $}
+
+                \State $ \Corridor \gets
+                    \begin{bmatrix}
+                        \EdgeParallelA.start &
+                        \EdgeParallelA.end \\
+
+                        \EdgeParallelB.start &
+                        \EdgeParallelB.end \\
+                    \end{bmatrix} $
+
+                \ForEach{$ \nodeInCorridor $ inside $ \Corridor $}
+
+                    \State $ \Vector{\nodeInCorridor ~\nodeInCorridorT} = \nodeVectorTC $
+
+                    \State $ \nodeInCorridor(x,y) \gets \nodeInCorridorT(x,y) $
+
+                \EndFor
+            \EndProcedure
+        \end{algorithmic}
+    \end{algorithm}
+
+    %\end{noindent}
+\section{Procedure: TestIntersection}
+% \begin{noindent}
+    \begin{algorithm}[H]
+        \caption{Procedure to test if a node's translation path, $ \nodeLineNV $ intersects a river.}\label{alg:check river intersection}
+        \textbf{Input:} \\
+        $ \nodeLineNV \gets $ the node's translation path \\
+        $ \river \gets $ a river feature \\
+
+        \textbf{Output:} \\
+        Returns $ True $ if the node crosses a river. \\
+
+        \textbf{Local variables:} \\
+        $ \nodeBoundingBox, \riverEdgeBoundingBox \gets $ the bounding boxes for $ \nodeLineNV $ and $ \Edge $ \\
+        $ \Edge \gets $ an edge of $ \river $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{TestIntersection}{$ \nodeLineNV $, $ \river $}
+
+            \ForEach{$ \Edge \in \river $}
+                \State $ \nodeBoundingBox \gets $ GetBoundingBox ($ \nodeLineNV $)
+                \State $ \riverEdgeBoundingBox \gets $ GetBoundingBox ($ \Edge $)
+
+                \If{$ \nodeBoundingBox ~intersect~ \riverEdgeBoundingBox = True $}
+                    \State \Return{$ \nodeLineNV ~intersect~ \Edge $}
+                    \EndIf
+            \EndFor
+
+            \State \Return{$ False $}
+            \EndProcedure
+        \end{algorithmic}
+    \end{algorithm}
+    %\end{noindent}
+
+\section{Procedure: DerivePoint}
+
+% \begin{noindent}
+
+    \begin{algorithm}[H]
+        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
+        \textbf{Input:} \\
+        $ \Edge \gets $ the edge used to derive the new point \\
+        $ \Distance \gets $ the distance between $ \PointP $ and $ \EdgeStart $ \\
+
+        \textbf{Output:} \\
+        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\
+
+        \textbf{Local variables:} \\
+        $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
+
+        \begin{algorithmic}[1]
+            \Procedure{DerivePoint}{$ \Edge $, $ \Distance $}
+                \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
+
+                \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
+
+                \State $ \PointP.x \gets \frac{\dx}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
+
+                \State $ \PointP.y \gets \frac{\dy}{\sqrt{\dx^2 + \dy^2}} \cdot \Distance $
+
+            \State \Return{$ \PointP $}
+
+            \EndProcedure
+
+        \end{algorithmic}
+    \end{algorithm}
+
+%\end{noindent}
+\vfill\eject
+\section{Procedure: DeriveParallelEdge}
+
+% \begin{noindent}
+
+\begin{algorithm}[H]
+    \caption{Procedure to derive an edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.}\label{alg:derive corridor edge}
+
+    \textbf{Input:} \\
+    $ \Edge \gets $ the edge used to derive the parallel edge $ \EdgeParallel $ \\
+    $ \Distance \gets $ the shortest distance between $ \Edge $ and $ \EdgeParallel $ \\
+
+    \textbf{Output:} \\
+    An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $. \\
+
+    \textbf{Local variables:} \\
+    $ \dx, \dy \gets $ the differences in $ x, y $ for $ \EdgeStart $ and $ \EdgeEnd $ \\
+    $ \Scale \gets $ the scale of $ \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $ \\
+
+    \begin{algorithmic}[1]
+        \Procedure{DeriveParallelEdge}{$ \Edge $, $ \Distance $}
+            \State $ \dx \gets \EdgeStart.x - \EdgeEnd.x $
+
+            \State $ \dy \gets \EdgeStart.y - \EdgeEnd.y $
+
+            \State $ \Scale \gets \frac{\Distance}{\sqrt{\dx^2 + \dy^2}} $
+
+            \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \EdgeStart.x $
+
+            \State $ \EdgeParallel.start.y \gets \Scale \cdot \dx + \EdgeStart.y $
+
+            \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \EdgeEnd.x $
+
+            \State $ \EdgeParallel.end.y \gets \Scale \cdot \dx + \EdgeEnd.y $
+
+        \State \Return{$ \EdgeParallel $}
+
+        \EndProcedure
+
+    \end{algorithmic}
+\end{algorithm}
+
+%\end{noindent}
 

### 2023-08-18 paper/main.tex

@@
 \documentclass[Afour,sagev,times]{sagej}
+
+\newcommand{\new}[1]{\textcolor{blue}{#1}}
 
@@
     Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique.
-A relative of the Dorling cartogram, the Demers cartogram, utilizing squares instead of circles to represent regions.
+\new{As a variant of} the Dorling cartogram, the Demers cartogram utilizes squares instead of circles to represent regions.
 This alternative approach allows for a more intuitive comparison of regions, utilizing screen space more efficiently.
@@
 One of the major challenges involved is developing a layout algorithm that handles different shapes.
-In other words, the hybrid layout algorithm is novel because it handles different types of elements: square representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve stalemate situations introduced by rivers while minimizing error.
+In other words, the hybrid layout algorithm is novel because it handles different types of elements: square node representing regions and polylines representing rivers.
+Another challenge we overcome in developing the algorithm is to resolve \new{stalemate situations where nodes become congested due to constraints imposed by rivers, while ensuring error minimization.}
 
@@
                     \hline
-                    \citet{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
-                    \hline
-                    \citet{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
+                    \citet{nusrat2020Recognition} & Contiguous, Dorling & US & 49 & \citeyear{nusrat2020Recognition} \\
+                    \hline
+                    \citet{nickel2022Multicriteria} & Non-contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
                     \hline
@@
 \thumbnail{dorling}{
-    Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
-    A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
+    Dorling cartograms, as a variant of non-contiguous cartograms, generally do not preserve geography and topology.
+    A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle area \cite{dorling2011Area}.
     In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
@@
 
-Obtaining heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
+Processing heterogeneous data can be challenging, especially when an EHR dataset is involved, because the data comes from multiple sources \cite{wang2021EHR}.
 The first step is to obtain both geospatial boundaries and EHR data.
@@
 If a node oscillates across a river, we identify this as a stalemate.
-One iteration of the layout ends when 1) no node overlap is present; 2) no nodes cross a river.
+One iteration of the layout ends when 1) no node overlap is present; \new{AND} 2) no nodes cross a river.
 We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
@@
 If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion.
-A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelA \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelB \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
 All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
@@
             \centering
-            \includegraphics[width=0.7\textwidth,keepaspectratio]{UI.png}
+            \includegraphics[width=0.9\textwidth,keepaspectratio]{UI.png}
             \caption{A screenshot of the user interface.
@@
 The instructions are designed to help participants understand the concepts used in the tasks.
-Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+Instructional videos are available at \url{https://www.youtube.com/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
 
@@
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the case of river-river intersections and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
+We would like to explore the case of river-river intersections \new{(or confluence)} and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
 We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
@@
 
-                \State $ \EdgeParallel.start.x \gets \Scale \cdot -\dy + \EdgeStart.x $
+                \State $ \EdgeParallel.start.x \gets \Scale \cdot (-\dy) + \EdgeStart.x $
 
@@
 
-                \State $ \EdgeParallel.end.x \gets \Scale \cdot -\dy + \EdgeEnd.x $
+                \State $ \EdgeParallel.end.x \gets \Scale \cdot (-\dy) + \EdgeEnd.x $
 

### 2023-08-21 paper/main.tex
chore: increase figure size
draft: IV reviews

@@
     However, a drawback of the Dorling cartogram and its variants lies in the potential displacement of regions from their original positions, ultimately compromising legibility, readability, and accuracy.
-To tackle this limitation, we propose a novel hybrid cartogram layout algorithm that incorporates dynamic topological elements, such as rivers, into Demers cartograms.
+    To tackle this limitation, we propose a novel hybrid cartogram layout algorithm that incorporates topological elements, such as rivers, into Demers cartograms.
     The presence of rivers significantly impacts both the layout and visual appearance of the cartograms.
@@
 Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
-Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
+\new{Also, the comparison of magnitudes becomes an area estimation task, which is effective for a numeric data encoding} \cite{munzner2014Visualization}.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
-We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions.
+Standard Demers cartograms are composed of square nodes that \new{representing geographic enumeration units}.
+As such, this can reduce their legibility.
+We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes \new{representing geographic enumeration units, in this case, a Clinical Commissioning Group (CCG)}.
 We hypothesize that introducing rivers improves the overall legibility of a cartogram.
@@
                 \textbf{Geographic region} is the geographic region depicted by the cartogram.
-\textbf{Number of nodes} is the number of nodes depicted in the cartogram.
+                \textbf{Number of nodes} is the number of nodes \new{(representing geographic enumeration units)} depicted in the cartogram.
             }
@@
     In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
-    Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
+    Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
+    Here we focus on Demers cartograms as we use squares to depict regions.
     This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
@@
 
-\bobgraph{Peer-reviewed Applications: }\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\bobgraph{Peer-reviewed Applications: } \new{There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work}.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
 \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
 \citet{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China.
-\citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
+\citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by CCG regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy.
 \citet{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US.
@@
 
-\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news.
+\bobgraph{Cartograms in Media: }Cartograms \new{are an engaging visual representation and therefore they are a} popular choice of representation in covering various topics by the media.
 The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
@@
         \centering
-        \includegraphics[width=0.7\columnwidth,keepaspectratio]{river_resolution.png}
+        \includegraphics[width=0.8\columnwidth,keepaspectratio]{river_resolution.png}
         \caption{The resolution of rivers can be dynamically adjusted by the user.
@@
     \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
-        \caption{A flowchart illustration of stalemate processing. See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+    \caption{A flowchart illustration of stalemate processing.
+    See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
     \label{fig:flowchart-stalemate}
@@
 where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
+\new{For a detailed derivation of the formula, refer to the work by \citet{alam2015Quantitative}.}
 Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
@@
 
-\Cref{fig:overview} presents an overview of the application, including user options. The user can adjust the following parameters:
+\Cref{fig:overview} presents an overview of the application, including user options.
+The user can adjust the following parameters:
 
@@
 We test these hypotheses using location-based tasks.
-Given a standard choropleth on the left side of the screen and the equivalent cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
+Given a standard choropleth on the left side of the screen and the \new{corresponding} cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
 See \Cref{fig:task}.
@@
 In this section we describe the user study participants, datasets, and the experimental procedure.
+The study was approved by the University of Nottingham's Research Ethics Committee (Ref: 2021-2022-001).
 
@@
 
-\subsubsection{Procedure}
-\label{subsec:{Procedure}}
-
-The user study is designed to be conducted online, and participants are asked to complete the tasks using their own hardware. The user study includes four parts:
+\subsection{Procedure}\label{subsec:{Procedure}}
+
+\new{Due to pandemic restrictions, we designed the user study to be carried out online.
+Participants are expected to use their personal hardware for the tasks.
+However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.}
+The user study includes four parts:
 
@@
     \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
-    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
+    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants.
+        The x-axis shows the target CCGs.}
     \label{fig:task-acc}
@@
             \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
-            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants. The x-axis shows the target CCGs.}
+            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants.
+                The x-axis shows the target CCGs.}
             \label{fig:task-rt}
@@
 Because this is a new concept, it opens the door for many future research directions.
-The first limitation is the color map that we use to depict the data in our cartograms.
+
+\noindent\textbf{Color map choice:} The first limitation is the color map that we use to depict the data in our cartograms.
     We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
@@
     In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
-See \Cref{fig:extreme} for an example.
-We plan to explore the impact of different color maps on the legibility of cartograms in future work.
+    See \Cref{fig:extreme} for an example. We plan to explore the impact of different color maps on the legibility of cartograms in future work.
 
@@
             \centering
-            \includegraphics[width=0.7\columnwidth,keepaspectratio]{extreme.png}
+            \includegraphics[width=0.8\columnwidth,keepaspectratio]{extreme.png}
             \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
@@
 
-Another limitation is the overlap removal algorithm (FNOR) we use.
+\noindent\textbf{Overlap removal algorithm choice:} Another limitation is the algorithm (FNOR) we use.
 We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
 Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
-
-Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
+\new{ When the quantity of nodes and features increases, generating the optimal layout demands several hundred iterations.}
+
+\noindent\textbf{Generalizability:} Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the case of river-river intersections \new{(or confluence)} and testing out more geographies such as the U.S. and Europe, which have more complex rivers.
+We would like to explore the case of river-river intersections \new{(or confluence)} and testing out more geographies such as the U.S.
+and Europe, which have more complex rivers.
 We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
-Furture work also includes adding rivers to dorling-style cartograms.
+
+\noindent\textbf{Improved User Study:} \new{Due to pandemic restrictions, the user study did not take place in a controlled environment.
+In the future, we aim to execute a more structured study, ensuring a controlled environment and standardized hardware for participants.
+}
 
@@
 
-We then use QGIS to unify the CRS, and export both layers in GeoJSON format. See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}. The unified layer is shown in \Cref{fig:unify_crs}.
+We then use QGIS to unify the CRS, and export both layers in GeoJSON format.
+See \Cref{fig:export_rivers} and \Cref{fig:export_ccgs}.
+The unified layer is shown in \Cref{fig:unify_crs}.
 
@@
 
-\Cref{table:pre-processing_result} shows the preprocessing result. The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+\Cref{table:pre-processing_result} shows the preprocessing result.
+The reduction in file size is significant and greatly reduces the initialization time of our implementation.
 
@@
         \textbf{Output:} \\
-        An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $. \\
+        An edge, $ \EdgeParallel $, that is parallel to $ \Edge $ with a distance of $ \Distance $.
+\\
 
@@
 \begin{algorithm}[H]
-    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers. See \Cref{sec:main}.}\label{alg:UpdateLayout}
+    \caption{Procedure to adjust river positions, remove node overlap and prevent nodes from crossing rivers.
+        See \Cref{sec:main}.}\label{alg:UpdateLayout}
     \textbf{Global variables:} \\
@@
         \textbf{Output:} \\
-        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\
+        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $.
+\\
 
@@
         \textbf{Output:} \\
-        Returns $ True $ if the node crosses a river. \\
+        Returns $ True $ if the node crosses a river.
+\\
 

### 2023-09-14 paper/main.tex
draft: blue text
1

@@
 Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
-\new{Also, the comparison of magnitudes becomes an area estimation task, which is effective for a numeric data encoding} \cite{munzner2014Visualization}.
+\new{In other words, Demers cartograms are useful when a data dimension is not describing the geography of the region it represents but is tied to something else, for example the health of its population.}
+Also, the comparison of magnitudes becomes an area estimation task, which is effective for a numeric data encoding \cite{munzner2014Visualization}.
+\new{See \citea{nickel2022Multicriteria} for a more complete description of the advantages that Demers cartograms offer.}
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.

### 2023-09-26 paper/main.tex
draft: review

--- 2023-08-21 paper/main.tex
@@
     Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique.
-    \new{As a variant of} the Dorling cartogram, the Demers cartogram utilizes squares instead of circles to represent regions.
+    As a variant of the Dorling cartogram, the Demers cartogram utilizes squares instead of circles to represent regions.
     This alternative approach allows for a more intuitive comparison of regions, utilizing screen space more efficiently.
@@
 Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes.
-\new{Also, the comparison of magnitudes becomes an area estimation task, which is effective for a numeric data encoding} \cite{munzner2014Visualization}.
+In other words, Demers cartograms are useful when a data dimension is not describing the geography of the region it represents but is tied to something else, for example the health of its population.
+Also, the comparison of magnitudes becomes an area estimation task, which is effective for a numeric data encoding \cite{munzner2014Visualization}.
+See \citea{nickel2022Multicriteria} for a more complete description of the advantages that Demers cartograms offer.
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-Standard Demers cartograms are composed of square nodes that \new{representing geographic enumeration units}.
+Standard Demers cartograms are composed of square nodes that representing geographic enumeration units.
 As such, this can reduce their legibility.
-We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes \new{representing geographic enumeration units, in this case, a Clinical Commissioning Group (CCG)}.
+We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geographic enumeration units, in this case, a Clinical Commissioning Group (CCG).
 We hypothesize that introducing rivers improves the overall legibility of a cartogram.
@@
 In other words, the hybrid layout algorithm is novel because it handles different types of elements: square node representing regions and polylines representing rivers.
-Another challenge we overcome in developing the algorithm is to resolve \new{stalemate situations where nodes become congested due to constraints imposed by rivers, while ensuring error minimization.}
+Another challenge we overcome in developing the algorithm is to resolve stalemate situations where nodes become congested due to constraints imposed by rivers, while ensuring error minimization.
 
@@
                 \textbf{Geographic region} is the geographic region depicted by the cartogram.
-                \textbf{Number of nodes} is the number of nodes \new{(representing geographic enumeration units)} depicted in the cartogram.
+                \textbf{Number of nodes} is the number of nodes (representing geographic enumeration units) depicted in the cartogram.
             }
@@
 
-\bobgraph{Peer-reviewed Applications: } \new{There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work}.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\bobgraph{Peer-reviewed Applications: } There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
 \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
@@
 
-\bobgraph{Cartograms in Media: }Cartograms \new{are an engaging visual representation and therefore they are a} popular choice of representation in covering various topics by the media.
+\bobgraph{Cartograms in Media: }Cartograms are an engaging visual representation and therefore they are a popular choice of representation in covering various topics by the media.
 The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
@@
 If a node oscillates across a river, we identify this as a stalemate.
-One iteration of the layout ends when 1) no node overlap is present; \new{AND} 2) no nodes cross a river.
+One iteration of the layout ends when 1) no node overlap is present; AND 2) no nodes cross a river.
 We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
@@
 where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
-\new{For a detailed derivation of the formula, refer to the work by \citet{alam2015Quantitative}.}
+For a detailed derivation of the formula, refer to the work by \citet{alam2015Quantitative}.
 Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
@@
 
-We conduct a user study to evaluate the effectiveness of our approach.
+We conduct a \new{within-subject} user study to evaluate the effectiveness of our approach.
 We chose this type of evaluation because legibility is a human-centred characteristic.
@@
 We test these hypotheses using location-based tasks.
-Given a standard choropleth on the left side of the screen and the \new{corresponding} cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
+Given a standard choropleth on the left side of the screen and the corresponding cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
 See \Cref{fig:task}.
@@
 \bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. 
-We set this parameter to render rivers or not.
+\new{The identical set of rivers is either rendered or hidden collectively.}
 
@@
 
-\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green, to depict the data in our cartograms.
+\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a color scheme of red, yellow, and green, to depict the data in our cartograms.
 \textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find.
@@
 We then generated another view on the right using cartograms.
-The color of both visual designs is mapped to the dataset.
+The color of both visual designs is mapped to a data dimension.
+\new{No cartograms were used more than once per participant.}
 See \Cref{fig:task} for an example.
@@
 
-\new{Due to pandemic restrictions, we designed the user study to be carried out online.
-Participants are expected to use their personal hardware for the tasks.
-However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.}
-The user study includes four parts:
+Due to pandemic restrictions, we designed the user study to be carried out online.
+Participants are expected to use their \new{retail} hardware for the tasks.
+However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.
+The \new{within-subject} user study includes four parts:
 
@@
 \textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
+A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials \footnotemark{}.
+\new{The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
+The results from these practice tasks are not included in the final analysis.}
 These tasks also serve as a quality test to see whether users are taking the study seriously.
@@
 
-\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 CCGs.
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 targetCCGs.
+\new{See the online materials for exact locations of target CCGs \footnotemark{}.}
 Response and reaction time are recorded.
@@
 
-\textbf{P4:} The participants are asked to complete a questionnaire that consists of Likert Scale questions and open-ended questions.
-
-The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.
+\textbf{P4:} The participants are asked to complete a questionnaire that consists of \new{5-Point} Likert Scale questions and open-ended questions.
+
+\footnotetext{The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.}
 
@@
     \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants.
-        The x-axis shows the target CCGs.}
+        The x-axis shows the four target CCGs, as described in Procedure P3.}
     \label{fig:task-acc}
@@
             \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants.
-                The x-axis shows the target CCGs.}
+                The x-axis shows the four target CCGs, as described in Procedure P3.}
             \label{fig:task-rt}
@@
 Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
-\new{ When the quantity of nodes and features increases, generating the optimal layout demands several hundred iterations.}
+When the quantity of nodes and features increases, generating the optimal layout demands several hundred iterations.
 
@@
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the case of river-river intersections \new{(or confluence)} and testing out more geographies such as the U.S.
+We would like to explore the case of river-river intersections (or confluence) and testing out more geographies such as the U.S.
 and Europe, which have more complex rivers.
@@
 
-\noindent\textbf{Improved User Study:} \new{Due to pandemic restrictions, the user study did not take place in a controlled environment.
-In the future, we aim to execute a more structured study, ensuring a controlled environment and standardized hardware for participants.
-}
+\noindent\textbf{Improved User Study:} Due to pandemic restrictions, the user study did not take place in a controlled environment.
+In the future, we aim to execute a more controlled laboratory study, ensuring a controlled environment and standardized hardware for participants.
 

### 2023-09-27 paper/main.tex
fix: footnote

@@
 \textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
-A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials \footnotemark{}.
+A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials\footnotemark{}.
 \new{The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
@@
 \textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 targetCCGs.
-\new{See the online materials for exact locations of target CCGs \footnotemark{}.}
+\new{See the online materials for exact locations of target CCGs\footnotemark[\value{footnote}].}
 Response and reaction time are recorded.
@@
 
-\textbf{P4:} The participants are asked to complete a questionnaire that consists of \new{5-Point} Likert Scale questions and open-ended questions.
+\textbf{P4:} The participants are asked to complete a questionnaire\footnotemark[\value{footnote}] that consists of \new{5-Point} Likert Scale questions and open-ended questions.
 

### 2023-09-28 paper/elsarticle-template-num.tex
feat: submitted

--- 2023-09-26 paper/main.tex
@@
-\documentclass[Afour,sagev,times]{sagej}
-
-\newcommand{\new}[1]{\textcolor{blue}{#1}}
-
-\usepackage{moreverb,url}
-\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow, multicol}
-
-\newcommand{\algorithmautorefname}{Algorithm}
-
+%% 
+%% Copyright 2007-2020 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
+
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
+%%
+%% 
+%%
+%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
+%%
+%%
+% \documentclass[preprint,12pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+\documentclass[5p,times]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
 \usepackage[table, dvipsnames]{xcolor}
-
-% \include{preamble/preamble.main}
 
@@
 
-\newcommand{\bobgraph}[1]{\noindent\textbf{#1}}
-
 % cite with author names
 \newcommand{\citea}[1]{\citeauthor{#1}~\cite{#1}}
+\newcommand{\bobgraph}[1]{\noindent\textbf{#1}}
 
@@
         \begin{center}
-          \includegraphics[width=0.38\columnwidth]{"#1".png}
+          \includegraphics[width=0.35\columnwidth]{"#1".png}
         \end{center}
@@
 
+% list with styles
+\usepackage{enumitem}
+
+%% The amssymb package provides various useful mathematical symbols
+\usepackage{amsmath,bm}
+\usepackage{amssymb}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
+
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
+
+\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow,multicol}
+
+\usepackage{moreverb,url}
+
+\newcommand{\algorithmautorefname}{Algorithm}
 \usepackage{algpseudocode}
@@
 \lstset{escapechar=@,style=customc}
-
-% list with styles
-\usepackage{enumitem}
-% \include{preamble/preamble.main}
-
-% \include{preamble/preamble.math}
-\usepackage{amsmath,bm}
-\usepackage{amssymb}
 
@@
 \newcommand{\RxN}{\node_{r_{x}n}}
-% \include{preamble/preamble.math}
-
-\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}
+
+\usepackage[nameinlink]{cleveref}
 
@@
 
-\usepackage[nameinlink]{cleveref}
-
-\setcounter{secnumdepth}{3}
+% \journal{Nuclear Physics B}
 
 \begin{document}
+
+\begin{frontmatter}
+
+    %% Title, authors and addresses
+
+    %% use the tnoteref command within \title for footnotes;
+    %% use the tnotetext command for theassociated footnote;
+    %% use the fnref command within \author or \address for footnotes;
+    %% use the fntext command for theassociated footnote;
+    %% use the corref command within \author for corresponding author footnotes;
+    %% use the cortext command for theassociated footnote;
+    %% use the ead command for the email address,
+    %% and the form \ead[url] for the home page:
 
@@
         The obtained results illustrate that this approach successfully retains key aspects of the original cartogram while enhancing legibility, readability, and overall accuracy.
+
     \end{abstract}
 
-\keywords{Cartogram, Demers Cartogram}
-
-\maketitle
-
+    \begin{keyword}
+        Geospatial visualization \sep Cartogram \sep Demers Cartogram
+
+    \end{keyword}
+
+\end{frontmatter}
+
+%% \linenumbers
+
+%% main text
 \section{Introduction and Motivation}
-
-\begin{samepage}
-
+\label{}
 Cartograms are representations of geographical and abstract data based on a value-by-area mapping combining statistical and geographical information \cite{dent2009Cartography,inoue2011New}.
@@
 
-    \begin{figure}[b!]
-        \centering
-        \includegraphics[width=\linewidth,keepaspectratio]{cover.png}
+\begin{figure}[htpb]
+    \centering
+    \includegraphics[width=0.8\linewidth,keepaspectratio]{cover.png}
     \caption{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
@@
 \end{figure}
-
-\end{samepage}
 
@@
 
-\bobgraph{Definitions: }While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms preserve topology, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
+\subsection{Definitions}
+
+While we focus on rectangular cartogram variants, we start with brief definitions of contiguous and non-contiguous cartograms: Contiguous cartograms preserve topology, maintain connectivity with their adjacent neighbors, but are also subject to distortion in shape.
 Non-contiguous cartograms sacrifice topological connectivity with neighbors to enable expansion or reduction in size while maintaining their polygonal shape \cite{dent2009Cartography}.
@@
                     \textbf{Number of Nodes} &
-                    \multicolumn{1}{c}{\textbf{Year}} \\
+                    \multicolumn{1}{c|}{\textbf{Year}} \\
                     \hline
@@
 \thumbnail{dorling}{
-    Dorling cartograms, as a variant of non-contiguous cartograms, generally do not preserve geography and topology.
+    \bobgraph{Dorling cartograms}, as a variant of non-contiguous cartograms, generally do not preserve geography and topology.
     A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle area \cite{dorling2011Area}.
@@
 \thumbnail{rectangular}{
-    Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
+    \bobgraph{Rectangular cartograms} are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}.
     Depending on the variant, a rectangular cartogram may trade-off between statistical and topological accuracy.
@@
 \thumbnail{mosaic}{
-    Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
+    \bobgraph{Mosaic cartograms} usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}.
     Some variants are able to preserve topological accuracy as well.
@@
 
-\bobgraph{Peer-reviewed Applications: } There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\subsection{Peer-reviewed Applications}
+
+There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
 \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
@@
 
-\bobgraph{Cartograms in Media: }Cartograms are an engaging visual representation and therefore they are a popular choice of representation in covering various topics by the media.
+\subsection{Cartograms in Media}
+
+Cartograms are an engaging visual representation and therefore they are a popular choice of representation in covering various topics by the media.
 The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
@@
 
-\bobgraph{Choropleth Shapefile:} Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
+\subsection{Choropleth Shapefile}
+
+Clinical Commissioning Groups (CCGs) are the primary administrative and geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}.
 The number of CCGs changes over time due to NHS reorganization.
@@
 
-\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England.
+\subsection{River Shapefiles}
+
+We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England.
 These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues.
@@
 
-\bobgraph{EHR Data: }We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
+\subsection{EHR Data}
+
+We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
 The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.
@@
 
-\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
+\subsection{Initialization with Rivers}
+
+We first load and (optionally) render the CCG geospatial boundaries.
 For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel.
@@
 
-\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
+\subsection{Node Layout and Overlap Removal}
+
+We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
 We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}.
@@
 
-We conduct a \new{within-subject} user study to evaluate the effectiveness of our approach.
+We conduct a within-subject user study to evaluate the effectiveness of our approach.
 We chose this type of evaluation because legibility is a human-centred characteristic.
@@
 \bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. 
-\new{The identical set of rivers is either rendered or hidden collectively.}
+The identical set of rivers is either rendered or hidden collectively.
 
@@
 The color of both visual designs is mapped to a data dimension.
-\new{No cartograms were used more than once per participant.}
+No cartograms were used more than once per participant.
 See \Cref{fig:task} for an example.
@@
 Due to pandemic restrictions, we designed the user study to be carried out online.
-Participants are expected to use their \new{retail} hardware for the tasks.
+Participants are expected to use their retail hardware for the tasks.
 However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.
-The \new{within-subject} user study includes four parts:
+The within-subject user study includes four parts:
 
@@
 The instructions are designed to help participants understand the concepts used in the tasks.
-Instructional videos are available at \url{https://www.youtube.com/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+An instructional video was provided \footnote[1]{Instructional video: \url{https://www.youtube.com/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}}.
 
 \textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
-A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials \footnotemark{}.
-\new{The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
-The results from these practice tasks are not included in the final analysis.}
+A sample task \footnote[2]{A sample task is accessible at \url{https://ghr.wangqiru.com/#/P1}.} is shown in \Cref{fig:task} and included in the online materials\footnotemark[3].
+The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
+The results from these practice tasks are not included in the final analysis.
 These tasks also serve as a quality test to see whether users are taking the study seriously.
-A demonstration of three sample tasks is also included in the instructional video.
-
-\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 targetCCGs.
-\new{See the online materials for exact locations of target CCGs \footnotemark{}.}
+A demonstration of three sample tasks is also included in the instructional video\footnotemark[1].
+
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 target CCGs.
+See the online materials for exact locations of target CCGs\footnotemark[3].
 Response and reaction time are recorded.
@@
 
-\textbf{P4:} The participants are asked to complete a questionnaire that consists of \new{5-Point} Likert Scale questions and open-ended questions.
-
-\footnotetext{The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.}
+\textbf{P4:} The participants are asked to complete a questionnaire\footnote[3]{The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.} that consists of 5-Point Likert Scale questions and open-ended questions.
 
@@
 
-\noindent\textbf{Color map choice:} The first limitation is the color map that we use to depict the data in our cartograms.
+\subsection{Color map choice}
+
+The first limitation is the color map that we use to depict the data in our cartograms.
     We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
@@
 
-\noindent\textbf{Overlap removal algorithm choice:} Another limitation is the algorithm (FNOR) we use.
+\subsection{Overlap removal algorithm choice}
+
+Another limitation is the algorithm (FNOR) we use.
 We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
@@
 
-\noindent\textbf{Generalizability:} Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
+\subsection{Generalizability}
+
+Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
@@
 
-\noindent\textbf{Improved User Study:} Due to pandemic restrictions, the user study did not take place in a controlled environment.
+\subsection{Improved User Study}
+
+Due to pandemic restrictions, the user study did not take place in a controlled environment.
 In the future, we aim to execute a more controlled laboratory study, ensuring a controlled environment and standardized hardware for participants.
@@
 
-\urlstyle{same}
-\bibliographystyle{plainnat}
+\section{Acknowledgements}
+This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC).
+EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.
+
+\bibliographystyle{elsarticle-num-names}
 \bibliography{references}
-
 \cleardoublepage
 
-\section{Appendix}
 \appendix
@@
 \end{lstlisting}
+
 \section{Pre-processing Shapefiles}
@@
 \end{document}
+\endinput
+%%
+%% End of file `elsarticle-template-num.tex'.

### 2024-04-28 paper/main.tex
fix: table border
chore: add keywords
feat: JOVI latex template

@@
-\documentclass[Afour,sagev,times]{sagej}
+\documentclass[iicol,sn-basic]{sn-jnl}% Basic Springer
+
+\usepackage{graphicx}%
+\usepackage{amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithmicx}%
+%%%%
 
@@
 \usepackage[table, dvipsnames]{xcolor}
-
-% \include{preamble/preamble.main}
 
@@
         \begin{center}
-          \includegraphics[width=0.38\columnwidth]{"#1".png}
+          \includegraphics[width=0.38\columnwidth]{figure/"#1".png}
         \end{center}
@@
 % \include{preamble/preamble.math}
-\usepackage{amsmath,bm}
-\usepackage{amssymb}
+\usepackage{amsmath,amssymb,bm}
 
@@
 
-\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}
-
 \newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
@@
 
+\usepackage{hyperref}
+
 \usepackage[nameinlink]{cleveref}
@@
 
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+%%\unnumbered% uncomment this for unnumbered level heads
+
 \begin{document}
@@
 
-\begin{abstract}
+%%=============================================================%%
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% \author*[1,2]{\fnm{Joergen W.} \spfx{van der} \sur{Ploeg} 
+%%  \sfx{IV}}\email{iauthor@gmail.com}
+%%=============================================================%%
+
+\author*[1]{\fnm{Qiru} \sur{Wang}}\email{qiru.wang1@nottingham.ac.uk}
+
+\author[1]{\fnm{Kai} \sur{Xu}}\email{kai.xu@nottingham.ac.uk}
+\equalcont{These authors contributed equally to this work.}
+
+\author[1]{\fnm{Robert} \sur{Laramee}}\email{robert.laramee@nottingham.ac.uk}
+\equalcont{These authors contributed equally to this work.}
+
+\affil[1]{\orgdiv{School of Computer Science}, \orgname{University of Nottingham}, \orgaddress{\street{University Park}, \city{Nottingham}, \postcode{NG7 2RD}, \state{England}, \country{United Kingdom}}}
+
+%%==================================%%
+%% Sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{
     Cartograms serve as representations of geographical and abstract data, employing a value-by-area mapping technique.
@@
     The obtained results illustrate that this approach successfully retains key aspects of the original cartogram while enhancing legibility, readability, and overall accuracy.
-\end{abstract}
-
-\keywords{Cartogram, Demers Cartogram}
+}
+
+%%================================%%
+%% Sample for structured abstract %%
+%%================================%%
+
+\keywords{Cartogram, Demers Cartogram, EHR Visualization, Visual Analytics}
+
+%%\pacs[JEL Classification]{D8, H51}
+
+%%\pacs[MSC Classification]{35A01, 65L10, 65L12, 65L20, 65L70}
 
@@
 \section{Introduction and Motivation}
-
-\begin{samepage}
 
@@
     \centering
-        \includegraphics[width=\linewidth,keepaspectratio]{cover.png}
+    \includegraphics[width=\linewidth,keepaspectratio]{figure/cover.png}
     \caption{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
@@
       \centering
-      \includegraphics[width=0.6\columnwidth]{ccg.png}
+      \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
       \caption{A map of 135 CCGs in England as of 2020, obtained from the Open Geography Portalx \cite{opengeographyportalxOpen} with EPSG:4326 (WGS84 - World Geodetic System) as the Coordinate Reference System (CRS).}
@@
         \centering
-        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{flowchart.png}
+        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
         \caption{An overview of our hybrid layout algorithm incorporating rivers.
@@
         \centering
-        \includegraphics[width=0.8\columnwidth,keepaspectratio]{river_resolution.png}
+        \includegraphics[width=0.8\columnwidth,keepaspectratio]{figure/river_resolution.png}
         \caption{The resolution of rivers can be dynamically adjusted by the user.
@@
 
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{figure/flowchart stalemate.png}
+    \caption{A flowchart illustration of stalemate processing.
+    See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+    \label{fig:flowchart-stalemate}
+\end{figure}
+
 \subsection{Process Stalemates}
@@
 
-{
 \begin{figure}[tb!]
     \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
-    \caption{A flowchart illustration of stalemate processing.
-    See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
-    \label{fig:flowchart-stalemate}
-\end{figure}
-}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{stalemate.png}
+    \includegraphics[width=\columnwidth,keepaspectratio]{figure/stalemate.png}
     \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
@@
 \end{figure}
-}
-
-{
+
+\subsection{Terminating the Algorithm}
+\label{subsec:{Terminating the Algorithm}}
+
 \begin{figure}[tb!]
     \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{corridor.png}
+    \includegraphics[width=\columnwidth,keepaspectratio]{figure/corridor.png}
     \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A).
@@
 \end{figure}
-}
-
-\subsection{Terminating the Algorithm}
-\label{subsec:{Terminating the Algorithm}}
 
@@
             \centering
-            \includegraphics[width=0.9\textwidth,keepaspectratio]{UI.png}
+            \includegraphics[width=0.9\textwidth,keepaspectratio]{figure/UI.png}
             \caption{A screenshot of the user interface.
@@
             \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{task.png}
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/task.png}
             \caption{A sample location task for participants.
@@
     \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
+    \includegraphics[width=\columnwidth,keepaspectratio]{figure/accuracy.png}
     \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants.
@@
             \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
+            \includegraphics[width=\columnwidth,keepaspectratio]{figure/rt.png}
             \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants.
@@
     \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{likert.png}
+    \includegraphics[width=\columnwidth,keepaspectratio]{figure/likert.png}
     \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
@@
             \centering
-            \includegraphics[width=0.8\columnwidth,keepaspectratio]{extreme.png}
+            \includegraphics[width=0.8\columnwidth,keepaspectratio]{figure/extreme.png}
             \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
@@
 
-\urlstyle{same}
-\bibliographystyle{plainnat}
-\bibliography{references}
-
-\cleardoublepage
-
-\section{Appendix}
-\appendix
+\begin{appendices}
 
@@
         \centering
-    \includegraphics[width=\columnwidth]{import_rivers.png}
+        \includegraphics[width=\columnwidth]{figure/import_rivers.png}
         \caption{QGIS interface, with River Trent, River Great Ouse, and River Thames (from top to bottom) imported.}
@@
         \centering
-    \includegraphics[width=\columnwidth]{import_ccgs.png}
+        \includegraphics[width=\columnwidth]{figure/import_ccgs.png}
         \caption{QGIS interface, with all NHS CCGs imported.}
@@
         \centering
-    \includegraphics[width=\columnwidth]{unify_crs.png}
+        \includegraphics[width=\columnwidth]{figure/unify_crs.png}
         \caption{QGIS interface, showing the unified CRS (OSGB36) for both layers.}
@@
         \centering
-    \includegraphics[width=\columnwidth]{export_rivers.png}
+        \includegraphics[width=\columnwidth]{figure/export_rivers.png}
         \caption{QGIS interface, exporting all rivers using the OSGB36 CRS in GeoJSON.}
@@
         \centering
-    \includegraphics[width=\columnwidth]{export_ccgs.png}
+        \includegraphics[width=\columnwidth]{figure/export_ccgs.png}
         \caption{QGIS interface, exporting all NHS CCGs using the OSGB36 CRS in GeoJSON.}
@@
         \centering
-    \includegraphics[width=\columnwidth]{export_topojson.png}
+        \includegraphics[width=\columnwidth]{figure/export_topojson.png}
         \caption{Mapshaper interface, merging all rivers with NHS CCGs into one layer, and export the merged layer in TopoJSON.}
@@
     \Cref{table:pre-processing_result} shows the preprocessing result.
-The reduction in file size is significant and greatly reduces the initialization time of our implementation.
+    The reduction in file size is significant and greatly reduces the initialization time of our implementation. See \Cref{table:pre-processing_result}.
 
@@
     \renewcommand{\arraystretch}{1.5}
-        \begin{table}[H]
+    \begin{table}[!tb]
         \centering
@@
 
-% \clearpage
     \section{Procedure: DeriveParallelEdge}
@@
     \end{algorithm}
-%\end{noindent}
 
@@
     \end{algorithm}
-\vfill\eject
 
@@
         \textbf{Output:} \\
-        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $.
-\\
+        A point, $ \PointP $, that is distance $ \Distance $ away from $ \EdgeStart $. \\
 
@@
     \end{algorithm}
-
-%\end{noindent}
-% \vfill\eject
 
@@
     \end{algorithm}
-
-    %\end{noindent}
-% \vfill\eject
 
@@
         \textbf{Output:} \\
-        Returns $ True $ if the node crosses a river.
-\\
+        Returns $ True $ if the node crosses a river. \\
 
@@
     \end{algorithm}
-    %\end{noindent}
-% \vfill\eject
+
+\end{appendices}
+
+\bibliography{references}
 

### 2024-07-02 paper/main.tex
draft: revision

--- 2023-09-27 paper/main.tex
@@
 
-\keywords{Cartogram, Demers Cartogram}
+\keywords{Cartogram, Demers Cartogram, EHR Visualization, Visual Analytics}
 
@@
 See \citea{nickel2022Multicriteria} for a more complete description of the advantages that Demers cartograms offer.
+\new{One of the drawbacks of Demers' cartograms is that they may become more difficult to read when a region becomes displaced from its geospatial origin through the node layout process.
+The layout of these more abstract shapes may simultaneously reduce the map's legibility and increase error.
+See \citet{tong2018Cartograms} for a more detailed explanation.}
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
@@
 In previous work, the river is static and serves merely as a boundary.
+\new{To illustrate this more clearly, we refer to the video demonstration published by \citet{tong2018Cartograms}.
+Because the behavior of the layout algorithm is dynamic, a video is more appropriate to convey the motivation of dynamic features of interest.
+We can observe the cartographic layout algorithm at 1:30 and 2:40 of the video.
+During the update process, we can clearly observe nodes cross the Thames River as the size expands.
+As the size of all the nodes expand, a number of nodes are pushed south in order to make more use of screen space.
+If we want to make the most efficient use of screen space, we need to translate the Thames River further south during the layout process in order to prevent nodes from crossing it.
+This is one of the main motivations for introducing dynamic rivers. }
 Third, we improve the algorithm to resolve stalemates.
@@
             See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
-            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
+            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.
+\new{All squares are the same size in this figure for illustration purposes.}}
         \label{fig:flowchart}
@@
 A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials\footnotemark{}.
-\new{The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
-The results from these practice tasks are not included in the final analysis.}
+The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
+The results from these practice tasks are not included in the final analysis.
 These tasks also serve as a quality test to see whether users are taking the study seriously.
@@
 
-\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 targetCCGs.
-\new{See the online materials for exact locations of target CCGs\footnotemark[\value{footnote}].}
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 target CCGs.
+See the online materials for exact locations of target CCGs.
 Response and reaction time are recorded.
@@
 
-\textbf{P4:} The participants are asked to complete a questionnaire\footnotemark[\value{footnote}] that consists of \new{5-Point} Likert Scale questions and open-ended questions.
+\textbf{P4:} The participants are asked to complete a questionnaire that consists of 5-Point Likert Scale questions and open-ended questions.
 

### 2024-07-02 paper/main.tex
fix: typo

@@
 Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
-Standard Demers cartograms are composed of square nodes that representing geographic enumeration units.
+Standard Demers cartograms are composed of square nodes that represent geographic enumeration units.
 As such, this can reduce their legibility.
@@
             See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
-            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.
-\new{All squares are the same size in this figure for illustration purposes.}}
+            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation, \new{and set all squares to the same size for clarity.}}
         \label{fig:flowchart}
@@
             \caption{A screenshot of the user interface.
-            User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers.
+            User options are provided to adjust the terminating conditions (size and error), color mapping, and visibility of nodes and rivers.
             Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
@@
 A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials\footnotemark{}.
-The practice tasks are identital to the actual tasks, accompanied by instructions provided before the actual tasks commence.
+The practice tasks are identical to the actual tasks, accompanied by instructions provided before the actual tasks commence.
 The results from these practice tasks are not included in the final analysis.
@@
 We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the case of river-river intersections (or confluence) and testing out more geographies such as the U.S.
+We would like to explore the case of river-river intersections (or confluence) and test out more geographies such as the U.S.
 and Europe, which have more complex rivers.

### 2024-07-02 paper/main.tex
Updates from Overleaf

@@
 See \citea{nickel2022Multicriteria} for a more complete description of the advantages that Demers cartograms offer.
-\new{One of the drawbacks of Demers' cartograms is that they may become more difficult to read when a region becomes displaced from its geospatial origin through the node layout process.
+\new{One of the drawbacks of Demers cartograms is that they may become more difficult to read when a region becomes displaced from its geospatial origin through the node layout process.
 The layout of these more abstract shapes may simultaneously reduce the map's legibility and increase error.
@@
 
-\bobgraph{Peer-reviewed Applications: } There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work.\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
+\bobgraph{Peer-reviewed Applications: } There has been a substantial amount of research done in this area, here we review some of the important work that has inspired our work.
+\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US.
 \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms.
@@
 We can observe the cartographic layout algorithm at 1:30 and 2:40 of the video.
-During the update process, we can clearly observe nodes cross the Thames River as the size expands.
-As the size of all the nodes expand, a number of nodes are pushed south in order to make more use of screen space.
-If we want to make the most efficient use of screen space, we need to translate the Thames River further south during the layout process in order to prevent nodes from crossing it.
+During the update process, we can clearly observe nodes cross the River Thames as the size expands.
+As the size of all the nodes expands, some nodes are pushed south to make more use of screen space.
+If we want to make the most efficient use of screen space, we need to translate the River Thames further south during the layout process to prevent nodes from crossing it.
 This is one of the main motivations for introducing dynamic rivers. }
@@
 \bobgraph{Cartograms in Media: }Cartograms are an engaging visual representation and therefore they are a popular choice of representation in covering various topics by the media.
-The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
+The Washington Post uses cartograms to visualize the US overseas economic assistance, in arms sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}.
 National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}.
@@
 
-\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the Thames River, the Trent River, and the Great Ouse River in England.
+\bobgraph{River Shapefiles:} We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for the River Thames, the Trent River, and the Great Ouse River in England.
 These rivers were chosen as they are well-known rivers, pass through regions with dense populations, and provide informative geographical and topological cues.
@@
 \bobgraph{EHR Data: }We obtained the Clinical Commissioning Group Outcomes Indicator Set (CCG OIS) from NHS Digital \cite{nhsdigitalClinical}.
+\new{See \ref{dataset} for a detailed description of the datasets used.}
 The OIS is a set of indicators that are used to measure the quality of care and the associated health outcomes in the NHS.
-Some datasets include:
+Each CCG has a unique ONS code, which is used to link the CCG shapefile with the statistical data.
+
+\section{Demers Cartogram with Rivers}
+\label{sec:main}
+
+{
+    \begin{figure*}[tb!]
+        \centering
+        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{flowchart.png}
+        \caption{An overview of our hybrid layout algorithm incorporating rivers.
+            See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail.
+            See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
+            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation, \new{and use the same size for all squares for clarity.}}
+        \label{fig:flowchart}
+    \end{figure*}
+}
+
+{
+    \begin{figure}[tb!]
+        \centering
+        \includegraphics[width=0.8\columnwidth,keepaspectratio]{river_resolution.png}
+        \caption{The resolution of rivers can be dynamically adjusted by the user.
+            (A) shows River Thames at its original resolution with 10,170 edges.
+            (B) shows the river at a reduced resolution of 49 edges.
+            We further smooth the river by removing 19 vertices in dense areas, as shown in (C).
+            The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
+        \label{fig:river resolution}
+    \end{figure}
+}
+
+\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the hybrid cartogram layout process that includes rivers.
+
+\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
+For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel.
+We then load the river shapefiles and render the rivers.
+Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex.
+We do not need the original river resolution to incorporate them into cartograms.
+We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests.
+This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}.
+We further apply simplification by removing vertices that are too close to each other.
+The initialization procedure is a one-time process that can be saved for reuse.
+
+\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
+We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}.
+Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions.
+An increase in $ \nodeSize $ can cause the nodes to overlap.
+During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position.
+Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position.
+If a node oscillates across a river, we identify this as a stalemate.
+One iteration of the layout ends when 1) no node overlap is present; AND 2) no nodes cross a river.
+We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
+The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.
+
+\subsection{River Intersection Testing}
+\label{subsec:{River Intersection Testing}}
+
+The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}.
+When a node's position changes, we test if the node's trajectory intersects any segment of a river.
+See \Cref{alg:check river intersection}.
+A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
+Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
+We label these nodes, $ \NxR $.
+
+\subsection{Translating Rivers}
+\label{subsec:{Translating Rivers}}
+
+For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $.
+Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
+We then use a heuristic to translate $ \river $ using the average vector of node intersection $$ \vectorAvg = \sum_{i=1}^{\nodeTopologicalError} \frac{\Vector{\node_i\node_{i_t}}}{\nodeTopologicalError} $$, where $ \nodeTopologicalError $ is the number of nodes intersecting the river.
+This step intends to create space for the next iteration of node translation without crossing a river.
+The detailed procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
+
+When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $.
+As a heuristic, we also translate these nodes by $ \vectorAvg $.
+The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together.
+In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
+Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm.
+We can use this to ensure smooth transitions between iterations.
+
+\subsection{Process Stalemates}
+\label{subsec:{Process Stalemates}}
+
+As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}.
+If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion.
+A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelB \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
+All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
+The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
+    \caption{A flowchart illustration of stalemate processing.
+    See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
+    \label{fig:flowchart-stalemate}
+\end{figure}
+}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{stalemate.png}
+    \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
+        The node is translated back to its previous position (iteration 2).
+        A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
+    \label{fig:stalemate}
+\end{figure}
+}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{corridor.png}
+    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A).
+        To address this, we derive a corridor (orange rectangle in (E)) based on $ \node $ and $ \nodeFNOR $.
+        All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
+        For clarity in the illustration, we place nodes sparsely in this figure.}
+    \label{fig:corridor}
+\end{figure}
+}
+
+\subsection{Terminating the Algorithm}
+\label{subsec:{Terminating the Algorithm}}
+
+The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
+We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
+$$
+
+where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
+For a detailed derivation of the formula, refer to the work by \citet{alam2015Quantitative}.
+Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
+We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers.
+We express $ \nodeTopologicalError $ in the normalized range $ \nodeTopologicalError \in [0,1] $, where $ \nodeTopologicalErrorMax $ is the case where all nodes cross a river.
+
+When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
+Every node remains on the same side of the river as its centroid.
+
+\subsection{User Options}\label{subsec:{UserOptions}}
+
+\Cref{fig:overview} presents an overview of the application, including user options.
+The user can adjust the following parameters:
+
+\bobgraph{Rendering Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
+
+\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR attributes or set to uniform.
+
+\bobgraph{Overlap Removal Speed:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.
+
+\bobgraph{Maximum Cartographic Error:} The user can adjust the maximum cartographic error, $ \nodeCartographicErrorMax $, which is used to terminate the algorithm.
+
+\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers.
+Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
+Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.
+
+\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations.
+A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.
+The default corridor length is three times the max node size.
+
+\bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
+Similarly, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
+
+    {
+        \begin{figure*}[tb!]
+            \centering
+            \includegraphics[width=0.9\textwidth,keepaspectratio]{UI.png}
+            \caption{A screenshot of the user interface.
+            User options are provided to adjust the terminating conditions (size and error), color mapping, and visibility of nodes and rivers.
+            Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
+            See \Cref{subsec:{UserOptions}} for more on user options.
+            In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
+            \label{fig:overview}
+        \end{figure*}
+    }
+
+\newcommand{\pCount}{24 }
+\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}
+
+We conduct a within-subject user study to evaluate the effectiveness of our approach.
+We chose this type of evaluation because legibility is a human-centred characteristic.
+Many different types of statistical error metrics have been evaluated in previous work \cite{nusrat2016State}, however, our focus is more user-centric in nature.
+
+\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}
+
+We formulate the following hypotheses to motivate our user study:
+
+\textbf{H1:} The introduction of rivers can improve the legibility and recognizability of cartograms.
+
+\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.
+
+\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.
+
+We test these hypotheses using location-based tasks.
+Given a standard choropleth on the left side of the screen and the corresponding cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
+See \Cref{fig:task}.
+
+\subsection{User Study Variables}\label{subsec:{User Study Variables}}
+
+In this section, we discuss the variables of our user study.
+
+\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. 
+The identical set of rivers is either rendered or hidden collectively.
+
+\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
+Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants.
+\textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.
+
+\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a color scheme of red, yellow, and green, to depict the data in our cartograms.
+\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find.
+The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
+
+\subsection{User Study Design}\label{subsec:{User Study Design}}
+
+In this section we describe the user study participants, datasets, and the experimental procedure.
+The study was approved by the University of Nottingham's Research Ethics Committee (Ref: 2021-2022-001).
+
+\bobgraph{Participants:} We recruited \pCount participants.
+
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Gender: 10 females and 14 males
+    \item Age Group: 18-24 (16), 25-29 (8)
+    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
+\end{itemize}
+
+\bobgraph{Datasets:} We used the following EHR datasets for our evaluation:
+
+\begin{itemize}
+    \setlength\itemsep{0px}
+    \item Population
+    \item Under 75 mortality from cardiovascular disease
+    \item Emergency admissions for alcohol-related liver disease
+    \item Alcohol-specific admission and readmission
+\end{itemize}
+
+135 CCGs are rendered on the screen as a choropleth on the left.
+We then generated another view on the right using cartograms.
+The color of both visual designs is mapped to a data dimension.
+No cartograms were used more than once per participant.
+See \Cref{fig:task} for an example.
+
+    {
+        \begin{figure}[tb!]
+            \centering
+            \includegraphics[width=\columnwidth,keepaspectratio]{task.png}
+            \caption{A sample location task for participants.
+                The left shows the choropleth map, and the right shows the corresponding cartogram.
+                Both images show the three longest rivers in England, and color is mapped to CCG population.
+                The target CCG blinks on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
+                In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated.
+            }
+            \label{fig:task}
+        \end{figure}
+    }
+
+\subsection{Procedure}\label{subsec:{Procedure}}
+
+Due to pandemic restrictions, we designed the user study to be carried out online.
+Participants are expected to use their retail hardware for the tasks.
+However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.
+The within-subject user study includes four parts:
+
+\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos.
+The instructions are designed to help participants understand the concepts used in the tasks.
+Instructional videos are available at \url{https://www.youtube.com/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
+
+\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
+A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials\footnotemark{}.
+The practice tasks are identical to the actual tasks, accompanied by instructions provided before the actual tasks commence.
+The results from these practice tasks are not included in the final analysis.
+These tasks also serve as a quality test to see whether users are taking the study seriously.
+A demonstration of three sample tasks is also included in the instructional video.
+
+\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 target CCGs.
+See the online materials for exact locations of target CCGs.
+Response and reaction time are recorded.
+These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
+
+\textbf{P4:} The participants are asked to complete a questionnaire that consists of 5-Point Likert Scale questions and open-ended questions.
+
+\footnotetext{The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.}
+
+\subsection{User Study Results}
+\label{subsec:{User Study Results}}
+
+{
+\begin{figure}[b!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
+    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants.
+        The x-axis shows the four target CCGs, as described in Procedure P3.}
+    \label{fig:task-acc}
+\end{figure}
+}
+
+In this section, we delve into the analysis of the preliminary user study conducted.
+Our study involved the recruitment of \pCount participants, each tasked with completing 16 location-related tasks.
+As a result, we collected a total of 384 responses.
+However, we excluded 24 responses from our analysis as they took over 60 seconds to complete.
+We hypothesize that these prolonged response times indicate potential distractions experienced by participants during the study.
+
+\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
+The chart shows that the introduction of rivers has improved the accuracy of two CCGs.
+The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002.
+This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.
+
+\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs.
+The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.
+The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38.
+However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs in the study.
+
+    {
+        \begin{figure}[t!]
+            \centering
+            \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
+            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants.
+                The x-axis shows the four target CCGs, as described in Procedure P3.}
+            \label{fig:task-rt}
+        \end{figure}
+    }
+
+\bobgraph{Participant Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
+\begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
+    \item 87.5\% of participants agree that including rivers is useful.
+    \item 83.3\% of participants agree that rivers increase the legibility of a cartogram.
+    \item 95.8\% of participants agree that including rivers makes cartograms easier to understand.
+    \item 62.5\% of participants agree that including rivers makes CCGs easier to locate.
+    \item 87.5\% of participants agree that including rivers adds value to the standard cartogram.
+\end{enumerate}
+
+{
+\begin{figure}[tb!]
+    \centering
+    \includegraphics[width=\columnwidth,keepaspectratio]{likert.png}
+    \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
+    \label{fig:likert}
+\end{figure}
+}
+
+The qualitative results clearly support our hypothesis H1 that rivers generally increase the legibility of cartograms.
+
+\section{Limitations and Future Work}
+\label{sec:{Limitations and Future Work}}
+
+In this section, we discuss some limitations of our work and future research directions.
+Because this is a new concept, it opens the door for many future research directions.
+
+\noindent\textbf{Color map choice:} The first limitation is the color map that we use to depict the data in our cartograms.
+    We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
+    However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
+    In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
+    See \Cref{fig:extreme} for an example. We plan to explore the impact of different color maps on the legibility of cartograms in future work.
+
+    {
+        \begin{figure}[tb!]
+            \centering
+            \includegraphics[width=0.8\columnwidth,keepaspectratio]{extreme.png}
+            \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
+            \label{fig:extreme}
+        \end{figure}
+    }
+
+\noindent\textbf{Overlap removal algorithm choice:} Another limitation is the algorithm (FNOR) we use.
+We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
+Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
+When the quantity of nodes and features increases, generating the optimal layout demands several hundred iterations.
+
+\noindent\textbf{Generalizability:} Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
+We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
+We would like to explore the case of river-river intersections (or confluence) and test out more geographies such as the U.S.
+and Europe, which have more complex rivers.
+We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
+
+\noindent\textbf{Improved User Study:} Due to pandemic restrictions, the user study did not take place in a controlled environment.
+In the future, we aim to execute a more controlled laboratory study, ensuring a controlled environment and standardized hardware for participants.
+
+\section{Conclusions}\label{sec:{Conclusions}}
+
+In this paper, we present a novel hybrid layout approach to generate non-contiguous cartograms with rivers.
+We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers.
+We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
+We also discuss the limitations of our work and future research directions.
+
+\urlstyle{same}
+\bibliographystyle{plainnat}
+\bibliography{references}
+
+\cleardoublepage
+
+\section{Appendix}
+\appendix
+
+\section{Dataset Description}
+\label{dataset}
 \begin{itemize}
@@
 
-Each CCG has a unique ONS code, which is used to link the CCG shapefile with the statistical data.
-
-\section{Demers Cartogram with Rivers}
-\label{sec:main}
-
-{
-    \begin{figure*}[tb!]
-        \centering
-        \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{flowchart.png}
-        \caption{An overview of our hybrid layout algorithm incorporating rivers.
-            See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail.
-            See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
-            For illustration purposes, we show the rendered views alongside the logical views representing the actual computation, \new{and set all squares to the same size for clarity.}}
-        \label{fig:flowchart}
-    \end{figure*}
-}
-
-{
-    \begin{figure}[tb!]
-        \centering
-        \includegraphics[width=0.8\columnwidth,keepaspectratio]{river_resolution.png}
-        \caption{The resolution of rivers can be dynamically adjusted by the user.
-            (A) shows River Thames at its original resolution with 10,170 edges.
-            (B) shows the river at a reduced resolution of 49 edges.
-            We further smooth the river by removing 19 vertices in dense areas, as shown in (C).
-            The reduced resolution preserves the majority of River Thames' original shape and improves the performance of our river intersection tests.}
-        \label{fig:river resolution}
-    \end{figure}
-}
-
-\Cref{alg:UpdateLayout} and \Cref{fig:flowchart} provide an overview of the hybrid cartogram layout process that includes rivers.
-
-\bobgraph{Initialization with Rivers:} We first load and (optionally) render the CCG geospatial boundaries.
-For each CCG we compute the centroid and represent it using a square node, $ \node $, with the initial size, $ \nodeSize $ =1 pixel.
-We then load the river shapefiles and render the rivers.
-Since the vertices of the river in the shapefiles are not in sequential order, we first render the starting vertex, followed by the next nearest vertex.
-We do not need the original river resolution to incorporate them into cartograms.
-We reduce their resolution to match that of the cartogram nodes in order to facilitate node-river intersection tests.
-This rendering approach enables us to adjust the river resolution as shown in \Cref{fig:river resolution}.
-We further apply simplification by removing vertices that are too close to each other.
-The initialization procedure is a one-time process that can be saved for reuse.
-
-\bobgraph{Node Layout and Overlap Removal:} We first apply the Fast Node Overlap Removal (FNOR) algorithm that solves the Variable Placement with Separation Constraints (VPSC) problem \cite{dwyer2006Fast} in order to remove overlap between square nodes.
-We chose FNOR over other node overlap removal algorithms because FNOR is capable of minimizing spread and node movement while maintaining a good level of global shape preservation \cite{chen2020Node}.
-Initialized with a pixel size of unity, we gradually increase the node size by one unit at a time to ensure smooth transitions.
-An increase in $ \nodeSize $ can cause the nodes to overlap.
-During overlap removal, we compute node trajectories (See \Cref{alg:TranslateNode}) and translate nodes to their new position.
-Nodes that cross a river, denoted $ \NxR $, are translated back to their previous position.
-If a node oscillates across a river, we identify this as a stalemate.
-One iteration of the layout ends when 1) no node overlap is present; AND 2) no nodes cross a river.
-We then increase $ \nodeSize $ by one unit and repeat the algorithm until the average cartographic error, $ \nodeCartographicError $, a measure set by the user, reaches its maximum value $ \nodeCartographicErrorMax $.
-The gradual size increase process provides stability to the layout, as can be seen in the accompanying video.
-
-\subsection{River Intersection Testing}
-\label{subsec:{River Intersection Testing}}
-
-The logic for translating the position of a node is detailed in \Cref{alg:TranslateNode}.
-When a node's position changes, we test if the node's trajectory intersects any segment of a river.
-See \Cref{alg:check river intersection}.
-A bounding box intersection test between the edge defined by node translation and river edges can be performed to reduce the number of edge intersection tests required.
-Using the intersection test, we identify all nodes that cross the river as a result of the initial FNOR algorithm.
-We label these nodes, $ \NxR $.
-
-\subsection{Translating Rivers}
-\label{subsec:{Translating Rivers}}
-
-For all nodes $ \NxR $ that cross a river, $ \river $, we compute an average vector $ \vectorAvg $ used to translate $ \river $.
-Whenever a node, $ \node $, crosses a river, we store a vector $ \nodeVectorNNn $ that points in the direction of the translation.
-We then use a heuristic to translate $ \river $ using the average vector of node intersection $$ \vectorAvg = \sum_{i=1}^{\nodeTopologicalError} \frac{\Vector{\node_i\node_{i_t}}}{\nodeTopologicalError} $$, where $ \nodeTopologicalError $ is the number of nodes intersecting the river.
-This step intends to create space for the next iteration of node translation without crossing a river.
-The detailed procedure for translating rivers is provided by \Cref{alg:TranslateRiver}.
-
-When a river, $ \river $, is translated by $ \vectorAvg $, this can trigger a scenario where nodes are crossed by a translated river, denoted $ \RxN $.
-As a heuristic, we also translate these nodes by $ \vectorAvg $.
-The reasoning behind this is that $ \vectorAvg $ indicates which direction the river needs to be translated to create space for the nodes that are too crowded together.
-In practice, $ \vectorAvg $ is multiplied by a scaling factor $ \propto\vectorAvg $.
-Thus, we can influence how far $ \river $ is translated in each iteration of the layout algorithm.
-We can use this to ensure smooth transitions between iterations.
-
-\subsection{Process Stalemates}
-\label{subsec:{Process Stalemates}}
-
-As the FNOR always attempts to produce an optimal node layout where node distribution and translation are minimized, a node's translation path can repeatedly intersect a river due to congestion, creating a stalemate situation, as shown in \Cref{fig:stalemate}.
-If a node is translated between two positions, $ \node $ and $ \nodeFNOR $, for $ \stalemateMax $ iterations (a user-adjustable parameter), we introduce a heuristic solution: constructing a corridor to alleviate congestion.
-A corridor, $ \Corridor $, is a rectangle with a width of $ \CorridorWidth $ and a length of $ \CorridorLength $, formed by deriving two edges $ \EdgeParallelA $ and $ \EdgeParallelB $ such that $ \EdgeParallelA \parallel \EdgeParallelB \parallel \nodeLineNtNc $ (See \Cref{fig:corridor}C and D).
-All nodes enclosed by $ \Corridor $ are then translated by $ \nodeVectorTC $ to alleviate the congestion (See \Cref{fig:corridor}E).
-The procedure for constructing corridors is provided by \Cref{alg:derive corridor}.
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{flowchart stalemate.png}
-    \caption{A flowchart illustration of stalemate processing.
-    See \Cref{subsec:{Process Stalemates}}, \Cref{fig:stalemate}, and \Cref{fig:corridor} for more detail.}
-    \label{fig:flowchart-stalemate}
-\end{figure}
-}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{stalemate.png}
-    \caption{A stalemate: when a node's translation path $ \nodeVectorCT $ (iteration 1) intersects a river $ \stalemateMax $ times.
-        The node is translated back to its previous position (iteration 2).
-        A stalemate can occur when the area is congested and the node cannot translate to a new position without intersecting a river.}
-    \label{fig:stalemate}
-\end{figure}
-}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{corridor.png}
-    \caption{A stalemate occurs when a node's translation path $ \nodeVectorCT $ intersects a river for $ \stalemateMax $ times, as shown in (A).
-        To address this, we derive a corridor (orange rectangle in (E)) based on $ \node $ and $ \nodeFNOR $.
-        All nodes within the corridor are translated based on $ \nodeVectorTC $, such that $ \nodeVectorNNn = \nodeVectorNinNinn $.
-        For clarity in the illustration, we place nodes sparsely in this figure.}
-    \label{fig:corridor}
-\end{figure}
-}
-
-\subsection{Terminating the Algorithm}
-\label{subsec:{Terminating the Algorithm}}
-
-The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user.
-We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
-$$
-
-where $ \node_i $ and $ \node_{i_t} $ are the initial and translated regions in the cartogram, $ \nodeList $ represent the list of regions, and $ \nodeCartographicErrorMax $ is a normalized value that we express as a percentage.
-For a detailed derivation of the formula, refer to the work by \citet{alam2015Quantitative}.
-Because the algorithm processes node-river intersections, we can measure a novel kind of error, namely, topological error $ \nodeTopologicalError $.
-We maintain $ \nodeTopologicalError = 0 $, however, we can count how many nodes would have crossed a river if we did not test for this and simply let nodes cross rivers.
-We express $ \nodeTopologicalError $ in the normalized range $ \nodeTopologicalError \in [0,1] $, where $ \nodeTopologicalErrorMax $ is the case where all nodes cross a river.
-
-When the algorithm terminates, the node layout is considered optimal where no nodes have crossed or crossed a river (denoted $ \nodeTopologicalError $ = 0 and $ \nodeCartographicError < \nodeCartographicErrorMax $).
-Every node remains on the same side of the river as its centroid.
-
-\subsection{User Options}\label{subsec:{UserOptions}}
-
-\Cref{fig:overview} presents an overview of the application, including user options.
-The user can adjust the following parameters:
-
-\bobgraph{Rendering Visibility:} The rendering visibility of various elements, including the choropleth shapefile, rivers, nodes, and node centroids, can be toggled on and off.
-
-\bobgraph{Node Mapping:} Both size and color of the nodes can be mapped to different EHR attributes or set to uniform.
-
-\bobgraph{Overlap Removal Speed:} The overlap removal process can be observed in a step-by-step manner, or the algorithm can be run automatically.
-
-\bobgraph{Maximum Cartographic Error:} The user can adjust the maximum cartographic error, $ \nodeCartographicErrorMax $, which is used to terminate the algorithm.
-
-\bobgraph{River Translation:} The behavior of rivers during the process can also be adjusted by the user: 1) Enable river crossing: this option allows nodes to cross rivers.
-Nodes cannot cross rivers by default; 2) Disable river translation: this option disables the translation of rivers, rivers are translated by default.
-Both options are useful for generating different layouts and to observe the behavior of the hybrid layout algorithm.
-
-\bobgraph{Corridor Length:} The user can define the length of a corridor that is used to resolve stalemate situations.
-A longer corridor length allows more nodes to be translated when a stalemate occurs, but may result in a less accurate layout.
-The default corridor length is three times the max node size.
-
-\bobgraph{River Thickness and Resolution:} Increasing the thickness of rivers may improve the recognizability of the cartogram.
-Similarly, increasing the resolution of rivers, at the expense of the speed of node-river intersection test, may produce a layout with higher legibility.
-
-    {
-        \begin{figure*}[tb!]
-            \centering
-            \includegraphics[width=0.9\textwidth,keepaspectratio]{UI.png}
-            \caption{A screenshot of the user interface.
-            User options are provided to adjust the terminating conditions (size and error), color mapping, and visibility of nodes and rivers.
-            Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic.
-            See \Cref{subsec:{UserOptions}} for more on user options.
-            In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
-            \label{fig:overview}
-        \end{figure*}
-    }
-
-\newcommand{\pCount}{24 }
-\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}
-
-We conduct a within-subject user study to evaluate the effectiveness of our approach.
-We chose this type of evaluation because legibility is a human-centred characteristic.
-Many different types of statistical error metrics have been evaluated in previous work \cite{nusrat2016State}, however, our focus is more user-centric in nature.
-
-\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}
-
-We formulate the following hypotheses to motivate our user study:
-
-\textbf{H1:} The introduction of rivers can improve the legibility and recognizability of cartograms.
-
-\textbf{H1.1:} The introduction of rivers can increase the accuracy of the target location.
-
-\textbf{H1.2:} The introduction of rivers can reduce the time required to locate a target region in a cartogram.
-
-We test these hypotheses using location-based tasks.
-Given a standard choropleth on the left side of the screen and the corresponding cartogram on the right side, we ask user study participants to locate the given region in the choropleth (on the left) in the cartogram (on the right).
-See \Cref{fig:task}.
-
-\subsection{User Study Variables}\label{subsec:{User Study Variables}}
-
-In this section, we discuss the variables of our user study.
-
-\bobgraph{Independent Variables: }The primary independent variable is the presence of rivers in the cartograms, which directly impacts the final layout of the cartogram. 
-The identical set of rivers is either rendered or hidden collectively.
-
-\bobgraph{Dependent Variables: }\textit{Accuracy:} Given a CCG location in the choropleth on the left, we ask the participant to locate the corresponding node in the cartogram.
-Accuracy as the primary dependent variable is measured by the number of correct CCGs located and chosen by the participants.
-\textit{Response time:} Another dependent variable is the time it takes the participants to complete each task.
-
-\bobgraph{Control Variables: }\textit{Choice of color map:} We use D3's built-in interpolateRdYlGn color map, a color scheme of red, yellow, and green, to depict the data in our cartograms.
-\textit{Communicating the target CCG to the user:} We inform the participant about the target CCG they need to find.
-The target CCG is shown in the form of a nonstop blinking (between its original color and black) area on the screen, each blink is given a duration of 2 seconds.
-
-\subsection{User Study Design}\label{subsec:{User Study Design}}
-
-In this section we describe the user study participants, datasets, and the experimental procedure.
-The study was approved by the University of Nottingham's Research Ethics Committee (Ref: 2021-2022-001).
-
-\bobgraph{Participants:} We recruited \pCount participants.
-
-\begin{itemize}
-    \setlength\itemsep{0px}
-    \item Gender: 10 females and 14 males
-    \item Age Group: 18-24 (16), 25-29 (8)
-    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
-\end{itemize}
-
-\bobgraph{Datasets:} We used the following EHR datasets for our evaluation:
-
-\begin{itemize}
-    \setlength\itemsep{0px}
-    \item Population
-    \item Under 75 mortality from cardiovascular disease
-    \item Emergency admissions for alcohol-related liver disease
-    \item Alcohol-specific admission and readmission
-\end{itemize}
-
-135 CCGs are rendered on the screen as a choropleth on the left.
-We then generated another view on the right using cartograms.
-The color of both visual designs is mapped to a data dimension.
-No cartograms were used more than once per participant.
-See \Cref{fig:task} for an example.
-
-    {
-        \begin{figure}[tb!]
-            \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{task.png}
-            \caption{A sample location task for participants.
-                The left shows the choropleth map, and the right shows the corresponding cartogram.
-                Both images show the three longest rivers in England, and color is mapped to CCG population.
-                The target CCG blinks on the choropleth (shown in black), and participants are asked to identify this CCG on the cartogram.
-                In this figure, $ \nodeCartographicErrorMax $ = 1.875\% , and an $ \nodeTopologicalError $ of 6.67\% is eliminated.
-            }
-            \label{fig:task}
-        \end{figure}
-    }
-
-\subsection{Procedure}\label{subsec:{Procedure}}
-
-Due to pandemic restrictions, we designed the user study to be carried out online.
-Participants are expected to use their retail hardware for the tasks.
-However, devices with small screens are not recommended because of the large size of the cartograms that are rendered.
-The within-subject user study includes four parts:
-
-\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos.
-The instructions are designed to help participants understand the concepts used in the tasks.
-Instructional videos are available at \url{https://www.youtube.com/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.
-
-\textbf{P2:} The participants are given three practice tasks to familiarize themselves with the user study design.
-A sample task is shown in \Cref{fig:task}, and accessible at \url{https://ghr.wangqiru.com/#/P1} and included in the online materials\footnotemark{}.
-The practice tasks are identical to the actual tasks, accompanied by instructions provided before the actual tasks commence.
-The results from these practice tasks are not included in the final analysis.
-These tasks also serve as a quality test to see whether users are taking the study seriously.
-A demonstration of three sample tasks is also included in the instructional video.
-
-\textbf{P3:} The participants are asked to complete 16 location tasks that involve 4 target CCGs.
-See the online materials for exact locations of target CCGs.
-Response and reaction time are recorded.
-These 4 CCGs are selected to avoid extreme cases (thus bias the result), in terms of size, color, and location.
-
-\textbf{P4:} The participants are asked to complete a questionnaire that consists of 5-Point Likert Scale questions and open-ended questions.
-
-\footnotetext{The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.}
-
-\subsection{User Study Results}
-\label{subsec:{User Study Results}}
-
-{
-\begin{figure}[b!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{accuracy.png}
-    \caption{In the bar chart, the y-axis shows the average accuracy rate of locating CCGs with and without rivers for user study participants.
-        The x-axis shows the four target CCGs, as described in Procedure P3.}
-    \label{fig:task-acc}
-\end{figure}
-}
-
-In this section, we delve into the analysis of the preliminary user study conducted.
-Our study involved the recruitment of \pCount participants, each tasked with completing 16 location-related tasks.
-As a result, we collected a total of 384 responses.
-However, we excluded 24 responses from our analysis as they took over 60 seconds to complete.
-We hypothesize that these prolonged response times indicate potential distractions experienced by participants during the study.
-
-\bobgraph{Accuracy:} \Cref{fig:task-acc} shows the accuracy measured by the number of correct CCGs chosen by the participants.
-The chart shows that the introduction of rivers has improved the accuracy of two CCGs.
-The accuracy for the two groups differs significantly according to the two-sample unequal variance $t$-test performed, $t$(262.98) = 3.76, $p$ = 0.0002.
-This supports our hypothesis H1.1 that the introduction of rivers increases target location accuracy.
-
-\bobgraph{Response Time:} \Cref{fig:task-rt} shows the average time to locate a CCG measured for four CCGs.
-The chart indicates that the introduction of rivers has slightly reduced the time needed to locate CCGs.
-The two groups show no statistically significant differences according to the two-sample unequal variance $t$-test performed, $t$(356.03) = 0.87, $p$ = 0.38.
-However, the average time to locate a given CCG is reduced for 3 out of the 4 CCGs users in the study.
-
-    {
-        \begin{figure}[t!]
-            \centering
-            \includegraphics[width=\columnwidth,keepaspectratio]{rt.png}
-            \caption{In the bar chart, the y-axis shows the average response time for locating CCGs with and without rivers for user study participants.
-                The x-axis shows the four target CCGs, as described in Procedure P3.}
-            \label{fig:task-rt}
-        \end{figure}
-    }
-
-\bobgraph{Participant Feedback:} \Cref{fig:likert} shows the results of the following Likert Scale questions:
-\begin{enumerate}[label=(\Alph*),align=left,leftmargin=*,labelindent=1em]
-    \item 87.5\% of participants agree that including rivers is useful.
-    \item 83.3\% of participants agree that rivers increase the legibility of a cartogram.
-    \item 95.8\% of participants agree that including rivers makes cartograms easier to understand.
-    \item 62.5\% of participants agree that including rivers makes CCGs easier to locate.
-    \item 87.5\% of participants agree that including rivers adds value to the standard cartogram.
-\end{enumerate}
-
-{
-\begin{figure}[tb!]
-    \centering
-    \includegraphics[width=\columnwidth,keepaspectratio]{likert.png}
-    \caption{The stacked bar chart shows the user study participant responses of Likert Scale questions.}
-    \label{fig:likert}
-\end{figure}
-}
-
-The qualitative results clearly support our hypothesis H1 that rivers generally increase the legibility of cartograms.
-
-\section{Limitations and Future Work}
-\label{sec:{Limitations and Future Work}}
-
-In this section, we discuss some limitations of our work and future research directions.
-Because this is a new concept, it opens the door for many future research directions.
-
-\noindent\textbf{Color map choice:} The first limitation is the color map that we use to depict the data in our cartograms.
-    We use D3's built-in interpolateRdYlGn color map, a diverging color scheme of red, yellow, and green.
-    However, we believe that the choice of color map can have a significant impact on the legibility of cartograms.
-    In the user study, we carefully avoid extreme values where the location or color of the CCG makes it easier to locate the target.
-    See \Cref{fig:extreme} for an example. We plan to explore the impact of different color maps on the legibility of cartograms in future work.
-
-    {
-        \begin{figure}[tb!]
-            \centering
-            \includegraphics[width=0.8\columnwidth,keepaspectratio]{extreme.png}
-            \caption{Due to color and relative location, we believe the CCGs in the black circle are easier to locate.}
-            \label{fig:extreme}
-        \end{figure}
-    }
-
-\noindent\textbf{Overlap removal algorithm choice:} Another limitation is the algorithm (FNOR) we use.
-We believe that developing a new algorithm with built-in constraint support can significantly reduce the time required to generate cartograms with rivers.
-Currently, the runtime of our layout algorithm is approximately 30 milliseconds for each iteration.
-When the quantity of nodes and features increases, generating the optimal layout demands several hundred iterations.
-
-\noindent\textbf{Generalizability:} Future work also includes generalizations and extensions of the algorithm, e.g., the use of other features in the cartogram layout such as additional rivers, major highways, lakes, and coastlines, etc.
-We also consider whether increasing the length of the rivers as the size of the nodes increases would be a useful option.
-We would like to explore the case of river-river intersections (or confluence) and test out more geographies such as the U.S.
-and Europe, which have more complex rivers.
-We also considered the idea of deforming the rivers as part of the layout algorithm, however, this idea is open to future work.
-
-\noindent\textbf{Improved User Study:} Due to pandemic restrictions, the user study did not take place in a controlled environment.
-In the future, we aim to execute a more controlled laboratory study, ensuring a controlled environment and standardized hardware for participants.
-
-\section{Conclusions}\label{sec:{Conclusions}}
-
-In this paper, we present a novel hybrid layout approach to generate non-contiguous cartograms with rivers.
-We first propose a new algorithm to generate cartograms with rivers, and then we present a prototype to support the exploration of cartograms with rivers.
-We evaluate our approach with a preliminary user study, and the results support our hypotheses: the introduction of rivers improves the legibility and recognizability of cartograms, although a deeper analysis is necessary to reach more confident conclusions.
-We also discuss the limitations of our work and future research directions.
-
-\urlstyle{same}
-\bibliographystyle{plainnat}
-\bibliography{references}
-
-\cleardoublepage
-
-\section{Appendix}
-\appendix
-
-\section{Obtaining Shapefile}
+\section{Obtaining Shapefiles}
 \label{obtaining-shapefile}
@@
     \begin{algorithm}[H]
-        \caption{Procedure to derive a point based an edge and a distance.}\label{alg:derive corridor point}
+        \caption{Procedure to derive a point based on an edge and a distance.}\label{alg:derive corridor point}
         \textbf{Input:} \\
@@
     \begin{algorithm}[H]
-        \caption{Procedure to derive a corridor to resolve stalemate.
+        \caption{Procedure to derive a corridor to resolve stalemates.
 We use an SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending downwards (See \Cref{fig:stalemate}).}\label{alg:derive corridor}
